### 2016-07-15 article.tex
first commit

\documentclass[journal=esthag,manuscript=article]{achemso}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage[utf8]{inputenc}
\usepackage{todonotes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Eduard Szöcs}
\affiliation[Institute for Environmental Sciences]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}
\email{szoecs@uni-landau.de}
\phone{+49 (0)6341 280 31552}

\author{Marvin Brinke}
\affiliation[German Federal Institute of Hydrology]{German Federal Institute of Hydrology (BfG), Koblenz, Germany}

\author{Bilgin Karaoglan}
\affiliation[German Federal Environmental Agency]{Federal Environmental Agency (UBA), Dessau-Roßlau, Germany}

\author{Ralf B. Schäfer}
\affiliation[University Koblenz-Landau]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[Pesticides small streams]{Pesticides pollution of small streams in Germany}
\abbreviations{mo, pest, ger, tu}
\keywords{Monitoring, Pesticides, Germany, Toxic Units, Freshwater}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Fehlt noch...
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Small waterbodies

The aim of this study was 
(i) to compile monitoring data on a national scale and to answer the questions: \\
(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
(iii)  How polluted are small streams and which pesticides are responsible?

\section{Methods}
\subsection{Data compilation}

We queried chemical monitoring data of pesticides from sampling sites with catchment size $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
Additionally, we compiled data available from previous studies and searched online databases.
This yielded to a total of more then 30 datasets of different formats.

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.

\subsection{Characterization of chemical pollution}
We characterized chemical pollution using three indicators:

\begin{enumerate}
\item National and international Environmental Quality Standards (EQS) \citep{ogewv_verordnung_2011,european_union_directive_2013}:
We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.

\item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
This is the lowest concentration at which no acceptable biological effects are expected. 
These are derived during regulatory approval  process of pesticides and contains a uncertainty factor.
The German Federal Environmental Agency provided RACs for this study.

\item Maximum Toxic Units ($\mathrm{TU_{max}}$)  \citep{sprague_measurement_1970}: 
\begin{equation}
TU_{max} = max(\frac{C_i}{EC_{50, D.magna, i}})
\end{equation}

Were $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{pesticide_action_network_pan_2015,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
\end{enumerate}

\subsection{Characterization of catchments}
We delineated catchments upstream of the sampling sites using a digital elevation model \citep{eea_digital_2013} and a multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
Catchment delineation has been manually checked for accuracy. 
In areas with low relief energy the delineation algorithm did not produce accurate results and we used river catchments provide by federal state authorities in these cases.
For each catchment we calculated the relative coverage (\%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities.

\subsection{Statistical analyses}

\section{Results}
\subsection{Overview on compiled data}

The compile dataset comprised only a small fraction of stand waters and most of the samples where sampled via grab sampling.
Therefore, we report  only results from streams and grab sampling.
We compiled a national scale dataset  comprising 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
We found big differences in the number of sampling sites between federal states.

The dataset include 484

\begin{figure}
  \includegraphics[width=.8\textwidth]{fig/fig1.pdf}
  \caption{Map of the 3109 sampling sites. Colour codes different federal states.}
  \label{fig:fig1}
\end{figure}

\subsection{Are small agricultural waters more polluted compared to bigger streams?}

\begin{figure}
  \includegraphics[width=.8\textwidth]{fig/ezg_lu.pdf}
  \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
  Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
  Colour codes the 2-dimensional density of points
  }
  \label{fig:fig_ezg_lu}
\end{figure}

\subsection{Pesticide pollution of small streams}

\section{Discussion}

Vergleich mit der Schweiz.....

\subsection{Subsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
The authors thank the authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding this project.
\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}
The following files are available free of charge.
\begin{itemize}
  \item Supplemental\_Materials.pdf : Supplemental Materials
\end{itemize}
\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{references}

\end{document}

### 2016-07-15 supplement/Supplemental_Materials.tex
first commit

\documentclass[pdftex,a4paper]{scrreprt}
\usepackage[utf8]{inputenc}       
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[
	round,	%(defaultage in the main file and \input ) for round parentheses;
	authoryear,% (default) for author-year citations;
	sort,		% orders multiple citations into the sequence in which they 
]{natbib}	
\usepackage{csvsimple}  % read csv files into Latex

\title{Supplemental Materials for the paper: \\Pesticides pollution of small streams in
Germany}
\author{Eduard Szöcs, Marvin Brinke, Bilgin Karaoglan, Ralf B. Schäfer}

\begin{document}
\maketitle

\chapter{Data Cleaning}
Each of more then 30 datasets have been cleaned and homogenized separately, before combing in a common database.
Cleaning steps comprised (Figure~\ref{fig:data_cleaning} gives a graphical overview).

\begin{enumerate}
	\item Structure: Structure has been adjusted to the database structure.
	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467) and duplicates merged.
	\item Chemicals: Chemical names and identifiers have been unified using the webchem package \citep{szocs_webchem:_2016}.
	\item  Identifiers: Unique identifiers have been assigned.
	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification have be set to zero.
	\item Other meta-data: meta-data has been standardised.
	\item Temporal resolution: The temporal resolution of the database is 1 day. Date below this resolution has been aggregated by maximum.
	\item Validity Checks: Simple rules for validity checks have been implemented (e.g. no negative concentrations).
\end{enumerate}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{data_cleaning}
\caption{Overview on data cleaning steps. After cleaning data has been stored in a relational spatial PostgreSQL database.}
\label{fig:data_cleaning}
\end{figure}

\chapter{Overview on compiled data}

\begin{table}[htbp]
\caption{Overview on chemical samples. Only data from running waters and grab sampling is shown.}
\begin{tabular}{lllrrr}
state\textsuperscript{a} & begin & end & no. sites & no. samples & no. compounds\textsuperscript{b} \\
BW & 2005-01-03 & 2014-10-02 & 118 & 4569 & 127 \\ 
BY & 2006-04-19 & 2013-12-18 & 19 & 297 & 157 \\ 
HE & 2007-01-15 & 2014-12-18 & 68 & 2512 & 144 \\ 
MV & 2005-03-08 & 2014-12-17 & 135 & 1535 & 227 \\ 
NI & 2014-03-24 & 2014-10-13 & 3 & 17 & 226 \\ 
NW & 2005-01-11 & 2015-01-22 & 1320 & 10985 & 204 \\ 
RP & 2005-01-05 & 2013-12-18 & 44 & 1277 & 278 \\ 
SH & 2005-04-26 & 2014-11-26 & 273 & 1419 & 180 \\ 
SL & 2005-01-03 & 2013-12-09 & 6 & 420 & 57 \\ 
SN & 2005-01-02 & 2013-12-18 & 917 & 17052 & 173 \\ 
ST & 2005-01-10 & 2015-03-25 & 46 & 712 & 93 \\ 
TH & 2005-01-31 & 2014-12-10 & 100 & 1441 & 76 \\ \hline
\textbf{Total} & 2005-01-02 & 2015-03-25 & 3049 & 42236 & 484 \\  \hline
\end{tabular}
\\[1em]
\textsuperscript{a}: Abbreviation according to ISO 3166-2:DE .\\
\textsuperscript{b}: Including metabolites.
\label{tab:phch}
\end{table}

\bibliographystyle{apalike}      % basic style, author-year citations
\bibliography{references}

\end{document}

### 2016-07-18 article.tex
setup code / project

@@
 
-Small waterbodies
+% Small waterbodies
+
+% runoff during rainfall (Liess et al. 1999)
+
+% underestiamtion of grab sampling (Stehle et al. 2013).
 
@@
 (i) to compile monitoring data on a national scale and to answer the questions: \\
-(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
-(iii)  How polluted are small streams and which pesticides are responsible?
+(ii) Is the data a representative description of the pollution situation? \\
+(iii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
+(iv)  How polluted are small streams and which pesticides are responsible?
 
@@
 An overview of the data cleaning process is provided in the supplemental materials.  
+The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling.
+Therefore, we report  only results for grab sampling from streams.  \todo{move to results and say that nearly no information on standing waters is available}
+To assess whether  grab samples were taken during potential rainfall events we intersected sampling coordinates and date with daily precipitation data \citep{rauthe_central_2013}.
 
@@
 \section{Results}
-\subsection{Overview on compiled data}
+\subsection{Overview and representativeness of compiled data}
 
-The compile dataset comprised only a small fraction of stand waters and most of the samples where sampled via grab sampling.
-Therefore, we report  only results from streams and grab sampling.
 We compiled a national scale dataset  comprising 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
@@
 
-\subsection{Are small agricultural waters more polluted compared to bigger streams?}
-
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Are small agricultural waters more polluted compared to bigger streams?}
 

### 2016-07-18 src/functions.R

@@ -1,0 +1,1 @@
+/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project/src/functions.R

### 2016-07-18 src/load.R

@@ -1,0 +1,1 @@
+/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project/src/load.R

### 2016-07-18 supplement/Supplemental_Materials.tex
include tabular data overview

@@
 \usepackage{xcolor}
+\usepackage{booktabs}
+
 \usepackage[
@@
 
-\begin{table}[htbp]
-\caption{Overview on chemical samples. Only data from running waters and grab sampling is shown.}
-\begin{tabular}{lllrrr}
-state\textsuperscript{a} & begin & end & no. sites & no. samples & no. compounds\textsuperscript{b} \\
-BW & 2005-01-03 & 2014-10-02 & 118 & 4569 & 127 \\ 
-BY & 2006-04-19 & 2013-12-18 & 19 & 297 & 157 \\ 
-HE & 2007-01-15 & 2014-12-18 & 68 & 2512 & 144 \\ 
-MV & 2005-03-08 & 2014-12-17 & 135 & 1535 & 227 \\ 
-NI & 2014-03-24 & 2014-10-13 & 3 & 17 & 226 \\ 
-NW & 2005-01-11 & 2015-01-22 & 1320 & 10985 & 204 \\ 
-RP & 2005-01-05 & 2013-12-18 & 44 & 1277 & 278 \\ 
-SH & 2005-04-26 & 2014-11-26 & 273 & 1419 & 180 \\ 
-SL & 2005-01-03 & 2013-12-09 & 6 & 420 & 57 \\ 
-SN & 2005-01-02 & 2013-12-18 & 917 & 17052 & 173 \\ 
-ST & 2005-01-10 & 2015-03-25 & 46 & 712 & 93 \\ 
-TH & 2005-01-31 & 2014-12-10 & 100 & 1441 & 76 \\ \hline
-\textbf{Total} & 2005-01-02 & 2015-03-25 & 3049 & 42236 & 484 \\  \hline
-\end{tabular}
-\\[1em]
-\textsuperscript{a}: Abbreviation according to ISO 3166-2:DE .\\
-\textsuperscript{b}: Including metabolites.
-\label{tab:phch}
-\end{table}
+% table generate in do_overview.R
+\input{phchoverview}
 

### 2016-07-21 article.tex
update

@@
 
+More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
+In the year 2014 more the 45000 tonnes of 766 authorized pesticides were sold for application on these areas \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001}.
+Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.
+
 % Small waterbodies
@@
 Therefore, we report  only results for grab sampling from streams.  \todo{move to results and say that nearly no information on standing waters is available}
-To assess whether  grab samples were taken during potential rainfall events we intersected sampling coordinates and date with daily precipitation data \citep{rauthe_central_2013}.
+To assess whether  grab samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before .
 
@@
 This is the lowest concentration at which no acceptable biological effects are expected. 
-These are derived during regulatory approval  process of pesticides and contains a uncertainty factor.
+These are derived during authorization  process of pesticides and contain an uncertainty factor.
 The German Federal Environmental Agency provided RACs for this study.  
+We expressed  RAC as Risk Quotient (RQ)
+
+\begin{equation}
+RQ = \frac{C}{RAC}
+\end{equation}
+
+Where $C$ is the concentration of a compound in a sample.
 
@@
 
-Were $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
+Where $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
 We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{pesticide_action_network_pan_2015,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
 We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
+A table of all included compounds can be found in the supplement.
 \end{enumerate}
@@
 \subsection{Statistical analyses}
+
+We used Multidimensional Scaling (MDS) based on jeccard dissimilarity in conjunction with hierarchical clustering to display differences in the spectra of analysed compounds per federal state.
 
@@
   \includegraphics[width=.8\textwidth]{fig/fig1.pdf}
-  \caption{Map of the 3109 sampling sites. Colour codes different federal states.}
+  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states.}
   \label{fig:fig1}
@@
   Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
-  Colour codes the 2-dimensional density of points
+  Colour codes the 2-dimensional density of points.
   }

### 2016-07-21 src/clean.R

@@ -1,0 +1,140 @@
+if (!exists("prj")) {
+  stop("You need to create a object 'prj' that points to the top folder, 
+       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est'!")
+} else {
+  source(file.path(prj, "src", "load.R"))
+}
+
+### ----------------------------------------------------------------------------
+### Code to clean data from database
+
+
+### ----------------------------------------------------------------------------
+### Load data
+
+# Load from server
+drv <- dbDriver("PostgreSQL")
+con <- dbConnect(drv, dbname = DBname, user = DBuser, host = DBhost, port = DBport)
+
+
+# samples restricted to: newer than 2005 and pesticide
+psm_samples <- dbGetQuery(con, "SELECT phch_samples.sample_id,
+    phch_samples.site_id,
+    phch_samples.date,
+    phch_samples.variable_id,
+    phch_samples.qualifier,
+    phch_samples.value,
+    phch_samples.unit,
+    phch_samples.value_fin,
+    phch_samples.loq,
+    phch_samples.lod,
+    phch_samples.sample_type,
+    phch_samples.sample_duration,
+    phch_samples.fraction,
+    phch_samples.comment,
+    phch_samples.source
+   FROM phch.phch_samples,
+    phch.phch_variables
+  WHERE phch_variables.variable_id = phch_samples.variable_id 
+                          AND date_part('year'::text, phch_samples.date) >= 2005::double precision 
+                          AND phch_variables.psm = true")
+
+psm_sites <- dbGetQuery(con, 'SELECT * FROM phch.phch_sites')
+psm_sites_info <- dbGetQuery(con, 'SELECT * FROM phch.phch_sites_info')
+psm_variables <- dbGetQuery(con, 'SELECT * FROM phch.phch_variables')
+psm_maxtu <- dbGetQuery(con, 'SELECT * FROM views.paper_maxtu')
+
+dbDisconnect(con)
+dbUnloadDriver(drv)
+
+
+# make data.table
+psm_sites[ , geom := NULL]
+setDT(psm_sites)
+setDT(psm_samples)
+setDT(psm_sites_info)
+setDT(psm_variables)
+setDT(psm_maxtu)
+
+
+
+# Clean data --------------------------------------------------------------
+
+### filter data
+
+## filter based on sample data
+# sample type
[…]

### 2016-07-21 src/do_overview.R

@@ -1,0 +1,208 @@
+if (!exists("prj")) {
+  stop("You need to create a object 'prj' that points to the top folder, 
+       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est'!")
+} else {
+  source(file.path(prj, "src", "load.R"))
+}
+
+### ----------------------------------------------------------------------------
+### Code for Overview on compiled dataset 
+
+
+# Load data ---------------------------------------------------------------
+psm_sites <- fread(file.path(cachedir, 'psm_sites.csv'))
+psm_samples <- fread(file.path(cachedir, 'psm_samples.csv'))
+psm_sites_info <- fread(file.path(cachedir, 'psm_sites_info.csv'))
+psm_variables <- fread(file.path(cachedir, 'psm_variables.csv'))
+psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
+
+
+
+# Spatial distribution ----------------------------------------------------
+# (=Figure 1)
+coordinates(psm_sites) <- ~easting + northing
+proj4string(psm_sites) <- CRS("+init=epsg:31467")
+psm_sites <- spTransform(psm_sites, CRS('+init=epsg:4326'))
+psm_sites_bb <- psm_sites@bbox
+psm_sites <- cbind(coordinates(psm_sites), psm_sites@data)
+psm_sites$state_ab <- gsub('(.*?)_.*', '\\1', psm_sites$site_id)
+
+
+library(raster)
+adm1 <- raster::getData('GADM', country = 'DE', level = 1)
+adm1 <- fortify(adm1)
+p <- ggplot() +
+  geom_polygon(data = adm1, aes(x = long, y = lat, group = group), fill = "grey90") +
+  geom_path(data = adm1, aes(x = long, y = lat, group = group), size = .3) +
+  geom_point(data = psm_sites, aes(x = easting, y = northing, col = state_ab), size = 1) +
+  theme(legend.key = element_rect(fill = 'white')) +
+  guides(colour = FALSE) +
+  labs(x = 'Lon.', y = 'Lat.') +
+  scale_color_hue(l = 50) +
+  theme_bw() +
+  coord_equal()
+p
+# ggsave(file.path(prj, "/fig/fig1.svg"),
+#        p, width = 7, height = 7)
+
+
+
+
+# Tabular overview --------------------------------------------------------
+#  (= Tab. xx in Supplement)
+psm_samples[ , state := gsub('(.*?)_.*', '\\1', site_id)]
[…]

### 2016-07-25 article.tex
first commit on new pc

@@
 
-We used Multidimensional Scaling (MDS) based on jeccard dissimilarity in conjunction with hierarchical clustering to display differences in the spectra of analysed compounds per federal state.
+We used Multidimensional Scaling (MDS) based on jaccard dissimilarity in conjunction with hierarchical clustering to display differences in the spectra of analysed compounds per federal state.
 

### 2016-07-25 article.tex
some changes

@@
 More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more the 45000 tonnes of 766 authorized pesticides were sold for application on these areas \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
-The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001}.
+In the year 2014 more the 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001,liess_determination_1999}.
 Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.
 
+National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
+These monitoring programs produce huge amounts of data, which possibly can also be used to answer other questions.
+In Germany monitoring programs are setup by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and further state specific needs.
+However, currently there is no curated national-wide compilation of this data.
+
+Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enabling downstream colonisation of 
+At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
+
+% Although small waterbodies make up most
+
+% running  water  
+% with  at  least  10km2  catchment  area
+
+% However, very little is know about the pollution of such streams. 
+% The
+
 % Small waterbodies
+% underestimation of grab sampling (Stehle et al. 2013).
 
-% runoff during rainfall (Liess et al. 1999)
-
-% underestiamtion of grab sampling (Stehle et al. 2013).
-
-The aim of this study was 
-(i) to compile monitoring data on a national scale and to answer the questions: \\
-(ii) Is the data a representative description of the pollution situation? \\
-(iii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
-(iv)  How polluted are small streams and which pesticides are responsible?
+The aim of this study was to compile monitoring data on a national scale and to answer the questions: \\
+(i) Is the currently available monitoring data a representative description of the pollution situation? \\
+(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
+(iii)  How polluted are small streams and which pesticides are responsible?
 
@@
 An overview of the data cleaning process is provided in the supplemental materials.  
-The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling.
-Therefore, we report  only results for grab sampling from streams.  \todo{move to results and say that nearly no information on standing waters is available}
-To assess whether  grab samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before .
+To assess whether samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before.
 
@@
   We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
-
   \item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
@@
 
-We compiled a national scale dataset  comprising 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
+The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling. \todo{zahlen}
+Therefore, we report only results for grab sampling from streams.
+
+The analysed dataset comprised 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
 We found big differences in the number of sampling sites between federal states.
@@
 
-Vergleich mit der Schweiz.....
+% Vergleich mit der Schweiz.....
+
+%%% Risk assessment
+% findings from field studies on pesti-
+% cide exposure and effects must be used for a retrospec-
+% tive validation of the current EU regulatory risk assess-
+% ment, particularly for its future development.
+
+% validation of the risk assessment through
+% targeted chemical and biological post-authorization
+% monitoring programs must be implemented for com-
+% pounds of concern to ensure that their application
 

### 2016-07-25 src/do_overview.R

@@ -1,6 +1,7 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est'!")
+       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
 }
@@ -114,95 +115,102 @@ options(stringsAsFactors = FALSE)
 
 # Intersect samples with precipitation data -------------------------------
 
-# path to regnie data
-regpath <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project/data/regnie/'
-# unique samplings
-samps <- unique(psm_samples[ , list(site_id, sample_id, date)])
-# join samps with coordinates
-setkey(samps, "site_id")
-setkey(psm_sites, "site_id")
-samps <- samps[psm_sites[ , list(site_id, easting, northing)]]
-
-# dates to search
-dates <- unique(samps[, date])
-container_dates <-  vector("list", length(dates)) 
-for (i in seq_along(dates)) {
-  message('Processing date', i, ' out of ', length(dates))
-  take <- dates[i]
+if (run_precip) {
+  # path to regnie data
+  # regpath <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project/data/regnie/'
+  regpath <- '/home/user/Documents/projects_git/ms_est/data/regnie'
+ 
+   # unique samplings
+  samps <- unique(psm_samples[ , list(site_id, sample_id, date)])
+  # join samps with coordinates
+  setkey(samps, "site_id")
+  setkey(psm_sites, "site_id")
+  samps <- samps[psm_sites[ , list(site_id, easting, northing)]]
   
-  # read regnie data
-  r <- try(esmisc::read_regnie(file.path(regpath, file_regnie(take))))
-  if (inherits(r, 'try-error'))
-    next
-  # r <- projectRaster(from = r, crs = CRS('+init=epsg:31467'))
-  # projectRaster() is slow - use gdalwarp instear
-  tf <- tempfile(fileext = '.tif')
-  tf2 <- tempfile(fileext = '.tif')
-  writeRaster(r, tf)
[…]

### 2016-07-25 src/paper.R

@@ -102,212 +102,6 @@ p2
 
 
 
-# ### ----------------------------------------------------------------------------
-# ### EZG - Width
-# ### ----------
-# # TH has good data 
-# th <- setDT(read_excel(file.path(datadir, 'physico_chem/TH/new/', 'Thueringen_PSM_FG_2005-14.xlsx')))
-# th <- th[ , c(8,10), with = FALSE]
-# setnames(th, c('ezg', 'width'))
-# th <- th[complete.cases(th)]
-# th[ , dataset := 'TH']
-# 
-# ### ----------
-# ### data from voss 2015
-# voss <- fread(file.path(datadir, 'misc/AG/katha', 'totalData.csv'))
-# voss <- voss[ , c(5, 34), with = FALSE]
-# setnames(voss, c('width', 'ezg'))
-# voss[ , c('width', 'ezg') := list(width = as.numeric(gsub(',', '.', width)),
-#                                   ezg = as.numeric(gsub(',', '.', ezg)))]
-# voss[ , dataset := 'Voss2015']
-# 
-# ### ----------
-# ### data from fernandez 2015
-# diego <- fread(file.path(datadir, 'misc/AG/diego', 'diego.csv'))
-# diego  <- diego[ , c(3, 2), with = FALSE]
-# setnames(diego, c('width', 'ezg'))
-# diego[ , dataset := 'Fernandez2015']
-# 
-# ### ----------
-# ### data from ufz
-# ufz <- fread(file.path(datadir, 'misc/UFZ', '2_WIDTH_EZG.txt'))
-# ufz <- ufz[ , list(width = width/100, ezg = Shape_Area)]
-# ufz[ , dataset := 'ufz']
-# ufz <- ufz[!(duplicated(ufz) | is.na(width)), ]
-# # drop observation 15 (= outlier)
-# ufz <- ufz[-15, ]
-# 
-# 
-# # a function to fit power model and export predicts
-# foo <- function(data, n = 1000){
-#   loglm <- lm(log(width) ~ log(ezg), data = data)
-#   nlm <- nls(width ~ a*ezg^b, data = data, start = list(a = exp(coef(loglm)[1]), b = coef(loglm)[2]))
-#   pdat <- data.frame(ezg = seq(1, 10000, length.out = n))
-#   pdat$width <- predict(nlm, newdata = pdat)
-#   return(pdat)
-# }
-# 
-# ew <- rbind(th, voss, diego, ufz)
[…]

### 2016-07-26 article.tex
supplement + ms

@@
 
-Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enabling downstream colonisation of 
+Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enabling downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
 At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
-
-% Although small waterbodies make up most
-
-% running  water  
-% with  at  least  10km2  catchment  area
-
-% However, very little is know about the pollution of such streams. 
-% The
-
-% Small waterbodies
-% underestimation of grab sampling (Stehle et al. 2013).
+Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007} relatively little is know about their chemical and ecological status.
 
 The aim of this study was to compile monitoring data on a national scale and to answer the questions: \\
-(i) Is the currently available monitoring data a representative description of the pollution situation? \\
-(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds? \\
-(iii)  How polluted are small streams and which pesticides are responsible?
+(i) Can the currently available monitoring data used for a representative description of the pollution situation? \\
+(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds in these relationships? \\
+(iii) baHow polluted are small streams and which pesticides are responsible?
 
@@
 
-The compiled dataset comprised only a small fraction of standing waters and most of the samples where sampled via grab sampling. \todo{zahlen}
-Therefore, we report only results for grab sampling from streams.
-
-The analysed dataset comprised 42236 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement). 
+The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
+Therefore, we report only results of grab samples from streams. 
+The analysed dataset comprised 42899 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}). 
 We found big differences in the number of sampling sites between federal states.
 
-The dataset include 484
+484 compounds that could be classified as pesticides and their metobalites (Supplement, Table \todo{Set ref to table}) .
 
@@
 
+% underestimation of grab sampling (Stehle et al. 2013).
+
 \subsection{Subsection}

### 2016-07-26 src/do_overview.R

@@ -88,6 +88,42 @@ print(xtab,
       )
 
 
+var_tab <- psm_variables[ , list(name, cas, pgroup, wrrl_zhkuqn, rak_uba, ger_auth_2015, eu_auth_2015)]
+var_tab[ , ger_auth_2015 := as.character(ger_auth_2015)]
+var_tab[ger_auth_2015 == 'FALSE', ger_auth_2015 := NA]
+var_tab[ger_auth_2015 == 'TRUE', ger_auth_2015 := 'x']
+var_tab[ , eu_auth_2015 := as.character(eu_auth_2015)]
+var_tab[eu_auth_2015 == 'FALSE', eu_auth_2015 := NA]
+var_tab[eu_auth_2015 == 'TRUE', eu_auth_2015 := 'x']
+names(var_tab) <- c('Name', 'CAS', 'Group', 'MAC-EQS\\textsuperscript{a}', 'RAC \\textsuperscript{b}', 
+                    'Auth. GER\\textsuperscript{c}', 'Auth. EU\\textsuperscript{d}')
+var_tab$Group <- gsub('organics, psm, ', '', var_tab$Group)
+# var_tab$Name <- gsub('ä', 'ae', var_tab$Name)
+# fix bug with encoding (extra space or so....)
+var_tab[Name == 'Benzoesäure', CAS := '65-85-0']
+
+var_tab_x <- xtable(var_tab, 
+                    label = 'tab:phch_var',
+                    caption = 'Analysed chemical compounds. 
+                    \\textsuperscript{a} Maximum Anual Concentration Environmental Quality Standard [ug/L].
+                    \\textsuperscript{b} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).
+                    \\textsuperscript{c} Authorized in Germany (Source: BVL, 2015). 
+                    \\textsuperscript{d} Authorized in the EU (Source: EU).',
+                    align = 'lp{4cm}rlp{1cm}p{1.5cm}p{1.5cm}p{1cm}')
+                    
+print(var_tab_x, 
+      file = file.path(prj, 'supplement/phchvar.tex'),
+      tabular.environment="longtable",
+      floating = FALSE,
+      caption.placement = 'top',
+      comment = FALSE,
+      booktabs = TRUE,
+      hline.after = c(-1, 0),
+      sanitize.text.function = identity,
+      size="\\fontsize{8pt}{10pt}\\selectfont"
+)   
+
+
 
 
[…]

### 2016-07-27 src/clean.R
completed overview

@@ -49,14 +49,13 @@ dbUnloadDriver(drv)
 
 
 # make data.table
-psm_sites[ , geom := NULL]
 setDT(psm_sites)
 setDT(psm_samples)
 setDT(psm_sites_info)
 setDT(psm_variables)
 setDT(psm_maxtu)
 
-
+psm_sites[ , geom := NULL]
 
 # Clean data --------------------------------------------------------------
 
@@ -70,7 +69,7 @@ setDT(psm_maxtu)
 table(unique(psm_samples[, list(sample_id, sample_type)])[ , sample_type], useNA = 'always')
 23277 + 19624
 100 - 4357 / (4357 + 23277 + 19624) * 100
-#! only ~ 90% of samples are grabsamples.
+#!~ 90% of samples are grabsamples.
 #! NA assumed to be individual samples
 #! keep only grab samples
 psm_samples <- psm_samples[sample_type == 'individual' | is.na(sample_type)]
@@ -95,9 +94,10 @@ psm_sites <- psm_sites[site_id %chin% unique(psm_samples$site_id)]
 table(psm_sites$lake, useNA = 'always')
 #! 58 sites form lakes (NA assumed as streams)
 psm_sites <- psm_sites[is.na(lake) | lake == FALSE]
+psm_samples <- psm_samples[site_id %in% unique(psm_sites$site_id)]
 
 
-## restrict tables to data in samples
+## restrict tables to samples
 psm_sites <- psm_sites[site_id %chin% unique(psm_samples$site_id)]
 psm_variables <- psm_variables[variable_id %in%  unique(psm_samples$variable_id)]
 psm_sites_info <- psm_sites_info[site_id %chin% unique(psm_samples$site_id)]
@@ -117,7 +117,7 @@ all(unique(psm_sites[ , site_id]) %in% unique(psm_samples[ , site_id]))
 length(unique(psm_sites_info[ , site_id])) == length(unique(psm_sites[ , site_id]))
 #! Info not for all sites available
 length(unique(psm_sites[ , site_id])) - length(unique(psm_sites_info[ , site_id]))
-#! for 316 missing
+#! for 265 missing
 psm_sites[!psm_sites[ , site_id] %in% psm_sites_info[ , site_id]]
 table(psm_sites[!psm_sites[ , site_id] %in% psm_sites_info[ , site_id], state])
 #! These are the sites were none of the both algorithms worked & no data was supplied from the states
[…]

### 2016-07-27 src/do_overview.R

@@ -35,7 +35,8 @@ adm1 <- fortify(adm1)
 p <- ggplot() +
   geom_polygon(data = adm1, aes(x = long, y = lat, group = group), fill = "grey90") +
   geom_path(data = adm1, aes(x = long, y = lat, group = group), size = .3) +
-  geom_point(data = psm_sites, aes(x = easting, y = northing, col = state_ab), size = 1) +
+  geom_point(data = psm_sites, aes(x = easting, y = northing, col = state_ab), 
+             size = 1) +
   theme(legend.key = element_rect(fill = 'white')) +
   guides(colour = FALSE) +
   labs(x = 'Lon.', y = 'Lat.') +
@@ -43,14 +44,14 @@ p <- ggplot() +
   theme_bw() +
   coord_equal()
 p
-# ggsave(file.path(prj, "/fig/fig1.svg"),
-#        p, width = 7, height = 7)
+ggsave(file.path(prj, "/fig/fig1.svg"),
+       p, width = 7, height = 7)
 
 
 
 
 # Tabular overview --------------------------------------------------------
-#  (= Tab. xx in Supplement)
+#  Sites and samples (= Tab. xx in Supplement)
 psm_samples[ , state := gsub('(.*?)_.*', '\\1', site_id)]
 psm_samples_tab <- psm_samples[, list(begin = min(date),
                                       end = max(date),
@@ -74,7 +75,8 @@ require(xtable)
 xtab <- xtable(psm_samples_tab, 
        label = 'tab:phch_overview',
       caption = 'Overview on chemical samples. Only data from running waters and grab
-sampling is shown. \\textsuperscript{a}: Abbreviations according to IS 3166-2:DE. \\textsuperscript{b}: Including metabolites',
+sampling is shown. \\textsuperscript{a}: Abbreviations according to IS 3166-2:DE. 
+      \\textsuperscript{b}: Including metabolites',
       align = 'llllrrr'
 )
 print(xtab, 
@@ -88,17 +90,21 @@ print(xtab,
       )
 
 
-var_tab <- psm_variables[ , list(name, cas, pgroup, wrrl_zhkuqn, rak_uba, ger_auth_2015, eu_auth_2015)]
[…]

### 2016-07-27 article.tex

@@
 This yielded to a total of more then 30 datasets of different formats.
+In the following we will use the ISO 3166-2:DE standard abbreviations for federal states.
 
@@
 Therefore, we report only results of grab samples from streams. 
-The analysed dataset comprised 42899 samples from 3049 sampling sites (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}). 
-We found big differences in the number of sampling sites between federal states.
-
-484 compounds that could be classified as pesticides and their metobalites (Supplement, Table \todo{Set ref to table}) .
+The analysed dataset comprised 42236 samples from 3049 sampling sites.  %see do_overview.R for numbers.
+We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).
 
@@
 
+In total 484 different compounds that could be classified as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
+Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
+We found substantial differences of the spectra of analysed compounds (Figure \ref{fig:figvar}).
+Hierarchical clustering revealed three groups of states:
+i) with less then 100 compounds (SL, ST and TH), ii) with a medium sized spectra and iii) with a big and distinct spectra (RP and NI).
+Only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day.
+
 \begin{figure}
-  \includegraphics[width=.8\textwidth]{fig/ezg_lu.pdf}
+  \includegraphics[width=\textwidth]{fig/figvar.pdf}
+  \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
+  Colors according to three groups determined by hierarchical clustering (see Supplement Figure (xxx).}
+  \label{fig:figvar}
+\end{figure}
+
+The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
+The proportion of agriculture in the catchments decreased with increasing catchment size.
+
+\begin{figure}
+  \includegraphics[width=.8\textwidth]{fig/ezglu.pdf}
   \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
@@
   }
-  \label{fig:fig_ezg_lu}
+  \label{fig:figezglu}
 \end{figure}

### 2016-07-27 article.tex
added lc50 to supplement

@@
 \subsection{Characterization of chemical pollution}
-We characterized chemical pollution using three indicators:
+We characterized chemical pollution (excluding sum parameters) using three indicators:
 
@@
   We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
+  These were available for 29 compounds (Supplement, Table xxx\todo{ref}).
+
   \item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
@@
   These are derived during authorization  process of pesticides and contain an uncertainty factor.
-  The German Federal Environmental Agency provided RACs for this study.  
-  We expressed  RAC as Risk Quotient (RQ)
-
+  The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table xxx\todo{ref}).  
+  We expressed RAC as Risk Quotient (RQ):
   \begin{equation}
@@
   \end{equation}
-
   Where $C$ is the concentration of a compound in a sample.
@@
 
-We used Multidimensional Scaling (MDS) based on jaccard dissimilarity in conjunction with hierarchical clustering to display differences in the spectra of analysed compounds per federal state.
+All data-processing and analyses have been performed using R, version 3.3.1 \citep{r_core_team_r:_2016}.
+To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
 
@@
 \end{figure}
+\todo{prettify figure!}
 
+We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. \todo{Check why there are so many missing in BW, NW, RP, SN and TH = phchsitesinfo}.
 The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
@@
 \end{figure}
+
+\todo{1-2 worte das es nicht überall möglich war ezg größe zu bestimmen.}
 
@@
 
-\subsection{Subsection}
+\subsection{Representativeness of the data}
+
+\subsection{Influence of catchment area and agriculture}
+
+\subsection{Pollution of streams}
 

### 2016-07-27 src/clean.R

@@ -43,6 +43,10 @@ psm_sites <- dbGetQuery(con, 'SELECT * FROM phch.phch_sites')
 psm_sites_info <- dbGetQuery(con, 'SELECT * FROM phch.phch_sites_info')
 psm_variables <- dbGetQuery(con, 'SELECT * FROM phch.phch_variables')
 psm_maxtu <- dbGetQuery(con, 'SELECT * FROM views.paper_maxtu')
+var_props <- dbGetQuery(con, 'SELECT variable_id,
+lc50_dm_fin, 
+lc50_dm_fin_source
+                        FROM phch.phch_variables_prop')
 
 dbDisconnect(con)
 dbUnloadDriver(drv)
@@ -54,6 +58,7 @@ setDT(psm_samples)
 setDT(psm_sites_info)
 setDT(psm_variables)
 setDT(psm_maxtu)
+setDT(var_props)
 
 psm_sites[ , geom := NULL]
 
@@ -135,6 +140,9 @@ all(unique(psm_samples[ , variable_id]) %in% unique(psm_variables[ , variable_id
 head(psm_variables)
 psm_variables[ , rak_uba_comment := NULL]
 
+# var_prop <-> variables
+var_props <- var_props[variable_id %in% psm_variables$variable_id]
+
 
 
 # Save to cache -----------------------------------------------------------
@@ -149,3 +157,6 @@ write.table(psm_variables, file.path(cachedir, 'psm_variables.csv'),
             row.names = FALSE, sep = ';')
 write.table(psm_maxtu, file.path(cachedir, 'psm_maxtu.csv'), 
             row.names = FALSE, sep = ';')
+write.table(var_props, file.path(cachedir, 'var_props.csv'), 
+            row.names = FALSE, sep = ';')
+

### 2016-07-27 src/do_overview.R

@@ -16,7 +16,7 @@ psm_samples <- fread(file.path(cachedir, 'psm_samples.csv'))
 psm_sites_info <- fread(file.path(cachedir, 'psm_sites_info.csv'))
 psm_variables <- fread(file.path(cachedir, 'psm_variables.csv'))
 psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
-
+var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 
 # Spatial distribution ----------------------------------------------------
@@ -92,17 +92,30 @@ print(xtab,
 
 # Tabular overview --------------------------------------------------------
 #  Variables (= Tab. xx in Supplement)
-var_tab <- psm_variables[ , list(name, cas, pgroup, wrrl_zhkuqn, rak_uba, 
+var_tab <- psm_variables[ , list(variable_id, name, cas, pgroup, wrrl_zhkuqn, rak_uba, 
                                  ger_auth_2015, eu_auth_2015)]
+
+# join with tu-data
+setkey(var_tab, variable_id)
+setkey(var_props, variable_id)
+var_tab <- var_props[var_tab] 
+var_tab[ , variable_id := NULL]
+
 var_tab[ , ger_auth_2015 := as.character(ger_auth_2015)]
 var_tab[ger_auth_2015 == 'FALSE', ger_auth_2015 := NA]
 var_tab[ger_auth_2015 == 'TRUE', ger_auth_2015 := 'x']
 var_tab[ , eu_auth_2015 := as.character(eu_auth_2015)]
 var_tab[eu_auth_2015 == 'FALSE', eu_auth_2015 := NA]
 var_tab[eu_auth_2015 == 'TRUE', eu_auth_2015 := 'x']
-names(var_tab) <- c('Name', 'CAS', 'Group', 'MAC-EQS\\textsuperscript{a}', 
-                    'RAC \\textsuperscript{b}', 
-                    'Auth. GER\\textsuperscript{c}', 'Auth. EU\\textsuperscript{d}')
+setcolorder(var_tab, c(names(var_tab)[3:5], names(var_tab)[8:9], names(var_tab)[1:2], names(var_tab)[6:7]))
+names(var_tab) <- c('Name', 'CAS', 'Group', 
+                    'Auth. GER\\textsuperscript{a}', 
+                    'Auth. EU\\textsuperscript{b}',
+                    'LC50\\textsubscript{D.magna}\\textsuperscript{c}',
[…]

### 2016-07-27 article.tex
added TU to supplement and numbers in ms

@@
   Where $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
-  We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{pesticide_action_network_pan_2015,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
+  We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
+  We could compile $EC_{50, D.magna}$ values for 384 compounds (Supplement, Table xxx\todo{ref}))
   We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.

### 2016-07-27 src/do_overview.R

@@ -75,7 +75,7 @@ require(xtable)
 xtab <- xtable(psm_samples_tab, 
        label = 'tab:phch_overview',
       caption = 'Overview on chemical samples. Only data from running waters and grab
-sampling is shown. \\textsuperscript{a}: Abbreviations according to IS 3166-2:DE. 
+sampling is shown. \\textsuperscript{a}: Abbreviations according to ISO 3166-2:DE. 
       \\textsuperscript{b}: Including metabolites',
       align = 'llllrrr'
 )
@@ -127,10 +127,11 @@ var_tab_x <- xtable(var_tab,
                     \\textsuperscript{a} Authorized in Germany (Source: BVL, 2015). 
                     \\textsuperscript{b} Authorized in the EU (Source: EU).
                     \\textsuperscript{c} [ug/L].
-                    \\textsuperscript{d} chemprop: Read-Across \citep{schuurmann_quantitative_2011};
-                                         epa: US EPA \citep{u.s._epa_ecotoxicology_2015};
-                                          malaj:\citep{malaj_organic_2014},
-                                          ppdb: Pesticides Properties database
+                    \\textsuperscript{d} chemprop: Read-Across \\citep{schuurmann_quantitative_2011};
+                                         epa: US EPA \\citep{u.s._epa_ecotoxicology_2015};
+                                         malaj:\\citep{malaj_organic_2014};
+                                         ppdb: Pesticides Properties database \\citep{lewis_international_2016};
+                                         none: no LC50 could be found.
                     \\textsuperscript{e} Maximum Anual Concentration Environmental Quality Standard [ug/L].
                     \\textsuperscript{f} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).',
                     align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1.5cm}p{1cm}p{1cm}p{1cm}')
@@ -155,8 +156,8 @@ psm_variables[!is.na(rak_uba)]
 # 105 compunds with RAKS
 psm_variables[!is.na(wrrl_zhkuqn)]
 # 29 compounds with MAC-EQS
-
-
+sum(!is.na(var_tab[ , 6, with = FALSE]))
+# 394 compounds with LC50 DM
 
 
 

### 2016-07-28 src/do_overview.R
added negbin gam for rac exceedances.

@@ -41,7 +41,7 @@ p <- ggplot() +
   guides(colour = FALSE) +
   labs(x = 'Lon.', y = 'Lat.') +
   scale_color_hue(l = 50) +
-  theme_bw() +
+  mytheme +
   coord_equal()
 p
 ggsave(file.path(prj, "/fig/fig1.svg"),
@@ -227,7 +227,7 @@ ylab <- paste0('Dim2 (', evar[2], '%)')
 
 p_mds <- ggplot(pco_dat, aes(x = Dim1, y = Dim2)) +
   coord_equal() +
-  theme_bw() +
+  mytheme +
   scale_colour_discrete(guide = FALSE) +
   xlab(xlab)+
   ylab(ylab) +
@@ -251,11 +251,12 @@ vwm$site_id <- factor(vwm$site_id, levels = c('SL', 'TH', 'ST', 'BY', 'SN', 'BW'
 # plot
 p_bar <- ggplot(vwm, aes(x = variable, y = site_id, fill = cols)) +
   geom_tile() +
+  mytheme
   scale_fill_manual(values = cols, name = 'gemessen') +
   theme(axis.text.x = element_blank(),
         axis.ticks.x = element_blank()) +
   labs(x = 'Compound', y = 'State') +
-  guides(fill = FALSE)
+  guides(fill = FALSE) 
 p_bar
 
 h <- c(7, 9)

### 2016-07-28 src/do_sizeagri.R

@@ -9,6 +9,9 @@ if (!exists("prj")) {
 ### ----------------------------------------------------------------------------
 ### Code for Modelling influence of catchment size and agriculture
 
+library(devtools)
+source("https://gist.githubusercontent.com/EDiLD/c4b748eb57b71c027cbfe9714222a918/raw/4cf44afb401e715548e3b2fedb4cb7d30976c8df/Deriv.R")
+
 # Load data ---------------------------------------------------------------
 psm_sites <- fread(file.path(cachedir, 'psm_sites.csv'))
 psm_samples <- fread(file.path(cachedir, 'psm_samples.csv'))
@@ -57,7 +60,7 @@ setkey(psm_sites, site_id)
 rak_exceed <- psm_sites_wi[samples_exceed]
 rak_exceed
 # 2970 sites
-
+rm(samples_raks, samples_exceed)
 rak_exceed <- rak_exceed[!(is.na(agri_fin) | is.na(ezg_fin)), ]
 # 2343 sites with ezg and agri data
 
@@ -82,6 +85,7 @@ rak_exceed$logn <- log(rak_exceed$n)
 
 
 # model using gam 
+rak_exceed$agri_fin <- rak_exceed$agri_fin*100
 library(mgcv)
 mod_p <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed, 
            family = poisson)
@@ -96,19 +100,108 @@ sum(r^2) / (mod_p$df.res)
 
 # try negative binomial model
 # with offset, automatic theta search and REML
-mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed, 
+mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed,
            family = nb())
+# offset out of formula: =ignored in predict?
+#! Check
+# mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin), offset = rak_exceed$logn, data = rak_exceed, 
+#               family = nb())
 plot(mod_nb, pages = 1)
 plot(mod_nb, pages = 1, residuals = TRUE) 
 gam.check(mod_nb)
 # overdispersion
 r <- resid(mod_nb, type = "pearson")
 sum(r^2) / (mod_nb$df.res)
+# OK
 
-#! TODO: Model visualisation
-#! derivation function?
 
[…]

### 2016-07-28 article.tex
added negbin gam for rac exceedances.
more on gam

@@
 \usepackage{todonotes}
+\usepackage{amsmath}
 
@@
 To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \todo{citation fewster}.
+
+We modeled the number of RAC exceedances (No) as:
+
+\begin{align}
+\begin{split}
+  No_i \sim NB(\mu_i, k) \\
+  E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{k} \\
+  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i) + log(n_i)
+\end{split}
+\end{align}
+
+where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
+We used the number of samplings ($n$) as an offset to account different sampling efforts. 
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
+$f_1$ and $f_2$ are smoothing functions using thin plate regression splines.
+The degree of smoothness was estimated using restricted maximum likelihood \todo{cite and check wood 2011}.
+Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $log(TU_{max})$ (see Supplement for details).
+GAM were fitted using the mgcv package \todo{citation mgcv}.
 
@@
 \begin{figure}
-  \includegraphics[width=.8\textwidth]{fig/fig1.pdf}
+  \includegraphics[width=.7\textwidth]{fig/fig1.pdf}
   \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states.}
@@
 \subsection{Are small agricultural waters more polluted compared to bigger streams?}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{fig/figrac.pdf}
+  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant increases.
+  }
+  \label{fig:figrac}
+\end{figure}
 

### 2016-07-28 src/do_sizeagri.R

@@ -100,8 +100,10 @@ sum(r^2) / (mod_p$df.res)
 
 # try negative binomial model
 # with offset, automatic theta search and REML
-mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed,
-           family = nb())
+mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset(logn), 
+              data = rak_exceed,
+              family = nb(),
+              method = 'REML')
 # offset out of formula: =ignored in predict?
 #! Check
 # mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin), offset = rak_exceed$logn, data = rak_exceed, 
@@ -195,7 +197,7 @@ p <- ggplot(pdat, aes(x = value, y = fit, group = variable)) +
   xlab('Value') +
   ylab('No. RAC exceedances') +
   ylim(c(0, 1.3))
-  
+p
 ggsave(file.path(prj, "/fig/figrac.svg"),
        p, width = 8, height = 5)
 
@@ -224,6 +226,18 @@ samples_lc50[ , logtu := ifelse(value_fin > 0, log10(value_fin / lc50_dm_fin), -
 
 # calculate per sample max(logtu)
 logtumax <- samples_lc50[ , list(logtumax = max(logtu)), by = sample_id]
+
+
+# join with data from database
+setkey(logtumax, sample_id)
+setkey(psm_maxtu, sample_id)
+jj <- psm_maxtu[logtumax]
+jj[ , diff := log_maxtu - logtumax]
+range(jj$diff)
+jj[diff > 0.1]
+#! Why the differences???
+
+
 # join back site & date
 setkey(logtumax, sample_id)
 setkey(samples_lc50, sample_id)

### 2016-07-29 article.tex
added refs

@@
 where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
-We used the number of samplings ($n$) as an offset to account different sampling efforts. 
+We used the number of samplings ($n_i$) as an offset to account different sampling efforts. 
 We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
-$f_1$ and $f_2$ are smoothing functions using thin plate regression splines.
-The degree of smoothness was estimated using restricted maximum likelihood \todo{cite and check wood 2011}.
+$f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
+The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
 Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $log(TU_{max})$ (see Supplement for details).
-GAM were fitted using the mgcv package \todo{citation mgcv}.
+GAM were fitted using the mgcv package \citep{wood_fast_2011}.
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to three groups determined by hierarchical clustering (see Supplement Figure (xxx).}
+  Colors according to three groups determined by hierarchical clustering (see Supplement Figure xxx).}
   \label{fig:figvar}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{fig/figrac.pdf}
+  \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
   \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant increases.

### 2016-07-29 src/do_sizeagri.R

@@ -86,7 +86,6 @@ rak_exceed$logn <- log(rak_exceed$n)
 
 # model using gam 
 rak_exceed$agri_fin <- rak_exceed$agri_fin*100
-library(mgcv)
 mod_p <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed, 
            family = poisson)
 plot(mod_p, pages = 1)
@@ -96,7 +95,7 @@ gam.check(mod_p)
 # overdispersion
 r <- resid(mod_p, type = "pearson")
 sum(r^2) / (mod_p$df.res)
-#! present
+#! slightly present
 
 # try negative binomial model
 # with offset, automatic theta search and REML
@@ -154,8 +153,6 @@ pdat_ezg <- melt(pdat_ezg, measure.vars = 'ezg_fin')
 names(pdat_ezg) <- c('fit', 'up', 'low', 'variable', 'value')
 pdat_ezg$variable <- as.character(pdat_ezg$variable)
 pdat <- rbind(pdat_agri, pdat_ezg)
-
-
 
 # calculate derivatives
 # see http://www.fromthebottomoftheheap.net/2014/05/15/identifying-periods-of-change-with-gams/
@@ -235,7 +232,8 @@ jj <- psm_maxtu[logtumax]
 jj[ , diff := log_maxtu - logtumax]
 range(jj$diff)
 jj[diff > 0.1]
-#! Why the differences???
+# Difference is due to HCH_gesamt as sum parameter....
+
 
 
 # join back site & date

### 2016-07-29 src/load.R

@@ -61,6 +61,7 @@ require(webchem)
 
 # modelling
 require(vegan)
+library(mgcv)
 require(gamlss)
 require(gamlss.cens)
 require(gamlss.dist)

### 2016-07-29 article.tex
rerun all, checked for smooth interaction, add raw plot of rac exceedances

@@
 National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
-These monitoring programs produce huge amounts of data, which possibly can also be used to answer other questions.
+These monitoring programs produce large amounts of data, which possibly can also be used to answer other questions.
 In Germany monitoring programs are setup by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and further state specific needs.
@@
 To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
-We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \todo{citation fewster}.
+We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
 
@@
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $log(TU_{max})$ (see Supplement for details).
+Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $log(TU_{max})$ (see Supplement for details). \todo{TRUE?}
 GAM were fitted using the mgcv package \citep{wood_fast_2011}.

### 2016-07-29 supplement/Supplemental_Materials.tex
supplement + ms
completed overview
rerun all, checked for smooth interaction, add raw plot of rac exceedances

@@
 \usepackage{xcolor}
-\usepackage{booktabs}
+\usepackage{booktabs} % table rulers
+\usepackage{longtable} % multipage tables
+\usepackage{todonotes}
 
@@
 ]{natbib}
-\usepackage{csvsimple}  % read csv files into Latex
 
@@
 \maketitle
+
+\tableofcontents
+\listoffigures
+\listoftables
 
@@
 \chapter{Overview on compiled data}
+% Overview samples
+% table generated in do_overview.R
+\input{phchoverview}
 
-% table generate in do_overview.R
-\input{phchoverview}
+\begin{figure}
+	\centering
+	\includegraphics[width = 0.8\textwidth]{varclus}
+	\caption{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states. Abbreviations of state names according to ISO 3166-2:DE.}
+	\label{fig:varclus}
+\end{figure}
+
+% Overview variables
+% table generated in do_overview.R
+\input{phchvar}
+
+\begin{figure}
+	\centering
+	\includegraphics[width = 0.75\textwidth]{precip}
+	\caption{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
+	\label{fig:precip}
+\end{figure}
+
+\chapter{Are small agricultural waters more polluted compared to bigger streams?}
+
+\begin{figure}
+	\centering
+	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
+	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points are sites without any exceedance.}
+	\label{fig:ezgagrirac}
+\end{figure}
+\todo{update refs}
 

### 2016-07-29 src/do_sizeagri.R

@@ -9,7 +9,6 @@ if (!exists("prj")) {
 ### ----------------------------------------------------------------------------
 ### Code for Modelling influence of catchment size and agriculture
 
-library(devtools)
 source("https://gist.githubusercontent.com/EDiLD/c4b748eb57b71c027cbfe9714222a918/raw/4cf44afb401e715548e3b2fedb4cb7d30976c8df/Deriv.R")
 
 # Load data ---------------------------------------------------------------
@@ -83,6 +82,16 @@ ggplot(rak_exceed, aes(x = agri_fin, y = n_exceed)) +
 
 rak_exceed$logn <- log(rak_exceed$n)
 
+p_raw <- ggplot(rak_exceed, aes(y = agri_fin, x = ezg_fin, col = log(n_exceed))) +
+  geom_point(alpha = 0.5) +
+  mytheme +
+  scale_color_gradient(low = 'blue', high = 'red') +
+  ylab("Agriculture [%]") +
+  xlab("Catchment Size [km2]") +
+  ggtitle('RAC Exceedances')
+p_raw
+ggsave(file.path(prj, "supplement/ezgagrirac.pdf"),
+       p_raw)
 
 # model using gam 
 rak_exceed$agri_fin <- rak_exceed$agri_fin*100
@@ -103,6 +112,18 @@ mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset
               data = rak_exceed,
               family = nb(),
               method = 'REML')
+
+mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin) + offset(logn), 
+              data = rak_exceed,
+              family = nb(),
+              method = 'REML')
+plot(mod_nb_te)
+vis.gam(mod_nb_te, view = c('agri_fin', 'ezg_fin'))
+anova(mod_nb, mod_nb_te, test = 'Chisq') 
+# smoothing interaction not significant and can be omited
+
+
+
 # offset out of formula: =ignored in predict?
 #! Check
 # mod_nb <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin), offset = rak_exceed$logn, data = rak_exceed, 
@@ -225,14 +246,13 @@ samples_lc50[ , logtu := ifelse(value_fin > 0, log10(value_fin / lc50_dm_fin), -
 logtumax <- samples_lc50[ , list(logtumax = max(logtu)), by = sample_id]
 
 
[…]

### 2016-07-29 src/load.R

@@ -22,7 +22,7 @@ cachedir <- file.path(prj, "cache")	# intermediate data
 #! use developmental version of data.table
 #! CRAN version should also work...
 # install.packages('devtools')
-# require(devtools)
+require(devtools)
 # install_github("Rdatatable/data.table")
 require(data.table)
 require(xlsx)    # to read .xls files

### 2016-08-10 article.tex
added model to TU.

@@
 where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
-We used the number of samplings ($n_i$) as an offset to account different sampling efforts. 
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) and used the number of sampling events per site ($n_i$) as an offset to account different sampling efforts. 
 $f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $log(TU_{max})$ (see Supplement for details). \todo{TRUE?}
-GAM were fitted using the mgcv package \citep{wood_fast_2011}.
+Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement). 
+We used pointwise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
+GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
 
@@
 \end{figure}
-\todo{prettify figure!}
+\todo{prettify figure! check cowplot package.}
 
@@
   \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
-  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant increases.
+  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+Red line marks statistically significant increases.
+Dashed lines denote 95\% Confidence Intervals.
   }

### 2016-08-10 supplement/Supplemental_Materials.tex
added model to TU.

@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
@@
 
-\begin{figure}
+\section{RAC Exceedances}
+\begin{figure}[h]
 	\centering
-	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
+	\includegraphics[width = 0.6\textwidth]{ezgagrirac}
 	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points are sites without any exceedance.}
@@
 
+\pagebreak
+\section{Toxic Units}
+
+We modeled the 95th percentile of $TU_{max}$ on a logarithmic scale assuming a Normal distribution:
+
+\begin{align}
+\begin{split}
+  log(TU_{max}) \sim Normal(\mu_i, \sigma) \\
+  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i)
+\end{split}
+\end{align}
+
+where $TU_{max i}$ is 95th percentile of $TU_{max}$ at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site.  
+$f_1$ and $f_2$ are smoothing functions.
+
+Because of samples at low polluted sites contained only values below the limit of quantification (LOQ) the derived $TU_{max}$ were also left censored (at a value of $10^{-8.75}$), thus we have
+
+\begin{align}
+  log(TU_{max~i})=\begin{cases}
+    \beta_0 + f_1(Agri_i) + f_2(Size_i) + \epsilon_i, & \text{if $log(TU_{max}) > -8.75 $}\\
+    -8.75, & \text{if $log(TU_{max}) \le -8.75 $}.
+  \end{cases}
+\end{align}
+
+Instead of assuming a fixed value for the data \textless LOQ, we use a model taking this censoring into account \citep{helsel_fabricating_2006} with the expected value being a mixture of censored and uncensored data.
+
+\begin{align}
+\begin{split}
+	E(y_i | x_i) = P(uncensored|x_i) \times E(y_i | y_i > -8.75)) + P(censored|x_i) \times -8.75\\
+	Var(y_i) = \sigma
+\end{split}
+\end{align}
+
+With $y_i$ being $log(TU_{max~i})$ and $x_i$ the predictor functions $f_1(Agri_i) + f_2(Size_i)$.
+$P(censored|x_i)$ is the probability of censoring at given predictors.
+Further, we assumed constant variance.
+This model was fitted using the gamlss packages \citep{stasinopoulos_generalized_2007}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.6\textwidth]{ezgagritu}
+	\caption{Raw data of effect of agriculture and catchment size on the 95th percentile of log(TUmax).}
+	\label{fig:ezgagritu}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.9\textwidth]{ezgagritumodel}
+	\caption{Effect of agriculture within the catchment (left) and catchment size (rigth) on the 95th percentile of log(TUmax).
+The model took censored data into account (see description above).
+Dashed lines denote 95\% Confidence Intervals.}
+	\label{fig:ezgagritu_model}
+\end{figure}
+
 \bibliographystyle{apalike}      % basic style, author-year citations

### 2016-08-10 src/do_sizeagri.R

@@ -95,8 +95,8 @@ ggsave(file.path(prj, "supplement/ezgagrirac.pdf"),
 
 # model using gam 
 rak_exceed$agri_fin <- rak_exceed$agri_fin*100
-mod_p <- gam(n_exceed ~ s(agri_fin) + s(ezg_fin) + offset(logn), data = rak_exceed, 
-           family = poisson)
+mod_p <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset(logn), data = rak_exceed, 
+           family = poisson, method = 'REML')
 plot(mod_p, pages = 1)
 plot(mod_p, pages = 1, residuals = TRUE) 
 gam.check(mod_p)
@@ -113,7 +113,7 @@ mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset
               family = nb(),
               method = 'REML')
 
-mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin) + offset(logn), 
+mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin, bs = 'tp') + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
@@ -189,12 +189,18 @@ mod_nb.dsig_agri <- signifD(pdat$value[pdat$variable == 'agri_fin'],
                             mod_nb.dci_agri[['agri_fin']]$lower)
 
 mod_nb.dci_ezg <- confint(mod_nb.d, term = 'ezg_fin')
-mod_nb.dsig_ezg <- signifD(pdat$value[pdat$variable == 'ezg_fin'], d = mod_nb.d[['ezg_fin']]$deriv,
-                            mod_nb.dci_ezg[['ezg_fin']]$upper, mod_nb.dci_ezg[['ezg_fin']]$lower)
+mod_nb.dsig_ezg <- signifD(pdat$value[pdat$variable == 'ezg_fin'], 
+                           d = mod_nb.d[['ezg_fin']]$deriv,
+                            mod_nb.dci_ezg[['ezg_fin']]$upper, 
+                           mod_nb.dci_ezg[['ezg_fin']]$lower)
 
 # check if significant
-pdat$sig[pdat$variable == 'agri_fin'] <- ifelse(!is.na(mod_nb.dsig_agri$incr) | !is.na(mod_nb.dsig_agri$decr), TRUE, FALSE)
[…]

### 2016-08-10 article.tex
eqs model

@@
 \end{split}
+
 \end{align}
@@
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement). 
+Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
 We used pointwise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
   \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
-Red line marks statistically significant increases.
+Red line marks statistically significant changes.
 Dashed lines denote 95\% Confidence Intervals.

### 2016-08-10 supplement/Supplemental_Materials.tex
eqs model

@@
 \usepackage{amsmath}
+
+% Continuous figure counting (instead of per cahpter)
+\usepackage{chngcntr}
+\counterwithout{figure}{chapter}
+\counterwithout{table}{chapter}
+% Add 'S' before figure/table numbering
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 	\centering
-	\includegraphics[width = 0.75\textwidth]{precip}
+	\includegraphics[width = 0.6\textwidth]{precip}
 	\caption{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
@@
 Further, we assumed constant variance.
-This model was fitted using the gamlss packages \citep{stasinopoulos_generalized_2007}.
+This model was fitted using the gamlss package \citep{stasinopoulos_generalized_2007}.
 
@@
 	\includegraphics[width = 0.6\textwidth]{ezgagritu}
-	\caption{Raw data of effect of agriculture and catchment size on the 95th percentile of log(TUmax).}
+	\caption{Raw data of effect of agriculture and catchment size on the 95th percentile of $log(TU_max)$.}
 	\label{fig:ezgagritu}
@@
 Dashed lines denote 95\% Confidence Intervals.}
-	\label{fig:ezgagritu_model}
+	\label{fig:ezgagritumodel}
+\end{figure}
+
+\pagebreak
+\section{EQS Exceedances}
+We modeled the number of EQS exceedances (No) the same way as RAC exceedances (see main article).
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.6\textwidth]{ezgagrieqs}
+	\caption{Raw data used for modelling EQS exceedances. Color codes the number of EQS exceedances (on a log-scale). Grey points are sites without any exceedance.}
+	\label{fig:ezgagrieqs}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.8\textwidth]{ezgagrieqsmodel}
+  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of EQS exceedances.
+Red line marks statistically significant changes.
+Dashed lines denote 95\% Confidence Intervals.
+  }
+  \label{fig:ezgeqsagrimodel}
 \end{figure}

### 2016-08-10 src/do_sizeagri.R

@@ -89,13 +89,13 @@ p_raw <- ggplot(rak_exceed, aes(y = agri_fin, x = ezg_fin, col = log(n_exceed)))
   ylab("Agriculture [%]") +
   xlab("Catchment Size [km2]") +
   ggtitle('RAC Exceedances')
-p_raw
+# p_raw
 ggsave(file.path(prj, "supplement/ezgagrirac.pdf"),
        p_raw)
 
 # model using gam 
 rak_exceed$agri_fin <- rak_exceed$agri_fin*100
-mod_p <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset(logn), data = rak_exceed, 
+mod_p <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset(logn), data = rak_exceed, 
            family = poisson, method = 'REML')
 plot(mod_p, pages = 1)
 plot(mod_p, pages = 1, residuals = TRUE) 
@@ -108,12 +108,12 @@ sum(r^2) / (mod_p$df.res)
 
 # try negative binomial model
 # with offset, automatic theta search and REML
-mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'tp') + s(ezg_fin, bs = 'tp') + offset(logn), 
+mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
 
-mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin, bs = 'tp') + offset(logn), 
+mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin, bs = 'cr') + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
@@ -221,7 +221,7 @@ p <- ggplot(pdat, aes(x = value, y = fit, group = variable)) +
   xlab('Value') +
   ylab('No. RAC exceedances') +
   ylim(c(0, 1.3))
-p
+# p
 ggsave(file.path(prj, "/fig/figrac.svg"),
        p, width = 8, height = 5)
 
@@ -335,13 +335,14 @@ mod_logno_cens <-
                          family = cens(NO, type = 'left'))
 plot(mod_logno_cens)
 
-pdf(file.path(prj, "supplement/ezgagritumodel.pdf"))
[…]

### 2016-08-10 article.tex
first draft thresholds.

@@
 
-We queried chemical monitoring data of pesticides from sampling sites with catchment size $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
+We queried chemical monitoring data of pesticides from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
 Additionally, we compiled data available from previous studies and searched online databases.
@@
   E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{k} \\
-  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i) + log(n_i)
+  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i) + log(n_i) \\
 \end{split}
-
 \end{align}
@@
 \end{figure}
-\todo{prettify figure! check cowplot package.}
 
-We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. \todo{Check why there are so many missing in BW, NW, RP, SN and TH = phchsitesinfo}.
+We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. 
 The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
@@
 
-\todo{1-2 worte das es nicht überall möglich war ezg größe zu bestimmen.}
-
 \subsection{Are small agricultural waters more polluted compared to bigger streams?}
 
-\begin{figure}
+Modelling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
+Above it levels off with a slight increase towards completely agricultural catchments (Figure \ref{fig:figrac}, left).
+This indicates thresholds at 25\%, above chemical pollution from agriculture can be detected, and 75\% above pollution increases. \todo{discussion?}
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:figrac}, right).
+
+\begin{figure} [h]
   \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
@@
 
+The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed the same patterns (see Supplement Figure Sxxx - Sxxx). \todo{references}.
+
 \subsection{Pesticide pollution of small streams}
+Based on the results previous results and give the low amount of sample sites with catchment sizes \textless 10 km\textsuperscript{2} we divided the dataset into four groups to further explore the pollution of german streams.
+
+The four groups were:
+i) small, agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textgreater 40\% agriculture);
+ii) small, non-agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textless 10\% agriculture);
+iii) medium sized, agricultural streams (30 km\textsuperscript{2} $\ge$catchment size \textless 150 km\textsuperscript{2} and \textgreater 40\% agriculture)
+iv) medium sized, non-agricultural streams (30 km\textsuperscript{2} $\ge$catchment size \textless 150 km\textsuperscript{2} and \textless 10\% agriculture)
 

### 2016-08-10 src/do_overview.R

@@ -223,8 +223,8 @@ pco1 <- wcmdscale(dp, k = 2, eig = TRUE)
 rownames(pco1$points) <- unlist(vw[ , 1, with = FALSE])
 pco_dat <- data.frame(scores(pco1), state = rownames(scores(pco1)), group = bl_groups)
 evar <- round(pco1$eig / sum(pco1$eig), 2) * 100
-xlab <- paste0('Dim1 (', evar[1], '%)')
-ylab <- paste0('Dim2 (', evar[2], '%)')
+xlab <- paste0('Axis 1 (', evar[1], '%)')
+ylab <- paste0('Axis 2 (', evar[2], '%)')
 
 p_mds <- ggplot(pco_dat, aes(x = Dim1, y = Dim2)) +
   mytheme +
@@ -233,8 +233,9 @@ p_mds <- ggplot(pco_dat, aes(x = Dim1, y = Dim2)) +
   ylab(ylab) +
   geom_vline(xintercept = 0, linetype = 'dotted') +
   geom_hline(yintercept = 0, linetype = 'dotted') +
-  geom_text(aes(label = state, col = as.factor(group))) 
-p_mds
+  geom_text(aes(label = state, col = as.factor(group)), size = 6) #+
+  # coord_equal()
+# p_mds
 
 
 # tile plot / barcode
@@ -260,12 +261,17 @@ p_bar <- ggplot(vwm, aes(x = variable, y = site_id, fill = cols)) +
   guides(fill = FALSE) 
 p_bar
 
-h <- c(7, 9)
-h <- h/sum(h)
-p <- arrangeGrob(p_bar, p_mds, ncol = 2, respect = TRUE, heights = h)
+# install.packages('cowplot')
+library(cowplot)
+p <- plot_grid(p_bar, p_mds, rel_heights = c(3, 1))
 plot(p)
+
+# h <- c(7, 9)
+# h <- h/sum(h)
+# p <- arrangeGrob(p_bar, p_mds, ncol = 2, respect = TRUE, heights = h)
+
 ggsave(file.path(prj, "/fig/figvar.svg"),
-       p, width = 12, height = 8)
+       p, width = 11, height = 7)
 #  Need to be manually cropped in inkscape
 
 

### 2016-08-10 src/do_sizeagri.R

@@ -63,6 +63,9 @@ rm(samples_raks, samples_exceed)
 rak_exceed <- rak_exceed[!(is.na(agri_fin) | is.na(ezg_fin)), ]
 # 2343 sites with ezg and agri data
 
+# logn used as offset
+rak_exceed$logn <- log(rak_exceed$n)
+
 
 ggplot(rak_exceed, aes(x = ezg_fin, y = p_exceed)) +
   geom_point() +
@@ -80,7 +83,7 @@ ggplot(rak_exceed, aes(x = agri_fin, y = n_exceed)) +
   geom_point() +
   geom_smooth()
 
-rak_exceed$logn <- log(rak_exceed$n)
+
 
 p_raw <- ggplot(rak_exceed, aes(y = agri_fin, x = ezg_fin, col = log(n_exceed))) +
   geom_point(alpha = 0.5) +
@@ -218,7 +221,7 @@ p <- ggplot(pdat, aes(x = value, y = fit, group = variable)) +
   geom_line(aes(y = sig_value), colour = 'red', lwd = 1.5) +
   facet_wrap(~variable, scales = 'free_x', labeller = mylabeller) +
   mytheme +
-  xlab('Value') +
+  xlab('') +
   ylab('No. RAC exceedances') +
   ylim(c(0, 1.3))
 # p
@@ -553,7 +556,7 @@ p <- ggplot(pdat, aes(x = value, y = fit, group = variable)) +
   geom_line(aes(y = sig_value), colour = 'red', lwd = 1.5) +
   facet_wrap(~variable, scales = 'free_x', labeller = mylabeller) +
   mytheme +
-  xlab('Value') +
+  xlab('') +
   ylab('No. EQS exceedances') 
 # p
 ggsave(file.path(prj, "/supplement/ezgagrieqsmodel.pdf"),

### 2016-08-10 src/paper.R

@@ -1,168 +1,6 @@
 
 
-# ### ----------------------------------------------------------------------------
-# # Modelling influence of EZG & Landuse / threshold detection
-# 
-# # join sites with info
-# setkey(psm_sites, site_id)
-# setkey(psm_sites_info, site_id)
-# psm_sites_wi <- psm_sites[psm_sites_info[ , list(site_id, use, ezg_fin, agri_fin)]]
-# 
-# 
-# ### Endpoint: Number of RAK execeedances
-# raks <- psm_variables[!is.na(rak_uba), list(variable_id, name, cas, pgroup, rak_uba)]
-# 
-# samples_raks <- psm_samples[variable_id %in% raks$variable]
-# setkey(samples_raks, variable_id)
-# setkey(raks, variable_id)
-# samples_raks <- raks[samples_raks]
-# samples_raks[ , g_rak_uba := value_fin > rak_uba]
-# samples_raks <- samples_raks[month(date) %in% c(5, 6, 7)]
-# 
-# rak_thrs_sites <- samples_raks[!is.na(g_rak_uba) , list(sum_g_rak_uba_abs = sum(g_rak_uba),
-#                                   sum_g_rak_uba_prop = sum(g_rak_uba) / length(g_rak_uba),
-#                                   n_samp = length(g_rak_uba)), by = site_id][order(sum_g_rak_uba_prop, decreasing = TRUE)]
-# setkey(rak_thrs_sites, site_id)
-# setkey(psm_sites, site_id)
-# 
-# w3 <- psm_sites_wi[rak_thrs_sites]
-# w3 <- w3[ezg_fin < 100 & ezg_fin > 1 & n_samp > 10]
-# 
-# ggplot(w3, aes(x = ezg_fin, y = sum_g_rak_uba_prop)) +
-#   geom_point() +
-#   geom_smooth()
-# 
-# 
-# ggplot(w3, aes(x = agri_fin, y = sum_g_rak_uba_prop)) +
-#   geom_point() +
-#   geom_smooth()
-# 
-# 
-# # model using zero inflated beta GAM
-# 
-# 
-# mdata <- na.omit(w3[ , list(ezg_fin, agri_fin, sum_g_rak_uba_prop)])
-# df <- 4
-# method <- 'GAIC'
[…]

### 2016-08-10 article.tex
started pollution section

@@
 The four groups were:
-i) small, agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textgreater 40\% agriculture);
-ii) small, non-agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textless 10\% agriculture);
-iii) medium sized, agricultural streams (30 km\textsuperscript{2} $\ge$catchment size \textless 150 km\textsuperscript{2} and \textgreater 40\% agriculture)
-iv) medium sized, non-agricultural streams (30 km\textsuperscript{2} $\ge$catchment size \textless 150 km\textsuperscript{2} and \textless 10\% agriculture)
+i) small, agricultural streams (catchment size \textless 30 km\textsuperscript{2} and agriculture \textgreater 40\%, n = 842 sites);
+ii) small, non-agricultural streams (catchment size \textless 30 km\textsuperscript{2} and agriculture \textless 10\%, n = 254 sites);
+iii) medium sized, agricultural streams (30 km\textsuperscript{2} $\le$ catchment size \textless 150 km\textsuperscript{2} and agriculture \textgreater 40\%, n = 489 sites);
+iv) medium sized, non-agricultural streams (30 km\textsuperscript{2} $\le$ catchment size \textless 150 km\textsuperscript{2} and agriculture \textless 10\%, n = 160 sites).
 

### 2016-08-11 article.tex
add numbers to overview, finalized pollution, added s(x) + s(y) + ti(x, y) to test for interaction in GAM

--- 2016-08-10 article.tex
@@
 An overview of the data cleaning process is provided in the supplemental materials.  
-To assess whether samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before.
+To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before.
+\citep{rauthe_central_2013}.
 
@@
 Therefore, we report only results of grab samples from streams. 
-The analysed dataset comprised 42236 samples from 3049 sampling sites.  %see do_overview.R for numbers.
+The analysed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
 We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).
@@
 
-In total 484 different compounds that could be classified as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
+In total 484 different compounds used as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analysed compounds (Figure \ref{fig:figvar}).
-Hierarchical clustering revealed three groups of states:
-i) with less then 100 compounds (SL, ST and TH), ii) with a medium sized spectra and iii) with a big and distinct spectra (RP and NI).
-Only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day.
+We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:figvar}).
+Hierarchical clustering revealed three groups:
+i) with less then 100 compounds (SL, ST and TH), \\
+ii) with a medium sized spectra and \\
+iii) with a big and distinct spectra (RP and NI).
+
+Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day. \todo{ref supplement}.
 
@@
 
-\todo{1-2 worte das es nicht überall möglich war ezg größe zu bestimmen.}
-
 \subsection{Are small agricultural waters more polluted compared to bigger streams?}
 
-\begin{figure}
+Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
+Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:figrac}, left).
+
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:figrac}, right) and no interaction between these two predictors.
+
+\begin{figure} [h]
   \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
@@
 
-\subsection{Pesticide pollution of small streams}
+The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement Figure Sxxx - Sxxx). \todo{references}.
+
+\subsection{Pesticide pollution of small agricultural streams in germany}
+Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites).
+
+% EQS
+% RAC
+% TU
+% Mixture
 

### 2016-08-11 supplement/Supplemental_Materials.tex
add numbers to overview, finalized pollution, added s(x) + s(y) + ti(x, y) to test for interaction in GAM

@@
 
+\section{EQS Exceedances}
+We modeled the number of EQS exceedances (No) the same way as RAC exceedances (see main article).
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.6\textwidth]{ezgagrieqs}
+	\caption{Raw data used for modelling EQS exceedances. Color codes the number of EQS exceedances (on a log-scale). Grey points denote sites without any exceedance.}
+	\label{fig:ezgagrieqs}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.8\textwidth]{ezgagrieqsmodel}
+  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of EQS exceedances.
+Red line marks statistically significant changes.
+Dashed lines denote 95\% Confidence Intervals.
+  }
+  \label{fig:ezgeqsagrimodel}
+\end{figure}
+
 \section{RAC Exceedances}
@@
 	\includegraphics[width = 0.6\textwidth]{ezgagrirac}
-	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points are sites without any exceedance.}
+	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
 	\label{fig:ezgagrirac}
@@
 With $y_i$ being $log(TU_{max~i})$ and $x_i$ the predictor functions $f_1(Agri_i) + f_2(Size_i)$.
-$P(censored|x_i)$ is the probability of censoring at given predictors.
+$P(censored|x_i)$ is the probability of censoring at given predictors and $E(y_i | y_i > -8.75))$ the expectet value given non-censoring.
 Further, we assumed constant variance.
@@
 
-\pagebreak
-\section{EQS Exceedances}
-We modeled the number of EQS exceedances (No) the same way as RAC exceedances (see main article).
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width = 0.6\textwidth]{ezgagrieqs}
-	\caption{Raw data used for modelling EQS exceedances. Color codes the number of EQS exceedances (on a log-scale). Grey points are sites without any exceedance.}
-	\label{fig:ezgagrieqs}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=0.8\textwidth]{ezgagrieqsmodel}
-  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of EQS exceedances.
-Red line marks statistically significant changes.
-Dashed lines denote 95\% Confidence Intervals.
-  }
-  \label{fig:ezgeqsagrimodel}
-\end{figure}
-
 \bibliographystyle{apalike}      % basic style, author-year citations

### 2016-08-11 src/do_overview.R

@@ -18,6 +18,19 @@ psm_variables <- fread(file.path(cachedir, 'psm_variables.csv'))
 psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
 var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
+# some numbers
+# no. observations
+(tot <- nrow(psm_samples))
+# no. sites
+length(unique(psm_samples$site_id))
+# no. compounds
+length(unique(psm_samples$variable_id))
+# values > loq
+length(psm_samples[value_fin > 0, value_fin])
+length(psm_samples[value_fin > 0, value_fin]) / tot * 100 
+
+
+
 
 # Spatial distribution ----------------------------------------------------
 # (=Figure 1)
@@ -261,8 +274,6 @@ p_bar <- ggplot(vwm, aes(x = variable, y = site_id, fill = cols)) +
   guides(fill = FALSE) 
 p_bar
 
-# install.packages('cowplot')
-library(cowplot)
 p <- plot_grid(p_bar, p_mds, rel_heights = c(3, 1))
 plot(p)
 
@@ -271,7 +282,7 @@ plot(p)
 # p <- arrangeGrob(p_bar, p_mds, ncol = 2, respect = TRUE, heights = h)
 
 ggsave(file.path(prj, "/fig/figvar.svg"),
-       p, width = 11, height = 7)
+       p, width = 11, height = 6)
 #  Need to be manually cropped in inkscape
 
 

### 2016-08-11 src/do_pollution.R

@@ -20,16 +20,143 @@ var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 # Classify stream types ---------------------------------------------------
 
-# i) small, agricultural streams (catchment size < 30 km2 and  agriculture > 40%);
-# ii) small, non-agricultural streams (catchment size < 30 km2 and agriculture < 10%);
-# iii) medium sized, agricultural streams (30 km2 <= catchment size < 150 km2 and agriculture > 40%);
-# iv) medium sized, non-agricultural streams (30 km2 <= catchment size < 150 km2 and agriculture < 10%).
+# Analyze only small agricultural streams (catchment < 30km2 and >25% agriculture)
+take_sites <- psm_sites_info[ezg_fin < 30 & agri_fin > 0.25, site_id]  
+take_samples <- psm_samples[site_id %chin% take_sites]
 
-# classify stream types
-psm_sites_info[ezg_fin < 30 & agri_fin > 0.4, type := 'small, agricultural']  # small agricultural
-psm_sites_info[ezg_fin < 30 & agri_fin < 0.1, type := 'small, non-agricultural'] # small non - agricultural
-psm_sites_info[ezg_fin >= 30 & ezg_fin < 150 & agri_fin > 0.4, type := 'mid, agricultural'] # agricultural
-psm_sites_info[ezg_fin >= 30 & ezg_fin < 150 & agri_fin < 0.1, type := 'mid, non-agricultural'] # non-agricultural
+# no. observations
+(tot <- nrow(take_samples))
+# no. sites
+length(unique(take_samples$site_id))
+# no. compounds
+length(unique(take_samples$variable_id))
+# values > loq
+length(take_samples[value_fin > 0, value_fin])
 
-table(psm_sites_info$type, useNA = 'always') # 1040 sites ommitted.
+length(take_samples[variable_id == 457, value_fin])
 
+
+# EQS as endpoint ---------------------------------------------------------
+# restrict samples to variables with eqs
+eqs <- psm_variables[!is.na(wrrl_zhkuqn), list(variable_id, name, cas, pgroup, wrrl_zhkuqn)]
+# join eqs
+setkey(take_samples, variable_id)
+setkey(eqs, variable_id)
+take_eqs <- take_samples[eqs]
+
+# check if value is greater than EQS
+take_eqs[ , g_wrrl_zhkuqn := value_fin > wrrl_zhkuqn]
[…]

### 2016-08-11 src/do_sizeagri.R

@@ -116,14 +116,15 @@ mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset
               family = nb(),
               method = 'REML')
 
-mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin, bs = 'cr') + offset(logn), 
+# model with interaction
+mod_nb_ti <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') +  ti(agri_fin, ezg_fin, bs = 'cr') + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
 plot(mod_nb_te)
 vis.gam(mod_nb_te, view = c('agri_fin', 'ezg_fin'))
-anova(mod_nb, mod_nb_te, test = 'Chisq') 
-# smoothing interaction not significant and can be omited
+anova(mod_nb, mod_nb_ti, test = 'Chisq') 
+# smoothing interaction not of interest
 
 
 
@@ -465,13 +466,13 @@ r <- resid(mod_nb, type = "pearson")
 sum(r^2) / (mod_nb$df.res)
 #! better
 
-mod_nb_te <- gam(n_exceed ~ te(agri_fin, ezg_fin) + offset(logn),
+mod_nb_ti <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + ti(agri_fin, ezg_fin) + offset(logn),
                  data = eqs_exceed,
                  family = nb(),
                  method = 'REML')
-plot(mod_nb_te)
-vis.gam(mod_nb_te, view = c('agri_fin', 'ezg_fin'))
-anova(mod_nb, mod_nb_te, test = 'Chisq')
+plot(mod_nb_ti)
+vis.gam(mod_nb_ti, view = c('agri_fin', 'ezg_fin'))
+anova(mod_nb, mod_nb_ti, test = 'Chisq')
 # smoothing interaction not significant and can be omited
 
 

### 2016-08-11 src/functions.R

@@ -361,7 +361,6 @@ mytheme <- theme_bw(base_size = 12, base_family = "Helvetica") +
         axis.text = element_text(size = 12),
         axis.title.x = element_text(size = 14, face = "bold", vjust = 0),
         axis.title.y = element_text(size = 14, face = "bold", vjust = 1),
-        legend.position = "bottom",
         legend.key = element_blank(),
         strip.background = element_blank(),
         strip.text = element_text(size = 14, face = 'bold'))

### 2016-08-11 src/load.R

@@ -53,6 +53,7 @@ require(ggmap)
 require(ggplot2)
 require(gridExtra)
 require(ggExtra)
+library(cowplot)
 
 # misc
 require(pbapply)

### 2016-08-11 article.tex
finalized results

@@
 \abbreviations{mo, pest, ger, tu}
-\keywords{Monitoring, Pesticides, Germany, Toxic Units, Freshwater}
+\keywords{Monitoring, Neonicotinoid, Germany, Toxic Units, Freshwater}
 
@@
   We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
-  We could compile $EC_{50, D.magna}$ values for 384 compounds (Supplement, Table xxx\todo{ref}))
+  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table xxx\todo{ref}))
   We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
@@
 
-\begin{figure} [h]
+\begin{figure}[h]
   \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
@@
 \subsection{Pesticide pollution of small agricultural streams in germany}
-Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites).
+Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
+%EQS
+Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
+Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:figpol}A).
+% RAC
+Insecticides showed most exceedances of RAC. 
+For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:figpol}B). 
+Highest RQ was observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
+% TU
+In 33\% of the samples no compounds no compounds were detected. 
+The mean $log(TU_{max})$ for detects was -4.6.
+2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:figpol}C).
+%Mixtures
+In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:figpol}D).
 
-% EQS
-% RAC
-% TU
-% Mixture
+\begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{fig/pall.pdf}
+  \caption{Overview on pesticide pollution of small agricultural streams. 
+  A: 10 compounds with most frequent EQS exceedances. See B for color legend.
+  B: 15 compounds with highest risk quotients.
+  C: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples).
+  D: Distribution of the number of detected compounds in a sample.
+  }
+  \label{fig:figpol}
+\end{figure}
 
@@
 
+%%% Word count
+%% Figures 
+% 3 big = 3*600 = 1800
+% 2 small = 2*300 = 600
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{itemize}
-  \item Supplemental\_Materials.pdf : Supplemental Materials
+  \item Supplemental\_Materials.pdf : Supplemental Materials (Figures, Tables, Models).
 \end{itemize}

### 2016-08-11 src/do_overview.R

@@ -85,7 +85,7 @@ psm_samples_tab <- psm_samples_tab[order(psm_samples_tab$state), ]
 names(psm_samples_tab)[1] <- 'state \\textsuperscript{a}'
 names(psm_samples_tab)[6] <- 'no.compounds \\textsuperscript{b}'
 
-require(xtable)
+
 xtab <- xtable(psm_samples_tab, 
        label = 'tab:phch_overview',
       caption = 'Overview on chemical samples. Only data from running waters and grab
@@ -491,7 +491,7 @@ ggplot() +
              size = 1) +
   theme(legend.key = element_rect(fill = 'white')) +
   labs(x = 'Lon.', y = 'Lat.') +
-  theme_bw() +
+  mytheme +
   coord_equal()
 
 table(ss[all == FALSE, substr(site_id, 1, 2)], useNA = 'always')

### 2016-08-11 src/do_pollution.R

@@ -28,6 +28,8 @@ take_samples <- psm_samples[site_id %chin% take_sites]
 (tot <- nrow(take_samples))
 # no. sites
 length(unique(take_samples$site_id))
+# no. samples
+length(unique(take_samples$sample_id))
 # no. compounds
 length(unique(take_samples$variable_id))
 # values > loq
@@ -103,6 +105,8 @@ dd <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 # calculate media per variable and take only first 15 hits
 take_var <- dd[ , list(media = max(rq)) , by = variable_id][order(media, decreasing = TRUE)][1:15, variable_id]
 dd <- dd[variable_id %in% take_var]
+dd[ , list(media = max(rq)) , by = name][order(media, decreasing = TRUE)][1:15]
+
 
 prac <- ggplot(dd) +
   geom_violin(aes(x = reorder(name, rq, FUN = median), y = rq, fill = psm_type)) +
@@ -136,6 +140,18 @@ take_lc50[ , tu := value_fin / lc50_dm_fin]
 # calculate TUmax per sample
 tumax <- take_lc50[ , list(tumax = max(tu)), by = sample_id]
 
+# number of non-detects
+nrow(tumax[tumax == 0])
+nrow(tumax)
+nrow(tumax[tumax == 0]) / nrow(tumax) * 100
+
+# number of logTU > -2
+nrow(tumax[tumax > 0.01]) / nrow(tumax) * 100
+nrow(tumax[tumax > 0.01]) / nrow(tumax[tumax != 0]) * 100
+
+# mean logTU
+mean(log10(tumax[tumax != 0, tumax]))
+
 
 ptu <- ggplot(data = tumax, aes(x = log10(tumax))) +
   geom_histogram(fill = 'grey75') +
@@ -145,11 +161,26 @@ ptu <- ggplot(data = tumax, aes(x = log10(tumax))) +
 
 
 
+
 # Mixtures ----------------------------------------------------------------
 
 # calculate the number of compounds per sample
 mix <- take_samples[ , list(no_subs = sum(value_fin > 0)), by = list(sample_id, site_id, date)]
 
+(tot <- nrow(mix))
+# no detects
+mix[no_subs == 0] # 3147 
+nrow(mix[no_subs == 0]) / tot*100
+# one compound
+mix[no_subs == 1] # 1782
[…]

### 2016-08-11 src/load.R

@@ -59,6 +59,7 @@ library(cowplot)
 require(pbapply)
 require(esmisc) # https://github.com/EDiLD/esmisc
 require(webchem)
+require(xtable)
 
 # modelling
 require(vegan)

### 2016-08-12 article.tex
restructured figures

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Pesticides pollution of small streams in Germany}
+\title[Pesticides small streams]{Pesticides in small agricultural streams in Germany}
 \abbreviations{mo, pest, ger, tu}
 \keywords{Monitoring, Neonicotinoid, Germany, Toxic Units, Freshwater}
+% RAC,
 
@@
 
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
+\includegraphics[width=0.7\textwidth]{abstract.pdf}
 
@@
 
-\begin{figure}
-  \includegraphics[width=.7\textwidth]{fig/fig1.pdf}
+\begin{figure}[ht]
+  \includegraphics[width=0.6\textwidth]{figure1.pdf}
   \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states.}
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:figvar}).
+We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:fig2}).
 Hierarchical clustering revealed three groups:
@@
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{fig/figvar.pdf}
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{figure2.pdf}
   \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
   Colors according to three groups determined by hierarchical clustering (see Supplement Figure xxx).}
-  \label{fig:figvar}
+  \label{fig:fig2}
 \end{figure}
@@
 We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. 
-The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
+The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:fig3}).
 The proportion of agriculture in the catchments decreased with increasing catchment size.
 
-\begin{figure}
-  \includegraphics[width=.8\textwidth]{fig/ezglu.pdf}
+\begin{figure}[ht]
+  \includegraphics[width=.8\textwidth]{figure3.pdf}
   \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
@@
   }
-  \label{fig:figezglu}
+  \label{fig:fig3}
 \end{figure}
@@
 Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
-Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:figrac}, left).
+Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:fig4}, left).
 
-We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:figrac}, right) and no interaction between these two predictors.
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right) and no interaction between these two predictors.
 
-\begin{figure}[h]
-  \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
+\begin{figure}[ht]
+  \includegraphics[width=0.95\textwidth]{figure4.pdf}
   \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
@@
   }
-  \label{fig:figrac}
+  \label{fig:fig4}
 \end{figure}
@@
 Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
-Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:figpol}A).
+Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
 % RAC
 Insecticides showed most exceedances of RAC. 
-For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:figpol}B). 
+For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
 Highest RQ was observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
@@
 The mean $log(TU_{max})$ for detects was -4.6.
-2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:figpol}C).
+2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
 %Mixtures
-In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:figpol}D).
+In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).
 
-\begin{figure}[h]
-  \includegraphics[width=0.95\textwidth]{fig/pall.pdf}
+\begin{figure}[ht]
+  \includegraphics[width=0.95\textwidth]{figure5.pdf}
   \caption{Overview on pesticide pollution of small agricultural streams. 
-  A: 10 compounds with most frequent EQS exceedances. See B for color legend.
-  B: 15 compounds with highest risk quotients.
-  C: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples).
-  D: Distribution of the number of detected compounds in a sample.
+  \textbf{A}: 10 compounds with most frequent EQS exceedances. See B for color legend.
+  \textbf{B}: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples).
+  \textbf{C}: 15 compounds with highest risk quotients.
+  \textbf{D}: Distribution of the number of detected compounds in a sample.
   }
-  \label{fig:figpol}
+  \label{fig:fig5}
 \end{figure}

### 2016-08-12 fig/svgtopdf.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-for var in "$@"
-do
-	echo '*****'
-	s=${var##*/}
-	base=${s%.svg}
-	echo 'converting' $var
-
-	LD_LIBRARY_PATH=$HOME/cairo-git/lib inkscape -f $var -A $base.pdf
-	echo 'done!'
-done

### 2016-08-12 src/do_overview.R

@@ -42,8 +42,7 @@ psm_sites <- cbind(coordinates(psm_sites), psm_sites@data)
 psm_sites$state_ab <- gsub('(.*?)_.*', '\\1', psm_sites$site_id)
 setDT(psm_sites)
 
-
-library(raster)
+# get administrative borders
 adm1 <- raster::getData('GADM', country = 'DE', level = 1)
 adm1 <- fortify(adm1)
 p <- ggplot() +
@@ -58,8 +57,8 @@ p <- ggplot() +
   mytheme +
   coord_equal()
 p
-ggsave(file.path(prj, "/fig/fig1.svg"),
-       p, width = 7, height = 7)
+ggsave(file.path(prj, "figure1.pdf"),
+       p, width = 3.3, height = 3, units = 'in', dpi = 300, scale = 2)
 
 
 
@@ -272,17 +271,16 @@ p_bar <- ggplot(vwm, aes(x = variable, y = site_id, fill = cols)) +
         axis.ticks.x = element_blank()) +
   labs(x = 'Compound', y = 'State') +
   guides(fill = FALSE) 
-p_bar
 
 p <- plot_grid(p_bar, p_mds, rel_heights = c(3, 1))
-plot(p)
+# plot(p)
 
 # h <- c(7, 9)
 # h <- h/sum(h)
 # p <- arrangeGrob(p_bar, p_mds, ncol = 2, respect = TRUE, heights = h)
 
-ggsave(file.path(prj, "/fig/figvar.svg"),
-       p, width = 11, height = 6)
+ggsave(file.path(prj, "figure2.pdf"),
+       p, width = 7, height = 3.5, units = 'in', dpi = 300, scale = 1.5)
 #  Need to be manually cropped in inkscape
 
 
@@ -291,7 +289,6 @@ ggsave(file.path(prj, "/fig/figvar.svg"),
 
 # Catchment size and landuse distribution ---------------------------------
 options(stringsAsFactors = TRUE) # to fix bug in stat_density2d with polygons
-library(viridis)
 ezg_lu <- ggplot(psm_sites_info[ezg_fin < 150 & !is.na(agri_fin) & !is.na(ezg_fin)], 
                  aes(x = ezg_fin, y = agri_fin * 100)) +
   stat_density2d(aes(alpha = ..level.., fill = ..level..), geom = "polygon") +
[…]

### 2016-08-12 src/do_pollution.R

@@ -190,5 +190,5 @@ pmix <- ggplot(mix, aes(x = no_subs)) +
 
 pall <- plot_grid(peqs, ptu, prac, pmix, labels = c('A', 'C', 'B', 'D'), 
                   label_size = 20)
-ggsave(file.path(prj, "/fig/pall.svg"), pall, width = 11, height = 11)
-# ggsave(file.path(prj, "/fig/pall.pdf"), pall, width = 11, height = 11)
+ggsave(file.path(prj, "figure5.pdf"), pall, width = 7, height = 7, 
+       units = 'in', dpi = 300, scale = 1.5)

### 2016-08-12 src/do_sizeagri.R

@@ -28,7 +28,9 @@ nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))])
 # join sites with info
 setkey(psm_sites, site_id)
 setkey(psm_sites_info, site_id)
-psm_sites_wi <- psm_sites[ , list(site_id, state, easting, northing)][psm_sites_info[ , list(site_id, use, ezg_fin, agri_fin)]]
+psm_sites_wi <- psm_sites[ , list(site_id, state, easting, 
+                                  northing)][psm_sites_info[ , list(site_id, 
+                                                                    use, ezg_fin, agri_fin)]]
 
 
 
@@ -98,8 +100,9 @@ ggsave(file.path(prj, "supplement/ezgagrirac.pdf"),
 
 # model using gam 
 rak_exceed$agri_fin <- rak_exceed$agri_fin*100
-mod_p <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset(logn), data = rak_exceed, 
-           family = poisson, method = 'REML')
+mod_p <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset(logn), 
+             data = rak_exceed, 
+             family = poisson, method = 'REML')
 plot(mod_p, pages = 1)
 plot(mod_p, pages = 1, residuals = TRUE) 
 gam.check(mod_p)
@@ -117,7 +120,8 @@ mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset
               method = 'REML')
 
 # model with interaction
-mod_nb_ti <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') +  ti(agri_fin, ezg_fin, bs = 'cr') + offset(logn), 
+mod_nb_ti <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') +  
+                   ti(agri_fin, ezg_fin, bs = 'cr') + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
@@ -127,11 +131,6 @@ anova(mod_nb, mod_nb_ti, test = 'Chisq')
 # smoothing interaction not of interest
 
 
-
[…]

### 2016-08-12 src/load.R

@@ -54,6 +54,7 @@ require(ggplot2)
 require(gridExtra)
 require(ggExtra)
 library(cowplot)
+library(viridis)
 
 # misc
 require(pbapply)

### 2016-08-12 article.tex
intro

@@
 These monitoring programs produce large amounts of data, which possibly can also be used to answer other questions.
-In Germany monitoring programs are setup by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and further state specific needs.
-However, currently there is no curated national-wide compilation of this data.
+In Germany monitoring programs are setup independently by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and additional state specific needs.
+However, currently there is no curated national-wide compilation of this data available.
 
-Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enabling downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
+\citet{malaj_organic_2014} analysed data supplied to the European Union in the context of the Water Framework Directive (WFD). 
+However, this reflects only a small part of data from national scale monitoring programs.
+
+Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
 At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.

### 2016-08-12 article.tex
minor tweaks to figures + results.

@@
 
+\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
+However, this reflects only a small part of data from national scale monitoring programs.
 National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
@@
 However, currently there is no curated national-wide compilation of this data available.
-
-\citet{malaj_organic_2014} analysed data supplied to the European Union in the context of the Water Framework Directive (WFD). 
-However, this reflects only a small part of data from national scale monitoring programs.
+\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
+They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
 
@@
 
-The aim of this study was to compile monitoring data on a national scale and to answer the questions: \\
+The aim of this study was to compile all available chemical monitoring data in Germany and to answer the questions: \\
 (i) Can the currently available monitoring data used for a representative description of the pollution situation? \\
 (ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds in these relationships? \\
-(iii) baHow polluted are small streams and which pesticides are responsible?
+(iii) How polluted are small streams and which pesticides are most important?
 
@@
 Red line marks statistically significant changes.
-Dashed lines denote 95\% Confidence Intervals.
+Dashed lines denote 95\% pointwise Confidence Intervals.
   }
@@
 
-The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement Figure Sxxx - Sxxx). \todo{references}.
+The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement Figure Sxxx - Sxxx). \todo{references}
 
-\subsection{Pesticide pollution of small agricultural streams in germany}
+\subsection{Pollution of small agricultural streams}
 Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
@@
 % RAC
-Insecticides showed most exceedances of RAC. 
-For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
-Highest RQ was observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
+Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
+For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
+In 15\% of all samples (n = 6377) risc quotients higher then 1 were observed.
+Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
 % TU
-In 33\% of the samples no compounds no compounds were detected. 
+In 33\% of the samples no pesticides were detected. 
 The mean $log(TU_{max})$ for detects was -4.6.
@@
 In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).
+
+%% Add more numbers / percentages here => compare with knauer/stehle/Vijver
 
@@
 
-% Vergleich mit der Schweiz.....
+The compiled dataset of governmental monitoring data respresents the
+
+% Vergleich mit der Schweiz..... knauer / munz
+% niederland Pestizide atlas
+% USA WaterQuality Portal
 
@@
 % underestimation of grab sampling (Stehle et al. 2013).
+
+% Stehle nur insectizide
+% malaj nur WFD
+
+% maxtu vs sumTU vs most sensitive => check publication.
+
+% Stehle + Munz + Schulz kleine gewässer mehr belastet?
+% Munz auch landwirtschaft
+
+% kleinere prozent werte wegen saisonalität / Proben im winter
+
+% knauer auch RAC
 

### 2016-08-12 src/load.R

@@ -22,52 +22,53 @@ cachedir <- file.path(prj, "cache")	# intermediate data
 #! use developmental version of data.table
 #! CRAN version should also work...
 # install.packages('devtools')
-require(devtools)
+library(devtools)
 # install_github("Rdatatable/data.table")
-require(data.table)
-require(xlsx)    # to read .xls files
+library(data.table)
+library(xlsx)    # to read .xls files
 # Note this needs Java JDK (http://www.oracle.com/technetwork/java/javase/downloads/jdk8-downloads-2133151.html)
-require(reshape2)
-require(plyr)
-require(bit64)
-require(Hmisc)
-require(stringr)
+library(reshape2)
+library(plyr)
+library(bit64)
+library(Hmisc)
+library(stringr)
 # install.packages("devtools")
 # devtools::install_github("hadley/readxl")
-require(readxl)
-require(stringi)
-require(stringdist)
+library(readxl)
+library(stringi)
+library(stringdist)
 
 # DB
-require(RPostgreSQL)
+library(RPostgreSQL)
 
 # spatial
-require(sp)
-require(rgdal)
-require(maptools)
-require(rgeos)
-require(raster)
-require(ggmap)
+library(sp)
+library(rgdal)
+library(maptools)
+library(rgeos)
+library(raster)
+library(ggmap)
 
 # plotting
-require(ggplot2)
-require(gridExtra)
-require(ggExtra)
+library(ggplot2)
+library(gridExtra)
+library(ggExtra)
 library(cowplot)
 library(viridis)
 
 # misc
-require(pbapply)
-require(esmisc) # https://github.com/EDiLD/esmisc
-require(webchem)
-require(xtable)
+library(pbapply)
+library(esmisc) # https://github.com/EDiLD/esmisc
+library(webchem)
+library(xtable)
 
 # modelling
-require(vegan)
+library(vegan)
 library(mgcv)
-require(gamlss)
-require(gamlss.cens)
-require(gamlss.dist)
+library(gamlss)
+library(gamlss.cens)
+library(gamlss.dist)
+library(gamlss.add)
 
 
 ### ----------------------------------------------------------------------------

### 2016-08-16 article.tex
discussion, cover letter, minor tweaks to text

@@
   This is the lowest concentration at which no acceptable biological effects are expected. 
-  These are derived during authorization  process of pesticides and contain an uncertainty factor.
+  These are derived during authorization process of pesticides and incorporate also an uncertainty factor.
   The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table xxx\todo{ref}).  
@@
   \begin{equation}
-  RQ = \frac{C}{RAC}
+  RQ_i = \frac{C_i}{RAC_i}
   \end{equation}
-  Where $C$ is the concentration of a compound in a sample.
+  Where $C_i$ is the concentration of a compound $i$ bin a sample.
 
@@
   We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
-  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table xxx\todo{ref}))
+  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table xxx\todo{ref})).
   We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
+  Additionally, we also calculated the sum of toxic units ($TU_{sum}$).
   A table of all included compounds can be found in the supplement.
@@
 Catchment delineation has been manually checked for accuracy. 
-In areas with low relief energy the delineation algorithm did not produce accurate results and we used river catchments provide by federal state authorities in these cases.
+In areas with low relief energy the delineation algorithm did not produce accurate results and we used river catchments provided by federal state authorities in these cases.
 For each catchment we calculated the relative coverage (\%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities.
@@
 
-All data-processing and analyses have been performed using R, version 3.3.1 \citep{r_core_team_r:_2016}.
+All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
 To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
@@
 
-We modeled the number of RAC exceedances (No) as:
+We modeled the number of RAC exceedances as:
 
@@
 2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
+We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
+
 %Mixtures
@@
 
-\begin{figure}[ht]
-  \includegraphics[width=0.95\textwidth]{figure5.pdf}
+\begin{figure}[h]
+  \includegraphics[width=0.9\textwidth]{figure5.pdf}
   \caption{Overview on pesticide pollution of small agricultural streams. 
@@
 
-The compiled dataset of governmental monitoring data respresents the
+\subsection{Representativeness of the dataset}
+The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
+Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \todo{cite munz} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
 
-% Vergleich mit der Schweiz..... knauer / munz
-% niederland Pestizide atlas
-% USA WaterQuality Portal
+\todo{compare numbers of different atlases, number of sites / compounds}.
 
-%%% Risk assessment
-% findings from field studies on pesti-
-% cide exposure and effects must be used for a retrospec-
-% tive validation of the current EU regulatory risk assess-
-% ment, particularly for its future development.
+Nevertheless such nationwide compilations, may not only be useful for governmental surveillance, but also to answer other questions, like validation of exposure modeling \todo{cite knäbel?}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}, investigation of pesticide mixtures \todo{cite schreiner} or scientific research.
 
-% validation of the risk assessment through
-% targeted chemical and biological post-authorization
-% monitoring programs must be implemented for com-
-% pounds of concern to ensure that their application
+% current problems in monitoring and possible sollutions
+A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
+There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analysed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
+Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.
+
+\subsection{Influence of catchment area and agriculture}
+% agriculture
+We found a strong influence of agriculture on the pollution of streams.
+If there is more the 25\% agriculture within a catchment pesticides can be detected in streams and may have biological effects \todo{citation needed}.
+
+% size
+We did not find support for our hypothesis that small streams are more polluted then big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015}.
+This could be explained by the relatively small gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
+For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
+Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$(Figure \ref{fig:fig3}, top).
+
+% Cite Schreiner in review
+% combination with biological data.
 
@@
 % Stehle nur insectizide
-% malaj nur WFD
+% malaj nur WFD?
 
@@
 
-% knauer auch RAC
-
-\subsection{Representativeness of the data}
-
-\subsection{Influence of catchment area and agriculture}
-
 \subsection{Pollution of streams}
 
+% EQS
+
+% RAC
+\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances organophosphate insecticides. 
+Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
+This can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
+Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality.
+This is also reflected in the different sample sizes (1,566 vs. 2,918,604 measurements \todo{use number for substances with RAC}) and the high number of non-detects in the monitoring data.
+
+Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
+This might reflect differences in use between countries and RACs.
+
+% TU
+xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
+The high correlation of TU and the relatively low exceedance of $TU_sum$ reveals that in compound mixtures occurring in the field there most only one compound dominating the toxicity.
+
+% Mixtures
+Nevertheless, most pesticides did not occur individually but in mixtures.
+
+% TUsum vs TUmax => calculate
+
+% Mixtures
+
+% EQS & RAC
+
+% Conclusions
+Exceedances of RACs are undesirable
+
 %%% Word count
-%% Figures 
+%% Figures  = 2400
 % 3 big = 3*600 = 1800

### 2016-08-16 cover_letter/cover.tex
discussion, cover letter, minor tweaks to text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Formal Letter
% LaTeX Template
% Version 1.0 (28/12/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Brian Moses (http://www.ms.uky.edu/~math/Resources/Templates/LaTeX/)
% with extensive modifications by Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%   PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%------------------------------------------------------------------------------

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)

\usepackage{graphicx} % Required for including pictures
\usepackage{microtype} % Improves typography
\usepackage[T1]{fontenc} % Required for accented characters
\usepackage[utf8]{inputenc}

% Create a new command for the horizontal rule in the document which allows thickness specification
\makeatletter
\def\vhrulefill#1{\leavevmode\leaders\hrule\@height#1\hfill \kern\z@}
\makeatother

%----------------------------------------------------------------------------------------
%   DOCUMENT MARGINS
%----------------------------------------------------------------------------------------

\textwidth 6.75in
\textheight 9.25in
\oddsidemargin -.25in
\evensidemargin -.25in
\topmargin -1in
\parindent 0.4in

%------------------------------------------------------------------------------
%   SENDER INFORMATION
%------------------------------------------------------------------------------

\def\Who{Eduard Szöcs} % Your name
\def\Where{University of Koblenz-Landau\\ Institute for Environmental Sciences} % Your department/institution
\def\Address{Fortstr. 7} % Your address
\def\CityZip{76829 Landau - GERMANY} % Your city, zip code, country, etc
\def\Email{E-mail: szoecs@uni-landau.de} % Your email address
\def\TEL{Phone: +49 6321 280 31552} % Your phone number

%------------------------------------------------------------------------------
%   HEADER AND FROM ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\address{
    \hspace{\fill}\includegraphics[width=3in]{Uni-Logo-2c.jpg} % Include the logo 
    ~\\[-0.11in] % Reduce the whitespace above the horizontal rule
    \vhrulefill{1pt} \\ 
    \hspace{\fill}\parbox[t]{4in}{ % Create a box for your details underneath the horizontal rule on the right
        \vspace{0.6in}
        \textbf{\Who} \em \\ % Your name, all text after this will be italicized
        \Where\\ % Your department
        \Address\\ % Your address
        \CityZip\\ % Your city and zip code
        \TEL\\ % Your phone number
        \Email\\ % Your email address
    }
    \hspace{-1.4in} % Horizontal position of this block, increase to move left, decrease to move right
    \vspace{-1in} % Move the letter content up for a more compact look
}

%------------------------------------------------------------------------------
%   TO ADDRESS STRUCTURE
%------------------------------------------------------------------------------

\def\opening#1{
\thispagestyle{empty}
{\centering\fromaddress \vspace{1in} \\} 
{\raggedright \toname \\ \toaddress \par} % Print the to name and address
\vspace{2em}
% Subject line
\noindent
\textbf{Subject line} \hspace*{\fill}\today \\[2em]
\vspace{0.6in} % White space after the to address
\noindent #1 % Print the opening line
}

%----------------------------------------------------------------------------------------
%   SIGNATURE STRUCTURE
%----------------------------------------------------------------------------------------

\signature{\Who} % The signature is a combination of your name and title

\long\def\closing#1{
    \vspace{0.1in} % Some whitespace after the letter content and before the signature
    \noindent % Stop paragraph indentation
    \hspace*{\fill} % Move the signature right
    \parbox{\indentedwidth}{\raggedright
    #1 % Print the signature text
    \vskip 0.65in % Whitespace between the signature text and your name
    \fromsig}
} % Print your name and title

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%   TO ADDRESS
%----------------------------------------------------------------------------------------

\begin{letter}
{
\textbf{Environmental Science \& Technology} \\
\vspace{1em}
To the editor
}

%------------------------------------------------------------------------------
%   LETTER CONTENT
%------------------------------------------------------------------------------

\opening{Dear editor of Environmental Science \& Technology,}

xxxxx

\closing{Sincerely, }

%----------------------------------------------------------------------------------------

\end{letter}
\end{document}

### 2016-08-19 article.tex
Revert "restructured (moved latex file to separate folder)"
This reverts commit 3f4daa0bd542b0f2452e3dde6c7592b911bdce4a.

@@
 \citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
-However, this reflects only a small part of data from national scale monitoring programs.
-National monitoring programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
-These monitoring programs produce large amounts of data, which possibly can also be used to answer other questions.
-In Germany monitoring programs are setup independently by the federal states in compliance with the Water Framework Directive \citep{quevauviller_water_2008} and additional state specific needs.
+However, this study reflected only a small part of data that is available from national monitoring programs.
+These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
+Large amounts of data are generated, which possibly can also be used to answer other questions.
+In Germany monitoring programs are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
 However, currently there is no curated national-wide compilation of this data available.
+
 \citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
 They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
-
 Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
 At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
-Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007} relatively little is know about their chemical and ecological status.
+Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, relatively little is know about their chemical and ecological status.
 
@@
 We homogenized and unified these datasets into a common database.
-We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
+We implemented a robust and reproducible data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
 An overview of the data cleaning process is provided in the supplemental materials.  
@@
 We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
-
 %Mixtures
@@
 
-% current problems in monitoring and possible sollutions
+% current problems in monitoring and possible solutions
 A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
@@
 Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
 Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality.
@@
 xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
-The high correlation of TU and the relatively low exceedance of $TU_sum$ reveals that in compound mixtures occurring in the field there most only one compound dominating the toxicity.
-
+The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.
 % Mixtures
 Nevertheless, most pesticides did not occur individually but in mixtures.
+This highlights that
 

### 2016-08-19 src/do_overview.R

@@ -57,7 +57,7 @@ p <- ggplot() +
   mytheme +
   coord_equal()
 p
-ggsave(file.path(prj, "manuscript/figure1.pdf"),
+ggsave(file.path(prj, "figure1.pdf"),
        p, width = 3.3, height = 3, units = 'in', dpi = 300, scale = 2)
 
 
@@ -279,7 +279,7 @@ p <- plot_grid(p_bar, p_mds, rel_heights = c(3, 1))
 # h <- h/sum(h)
 # p <- arrangeGrob(p_bar, p_mds, ncol = 2, respect = TRUE, heights = h)
 
-ggsave(file.path(prj, "manuscript/figure2.pdf"),
+ggsave(file.path(prj, "figure2.pdf"),
        p, width = 7, height = 3.5, units = 'in', dpi = 300, scale = 1.5)
 #  Need to be manually cropped in inkscape
 
@@ -302,7 +302,7 @@ ezg_lu <- ggplot(psm_sites_info[ezg_fin < 150 & !is.na(agri_fin) & !is.na(ezg_fi
 # ezg_lu
 ezg_lu <- ggMarginal(ezg_lu, type = 'histogram', binwidth = 5)
 # ezg_lu
-ggsave(file.path(prj, 'manuscript/figure3.pdf'), 
+ggsave(file.path(prj, 'figure3.pdf'), 
        ezg_lu,
        width = 3.3, height = 3.3/1.2,
        units = 'in', dpi = 300, scale = 2.5)

### 2016-08-19 src/do_sizeagri.R

@@ -228,7 +228,7 @@ p <- ggplot() +
   ylab('No. RAC exceedances') +
   ylim(c(0, 1.3))
 # p
-ggsave(file.path(prj, "manuscript/figure4.pdf"),
+ggsave(file.path(prj, "figure4.pdf"),
        p, width = 7, height = 7/1.6,
        units = 'in', dpi = 300, scale = 1)
 

### 2016-08-19 article.tex
Update on Overleaf.

@@
 
-\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
+\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
 However, this study reflected only a small part of data that is available from national monitoring programs.
@@
 All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
-To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
@@
 Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
-We used pointwise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
+We used point-wise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
 GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
@@
 Therefore, we report only results of grab samples from streams. 
-The analysed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
+The analyzed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
 We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).
@@
 
-In total 484 different compounds used as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
+In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table \todo{Set ref to table}). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:fig2}).
+We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
 Hierarchical clustering revealed three groups:
@@
 For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
-In 15\% of all samples (n = 6377) risc quotients higher then 1 were observed.
+In 15\% of all samples (n = 6377) risk quotients higher then 1 were observed.
 Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
@@
 A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
-There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analysed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
+There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
 Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.
@@
 % TU
-xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshhold above biological effects have been shown \todo{add citation schaefer}. 
+xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshold above biological effects have been shown \todo{add citation schaefer}. 
 The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.

### 2016-08-19 cover_letter/cover.tex
cover letter + refs

@@
 \usepackage{graphicx} % Required for including pictures
-\usepackage{microtype} % Improves typography
 \usepackage[T1]{fontenc} % Required for accented characters
@@
 \evensidemargin -.25in
-\topmargin -1in
-\parindent 0.4in
+\topmargin -1.5in
+\parindent 0in
 
@@
 \address{
-    \hspace{\fill}\includegraphics[width=3in]{Uni-Logo-2c.jpg} % Include the logo 
+    % Include the logos
+    \includegraphics[width=3in]{Uni-Logo-2c.jpg}\hspace{\fill}\includegraphics[width=2in]{logo_institut.jpg}
     ~\\[-0.11in] % Reduce the whitespace above the horizontal rule
@@
     \hspace{\fill}\parbox[t]{4in}{ % Create a box for your details underneath the horizontal rule on the right
-        \vspace{0.6in}
+        \vspace{0.2in}
         \textbf{\Who} \em \\ % Your name, all text after this will be italicized
@@
 \thispagestyle{empty}
-{\centering\fromaddress \vspace{1in} \\} 
+{\centering\fromaddress \vspace{0.5in} \\} 
 {\raggedright \toname \\ \toaddress \par} % Print the to name and address
@@
 \noindent
-\textbf{Subject line} \hspace*{\fill}\today \\[2em]
+\textbf{Submission of article} \hspace*{\fill}\today \\[2em]
 \vspace{0.6in} % White space after the to address
@@
 
+% noindent
+\setlength\parindent{0pt}
+
 %------------------------------------------------------------------------------
@@
 
-\opening{Dear editor of Environmental Science \& Technology,}
+\opening{Dear Editor,}
+\vspace{-1em}\\
+enclosed is a manuscript to be considered for publication in Environmental Science \& Technology titled \\
 
-xxxxx
+\emph{Pesticides in small agricultural streams in Germany}\\
 
-\closing{Sincerely, }
+Pesticides enter freshwater ecosystems from agricultural areas and may have effects on biota.
+However, relatively little is know about their large-scale distribution.
+Especially, small streams may be at high risk because of adjacency to fields and small water columns.
+
+We compiled and homogenized monitoring data from all German Federal States with a focus on small to mid-size streams. 
+Our results show that there is lack of knowledge for small streams. 
+However, the data revealed the pesticides are commonly occurring.
+Especially, neonicotinoid pesticides may pose a risk to small agricultural freshwaters.
+
+Our study should be of general interest and add valuable information on chemical monitoring, pesticide occurrences and ecological risk arising from them.
+
+If you have any further questions do not hesitate to contact me.
+
+\closing{Yours sincerely, }
 

### 2016-08-19 article.tex
test

@@
 The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
-Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \todo{cite munz} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
+Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
 
@@
 
-Nevertheless such nationwide compilations, may not only be useful for governmental surveillance, but also to answer other questions, like validation of exposure modeling \todo{cite knäbel?}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}, investigation of pesticide mixtures \todo{cite schreiner} or scientific research.
+Nevertheless, such nationwide compilations, may not only be useful for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or investigation of pesticide mixtures \cite{schreiner_pesticide_2016}.
+
+% underestimation grab sampling
 
 % current problems in monitoring and possible solutions
-A nationwide assessment is hampered \todo{check verb} by the inhomogeneity of monitoring data between federal states:
+A nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
 There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
@@
 
+% all detects for neonics > RAC => need to lower LOQs!+-
+
 % TU
@@
 
+% WFD 10km² 0> brauchen auch kleinere.
+
 % Conclusions
-Exceedances of RACs are undesirable
+Exceedances of RACs are undesirable ...
 

### 2016-08-19 article.tex
more discussion
tusum vs tumax

@@
 In 33\% of the samples no pesticides were detected. 
-The mean $log(TU_{max})$ for detects was -4.6.
-2.6\% of the samples showed $log(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
-We found a high correlation of $log(TU_{max})$ and $TU_{sum}$ \todo{ausrechnen und r angeben}, with a maximum higher value of $TU_{sum}$ xx units.
+The mean $log_{10}(TU_{max})$ for detects was -4.6.
+2.6\% of the samples showed $log_{10}(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
+We found a high correlation of $log_{10}(TU_{max})$ and $log_{10}(TU_{sum})$ (r = 0.985, with a maximum higher value of $log_{10}(TU_{sum})$ of 0.75 units (Supplement) \todo{ref}.
+% see do_polluion.R for numbers
 %Mixtures
@@
       \textbf{A}: 10 compounds with most frequent EQS exceedances. See B for color legend.
-      \textbf{B}: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples).
-      \textbf{C}: 15 compounds with highest risk quotients.
+      \textbf{B}: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples). Non-detects are not shown due to the logarithmic axis.
+      \textbf{C}: 15 compounds with highest risk quotients. Non-detects are not shown due to the logarithmic axis.
       \textbf{D}: Distribution of the number of detected compounds in a sample.
@@
 \section{Discussion}
-
 \subsection{Representativeness of the dataset}
@@
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
-
-\todo{compare numbers of different atlases, number of sites / compounds}.
-
-Nevertheless, such nationwide compilations, may not only be useful for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or investigation of pesticide mixtures \cite{schreiner_pesticide_2016}.
-
-% underestimation grab sampling
+The data compiled in Germany is of similar quantity.
 
 % current problems in monitoring and possible solutions
-A nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
+Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
 There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
 Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.
+
+Given the fixed term monitoring design \todo{cite stehle 2012} and the low amount of samples taken at precipitation events that could lead to run-off, the data resents a strong underestimation of the real pollution and peak concentrations.
+Automatic water samplers triggered at precipitation or passive samplers \todo{cite moschet and fernandez}.
+
+Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
 
@@
 Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$(Figure \ref{fig:fig3}, top).
-
-% Cite Schreiner in review
-% combination with biological data.
-
-% underestimation of grab sampling (Stehle et al. 2013).
-
-% Stehle nur insectizide
-% malaj nur WFD?
-
-% maxtu vs sumTU vs most sensitive => check publication.
-
-% Stehle + Munz + Schulz kleine gewässer mehr belastet?
-% Munz auch landwirtschaft
-
-% kleinere prozent werte wegen saisonalität / Proben im winter
 
@@
 Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in use between countries and RACs.
+This might reflect differences in pesticide use between countries and also RACs.
 
-% all detects for neonics > RAC => need to lower LOQs!+-
+For three insecticides all detected values where above the RAC.
+This highlights that national monitoring programs need to improve chemical analytics and lower the LOQ for these highly toxic compounds.
 
 % TU
-xxxx\todo{add number} samples showed $TU_{max}$ values greater than -3, a threshold above biological effects have been shown \todo{add citation schaefer}. 
-The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with one dominant compound.
+8.2\% of samples showed $TU_{max}$ values greater than -3 (Figure \ref{fig:fig5}C), a threshold above biological effects have been shown \citep{schafer_thresholds_2012}.
+
+The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with mainly one compound dominating the toxicity.
 % Mixtures
 Nevertheless, most pesticides did not occur individually but in mixtures.
-This highlights that
-
-% TUsum vs TUmax => calculate
+This highlights that future risk assessment needs to take mixtures into account.
 
@@
 % WFD 10km² 0> brauchen auch kleinere.
+
+% combination with biological data.
+
+% maxtu vs sumTU vs most sensitive => check publication.
+
+% Stehle + Munz + Schulz kleine gewässer mehr belastet?
+% Munz auch landwirtschaft
+
+% kleinere prozent werte wegen saisonalität / Proben im winter
 

### 2016-08-22 article.tex

@@
 They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
-Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonisation of polluted streams \citep{liess_analyzing_2005}.
+Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonization of polluted streams \citep{liess_analyzing_2005}.
 At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
@@
 \section{Results}
-\subsection{Overview and representativeness of compiled data}
+\subsection{Overview of the compiled data}
 
@@
 \section{Discussion}
-\subsection{Representativeness of the dataset}
+\subsection{Overview of the compiled dataset}
 The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
-The data compiled in Germany is of similar quantity.
+The data compiled in Germany is of similar quantity and quality.
 
@@
 There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
-Future monitoring programs should aim at resolving these differences in order to enable a nationwide assessment.
+Our dataset comprised several thousand of stream sampling sites, the was nearly no data available for standing waters. 
+Although, standing waters are abundant in the agricultural landscape and may have a high contribution to biodiversity \citep{davies_comparison_2008}, nearly nothing is known on their pollution.
+Although, small streams comprise most of the total stream length \citep{nadeau_hydrological_2007} sampling sites with catchment sizes below $10km^2$ are underrepresented by the current monitoring scheme (Figure \ref{fig:fig3}).
+This arises because catchments under $10km^2$ are not specified by the WFD.
+Nevertheless, they should not be neglected.
+Future monitoring programs should aim at resolving these issues in order to enable a nationwide and representative assessment of freshwaters.
 
-Given the fixed term monitoring design \todo{cite stehle 2012} and the low amount of samples taken at precipitation events that could lead to run-off, the data resents a strong underestimation of the real pollution and peak concentrations.
-Automatic water samplers triggered at precipitation or passive samplers \todo{cite moschet and fernandez}.
-
+% underestimation
+Given the fixed term monitoring design \citep{stehle_probabilistic_2013} and the low amount of samples taken at precipitation events that could lead to run-off, the data resents a underestimation of the real pollution and peak concentrations.
+Automatic event-drive samplers and passive samplers may help overcome these shortcomings and provide a better representation \citep{fernandez_calibration_2014,moschet_evaluation_2015}.
 Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
@@
 This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
-Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality.
+Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, sldo during periods with lower pesticide usage.
 This is also reflected in the different sample sizes (1,566 vs. 2,918,604 measurements \todo{use number for substances with RAC}) and the high number of non-detects in the monitoring data.
@@
 % Mixtures
-Nevertheless, most pesticides did not occur individually but in mixtures.
+Nevertheless, most pesticides did not occur individually but in mixtures \cite{schreiner_pesticide_2016}.
 This highlights that future risk assessment needs to take mixtures into account.
-
-% Mixtures
 
@@
 
-% WFD 10km² 0> brauchen auch kleinere.
-
-% combination with biological data.
-
-% maxtu vs sumTU vs most sensitive => check publication.
-
-% Stehle + Munz + Schulz kleine gewässer mehr belastet?
 % Munz auch landwirtschaft
-
-% kleinere prozent werte wegen saisonalität / Proben im winter
 
@@
 \begin{acknowledgement}
-The authors thank the authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding this project.
+The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding this project.
 \end{acknowledgement}

### 2016-08-22 article.tex
more discussion

@@
 \begin{abstract}
+% 150-200 words
+
 Fehlt noch...
@@
 (ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds in these relationships? \\
-(iii) How polluted are small streams and which pesticides are most important?
+(iii) How polluted are small streams and which pesticides are most important? Is the current approval process protective for these streams?
 
@@
 We homogenized and unified these datasets into a common database.
-We implemented a robust and reproducible data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
+We implemented a robust and reproducible data cleaning work flow (Supplement, Figure S1) \citep{poisot_best_2015}, though parts of the dataset are proprietary.
 An overview of the data cleaning process is provided in the supplemental materials.  
@@
   We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
-  These were available for 29 compounds (Supplement, Table xxx\todo{ref}).
+  These were available for 29 compounds (Supplement, Table S2).
 
@@
   This is the lowest concentration at which no acceptable biological effects are expected. 
-  These are derived during authorization process of pesticides and incorporate also an uncertainty factor.
-  The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table xxx\todo{ref}).  
+  These are derived during authorization process of pesticides, incorporate also an uncertainty factor and aim to be protective especially for small streams.
+  The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table S2).  
   We expressed RAC as Risk Quotient (RQ):
@@
   We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
-  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table xxx\todo{ref})).
+  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table S2)).
   We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
@@
 The analyzed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
-We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).
+We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
 
@@
 
-In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table \todo{Set ref to table}). 
+In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
+We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
 Hierarchical clustering revealed three groups:
@@
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day. \todo{ref supplement}.
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
 
@@
 
-The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement Figure Sxxx - Sxxx). \todo{references}
+The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement, Figures S4-S8).
 
@@
 2.6\% of the samples showed $log_{10}(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
-We found a high correlation of $log_{10}(TU_{max})$ and $log_{10}(TU_{sum})$ (r = 0.985, with a maximum higher value of $log_{10}(TU_{sum})$ of 0.75 units (Supplement) \todo{ref}.
+We found a high correlation of $log_{10}(TU_{max})$ and $log_{10}(TU_{sum})$ (r = 0.985, with a maximum higher value of $log_{10}(TU_{sum})$ of 0.75 units (Supplement, Figure S9).
 % see do_polluion.R for numbers
@@
   \includegraphics[width=0.9\textwidth]{figure5.pdf}
-  \caption{Overview on pesticide pollution of small agricultural streams. 
+  \caption{Overview on pesticide pollution of small agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textgreater 25\% agriculture). 
       \textbf{A}: 10 compounds with most frequent EQS exceedances. See B for color legend.
@@
 \section{Discussion}
-\subsection{Overview of the compiled dataset}
+\subsection{Overview on the compiled dataset}
 The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
-The data compiled in Germany is of similar quantity and quality.
+The data compiled here for Germany is of similar quantity and quality.
 
@@
 There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
-Our dataset comprised several thousand of stream sampling sites, the was nearly no data available for standing waters. 
-Although, standing waters are abundant in the agricultural landscape and may have a high contribution to biodiversity \citep{davies_comparison_2008}, nearly nothing is known on their pollution.
-Although, small streams comprise most of the total stream length \citep{nadeau_hydrological_2007} sampling sites with catchment sizes below $10km^2$ are underrepresented by the current monitoring scheme (Figure \ref{fig:fig3}).
-This arises because catchments under $10km^2$ are not specified by the WFD.
-Nevertheless, they should not be neglected.
+An increase of the analyzed compound spectra to all registered compounds and a lowering of LOQ are desirable for a reliable assessment.
+Our dataset comprised several thousand stream sampling sites, but nearly no data could be compiled for standing waters. 
+Although, standing waters are abundant in the agricultural landscape and may have a high contribution to biodiversity \citep{davies_comparison_2008}, nearly nothing is known on their chemical status.
+Small streams comprise most of the total stream length \citep{nadeau_hydrological_2007}.
+However, sampling sites with catchment sizes below $10km^2$ are underrepresented by the current monitoring scheme (Figure \ref{fig:fig3}).
+This likely arises because catchments under $10km^2$ are not specified by the WFD.
+Nevertheless, small streams should not be neglected for ecological and chemical quality assessment.
 Future monitoring programs should aim at resolving these issues in order to enable a nationwide and representative assessment of freshwaters.
@@
 % underestimation
-Given the fixed term monitoring design \citep{stehle_probabilistic_2013} and the low amount of samples taken at precipitation events that could lead to run-off, the data resents a underestimation of the real pollution and peak concentrations.
+Given the fixed term monitoring design \citep{stehle_probabilistic_2013} and the low amount of samples taken at precipitation events that could lead to run-off, the data represents a underestimation of the real pollution and peak concentrations.
 Automatic event-drive samplers and passive samplers may help overcome these shortcomings and provide a better representation \citep{fernandez_calibration_2014,moschet_evaluation_2015}.
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides can be detected in streams and may have biological effects \todo{citation needed}.
+If there is more the 25\% agriculture within a catchment pesticides can generally be detected in streams with an increase in fully agricultural catchments (above 75 \% agriculture).
+To our knowledge this is the first study \todo{True? Could not find anything...} investigating such thresholds of pesticide occurrences.
+Our results suggest that only catchments with \textless 25\% agriculture may be regarded as unpolluted and used as reference sites.
+Slightly higher thresholds (20-40\%) have been found for biological responses of lotic assemblages to land-use \citep{waite_agricultural_2014, feld_response_2013}, though they might be differences between taxonomic groups and landscape \citep{feld_response_2013}.
 
 % size
-We did not find support for our hypothesis that small streams are more polluted then big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015}.
-This could be explained by the relatively small gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
+We did not find support for our hypothesis that small streams are more polluted then big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$(Figure \ref{fig:fig3}, top).
+Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
 
 \subsection{Pollution of streams}
-
-% EQS
-
 % RAC
@@
 Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here \todo{zahlen neonics} and shows that this class may pose a high risk to freshwaters \todo{anderes Wort}.
-Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, sldo during periods with lower pesticide usage.
-This is also reflected in the different sample sizes (1,566 vs. 2,918,604 measurements \todo{use number for substances with RAC}) and the high number of non-detects in the monitoring data.
-
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (n = 3516-4745) and shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
+Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage.
+This is also reflected in the different sample sizes (1,566 vs. 1,117,988 measurements of compounds with RAC) and the high number of non-detects in the monitoring data.
 Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in pesticide use between countries and also RACs.
-
+This might reflect differences in pesticide use between countries and defined RACs.
 For three insecticides all detected values where above the RAC.
@@
 8.2\% of samples showed $TU_{max}$ values greater than -3 (Figure \ref{fig:fig5}C), a threshold above biological effects have been shown \citep{schafer_thresholds_2012}.
+This suggests that pesticides may act as a common stressor to freshwaters and should not be neglected \citep{schafer_contribution_2016}.
+The high correlation of TU and the relatively low increase of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with mainly one compound dominating the toxicity.
+% Mixtures
+Nevertheless, most pesticides did not occur individually but in mixtures \cite{schreiner_pesticide_2016} and future risk assessment should take mixtures into account.
 
-The high correlation of TU and the relatively low exceedance of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with mainly one compound dominating the toxicity.
-% Mixtures
-Nevertheless, most pesticides did not occur individually but in mixtures \cite{schreiner_pesticide_2016}.
-This highlights that future risk assessment needs to take mixtures into account.
+% EQS
+For only a limited number of compounds EQS have been defined. 
+We found generally low exceedances of EQS with only three herbicides showing higher exceedances.
+However, compared to the results from RAC comparisons it is clear that the currently defined EQS cannot capture the full risk exposed by pesticides.
+In order to provide a reliable measure of a good chemical status EQS for additional compounds must be defined.
 
-% EQS & RAC
-
-% Munz auch landwirtschaft
-
-% Conclusions
-Exceedances of RACs are undesirable ...
+% Approval / Risk Assessment
+Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
+The high exceedances of RAC indicate that the approval process must be checked and refined.
 

### 2016-08-22 src/do_pollution.R

@@ -18,6 +18,15 @@ psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
 var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 
+## n neonics
+psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin')]
+neo_id <- psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin'), variable_id]
+psm_samples[variable_id %in% neo_id, length(unique(sample_id)), by = variable_id]
+
+# n measurements with RCA
+psm_samples[variable_id %in% psm_variables[!is.na(rak_uba), variable_id]]
+
+
 # Classify stream types ---------------------------------------------------
 
 # Analyze only small agricultural streams (catchment < 30km2 and >25% agriculture)

### 2016-08-22 article.tex
abstract

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+%% \newcommand*\mycommand[1]{\texttt{\emph{#1}}}
 
@@
 % 150-200 words
+Small streams are important refuge areas for biodiversity.
+Although, small agricultural streams may be a high risk of pesticide pollution, relatively little is know on their chemical status.
+Monitoring of chemical status is performed in Germany be the federal states for surveillance of chemical status and might shed some light also on small streams.
 
-Fehlt noch...
+We compiled available monitoring data from federal states with a focus on small streams, resulting in a dataset of 2,918,604 measurements from 42,236 samples from 3,049 sampling sites in the years 2005-2014.
+The dataset covered 484 different compounds that can be classified as pesticides.
+We found a lack of data for small streams \textless 10 km\textsuperscript{2} catchment size.
+Especially neonicotinoid insecticides and Chlorpyrifos showed exceedancas of regulatory acceptable concentrations (RACs), which could be related to agricultural land use.
+However, a nation-wide assessment was hampered by a high heterogeneity between states.
+
+Our result suggest that pesticides might be a major thread for aquatic life.
+However, future monitoring programs need to be unified between states to provide more reliable assessments.
+
+\todo{not finished yet}
+
+%% Word
 \end{abstract}
@@
 
-%%% Word count
-%% Figures  = 2400
-% 3 big = 3*600 = 1800
-% 2 small = 2*300 = 600
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{acknowledgement}
+
+%%% Word count
+% abstract    :
+% 3 big(600)  : 1800
+% 2 small(300): 600
+% text body   : 3000 
+% ackno       : 24
+% ======================================
+% total       :
 

### 2016-08-26 article.tex
first draft

@@
 \title[Pesticides small streams]{Pesticides in small agricultural streams in Germany}
-\abbreviations{mo, pest, ger, tu}
-\keywords{Monitoring, Neonicotinoid, Germany, Toxic Units, Freshwater}
+% \abbreviations{mo, neon, ra, tu, fw}
+\keywords{Monitoring, Neonicotinoid, Risk Assessment Toxic Units, Freshwater}
 % RAC,
@@
 
-Our result suggest that pesticides might be a major thread for aquatic life.
-However, future monitoring programs need to be unified between states to provide more reliable assessments.
+Our result suggest that pesticides originating from agricultural land use might be a major thread for aquatic ecosystems.
+The exceedances of RACs indicate risks originating from few compounds.
+However, future monitoring programs need to be unified between states to provide more reliable assessments. \todo{Letzter Absatz gefällt mir nicht....}
 
-\todo{not finished yet}
-
-%% Word
+%182 words
 \end{abstract}
@@
 Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
-The high exceedances of RAC indicate that the approval process must be checked and refined.
+The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
 

### 2016-08-28 article.tex
first draft
comments_ralf_1

--- 2016-08-22 article.tex
@@
 % 150-200 words
+Small streams are important refugia for biodiversity.
+In agricultural areas, they may be at high risk of pesticide pollution.
+However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
+In Germany, pesticide monitoring is performed by the federal states as part of water quality surveillance.
+The related data also cover small streams and may contribute to national-scale risk assessment.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T13:30:31.255Z:
+%
+% Chemical status is misleading because pesticides are largely riverbasin specific pollutants and therefore are dealt with as complementary parameter for the ecological status
+%
+% ^.
 
-Fehlt noch...
+We compiled monitoring data focusing on small streams, resulting in a data set of 2,918,604 measurements from 42,236 samples in 3,049 sampling sites covering the years 2005-2014.
+A total of 484 different compounds that can be classified as pesticides were measured.
+Especially neonicotinoid insecticides and Chlorpyrifos showed exceedances of regulatory acceptable concentrations (RACs).
+We found a lack of data for small streams \textless 10 km\textsuperscript{2} catchment size.
+This, and the heterogeneity between the federal state monitoring programs, hampered a nation-scale assessment.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T13:35:37.869Z:
+%
+% Mehr inhaltliche ERgebnisse wären gut
+%
+% ^ <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T15:17:05.687Z:
+%
+% Hm - widerspricht sich auch mit dem Scope oben/storyline... Würde ich positiver formulieren, Despite heterogeneity etc. the data show that...
+%
+% ^.
+Our result suggest that pesticides originating from agricultural land use might be a major threat for aquatic ecosystems.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T15:17:24.497Z:
+%
+% paar mehr details wären gut hier, fällt etwas vom himmel die schlussfolgerung
+%
+% ^.
+The exceedances of RACs indicate risks originating from few compounds.
+However, future monitoring programs need to be unified between states to provide more reliable assessments. \todo{Letzter Absatz gefällt mir nicht...}
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T15:18:34.357Z:
+%
+% letzer absatz sehr wirr. Schlage folgende Struktur vor: 3-4 SÄtze zu Ergebnissen, ohne Interpretation und dann einen Satz Schlussfolgerung
+%
+% ^.
+
+%182 words
 \end{abstract}
@@
 More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more the 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+In the year 2014 more than 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
 The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001,liess_determination_1999}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:54:54.730Z:
+%
+% da unten regenanalysen kommen, sollte hier noch was zu precipitation hinzugefügt werden
+%
+% ^.
 Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T10:58:27.621Z:
+%
+% besser zitieren: schäfer et al. 2012 thresholds ES&T - ist stärker, da mehr studien drin
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T10:39:38.645Z:
+%
+% Der Einschub "frequently detected in .. monitoring" passt hier nicht so richtig hin
+%
+% ^.
 
-\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
-However, this study reflected only a small part of data that is available from national monitoring programs.
-These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
+\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from organic toxicants, with pesticides playing a major role.
+However, this study reflected only a small part of data that is available in national monitoring programs.
+These programs are setup to determine the pesticide pollution of surface, ground and drinking water.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-26T15:24:18.962Z:
+%
+% hier schon spezifischer werden: Nur auf Pestizide beziehen
+%
+% ^.
 Large amounts of data are generated, which possibly can also be used to answer other questions.
@@
 However, currently there is no curated national-wide compilation of this data available.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T10:44:09.240Z:
+%
+% Letzter Abschnitt etwas schwach.
+Es wäre besser einer Struktur zu folgen: Malaj - Pestizide in EU.
+Nicht gesamte Daten.
+Situation Deutschland sollte noch enger verknüpft werden, warum SPrung  zu Deutschland.
+Eine mögliche Storyline wäre:  Malaj und Stehle & Schulz zeigen, dass vermutlich Belastung vorhanden.
+EU fordert keine einbussen an biodiversität (sustainable use directive).
+Unklar, wie repräsentativ bisherige Studien sind.
+Weitere Daten sind verfügbar.
+IN Bezug auf Biodiv spielen vor allem Kleingewässer eine Rolle usw.
+%
+% ^.
 
 \citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
-They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
+They found that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
 Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonization of polluted streams \citep{liess_analyzing_2005}.
-At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T10:59:21.077Z:
+%
+% besser orlinskyi 2015 STOTEN, das bezieht sich direkt auf wiederbesiedelung
+%
+% ^.
+At the same time they are at high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:01:01.458Z:
+%
+% schulz 2004 beinhaltet auch direkt das verhältnis von größe und exposition
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:00:32.608Z:
+%
+% sind ja keine 2 unterschiedlichen punkte - besser präzisieren
+%
+% ^.
 Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, relatively little is know about their chemical and ecological status.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:01:43.074Z:
+%
+% ich würde diese WFD Begriffe vermeiden - chemical status im WFD Kontext beinhaltet gerade nicht die meisten Pestizide
+%
+% ^.
 
-The aim of this study was to compile all available chemical monitoring data in Germany and to answer the questions: \\
-(i) Can the currently available monitoring data used for a representative description of the pollution situation? \\
+We compiled a large data set with chemical monitoring data for Germany and aimed to answer the following research questions: \\
+(i) Can the currently available monitoring data be used for a representative description of the pollution situation? \\
 (ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds in these relationships? \\
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:05:01.605Z:
+%
+% welche relationships? Wird hier nicht klar
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:04:12.011Z:
+%
+% wir haben ja gerade für große gewässer gar nicht alle daten gesammelt und können dsa doch gar nicht direkt beantworten
+%
+% ^.
 (iii) How polluted are small streams and which pesticides are most important? Is the current approval process protective for these streams?  
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:29:55.457Z:
+%
+% Ich muss mal weiter lesen, aber einige der Fragen scheinen mir nicht beantwortet werden zu können.
+Z.b.
+Approval process - da wäre ja die Frage, ob die RACs ökologisch protektiv sind, dazu können wir nichts sagen...
+Du meinst wahrscheinlich was anderes
+%
+% ^.
 
@@
 
-We queried chemical monitoring data of pesticides from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
+We queried pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
+We will refer to the states using the ISO 3166-2:DE standard abbreviations.
 Additionally, we compiled data available from previous studies and searched online databases.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:34:07.673Z:
+%
+% zu unpräzise - mittlerweile muss man schon fast die genauen search terms angeben und bleibt auch unklar, was für zusätzliche Datensätze gesucht wurden.
+Tauchen die überhaupt auf?
+Wenn es nur Monitoringdaten sind, ist das eigentlich nicht wichtig
+%
+% ^.
 This yielded to a total of more then 30 datasets of different formats.
-In the following we will use the ISO 3166-2:DE standard abbreviations for federal states.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:36:48.417Z:
+%
+% entweder übersicht erstellen oder weglassen
+%
+% ^.
 
@@
 We implemented a robust and reproducible data cleaning work flow (Supplement, Figure S1) \citep{poisot_best_2015}, though parts of the dataset are proprietary.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:37:21.589Z:
+%
+% was hat das cleaning mit dem proprietary zu tun?! Das soll eher sagen, dass man die daten nicht rausgeben kann und nicht für alle reproducible ist?
+%
+% ^.
 An overview of the data cleaning process is provided in the supplemental materials.  
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:38:14.386Z:
+%
+% spezfischere angabe notwendig - kann aber eh in vorherigen Satz (for details see ...)
+%
+% ^.
 To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before.
 \citep{rauthe_central_2013}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:52:22.627Z:
+%
+% Das müsste ausführlicher beschrieben werden.
+Zunächst vielleicht mal 2 Sätze wie All sites  were georeferenced and transformed to the same CRS .
+Subsequently, each site was assigned assigned rainfall data of the nearest weather station ???
+, given the relevance of precipitation in causing runoff events.
+%
+% Das müsste ausführlicher beschrieben werden.
+Zunächst vielleicht mal 2 Sätze wie All sites  were georeferenced and transformed to the same CRS .
+Subsequently, each site was assigned assigned rainfall data of the nearest weather station ???
+, given the relevance of precipitation in causing runoff events.
+%
+% ^.
 
 \subsection{Characterization of chemical pollution}
-We characterized chemical pollution (excluding sum parameters) using three indicators:
+We characterized pesticide pollution (excluding sum parameters such as ???) using three indicators:
 
@@
   \item National and international Environmental Quality Standards (EQS) \citep{ogewv_verordnung_2011,european_union_directive_2013}:
-  We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
+  We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization because ...
   These were available for 29 compounds (Supplement, Table S2).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T11:57:50.603Z:
+%
+% das ist kein indikator. Ich würde den Abschnitt allgemeiner beginnen. Sowas wie:
+%
+% ^ <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T12:14:45.182Z:
+%
+% we characterised pesticide pollution by dividing the detected concentration ..
+of ..
+in a sample  with respect to a benchmark concentration.
+Formula...
+We used three benchmarks, resulting in three indicators:  1.
+%
+% ^.
 
   \item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
-  This is the lowest concentration at which no acceptable biological effects are expected. 
-  These are derived during authorization process of pesticides, incorporate also an uncertainty factor and aim to be protective especially for small streams.
+  No unacceptable ecological effect are expected if the environmental concentration remains below this concentration. 
+  RACs are derived during pesticide authorization and should especially protect small streams from unacceptable effects.
   The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table S2).  
@@
   \end{equation}
-  Where $C_i$ is the concentration of a compound $i$ bin a sample.
+  Where $C_i$ is the concentration of a compound $i$ in a sample.
 
-  \item Maximum Toxic Units ($\mathrm{TU_{max}}$)  \citep{sprague_measurement_1970}: 
+  \item Maximum Toxic Unit ($\mathrm{TU_{max}}$)  \citep{sprague_measurement_1970}: 
   \begin{equation}
@@
   \end{equation}
-
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T12:21:18.460Z:
+%
+% ich würde vorschlagen die neue Terminologie zu nutzen - das wäre hier maxTU d.magna - oder zu schreiben, for reason of brevity, we removed the test organism from the acronym
+%
+% ^.
   Where $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
   We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T12:25:14.302Z:
+%
+% "or model prediction" - dann wurden sie ja nicht compiled sondern modelliert?! Oder haben andere es modelliert und sie wurden von dort genommen?
+%
+% ^.
   We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table S2)).
-  We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
-  Additionally, we also calculated the sum of toxic units ($TU_{sum}$).
-  A table of all included compounds can be found in the supplement.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T12:24:07.016Z:
+%
+% We used D.magna as benchmark for reason of consistency, because toxicity data to invertebrates was mainly available for d.magna.
+Außerdem muss noch begründet werden, warum man nicht auf Fische und Pflanzen schaut...
+%
+% ^.
+  We used the maximum TU per sample, as it is independent from the number of measured compounds and makes no assumptions on the mode of action.
+  Additionally, we also calculated the sum of toxic units ($TU_{sum}$), which assumes concentration addition... .
 \end{enumerate}
@@
 \subsection{Characterization of catchments}
-We delineated catchments upstream of the sampling sites using a digital elevation model \citep{eea_digital_2013} and a multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
-Catchment delineation has been manually checked for accuracy. 
-In areas with low relief energy the delineation algorithm did not produce accurate results and we used river catchments provided by federal state authorities in these cases.
-For each catchment we calculated the relative coverage (\%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities.
+We delineated catchments upstream of the sampling sites using a digital elevation model \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
+Catchment delineation was manually checked for accuracy. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:16:43.040Z:
+%
+% in bezug auf was? Halbsatz wie "by comparison with catchment delineation/stream network from governmental ..."
+%
+% ^.
+In areas with small differences in elevation the delineation algorithm did not produce accurate results and we used river catchments provided by federal state authorities in these cases.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:19:26.672Z:
+%
+% kann man hier mehr details geben? Z.B. with min to max elevation in the federal state < ..., und wie viele betraf das
+%
+% ^.
+For each catchment we calculated the relative cover (\%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:21:12.999Z:
+%
+% ref needed
+%
+% ^.
 
@@
 
-All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
+All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
 To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:52:27.469Z:
+%
+% short description of cluster methods used 
+%
+% ^.
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:53:09.673Z:
+%
+% of the three indicators?
+%
+% ^.
 
 We modeled the number of RAC exceedances as:
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:53:27.209Z:
+%
+% das könnte man vorne noch hinschreiben - für TU wird gesagt, dass Aggregation über Max läuft, bei den beiden anderen könntest Du noch schreiben, dass die Anzahl der exceedances genommen wurde (warum eigentlich nicht der anteil?)
+%
+% ^.
 
@@
 We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) and used the number of sampling events per site ($n_i$) as an offset to account different sampling efforts. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:55:47.901Z:
+%
+% bleibt so ziemlich unklar, was alles eingeht. Mehrer Samplings pro Site? warum dann nicht auch korrektur für anzahl der gemessenen substanzen? 
+%
+% ^.
 $f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
@@
 Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T13:57:59.914Z:
+%
+% nach log transformation?
+Die Verteilung von TUs in einer Stelle ist ja keinesfalls normal, so dass der 95% cutoff ziemlich biased sein kann und sicherhlich von der anzahl der messungen abhängt.
+wie gingen non-detects hier ein?
+%
+% ^.
 We used point-wise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 Therefore, we report only results of grab samples from streams. 
-The analyzed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
-We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:03:12.649Z:
+%
+% hmm, zumindest sollte gennant werden, was außerdem noch an samples vorlag und ggf. wäre ein vergleich sinnvoll
+%
+% ^.
+The analyzed dataset comprised 2,918,604 measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+We found large differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:05:06.994Z:
+%
+% was ist denn mit niedersachsen und MV? WUsste ich gar nicht, dass wir da gar nichts haben?!
+%
+% ^.
 
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
+We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
 Hierarchical clustering revealed three groups:
 i) with less then 100 compounds (SL, ST and TH), \\
-ii) with a medium sized spectra and \\
-iii) with a big and distinct spectra (RP and NI).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:06:20.935Z:
+%
+% gibt es irgendwo eine übersicht der bundesländer im ANhang? Sonst sind die abkürzungen kaum nachvollziebar. Sie könnten zudem in Fig 1 hinzugefügt werden
+%
+% ^.
+ii) with medium sized spectra and \\
+iii) with a big and distinct spectrum (RP and NI).
 
@@
 We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:09:28.040Z:
+%
+% präzisieren: d.h. für die ging der delineation algorithm? Oder für die anderen hatten auch die BL nichts und die wurden dann ausgeschlossen?
+%
+% ^.
 The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:fig3}).
@@
 The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement, Figures S4-S8). 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:16:26.707Z:
+%
+% Die fragen die hier beantwortet wird ist doch eher how is pesticide pollution connected with agricultural catchment cover and stream size?
+in der introduction wird auch wenig dazu gesagt, was jetzt das neue/innovative ist.
+Das wäre aber wichtig, wenn man es wie bei höheren journals wie ES&T unterbringen will
+%
+% ^.
 
 \subsection{Pollution of small agricultural streams}
-Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
+Based on the results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:56:26.821Z:
+%
+% unklar, warum das hier genannt wird - ist ja nirgendwo vorher gesagt worden, dass genau die betrachtet werden sollten...
+%
+% ^ <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T14:59:43.735Z:
+%
+% sieht auch eher so aus, als wenn - aber nicht signifikant) - im mittel die RAK überschreitungen stark zunehmen mit der Catchment size...
+Also doch eigentlich kein Hinweis, dass kleinere belastet sind?
+Würde es ggf.
+sinn machen bei der abbildung mit catchment size (Fig4.) bei 150 abzuschneiden, die Figure 3 geht auch nur bis 150)
+%
+% ^.
 %EQS
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:02:38.601Z:
+%
+% der ganze abschnitt bezieht sich auf Pollution, aber eigentlich geht es nur darum, welche Substanzen relevant sind und nicht wie häufig überschreitungen auftreten.
+Überhaupt fragt man sich, warum genau diese Gruppe analysiert wird, da ja ggf.
+die größeren Gewässer nach der Darstellung hier anscheinend größer belastet sind?!
+(liegt ggf an der interaktion landwirtschaft und size?)
+%
+% ^.
 Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:04:43.052Z:
+%
+% es wäre insgesamt auch ganz gut zu wissen, an wie viel % der stellen überhaupt irgendwas überschritten wurde
+%
+% ^.
 Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
 % RAC
-Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
+Neonicotinoid Insecticides and Chlorpyrifos showed highest risk quotients.
 For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
@@
 2.6\% of the samples showed $log_{10}(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:07:14.707Z:
+%
+% Figure 5 kann ich nicht sehen
+%
+% ^.
 We found a high correlation of $log_{10}(TU_{max})$ and $log_{10}(TU_{sum})$ (r = 0.985, with a maximum higher value of $log_{10}(TU_{sum})$ of 0.75 units (Supplement, Figure S9).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:07:36.663Z:
+%
+% wenn man das in den Risk assessment kontext stellen will, was ja schwierig ist, dann müsste in der einleitung das besser vorbereitet werden.
+%
+% ^.
 % see do_polluion.R for numbers
@@
 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:13:15.860Z:
+%
+% was ist denn mit den Regendaten, die tauchen gar nicht mehr auf
+%
+% ^.
+
 \section{Discussion}
@@
 The data compiled here for Germany is of similar quantity and quality.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:09:01.348Z:
+%
+% and analysed? 
+%
+% ^.
 
@@
 There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:11:28.611Z:
+%
+% klar - aber je nach fragestellung könnte man sich nur auf die Substanzen beziehen, die überall gemessen werden.
+Was mit repräsentativ überhaupt gemeint ist, ist nicht überzeugend im artikel dargelegt und warum das wichtig ist...
+ich denke, hier können wir mal diskutieren, was denn die hauptfrage ist: Kann man mit monitoringdaten repräsentatives assessment machen?
+Das scheitert ja alleine schon am grab sampling und ist somit von vornherein beantwortet.
+Oder welche fragen man nur mit großen datensätzen beantworten kann.
+%
+% ^.
 An increase of the analyzed compound spectra to all registered compounds and a lowering of LOQ are desirable for a reliable assessment.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:13:49.591Z:
+%
+% umgekehrt wird aber gesagt, dass eigentlich nur ein paar substanzen relevant sind - so richtig begründet erscheint diese forderung nicht...
+%
+% ^.
 Our dataset comprised several thousand stream sampling sites, but nearly no data could be compiled for standing waters. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:14:41.955Z:
+%
+% das ist für das projekt relevant, aber man fragt sich, was das hier soll. Ich würde die von vornherein rausschmeißen und sagen, dass das monitoring vor allem auf fließgewässer bezogen ist
+%
+% ^.
 Although, standing waters are abundant in the agricultural landscape and may have a high contribution to biodiversity \citep{davies_comparison_2008}, nearly nothing is known on their chemical status.
@@
 This likely arises because catchments under $10km^2$ are not specified by the WFD.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:15:49.749Z:
+%
+% hä?? Die sind nicht berichtspflichtig, aber die behörden gehen davon aus, dass die anforderungen des guten ökologischen zustands auch für diese gewässer gelten
+%
+% hä?? Die sind nicht berichtspflichtig, aber die behörden gehen davon aus, dass die anforderungen des guten ökologischen zustands auch für diese gewässer gelten
+%
+% ^.
 Nevertheless, small streams should not be neglected for ecological and chemical quality assessment.
 Future monitoring programs should aim at resolving these issues in order to enable a nationwide and representative assessment of freshwaters. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:16:47.488Z:
+%
+% ich denke, dass kann hier eher ein untergeordneter aspekt sein
+%
+% ^.
 
@@
 Given the fixed term monitoring design \citep{stehle_probabilistic_2013} and the low amount of samples taken at precipitation events that could lead to run-off, the data represents a underestimation of the real pollution and peak concentrations.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:20:23.131Z:
+%
+% das wird doch gar nicht gezeigt bzw. dargelegt, dass das underestimated ist... bzw. ist das kein ergebnis dieser studie nach aktueller lage
+%
+% ^ <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:21:08.769Z:
+%
+% also mit lage meine ich im text
+%
+% ^.
 Automatic event-drive samplers and passive samplers may help overcome these shortcomings and provide a better representation \citep{fernandez_calibration_2014,moschet_evaluation_2015}.
@@
 If there is more the 25\% agriculture within a catchment pesticides can generally be detected in streams with an increase in fully agricultural catchments (above 75 \% agriculture).
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:21:56.186Z:
+%
+% unklar, wo das genau gezeigt wird - die anzahl der detektionen gegen %catchment wird nicht gezeigt...
+%
+% ^.
 To our knowledge this is the first study \todo{True? Could not find anything...} investigating such thresholds of pesticide occurrences.
-Our results suggest that only catchments with \textless 25\% agriculture may be regarded as unpolluted and used as reference sites.
-Slightly higher thresholds (20-40\%) have been found for biological responses of lotic assemblages to land-use \citep{waite_agricultural_2014, feld_response_2013}, though they might be differences between taxonomic groups and landscape \citep{feld_response_2013}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:22:35.375Z:
+%
+% denke auch - aber man müsste hier 2-3 Studien nennen, die ich im Abschlussbericht zitiert habe, die ähnliche thresholds gefunden haben und könnte sagen, dass unsere studie insofern nahelegt, dass der faktor pestizide auch relevant ist
+%
+% denke auch - aber man müsste hier 2-3 Studien nennen, die ich im Abschlussbericht zitiert habe, die ähnliche thresholds gefunden haben und könnte sagen, dass unsere studie insofern nahelegt, dass der faktor pestizide auch relevant ist
+%
+% ^.
+Our results suggest that only catchments with \textless 25\% agriculture may be regarded as unpolluted and could be used as reference sites.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:24:08.697Z:
+%
+% das ist mir in fig.
+4 nicht ganz klar.
+Wenn das immer point estimates sind, die signifikant sind, dann wären ja schon 5% höher als 1%, weil ja die ganze linie bis 25% rot ist.
+Oder wie ist das zu lesen?
+Insofern kann man auch schlecht sagen auf basis des fast linearen anstiegst, dass 20% genauso ist wie 2%
+%
+% ^.
+Slightly higher thresholds (20-40\%) have been found for biological responses of lotic assemblages to land-use \citep{waite_agricultural_2014, feld_response_2013}, though there might be differences between taxonomic groups and landscape \citep{feld_response_2013}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:26:50.631Z:
+%
+% das würde genauerer Diskussion bedürfen, denn das sagt dann ja, dass für Diatomeen oder invertebraten und für die unterschiedlichen bundesländer unterschiedliche % zahlen relevant sein könnten.
+Könnte man ja eh überprüfen
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:26:07.465Z:
+%
+% land use ist doch gar nciht relevant - nur spezifisch agriculture
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:25:37.776Z:
+%
+% 20% ist nicht higher...
+%
+% 20% ist nicht higher...
+%
+% ^.
 
 % size
-We did not find support for our hypothesis that small streams are more polluted then big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+We did not find support for our hypothesis that small streams are more polluted than big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:28:35.191Z:
+%
+% kannst du mir knauer mal schicken?
+%
+% ^.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:28:05.039Z:
+%
+% ist nicht als hypothese formuliert vorher
+%
+% ^.
 This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:30:08.003Z:
+%
+% indeed - so richtig kann man das nicht testen und die definition von "big" dürfte einige verwundern
+%
+% ^.
 Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
@@
 % RAC
-\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances organophosphate insecticides. 
-Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
+\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
+By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
 This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (n = 3516-4745) and shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:31:58.251Z:
+%
+% unklar warum spanne
+%
+% ^.
 Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:34:37.233Z:
+%
+% wo???
+%
+% ^.
 This is also reflected in the different sample sizes (1,566 vs. 1,117,988 measurements of compounds with RAC) and the high number of non-detects in the monitoring data.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:35:13.107Z:
+%
+% unklar
+%
+% ^.
 Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
 This might reflect differences in pesticide use between countries and defined RACs.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:36:00.776Z:
+%
+% was überhaupt mal gebracht werden sollte ist der bezug von RAC zu biodiversität - dafür kann ja Stehle & Schulz dienen?
+%
+% ^.
 For three insecticides all detected values where above the RAC.
@@
 % TU
-8.2\% of samples showed $TU_{max}$ values greater than -3 (Figure \ref{fig:fig5}C), a threshold above biological effects have been shown \citep{schafer_thresholds_2012}.
+8.2\% of samples showed $TU_{max}$ values greater than -3 (Figure \ref{fig:fig5}C), a threshold above which ecological effects have been shown \citep{schafer_thresholds_2012}.
 This suggests that pesticides may act as a common stressor to freshwaters and should not be neglected \citep{schafer_contribution_2016}.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:37:12.651Z:
+%
+% by what or whom? The monitoring shows that they are taken serious?!
+%
+% ^.
 The high correlation of TU and the relatively low increase of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with mainly one compound dominating the toxicity.
@@
 Nevertheless, most pesticides did not occur individually but in mixtures \cite{schreiner_pesticide_2016} and future risk assessment should take mixtures into account.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:37:54.188Z:
+%
+% problem der bestimmung ist doch hier auch relevant.
+SIehe die Moschet studien.
+ob dazu gezielter was zu sagen, müsste man mal stellen nach regen und vielen substanzen mit anderen vergleichen, was die sumTox angeht
+%
+% ^.
 
@@
 In order to provide a reliable measure of a good chemical status EQS for additional compounds must be defined.
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:39:15.216Z:
+%
+% was trägt EQS eigentlich insgesamt zur Story bei? Ist mir nicht ganz klar. Gibt kaum Ergebnisse hierzu und für 30 von 400 substanzen was zu machen, ist auch nicht so überzeugend
+%
+% ^.
 
@@
 Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
-The high exceedances of RAC indicate that the approval process must be checked and refined.
-
-%%% Word count
-%% Figures  = 2400
-% 3 big = 3*600 = 1800
-% 2 small = 2*300 = 600
+The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
 
@@
 \begin{acknowledgement}
-The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding this project. 
+The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project. 
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:40:29.902Z:
+%
+% typischerwiese wird dann die projektnummer angegebn
+%
+% ^.
 \end{acknowledgement}
+
+%%% Word count
+% abstract    :
+% 3 big(600)  : 1800
+% 2 small(300): 600
+% text body   : 3000 
+% ackno       : 24
+% ======================================
+% total       :
 
@@
 \end{suppinfo}
+% * <ralf.schaefer@aggiemail.usu.edu> 2016-08-27T15:41:32.807Z:
+%
+% ok - jetzt habe ich erst fig 5 gefunden... macht ggf. den ein oder anderen kommentar überflüsig
+%
+% ^.
 

### 2016-08-29 supplement/Supplemental_Materials.tex
finalized results
more discussion
tusum vs tumax
more discussion
removed tu and eqs analyses

--- 2016-08-10 supplement/Supplemental_Materials.tex
@@
 \usepackage{amsmath}
+\usepackage{float} % for [H]
+
+% Continuous figure counting (instead of per cahpter)
+\usepackage{chngcntr}
+\counterwithout{figure}{chapter}
+\counterwithout{table}{chapter}
+% Add 'S' before figure/table numbering
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 	\centering
-	\includegraphics[width = 0.6\textwidth]{ezgagrirac}
-	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points are sites without any exceedance.}
+	\includegraphics[width = 0.7\textwidth]{ezgagrirac}
+	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
 	\label{fig:ezgagrirac}
@@
 \todo{update refs}
-
-\pagebreak
-\section{Toxic Units}
-
-We modeled the 95th percentile of $TU_{max}$ on a logarithmic scale assuming a Normal distribution:
-
-\begin{align}
-\begin{split}
-  log(TU_{max}) \sim Normal(\mu_i, \sigma) \\
-  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i)
-\end{split}
-\end{align}
-
-where $TU_{max i}$ is 95th percentile of $TU_{max}$ at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site.  
-$f_1$ and $f_2$ are smoothing functions.
-
-Because of samples at low polluted sites contained only values below the limit of quantification (LOQ) the derived $TU_{max}$ were also left censored (at a value of $10^{-8.75}$), thus we have
-
-\begin{align}
-  log(TU_{max~i})=\begin{cases}
-    \beta_0 + f_1(Agri_i) + f_2(Size_i) + \epsilon_i, & \text{if $log(TU_{max}) > -8.75 $}\\
-    -8.75, & \text{if $log(TU_{max}) \le -8.75 $}.
-  \end{cases}
-\end{align}
-
-Instead of assuming a fixed value for the data \textless LOQ, we use a model taking this censoring into account \citep{helsel_fabricating_2006} with the expected value being a mixture of censored and uncensored data.
-
-\begin{align}
-\begin{split}
-	E(y_i | x_i) = P(uncensored|x_i) \times E(y_i | y_i > -8.75)) + P(censored|x_i) \times -8.75\\
-	Var(y_i) = \sigma
-\end{split}
-\end{align}
-
-With $y_i$ being $log(TU_{max~i})$ and $x_i$ the predictor functions $f_1(Agri_i) + f_2(Size_i)$.
-$P(censored|x_i)$ is the probability of censoring at given predictors.
-Further, we assumed constant variance.
-This model was fitted using the gamlss packages \citep{stasinopoulos_generalized_2007}.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width = 0.6\textwidth]{ezgagritu}
-	\caption{Raw data of effect of agriculture and catchment size on the 95th percentile of log(TUmax).}
-	\label{fig:ezgagritu}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width = 0.9\textwidth]{ezgagritumodel}
-	\caption{Effect of agriculture within the catchment (left) and catchment size (rigth) on the 95th percentile of log(TUmax).
-The model took censored data into account (see description above).
-Dashed lines denote 95\% Confidence Intervals.}
-	\label{fig:ezgagritu_model}
-\end{figure}
 

### 2016-08-29 src/do_sizeagri.R

@@ -62,8 +62,13 @@ rak_exceed <- psm_sites_wi[samples_exceed]
 rak_exceed
 # 2970 sites
 rm(samples_raks, samples_exceed)
+# exclude NA (=sites wihtout ezg & agri info)
 rak_exceed <- rak_exceed[!(is.na(agri_fin) | is.na(ezg_fin)), ]
 # 2343 sites with ezg and agri data
+
+# exclude sites > 150km2
+rak_exceed <- rak_exceed[ezg_fin < 150]
+
 
 # logn used as offset
 rak_exceed$logn <- log(rak_exceed$n)
@@ -125,8 +130,8 @@ mod_nb_ti <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') +
               data = rak_exceed,
               family = nb(),
               method = 'REML')
-plot(mod_nb_te)
-vis.gam(mod_nb_te, view = c('agri_fin', 'ezg_fin'))
+plot(mod_nb_ti)
+vis.gam(mod_nb_ti, view = c('agri_fin', 'ezg_fin'))
 anova(mod_nb, mod_nb_ti, test = 'Chisq') 
 # smoothing interaction not of interest
 
@@ -233,338 +238,3 @@ ggsave(file.path(prj, "figure4.pdf"),
        units = 'in', dpi = 300, scale = 1)
 
 
-
-
-
-
-
-# Model log(TUmax) --------------------------------------------------------
-psm_maxtu # = view as calculated from database
-
-# calculate withing R (for verification)
-setkey(var_props, variable_id)
-setkey(psm_samples, variable_id)
-
-samples_lc50 <- var_props[ , list(variable_id, lc50_dm_fin)][psm_samples]
-# rm samples without lc50 data
-samples_lc50 <- samples_lc50[!is.na(lc50_dm_fin)]
-
-# calculate log(tu)
-# min value is -8.72
-range(samples_lc50[value_fin > 0 , log10(value_fin / lc50_dm_fin)])
-
-# logtu: set zero to -8.75
-samples_lc50[ , logtu := ifelse(value_fin > 0, log10(value_fin / lc50_dm_fin), -8.75)]
-# tu (use with log-normal dist)
-samples_lc50[ , tu := value_fin / lc50_dm_fin]
-
-# calculate per sample max(logtu)
-logtumax <- samples_lc50[ , list(logtumax = max(logtu)), by = sample_id]
-
-
-# join with data from database, to check.
[…]

### 2016-08-30 article.tex
started revision accoring to ralfs suggestions

--- 2016-08-26 article.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{enumerate}
+
 \usepackage{todonotes}
-\usepackage{amsmath}
 
@@
 % 150-200 words
-Small streams are important refuge areas for biodiversity.
-Although, small agricultural streams may be a high risk of pesticide pollution, relatively little is know on their chemical status.
-Monitoring of chemical status is performed in Germany be the federal states for surveillance of chemical status and might shed some light also on small streams.
-
-We compiled available monitoring data from federal states with a focus on small streams, resulting in a dataset of 2,918,604 measurements from 42,236 samples from 3,049 sampling sites in the years 2005-2014.
-The dataset covered 484 different compounds that can be classified as pesticides.
-We found a lack of data for small streams \textless 10 km\textsuperscript{2} catchment size.
-Especially neonicotinoid insecticides and Chlorpyrifos showed exceedancas of regulatory acceptable concentrations (RACs), which could be related to agricultural land use.
-However, a nation-wide assessment was hampered by a high heterogeneity between states.
-
-Our result suggest that pesticides originating from agricultural land use might be a major thread for aquatic ecosystems.
-The exceedances of RACs indicate risks originating from few compounds.
-However, future monitoring programs need to be unified between states to provide more reliable assessments. \todo{Letzter Absatz gefällt mir nicht....}
-
-%182 words
+Small streams are important refugia for biodiversity.
+In agricultural areas, they may be at high risk of pesticide pollution.
+However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
+In Germany, pesticide monitoring is performed by the federal states as part of water quality surveillance. 
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% -------------------------------------------------------------------------
 \section{Introduction}
+More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
+In the year 2014 more than 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
+Especially run-off after heavy precipitation events has been shown to be one of the major input routes \citep{schulz_field_2004}.
+Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}.
 
-More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more the 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
-The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage, with run-off being one of the major input routes \citep{schulz_comparison_2001,liess_determination_1999}.
-Once entered the surface waters pesticides are frequently detected in environmental monitoring \citep{malaj_organic_2014} and may have adverse effects on biota and ecosystem functioning \citep{schulz_field_2004, schafer_effects_2007}.
+\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that most European water bodies are at risk from pesticides.
+\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
+They found that many of these measurements exceed regulatory acceptable concentrations (RAC).
+Both studies indicate that pesticides might be a threat to biodiversity in the European union. 
+However, theses studies reflect only a small part of available data and it is unclear how representative they are:
+For Germany the study of \citet{malaj_organic_2014} lists only 175 sites and \citet{stehle_pesticide_2015} only 138 measurements, although there might my much more data available from national monitoring programs.
+%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
 
-\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that pesticides jeopardize the health of freshwater ecosystems.
-However, this study reflected only a small part of data that is available from national monitoring programs.
-These programs are setup for determination and surveillance of the chemical and ecological status of surface, ground and drinking water.
-Large amounts of data are generated, which possibly can also be used to answer other questions.
-In Germany monitoring programs are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
-However, currently there is no curated national-wide compilation of this data available.
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004}.
+It has been show that SWB are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SWB \todo{citation Lorenz et al.}, with only few large scale studies.
 
-\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
-They found that insecticides that many of these measurements exceed regulatory acceptable concentrations (RAC), especially in very small catchments \textless $1~km^2$.
-Small water bodies are important refuges of biodiversity \citep{davies_comparison_2008} and enable downstream colonization of polluted streams \citep{liess_analyzing_2005}.
-At the same time they may be exposed to a high risk of pesticide contamination from adjacent agricultural areas and low dilution effects \citep{liess_determination_1999}.
-Although small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, relatively little is know about their chemical and ecological status.
+We compiled a large data set with chemical monitoring data for Germany and aimed to answer the following research questions:
 
-The aim of this study was to compile all available chemical monitoring data in Germany and to answer the questions: \\
-(i) Can the currently available monitoring data used for a representative description of the pollution situation? \\
-(ii) Are small agricultural waters more polluted compared to bigger streams? Are there thresholds in these relationships? \\
-(iii) How polluted are small streams and which pesticides are most important? Is the current approval process protective for these streams?
+\begin{enumerate}[(i)]
+	\item Can the currently available monitoring data be used for a representative description of the pollution situation?
+	\item Are there thresholds for agricultural land use above pesticide pollution becomes apparent?
+	\item Are there thresholds of stream size below pesticide pollution becomes apparent
+	\item  How polluted are small agricultural streams in Germany and which pesticides are most important? 
+\end{enumerate}
 
+%% -------------------------------------------------------------------------
 \section{Methods}
 \subsection{Data compilation}
+We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see supplemental table S1 for the abbreviations of federal state names).
+We homogenized and unified all data from the states into a common database.
+We implemented a robust and reproducible data cleaning workflow (see supplement for details on the data processing workflow) \citep{poisot_best_2015}.
+Nevertheless, parts of the dataset are proprietary and cannot be shared here.
 
-We queried chemical monitoring data of pesticides from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
-Additionally, we compiled data available from previous studies and searched online databases.
-This yielded to a total of more then 30 datasets of different formats.
-In the following we will use the ISO 3166-2:DE standard abbreviations for federal states.
-
-We homogenized and unified these datasets into a common database.
-We implemented a robust and reproducible data cleaning work flow (Supplement, Figure S1) \citep{poisot_best_2015}, though parts of the dataset are proprietary.
-An overview of the data cleaning process is provided in the supplemental materials.  
-To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before.
-\citep{rauthe_central_2013}.
+Given the relevance of precipitation in causing runoff events, we identified chemical samples taken during heavy rainfall events.
+We performed a spatio-temporal intersection of sampling events with gridded daily precipitation data available the German Weather service \citep{rauthe_central_2013}. 
+We performed the intersection for the actual sampling date and the day before.
 
 \subsection{Characterization of chemical pollution}
-We characterized chemical pollution (excluding sum parameters) using three indicators:
+We characterized pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
+RACs are derived during pesticide authorization and no unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
+The German Federal Environmental Agency provided RACs for the 105 compounds with highest detection rates (Supplement, Table S2). 
+We expressed RACs as Risk Quotient (RQ):
 
-\begin{enumerate}
-  \item National and international Environmental Quality Standards (EQS) \citep{ogewv_verordnung_2011,european_union_directive_2013}:
-  We used only Maximum Annual Concentration EQS (MAC-EQS) for characterization.
-  These were available for 29 compounds (Supplement, Table S2).
-
-  \item Regulatory Acceptable Concentrations (RAC) \citep{brock_linking_2010}:
-  This is the lowest concentration at which no acceptable biological effects are expected. 
-  These are derived during authorization process of pesticides, incorporate also an uncertainty factor and aim to be protective especially for small streams.
-  The German Federal Environmental Agency provided RACs for 105 compounds (Supplement, Table S2).  
-  We expressed RAC as Risk Quotient (RQ):
 \begin{equation}
@@
 \end{equation}
-  Where $C_i$ is the concentration of a compound $i$ bin a sample.
 
-  \item Maximum Toxic Units ($\mathrm{TU_{max}}$)  \citep{sprague_measurement_1970}: 
-  \begin{equation}
-  TU_{max} = max(\frac{C_i}{EC_{50, D.magna, i}})
-  \end{equation}
-
-  Where $C_i$ is the concentration of compound $i$ in a sample and $EC_{50, D.magna, i}$ is the concentration of this compound where 50\% of the exposed animals showed after 48 hours an effect in a laboratory study.
-  We compiled $EC_{50, D.magna}$ values from literature \citep{malaj_organic_2014}, databases \citep{lewis_international_2016,u.s._epa_ecotoxicology_2015} or model predictions \citep{schuurmann_quantitative_2011}, where experimental data had priority.
-  We could compile $EC_{50, D.magna}$ values for 394 compounds (Supplement, Table S2)).
-  We used the maximum TU per sample, as it is independent of the number of measured compounds and makes no assumptions on the mode of action.
-  Additionally, we also calculated the sum of toxic units ($TU_{sum}$).
-  A table of all included compounds can be found in the supplement.
-\end{enumerate}
+Where $C_i$ is the concentration of a compound $i$ in a sample.
 
 \subsection{Characterization of catchments}
-We delineated catchments upstream of the sampling sites using a digital elevation model \citep{eea_digital_2013} and a multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
-Catchment delineation has been manually checked for accuracy. 
-In areas with low relief energy the delineation algorithm did not produce accurate results and we used river catchments provided by federal state authorities in these cases.
-For each catchment we calculated the relative coverage (\%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities.
+We compiled a total of 3,049 sampling sites with pesticide measurements.
+We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
+Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
+The delineation algorithm produced only for 30\% of the sites accurate results.
+For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
+For 10\% of the sites we were not able to compile catchment size data.
+For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
+For 78\% of the sites both, the proportion of agricultural landuse and catchment size were available.
+% see do_overview.R for numbers
 
 \subsection{Statistical analyses}
-
-All data-processing and analyses have been performed using R \citep{r_core_team_r:_2016}.
-To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
+To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
-
 We modeled the number of RAC exceedances as:
@@
 
-where $No_i$ is the observed number of exceedances at site $i$, $Agri_i$ the proportion of agriculture within the catchment and $Size_i$ the catchment size of the site. 
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) and used the number of sampling events per site ($n_i$) as an offset to account different sampling efforts. 
+where $No_i$ is the observed number of exceedances at site $i$. 
+The proportion of agriculture within the catchment ($Agri_i$) and the catchment size of the site ($Size_i$) were used as predictors. 
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
+The number of sampling events per site ($n_i$) was as an offset to account different sampling efforts at a site and is equivalent to modeling the rate of exceedances. 
 $f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-Similar models were fitted to the number of EQS-exceedances and the 95th percentile of $TU_{max}$ (see Supplement for details). 
 We used point-wise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
 GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
+\todo{precipitation stats?}
 
+%% -------------------------------------------------------------------------
 \section{Results}
@@
 
-The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
+The compiled dataset comprised only few standing waters (58 sites) and the majority of samples (91\%) where taken via grab sampling.  % see clean.R for numbers
+9\% of samples from 33 sites were taken as composite samples of different durations.
 Therefore, we report only results of grab samples from streams. 
-The analyzed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
-We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
+The analyzed dataset comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+We found large differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
 
@@
   \includegraphics[width=0.6\textwidth]{figure1.pdf}
-  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states.}
+  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states, see supplemental table S1 for abbreviations.}
   \label{fig:fig1}
@@
 
-In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2). 
+In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2 \todo{remove LC50 and MAC-EQS from table!}). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-We found substantial differences of the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
+Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
+We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
 Hierarchical clustering revealed three groups:
-i) with less then 100 compounds (SL, ST and TH), \\
-ii) with a medium sized spectra and \\
-iii) with a big and distinct spectra (RP and NI).
 
-Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
+\begin{enumerate}[i)]
+	\item with less then 100 compounds (SL, ST and TH)
+	\item with medium sized spectra
+	\item with a big and distinct spectrum (RP and NI)
+\end{enumerate}
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to three groups determined by hierarchical clustering (see Supplement Figure xxx).}
+  Colors according to three groups determined by hierarchical clustering (see Supplement Figure S2).}
   \label{fig:fig2}
@@
 
-We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. 
-The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:fig3}).
-The proportion of agriculture in the catchments decreased with increasing catchment size.
-
-\begin{figure}[ht]
-  \includegraphics[width=.8\textwidth]{figure3.pdf}
-  \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
-  Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
-  Colour codes the 2-dimensional density of points.
-  }
-  \label{fig:fig3}
-\end{figure}
-
-\subsection{Are small agricultural waters more polluted compared to bigger streams?}
-
-Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
-Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:fig4}, left).
-
-We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right) and no interaction between these two predictors.
-
-\begin{figure}[ht]
-  \includegraphics[width=0.95\textwidth]{figure4.pdf}
-  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
-Red line marks statistically significant changes.
-Dashed lines denote 95\% pointwise Confidence Intervals.
-  }
-  \label{fig:fig4}
-\end{figure}
-
-The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement, Figures S4-S8).
-
-\subsection{Pollution of small agricultural streams}
-Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites with 10591 samples and 36550 measurements).
-%EQS
-Out of the 29 compounds with EQS, the most frequent EQS exceedances were recorded for the herbicides Nicosulfuron (2.1\%, n = 1769), Flufenacet (1.1\%, n = 6301) and Isoproturon (0.7\%, n = 8380). 
-Other compounds show exceedances only at less then 0.5\% of all samples (Figure \ref{fig:fig5}A).
-% RAC
-Neonicotinoid Insecticides and Chlropyrifos showed highest risk quotients.
-For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1 (Figure \ref{fig:fig5}B). 
-In 15\% of all samples (n = 6377) risk quotients higher then 1 were observed.
-Highest RQ were observed for Chlorpyrifos (244), Dimoxystrobin(117) and Isoproturon (80). 
-% TU
-In 33\% of the samples no pesticides were detected. 
-The mean $log_{10}(TU_{max})$ for detects was -4.6.
-2.6\% of the samples showed $log_{10}(TU_{max})$ values greater then -2 (Figure \ref{fig:fig5}C).
-We found a high correlation of $log_{10}(TU_{max})$ and $log_{10}(TU_{sum})$ (r = 0.985, with a maximum higher value of $log_{10}(TU_{sum})$ of 0.75 units (Supplement, Figure S9).
-% see do_polluion.R for numbers
-%Mixtures
-In most samples (55.5\%) more then one compound was detected, with a maximum of 54 different compounds (Figure \ref{fig:fig5}D).
-
-%% Add more numbers / percentages here => compare with knauer/stehle/Vijver
-
-\begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{figure5.pdf}
-  \caption{Overview on pesticide pollution of small agricultural streams (catchment size \textless 30 km\textsuperscript{2} and \textgreater 25\% agriculture). 
-      \textbf{A}: 10 compounds with most frequent EQS exceedances. See B for color legend.
-      \textbf{B}: Distribution of $TU_{max}$ values. Values with $TU_{max} = 0$ are not displayed (3501 out of 10591 samples). Non-detects are not shown due to the logarithmic axis.
-      \textbf{C}: 15 compounds with highest risk quotients. Non-detects are not shown due to the logarithmic axis.
-      \textbf{D}: Distribution of the number of detected compounds in a sample.
-  }
-  \label{fig:fig5}
-\end{figure}
-
-\section{Discussion}
-\subsection{Overview on the compiled dataset}
-The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
-Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
-The data compiled here for Germany is of similar quantity and quality.
-
-% current problems in monitoring and possible solutions
-Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
-There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
-An increase of the analyzed compound spectra to all registered compounds and a lowering of LOQ are desirable for a reliable assessment.
-Our dataset comprised several thousand stream sampling sites, but nearly no data could be compiled for standing waters. 
-Although, standing waters are abundant in the agricultural landscape and may have a high contribution to biodiversity \citep{davies_comparison_2008}, nearly nothing is known on their chemical status.
-Small streams comprise most of the total stream length \citep{nadeau_hydrological_2007}.
-However, sampling sites with catchment sizes below $10km^2$ are underrepresented by the current monitoring scheme (Figure \ref{fig:fig3}).
-This likely arises because catchments under $10km^2$ are not specified by the WFD.
-Nevertheless, small streams should not be neglected for ecological and chemical quality assessment.
-Future monitoring programs should aim at resolving these issues in order to enable a nationwide and representative assessment of freshwaters.
-
-% underestimation
-Given the fixed term monitoring design \citep{stehle_probabilistic_2013} and the low amount of samples taken at precipitation events that could lead to run-off, the data represents a underestimation of the real pollution and peak concentrations.
-Automatic event-drive samplers and passive samplers may help overcome these shortcomings and provide a better representation \citep{fernandez_calibration_2014,moschet_evaluation_2015}.
-Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
-
-\subsection{Influence of catchment area and agriculture}
-% agriculture
-We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides can generally be detected in streams with an increase in fully agricultural catchments (above 75 \% agriculture).
-To our knowledge this is the first study \todo{True? Could not find anything...} investigating such thresholds of pesticide occurrences.
-Our results suggest that only catchments with \textless 25\% agriculture may be regarded as unpolluted and used as reference sites.
-Slightly higher thresholds (20-40\%) have been found for biological responses of lotic assemblages to land-use \citep{waite_agricultural_2014, feld_response_2013}, though they might be differences between taxonomic groups and landscape \citep{feld_response_2013}.
-
-% size
-We did not find support for our hypothesis that small streams are more polluted then big streams, though previous studies have show such a relationship \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
-This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
-For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
-
-\subsection{Pollution of streams}
-% RAC
-\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances organophosphate insecticides. 
-Our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (n = 3516-4745) and shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
-Our results show lower proportions of exceedances compared to \citet{stehle_pesticide_2015}, which can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage.
-This is also reflected in the different sample sizes (1,566 vs. 1,117,988 measurements of compounds with RAC) and the high number of non-detects in the monitoring data.
-Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in pesticide use between countries and defined RACs.
-For three insecticides all detected values where above the RAC.
-This highlights that national monitoring programs need to improve chemical analytics and lower the LOQ for these highly toxic compounds.
-
-% TU
-8.2\% of samples showed $TU_{max}$ values greater than -3 (Figure \ref{fig:fig5}C), a threshold above biological effects have been shown \citep{schafer_thresholds_2012}.
-This suggests that pesticides may act as a common stressor to freshwaters and should not be neglected \citep{schafer_contribution_2016}.
-The high correlation of TU and the relatively low increase of $TU_{sum}$ reveals that in compound mixtures occurring in the field there is a skewed distribution of toxicities with mainly one compound dominating the toxicity.
-% Mixtures
-Nevertheless, most pesticides did not occur individually but in mixtures \cite{schreiner_pesticide_2016} and future risk assessment should take mixtures into account.
-
-% EQS
-For only a limited number of compounds EQS have been defined. 
-We found generally low exceedances of EQS with only three herbicides showing higher exceedances.
-However, compared to the results from RAC comparisons it is clear that the currently defined EQS cannot capture the full risk exposed by pesticides.
-In order to provide a reliable measure of a good chemical status EQS for additional compounds must be defined.
-
-% Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
-The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
+\todo{precipitation section should be moved / more}
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
 
@@
 \begin{acknowledgement}
-The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding this project. 
+The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
 \end{acknowledgement}

### 2016-08-30 supplement/Supplemental_Materials.tex
started revision accoring to ralfs suggestions

@@
 \chapter{Data Cleaning}
-Each of more then 30 datasets have been cleaned and homogenized separately, before combing in a common database.
+More then 30 datasets have been cleaned and homogenized separately, before combing into a common database.
 Cleaning steps comprised (Figure~\ref{fig:data_cleaning} gives a graphical overview).
@@
 	\includegraphics[width = 0.8\textwidth]{data_cleaning}
-	\caption{Overview on data cleaning steps. After cleaning data has been stored in a relational spatial PostgreSQL database.}
+	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning data has been stored in a relational spatial PostgreSQL database.}
 	\label{fig:data_cleaning}
@@
 	\includegraphics[width = 0.8\textwidth]{varclus}
-	\caption{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states. Abbreviations of state names according to ISO 3166-2:DE.}
+	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.
+Abbreviations of state names according to ISO 3166-2:DE.}
 	\label{fig:varclus}

### 2016-08-30 src/do_overview.R

@@ -31,6 +31,28 @@ length(psm_samples[value_fin > 0, value_fin]) / tot * 100
 
 
 
+# overview on Catchment derivation
+table(psm_sites_info$ezg_fin_source, useNA = 'always')
+# NA = no inforamtion available
+# auth = from authorities direct
+# drain = from river segments
+# TRUE = from DEM
+psm_sites_info[is.na(ezg_fin_source)]
+nrow(psm_sites) # 3049 sites in total
+nrow(psm_sites) - sum(table(psm_sites_info$ezg_fin_source))
+# 293 (=10%) without any information
+# 1452 (=47%) with data from authorities
+# 909 (=30%) with data from dem
+# 395 (=13%) with data from stream segments
+
+table(psm_sites_info$agri_fin_source, useNA = 'always')
+# 573 (=18\%) from auhtorities
+nrow(psm_sites) - sum(table(psm_sites_info$agri_fin_source))
+# 650 (=(without information)
+
+nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))])
+# 2376 (=77%) with both informations
+
 
 # Spatial distribution ----------------------------------------------------
 # (=Figure 1)
@@ -51,14 +73,14 @@ p <- ggplot() +
   geom_point(data = psm_sites, aes(x = easting, y = northing, col = state_ab), 
              size = 1) +
   theme(legend.key = element_rect(fill = 'white')) +
-  guides(colour = FALSE) +
+  # guides(colour = FALSE) +
   labs(x = 'Lon.', y = 'Lat.') +
-  scale_color_hue(l = 50) +
+  scale_color_hue(name = 'state', l = 50) +
   mytheme +
   coord_equal()
 p
 ggsave(file.path(prj, "figure1.pdf"),
-       p, width = 3.3, height = 3, units = 'in', dpi = 300, scale = 2)
+       p, width = 3.5, height = 3, units = 'in', dpi = 300, scale = 2)
 
 
 
@@ -202,6 +224,18 @@ dp <- vegdist(vw[ , -1, with = FALSE], method = 'jaccard')
 # hierarchical clustering
 hc <- hclust(dp, method = 'complete')
 plot(hc, labels = vw$site_id)
+# # fusion level plot
+# plot(hc$height, nrow(vw[ , -1, with = FALSE]):2, type = 'S')
[…]

### 2016-08-31 article.tex
refined results, rm EQS and TU

@@
 	\item Can the currently available monitoring data be used for a representative description of the pollution situation?
-	\item Are there thresholds for agricultural land use above pesticide pollution becomes apparent?
-	\item Are there thresholds of stream size below pesticide pollution becomes apparent
+	\item Are there thresholds for agricultural land use and catchment size where pesticide pollution becomes apparent?
 	\item  How polluted are small agricultural streams in Germany and which pesticides are most important? 
 \end{enumerate}
+\todo{add precipitation to aims}
 
@@
 For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
+\todo{check numbers with new catchment calculations from NRW}
 For 10\% of the sites we were not able to compile catchment size data.
@@
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2} and Supplement Figure S2).
-Hierarchical clustering revealed three groups:
+We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
+Hierarchical clustering revealed three groups (see also Supplement Figure S2):
 
@@
   \label{fig:fig2}
+\end{figure}
+
+The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:fig3}).
+The proportion of agriculture in the catchments decreased with increasing catchment size.
+
+\begin{figure}[ht]
+  \includegraphics[width=.8\textwidth]{figure3.pdf}
+  \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
+  Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
+  Colour codes the 2-dimensional density of points.
+  }
+  \label{fig:fig3}
+\end{figure}
+
+\subsection{Thresholds for agricultural land use and catchment size}
+
+Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
+Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:fig4}, left).
+
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right) and no interaction between these two predictors.
+
+\begin{figure}[ht]
+  \includegraphics[width=0.95\textwidth]{figure4.pdf}
+  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+Red line marks statistically significant changes.
+Dashed lines denote 95\% pointwise Confidence Intervals.
+  }
+  \label{fig:fig4}
 \end{figure}

### 2016-08-31 src/do_overview.R

@@ -31,7 +31,21 @@ length(psm_samples[value_fin > 0, value_fin]) / tot * 100
 
 
 
-# overview on Catchment derivation
+# Temporal overview of samples --------------------------------------------
+
+month(psm_samples[2, date])
+monthly <- psm_samples[ , length(unique(sample_id)), by = month(date)]
+ggplot(monthly, aes(x = factor(month), y = V1)) + 
+  geom_bar(stat = 'identity')
+yearlymonthly <- psm_samples[ , length(unique(sample_id)), by = list(year(date), month(date))]
+p <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) + 
+  geom_bar(stat = 'identity') +
+  facet_grid(~year) +
+  labs(x = 'Month', y = 'No. samples')
+ggsave(file.path(prj, 'supplement/temporal.pdf'), p, width = 20, height = 7)
+
+
+# overview on Catchment derivation ----------------------------------------
 table(psm_sites_info$ezg_fin_source, useNA = 'always')
 # NA = no inforamtion available
 # auth = from authorities direct
@@ -224,9 +238,12 @@ dp <- vegdist(vw[ , -1, with = FALSE], method = 'jaccard')
 # hierarchical clustering
 hc <- hclust(dp, method = 'complete')
 plot(hc, labels = vw$site_id)
+
+# fit <- cascadeKM(scale(d, center = TRUE,  scale = TRUE), 1, 10, iter = 1000)
+
 # # fusion level plot
-# plot(hc$height, nrow(vw[ , -1, with = FALSE]):2, type = 'S')
-# text(hc$height, nrow(vw[ , -1, with = FALSE]):2, nrow(vw[ , -1, with = FALSE]):2, col = 'red')
+plot(hc$height, nrow(vw[ , -1, with = FALSE]):2, type = 'S')
+text(hc$height, nrow(vw[ , -1, with = FALSE]):2, nrow(vw[ , -1, with = FALSE]):2, col = 'red')
 #  #silhoutte
 # library(cluster)
 # nr <- nrow(vw[ , -1, with = FALSE])
@@ -234,7 +251,7 @@ plot(hc, labels = vw$site_id)
 # for(k in 2:(nr-1)) {
 #   sil <- silhouette(cutree(hc, k = k), dp)
 # }
-#! continue here
+# #! continue here
 
 # 3 distinct groups: NI+RP; ST+SL+TH; Rest
 bl_groups <- cutree(hc, k = 3)

### 2016-09-12 article.tex
added precipitation analyses & short temporal overview

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Pesticides in small agricultural streams in Germany}
+\title[Pesticides small streams]{Pesticides in small streams in Germany}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
-Especially run-off after heavy precipitation events has been shown to be one of the major input routes \citep{schulz_field_2004}.
+Especially run-off after heavy precipitation events has been shown to be one of the major input routes for pesticides \citep{schulz_field_2004}.
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
+Although, it is known that pesticides pollution and its effect increase with agricultural land-use \todo{citation needed}, studies investigating the shape of this relationships are missing.
 
@@
 Both studies indicate that pesticides might be a threat to biodiversity in the European union. 
-However, theses studies reflect only a small part of available data and it is unclear how representative they are:
-For Germany the study of \citet{malaj_organic_2014} lists only 175 sites and \citet{stehle_pesticide_2015} only 138 measurements, although there might my much more data available from national monitoring programs.
+However, theses studies reflect only a small amount of data and it is unclear how representative they are:
+For Germany the study of \citet{malaj_organic_2014} lists only 175 sites and \citet{stehle_pesticide_2015} only 138 measurements.
 %175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
+National monitoring programs are setup for the surveillance of water quality.
+In Germany these are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
+These programs may provide the most comprehensive data that is currently available for pesticide occurrences in German rivers.
+However, currently there is no curated national-wide compilation of this monitoring data available.
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small waters (SW) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004}.
-It has been show that SWB are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SWB \todo{citation Lorenz et al.}, with only few large scale studies.
+It has been shown that SW are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SW \todo{citation Lorenz et al.}, with only few large scale studies.
 
-We compiled a large data set with chemical monitoring data for Germany and aimed to answer the following research questions:
+In this study we try to fill this gap and analyze nation-wide chemical monitoring data from SWB from Germany.
+Using this large scale dataset we aimed to answer the following research questions:
 
 \begin{enumerate}[(i)]
-	\item Can the currently available monitoring data be used for a representative description of the pollution situation?
-	\item Are there thresholds for agricultural land use and catchment size where pesticide pollution becomes apparent?
-	\item  How polluted are small agricultural streams in Germany and which pesticides are most important? 
+  \item Is currently available monitoring data suitable for a description of pesticide pollution?
+	\item How does the relationship between agricultural land use and catchment size with pesticide pollution look like?
+  Are there thresholds for where pesticide pollution becomes apparent?
+  \item Are samples taken during or shortly after heavy rainfall events or different seasons more likely to indicate chemical pollution?
+	\item  How polluted are SWB in Germany and which pesticides currently pose the biggest threat to them?
 \end{enumerate}
-\todo{add precipitation to aims}
+
+Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.
 
@@
 We homogenized and unified all data from the states into a common database.
-We implemented a robust and reproducible data cleaning workflow (see supplement for details on the data processing workflow) \citep{poisot_best_2015}.
+We implemented a robust data cleaning workflow (see supplemental figure S1 for details on data processing) \citep{poisot_best_2015}.
 Nevertheless, parts of the dataset are proprietary and cannot be shared here.
@@
 Given the relevance of precipitation in causing runoff events, we identified chemical samples taken during heavy rainfall events.
-We performed a spatio-temporal intersection of sampling events with gridded daily precipitation data available the German Weather service \citep{rauthe_central_2013}. 
+We performed a spatio-temporal intersection of sampling events with gridded daily precipitation data available the German Weather service.
+This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
 We performed the intersection for the actual sampling date and the day before.
+
+\subsection{Characterization of catchments}
+We compiled a total of 3,049 sampling sites with pesticide measurements.
+We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
+Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
+The delineation algorithm produced only for 30\% of the sites accurate results.
+For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
+For 10\% of the sites we were not able to compile catchment size data.
+For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
+For 78\% of the sites both, the proportion of agricultural landuse and catchment size were available.
+% see do_overview.R for numbers
 
@@
 
-\subsection{Characterization of catchments}
-We compiled a total of 3,049 sampling sites with pesticide measurements.
-We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
-Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
-The delineation algorithm produced only for 30\% of the sites accurate results.
-For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
-\todo{check numbers with new catchment calculations from NRW}
-For 10\% of the sites we were not able to compile catchment size data.
-For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
-We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
-For 78\% of the sites both, the proportion of agricultural landuse and catchment size were available.
-% see do_overview.R for numbers
-
 \subsection{Statistical analyses}
@@
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances as:
+We modeled the number of RAC exceedances (RQ \textgreater 1) as:
 
@@
 \begin{split}
-  No_i \sim NB(\mu_i, k) \\
-  E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{k} \\
+  No_i \sim NB(\mu_i, \kappa) \\
+  % E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{\kappa} \\
   log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i) + log(n_i) \\
@@
 where $No_i$ is the observed number of exceedances at site $i$. 
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
 The proportion of agriculture within the catchment ($Agri_i$) and the catchment size of the site ($Size_i$) were used as predictors. 
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
-The number of sampling events per site ($n_i$) was as an offset to account different sampling efforts at a site and is equivalent to modeling the rate of exceedances. 
-$f_1$ and $f_2$ are smoothing functions using thin plate regression splines \citep{wood_thin_2003}.
+$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-We used point-wise 95\% Confidence Intervals of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
+The number of sampling events per site ($n_i$) was used as an offset to account different sampling efforts at a site and is equivalent to modeling the rate of exceedances. 
+We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
 GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
-\todo{precipitation stats?}
+
+\todo{Überleitung}
+RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
+Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \todo{Cite rigby 2010}:
+
+\begin{align}
+y \sim ZAGA(\mu, \pi) = 
+  \begin{cases}
+    (1 - \pi)   & \quad  \text{if } y < LOQ \\
+    \pi \times f_{Gamma} (\mu) & \quad \text{if } y > LOQ \\
+  \end{cases}
+  \label{eqn:eqn3}
+\end{align}
+
+$\pi$ denotes the probability of a observation being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ.
+We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for the mean RQ ($\mu$) and the $\pi$:
+
+\begin{align}
+\begin{split}
+\log(\mu) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
+logit(\pi) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
+\end{split}
+\label{eqn:eqn4}
+\end{align}
+
+Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\pi$ can be interpreted as changes in the probability that there is any risk.
+To account for temporal autocorrelation and differences between federal states we used the site nested within state as random factors.
+We fitted this model separatedly to each compound with a RAC, with at least 500 samples and more then 5\% of values above quantification limit.
+
+\todo{check formulas, add i, add random effect? }
+
+\todo{1 - remove columns from appendix table, 
+2- keep only those with RACs,
+3- indicate which where used for modeling here
+}
+
+\todo{Add definition of swb.}
 
@@
 9\% of samples from 33 sites were taken as composite samples of different durations.
-Therefore, we report only results of grab samples from streams. 
+Therefore, we restricted the analyses to grab samples from streams. 
 The analyzed dataset comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
+Most samples were take in the months April till October, with less samples in the winter (see Supplemental Figure S2).
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
@@
   \includegraphics[width=\textwidth]{figure2.pdf}
-  \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
+  \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
   Colors according to three groups determined by hierarchical clustering (see Supplement Figure S2).}
@@
   Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
-  Colour codes the 2-dimensional density of points.
-  }
+  Colour codes the 2-dimensional density of points.}
   \label{fig:fig3}
@@
 \subsection{Thresholds for agricultural land use and catchment size}
-
 Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
 Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:fig4}, left).
-
-We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right) and no interaction between these two predictors.
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right).
+We also found no statistically significant interaction between catchment size and agriculture.
 
@@
 
-\todo{precipitation section should be moved / more}
+\subsection{Effect of precipitation on exposure}
+
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
+
+\begin{figure}[ht]
+  \includegraphics[width=0.95\textwidth]{figure5.pdf}
+  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4}.
+Left column: Effect on mean RQ; Right column: effect on the probability that RQ \textgreater 0.
+  Coefficient where the CI does not encompasses zero are shown in black.
+  }
+  \label{fig:fig5}
+\end{figure}
+
+\subsection{Pesticide pollution of small water bodies}
+
+\section{Discussion}
+\subsection{Overview on the compiled dataset}
+The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
+Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
+The data compiled and analysed here for Germany is of similar quantity and quality.
 

### 2016-09-12 src/do_overview.R

@@ -180,7 +180,7 @@ var_tab_x <- xtable(var_tab,
                                          epa: US EPA \\citep{u.s._epa_ecotoxicology_2015};
                                          malaj:\\citep{malaj_organic_2014};
                                          ppdb: Pesticides Properties database \\citep{lewis_international_2016};
-                                         none: no LC50 could be found.
+                                         none: no LC50 available.
                     \\textsuperscript{e} Maximum Anual Concentration Environmental Quality Standard [ug/L].
                     \\textsuperscript{f} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).',
                     align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1.5cm}p{1cm}p{1cm}p{1cm}')

### 2016-09-12 src/do_pollution.R

@@ -126,5 +126,5 @@ pmix <- ggplot(mix, aes(x = no_subs)) +
 
 pall <- plot_grid(peqs, ptu, prac, pmix, labels = c('A', 'B', 'C', 'D'), 
                   label_size = 20)
-ggsave(file.path(prj, "figure5.pdf"), pall, width = 7, height = 6.5, 
+ggsave(file.path(prj, "figure6.pdf"), pall, width = 7, height = 6.5, 
        units = 'in', dpi = 300, scale = 1.5)

### 2016-09-12 src/load.R

@@ -69,7 +69,8 @@ library(gamlss)
 library(gamlss.cens)
 library(gamlss.dist)
 library(gamlss.add)
-
+library(lme4)
+library(nlme)
 
 ### ----------------------------------------------------------------------------
 ### Source defined functions

### 2016-09-13 article.tex
fixed model description

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Pesticides in small streams in Germany}
+\title[Pesticides small streams]{Pesticide exposure in small streams in Germany}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 
-In this study we try to fill this gap and analyze nation-wide chemical monitoring data from SWB from Germany.
-Using this large scale dataset we aimed to answer the following research questions:
+In this study we try to fill this gap and analyze large scale nation-wide chemical monitoring data from Germany.
+First, we revise the available monitoring data if it is suitable for a large scale description of pesticide pollution.
+Then we analyse the relationship between agricultural land use and catchment size with pesticide pollution and investigate potential thresholds.
+We also investigate whether samples taken during or shortly after heavy rainfall events or different seasons are more likely to indicate chemical pollution.
+Finally, we give an overview on the pollution of SWB in Germany.
 
-\begin{enumerate}[(i)]
-  \item Is currently available monitoring data suitable for a description of pesticide pollution?
-	\item How does the relationship between agricultural land use and catchment size with pesticide pollution look like?
-  Are there thresholds for where pesticide pollution becomes apparent?
-  \item Are samples taken during or shortly after heavy rainfall events or different seasons more likely to indicate chemical pollution?
-	\item  How polluted are SWB in Germany and which pesticides currently pose the biggest threat to them?
-\end{enumerate}
-
-Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.
+% Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.
 
@@
 We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
-For 78\% of the sites both, the proportion of agricultural landuse and catchment size were available.
+For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
 % see do_overview.R for numbers
@@
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of sampling events per site ($n_i$) was used as an offset to account different sampling efforts at a site and is equivalent to modeling the rate of exceedances. 
+The number of samples per site ($n_i$) was used as an offset to account different sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 
-\todo{Überleitung}
+While agricultural land use and catchment size vary only between sites, this is not the case for precipitation that changes also with time.
+Therefore, we modeled the effects of precipitation in a separate model.
 RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
@@
 \begin{align}
-y \sim ZAGA(\mu, \pi) = 
+RQ_i \sim ZAGA(\mu_i, \pi_i) = 
   \begin{cases}
-    (1 - \pi)   & \quad  \text{if } y < LOQ \\
-    \pi \times f_{Gamma} (\mu) & \quad \text{if } y > LOQ \\
+    (1 - \pi_i)   & \quad  \text{if } y < LOQ \\
+    \pi_i \times f_{Gamma} (\mu_i) & \quad \text{if } y \ge LOQ \\
   \end{cases}
@@
 
-$\pi$ denotes the probability of a observation being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ.
-We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for the mean RQ ($\mu$) and the $\pi$:
+$\pi_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, where $\mu$ is the mean of the gamma function.
+We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\pi$.
+To account for temporal autocorrelation and differences between federal states we used the site nested within state as random intercepts:
 
@@
 \begin{split}
-\log(\mu) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
-logit(\pi) = \beta_0 + \beta_1 \times prec_0 + \beta_2 \times prec_{-1} + \beta_{3-5} \times season_{Q2-Q4} \\
+\log(\mu_i) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+logit(\pi_{ijk}) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+state_{j} \sim N(0, \sigma^2_{state}) \\
+site_{jk} \sim N(0, \sigma^2_{site})
 \end{split}
@@
 Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\pi$ can be interpreted as changes in the probability that there is any risk.
-To account for temporal autocorrelation and differences between federal states we used the site nested within state as random factors.
-We fitted this model separatedly to each compound with a RAC, with at least 500 samples and more then 5\% of values above quantification limit.
-
-\todo{check formulas, add i, add random effect? }
-
-\todo{1 - remove columns from appendix table, 
-2- keep only those with RACs,
-3- indicate which where used for modeling here
-}
+We fitted this model separately to each compound with a RAC, with at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
+We implemented the model using the gamlss packages \cite{}
 
@@
 
-In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2 \todo{remove LC50 and MAC-EQS from table!}). 
+In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
@@
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
+Precipitation at the day before sampling increases $\mu$ and also $\pi$ for nearly all modeled compounds. 
+The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling (Figure \ref{fig:fig5}, top).
+
+Quarters 2-4 showed higher RQ than in Q1, except for the compounds  (Figure \ref{fig:fig5}, bottom)
 
@@
   \includegraphics[width=0.95\textwidth]{figure5.pdf}
-  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4}.
+  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Table .
 Left column: Effect on mean RQ; Right column: effect on the probability that RQ \textgreater 0.
-  Coefficient where the CI does not encompasses zero are shown in black.
+  Coefficients where the CI encompasses zero are shown in grey.
   }

### 2016-09-13 src/do_overview.R

@@ -150,40 +150,36 @@ setkey(var_props, variable_id)
 var_tab <- var_props[var_tab] 
 var_tab[ , variable_id := NULL]
 
+setcolorder(var_tab, c(names(var_tab)[3:5], names(var_tab)[8:9], names(var_tab)[1:2], names(var_tab)[6:7]))
+var_tab <- var_tab[ , c(1,2,3,4, 5, 9), with = FALSE]
+
 var_tab[ , ger_auth_2015 := as.character(ger_auth_2015)]
 var_tab[ger_auth_2015 == 'FALSE', ger_auth_2015 := NA]
 var_tab[ger_auth_2015 == 'TRUE', ger_auth_2015 := 'x']
 var_tab[ , eu_auth_2015 := as.character(eu_auth_2015)]
 var_tab[eu_auth_2015 == 'FALSE', eu_auth_2015 := NA]
 var_tab[eu_auth_2015 == 'TRUE', eu_auth_2015 := 'x']
-setcolorder(var_tab, c(names(var_tab)[3:5], names(var_tab)[8:9], names(var_tab)[1:2], names(var_tab)[6:7]))
+
 names(var_tab) <- c('Name', 'CAS', 'Group', 
                     'Auth. GER\\textsuperscript{a}', 
                     'Auth. EU\\textsuperscript{b}',
-                    'LC50\\textsubscript{D.magna}\\textsuperscript{c}',
-                    'Source LC50\\textsuperscript{d}',
-                    'MAC-EQS\\textsuperscript{e}', 
-                    'RAC \\textsuperscript{f}'
+                    'RAC \\textsuperscript{c}'
                     )
 var_tab$Group <- gsub('organics, psm, ', '', var_tab$Group)
 
 # fix bug with encoding (extra space or so....)
 var_tab[Name == 'Benzoesäure', CAS := '65-85-0']
 
+
+
+
 var_tab_x <- xtable(var_tab, 
                     label = 'tab:phch_var',
-                    caption = 'Analysed chemical compounds. \\
+                    caption = 'Overview on pesticides in the database. \\
                     \\textsuperscript{a} Authorized in Germany (Source: BVL, 2015). 
                     \\textsuperscript{b} Authorized in the EU (Source: EU).
-                    \\textsuperscript{c} [ug/L].
-                    \\textsuperscript{d} chemprop: Read-Across \\citep{schuurmann_quantitative_2011};
-                                         epa: US EPA \\citep{u.s._epa_ecotoxicology_2015};
[…]

### 2016-09-13 src/do_precip.R

@@ -100,9 +100,36 @@ hist(props[tot < 1000, tot])
 keep <- props[prop > 0.05 & tot > 1000]
 # 25 compounds left
 
+# export overview for supplement
+keep
+setkey(psm_variables, variable_id)
+setkey(keep, variable_id)
+keep_tab <- psm_variables[ , list(variable_id, name, cas, psm_type)][keep]
+keep_tab[ , prop := round(prop*100, 2)]
+setnames(keep_tab, c('id', 'Name', 'CAS', 'Group', '\\%>LOQ', 'no. > LOQ', 'total no.'))
+keep_tab$id <- NULL
+keep_tab
+
+keep_tab_x <- xtable(keep_tab, 
+                    label = 'tab:var_model',
+                    caption = '24 pesticides for which we modelled the relationship with precipitation and seasonality.',
+                    align = 'lp{3cm}rlp{1cm}p{1cm}p{1cm}')
+
+print(keep_tab_x, 
+      file = file.path(prj, 'supplement/keeptab.tex'),
+      tabular.environment="longtable",
+      floating = FALSE,
+      caption.placement = 'top',
+      comment = FALSE,
+      booktabs = TRUE,
+      hline.after = c(-1, 0),
+      sanitize.text.function = identity,
+      size="\\fontsize{8pt}{10pt}\\selectfont"
+)
+
+
+
 take <- take[variable_id %in% keep$variable_id]
-
-
 psm_variables[variable_id %in% keep$variable_id, list(variable_id, name, psm_type)]
 
 
@@ -248,9 +275,35 @@ plot(p)
 ggsave(file.path(prj, "figure5.pdf"), p, width = 10, height = 9)
 
 
+resdf
 
 
 
+
+
+# extract random effect variances
+
+# function to extract the needed model components, 
+ra_extr <- function(file){
+  message('Working on file: ', file)
+  mod <- readRDS(file)
+  lme_out <- mod$mu.coefSmo[[1]]
+  vc <- VarCorr(lme_out)
+  suppressWarnings(storage.mode(vc) <- 'numeric')
+  vc <- vc[c(2, 4), 'StdDev']
+  names(vc) <- c('state/site', 'state')
+  vc <- c(vc, var = gsub('mod_(.*)\\.rds', '\\1', basename(file)))
+   # generalized rsq
+  vc <- c(vc, rsq = Rsq(mod, type = "Cox Snell"))
[…]

### 2016-09-14 article.tex
finalised precipitation and size width relationships

@@
 \section{Introduction}
+\todo{Check all references to supplement}.
 More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
@@
 It has been shown that SW are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SW \todo{citation Lorenz et al.}, with only few large scale studies.
+Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SW, with only few large scale studies \citep{lorenz_specifics_2016}.
 
@@
 
-\subsection{Characterization of chemical pollution}
+A clear definition of SWB in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
+The WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further consideration of streams \textless 10km\textsuperscript{2}. 
+\citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
+Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure \ref{fig:fig3}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
+This catchment size corresponds to a stream width of approximately 2~meters (see Supplement Figure S2).
+
+\subsection{Characterization of pesticide pollution}
 We characterized pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
@@
 RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \todo{Cite rigby 2010}:
+Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
 
@@
 We fitted this model separately to each compound with a RAC, with at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
-We implemented the model using the gamlss packages \cite{}
-
-\todo{Add definition of swb.}
+We implemented the model using the gamlss package \cite{stasinopoulos_generalized_2007}.
 
@@
 
-\subsection{Effect of precipitation on exposure}
+\subsection{Effect of precipitation on pesticide exposure}
 
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
-Precipitation at the day before sampling increases $\mu$ and also $\pi$ for nearly all modeled compounds. 
-The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling (Figure \ref{fig:fig5}, top).
+Precipitation at the day before ($precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all modeled compounds. 
+The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling ($precip_{0}$)(Figure \ref{fig:fig5}, top).
 
-Quarters 2-4 showed higher RQ than in Q1, except for the compounds  (Figure \ref{fig:fig5}, bottom)
+Quarters 2-4 showed generally higher $\mu$ than in Q1. 
+Exceptions were Diflufenican, that showed lower RQ in Q2-Q4 compared to Q1.
+Smaller negative effects were found for Clothianidin (Q2) and Isoproturon (Q3) (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
+We found similar effects on $\pi$.
 
-\begin{figure}[ht]
+\begin{figure}[h]
   \includegraphics[width=0.95\textwidth]{figure5.pdf}
-  \caption{Estimated coefficients and their 95\% CI for the model describe in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Table .
-Left column: Effect on mean RQ; Right column: effect on the probability that RQ \textgreater 0.
-  Coefficients where the CI encompasses zero are shown in grey.
+  \caption{Estimated coefficients and their 95\% CI for the model described in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Tables S3 and S4 for a list of compounds from top to bottom and their respective coefficients).
+Left column: Effect on mean RQ ($\mu$); Right column: effect on the probability that RQ \textgreater 0 ($\pi$).
+  Coefficients where the CI encompasses zero are shown in gray colour.
   }
@@
 
-\subsection{Pesticide pollution of small water bodies}
+\subsection{Pesticides in small water bodies}
 

### 2016-09-14 supplement/Supplemental_Materials.tex
fixed model description
finalised precipitation and size width relationships

@@
 
+\chapter{Catchment size - stream width relationships}
+We studied the relationship between catchment size based on three datasets containing this informations:
+Data delivered by the federal state Thuringia, \citet{vos_organic_2015} and \citet{fernandez_effects_2015} (both from Rhineland-Palatinate).
+We fitted to each dataset separately and to the combined dataset a power-function.
+The resulting models are shown in Figure \ref{fig:size_width}.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{width_size}
+	\caption[Relationship between catchment size and stream width.]{Relationship between catchment size and stream width.
+A power function has been fitted to each dataset separately and the combined dataset (black line and equation).
+The gray rectangle marks the estimated with for a catchment size of 25km\textsuperscript{2}.}
+	\label{fig:size_width}
+\end{figure}
+
 \chapter{Overview on compiled data}
@@
 \input{phchoverview}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = \textwidth]{temporal}
+	\caption{Number of sampling occasions per year and month.}
+	\label{fig:temporal}
+\end{figure}
 
@@
 
+\chapter{Effect of precipitation and season on RQ}
 \begin{figure}[h]
@@
 
-\chapter{Are small agricultural waters more polluted compared to bigger streams?}
+\input{keeptab}
+\input{keeptab2}
 
-\section{RAC Exceedances}
+\chapter{Pollution of small streams}
 \begin{figure}[h]

### 2016-09-14 src/do_precip.R

@@ -106,14 +106,15 @@ setkey(psm_variables, variable_id)
 setkey(keep, variable_id)
 keep_tab <- psm_variables[ , list(variable_id, name, cas, psm_type)][keep]
 keep_tab[ , prop := round(prop*100, 2)]
-setnames(keep_tab, c('id', 'Name', 'CAS', 'Group', '\\%>LOQ', 'no. > LOQ', 'total no.'))
+setnames(keep_tab, c('id', 'Compound', 'CAS', 'Group', '\\%>LOQ', 'no. > LOQ', 'total no.'))
 keep_tab$id <- NULL
 keep_tab
 
 keep_tab_x <- xtable(keep_tab, 
                     label = 'tab:var_model',
-                    caption = '24 pesticides for which we modelled the relationship with precipitation and seasonality.',
-                    align = 'lp{3cm}rlp{1cm}p{1cm}p{1cm}')
+                    caption = '24 pesticides for which we modelled the relationship with precipitation and seasonality.
+                    Order is the same as in Figure 5 of the articles. See Table \\ref{tab:var_model_coef} for model coefficients.',
+                    align = 'lp{2.5cm}rlp{1.5cm}p{2cm}p{2cm}')
 
 print(keep_tab_x, 
       file = file.path(prj, 'supplement/keeptab.tex'),
@@ -122,7 +123,7 @@ print(keep_tab_x,
       caption.placement = 'top',
       comment = FALSE,
       booktabs = TRUE,
-      hline.after = c(-1, 0),
+      hline.after = c(-1, 0, nrow(keep_tab)),
       sanitize.text.function = identity,
       size="\\fontsize{8pt}{10pt}\\selectfont"
 )
@@ -217,32 +218,88 @@ resdf$cisig <- ifelse(sign(resdf$upci)  == sign(resdf$lowci), 'cisig', 'nocisig'
 
 resdf$termind <- substr(resdf$term, 1, 2)
 
-# display as tile plot
-ggplot() +
-  geom_tile(data = resdf[!grepl('Intercept|si|mu.pre|nu', resdf$term, )], aes(x = term, y = factor(variable), fill = estsig)) +
-  scale_fill_gradient2(na.value = "white") +
-  ggtitle('mu.season')
+# export other table
+keep_tab2 <- resdf[!grepl('Intercept', term) & term_type != 'sigma', 
[…]

### 2016-09-14 src/do_size_width.R

@@ -1,6 +1,7 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project'!")
+       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
 }
@@ -12,116 +13,74 @@ if (!exists("prj")) {
 
 ### ----------
 # TH has good data 
-#! interessting data about ezg - width
-th <- setDT(read_excel(file.path(datadir, 'physico_chem/TH/new/', 'Thueringen_PSM_FG_2005-14.xlsx')))
-th <- th[ , c(8,10), with = FALSE]
+th <- fread(file.path(cachedir, 'TH_2005-14.csv'))
+th <- th[ , c(4, 5), with = FALSE]
 setnames(th, c('ezg', 'width'))
 th <- th[complete.cases(th)]
 th[ , dataset := 'TH']
 
-ggplot(th, aes(x = ezg, y = width)) +
-  geom_point() +
-  scale_x_log10() +
-  scale_y_log10() 
-
-# log-log linear model
-loglm <- lm(log(width) ~ log(ezg), data = th)
-coef(loglm)
-summary(loglm)
-# sum of squared error
-sum((th$width - exp(fitted(loglm)))^2)
-
-# power model
-nlm <- nls(width ~ a*ezg^b, data = th, start = list(a = exp(coef(loglm)[1]), b = coef(loglm)[2]))
-summary(nlm)
-sum((th$width - fitted(nlm))^2)
-# better odel
-
-pdat <- data.frame(ezg = seq(1, 10000, length.out = 1000))
-pdat$ploglm <- predict(loglm, newdata = pdat)
-pdat$pnlm <- predict(nlm, newdata = pdat)
-
-ggplot() +
-  geom_point(data = th, aes(x = ezg, y = width)) +
-  scale_x_log10() +
-  scale_y_log10() +
-  geom_line(data = pdat, aes(x = ezg, y = pnlm), col = 'blue') +
-  geom_hline(aes(yintercept = 2)) +
-  mytheme
-
-
 ### ----------
 ### data from voss 2015
[…]

### 2016-09-14 article.tex
added detection rates and mixtures

@@
 \subsection{Effect of precipitation on pesticide exposure}
-
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
@@
 \subsection{Pesticides in small water bodies}
+The dataset comprised 12,887 samples from 1313 small water bodies.
+1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher then 1 were observed.
+Neonicotinoid Insecticides and Chlorpyrifos showed highest risk quotients (Figure \ref{fig:fig6}).
+For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1. 
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin showed also high exceedances of RQ (29.5/14.2/21.4 \% of all samples with detects).
+Highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
+Metabolites were most commonly detected in samples (Metazachlor sulfonix acid was in 82\% of all samples were it was analysed) (Supplement Figure S8)
+Glyphosate was the compound with highest detection rates, followed by Boscalid and Isoproturon. 
+However, only the latter showed RQ exceedances (Figure \ref{fig:fig6}).
+In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one samples (supplemental figure S7).
 
+\begin{figure}[h]
+  \includegraphics[width=0.6\textwidth]{figure6.pdf}
+  \caption{15 compounds with highest risk quotients in SWB.
+Non-detects are not shown due to the logarithmic axis.
+Numbers on the right give the percentage of RQ \textgreater 0 and total number of samples.
+  }
+  \label{fig:fig6}
+\end{figure}
+
+%% -------------------------------------------------------------------------
 \section{Discussion}

### 2016-09-14 src/do_pollution.R

@@ -23,14 +23,14 @@ psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin')]
 neo_id <- psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin'), variable_id]
 psm_samples[variable_id %in% neo_id, length(unique(sample_id)), by = variable_id]
 
-# n measurements with RCA
+# n measurements with RAC
 psm_samples[variable_id %in% psm_variables[!is.na(rak_uba), variable_id]]
 
 
 # Classify stream types ---------------------------------------------------
 
-# Analyze only small agricultural streams (catchment < 30km2 and >25% agriculture)
-take_sites <- psm_sites_info[ezg_fin < 30 & agri_fin > 0.25, site_id]  
+# Analyze only small streams (catchment < 25km2)
+take_sites <- psm_sites_info[ezg_fin < 25, site_id]  
 take_samples <- psm_samples[site_id %chin% take_sites]
 
 # no. observations
@@ -50,10 +50,10 @@ length(take_samples[variable_id == 457, value_fin])
 # Risk Quotients ----------------------------------------------------------
 # restrict samples to variables with rac
 rac <- psm_variables[!is.na(rak_uba), list(variable_id, name, cas, pgroup, rak_uba)]
-# join eqs
+# join rac with samples
 setkey(take_samples, variable_id)
 setkey(rac, variable_id)
-take_rac <- take_samples[value_fin > 0][rac]
+take_rac <- take_samples[rac]
 # rm left over from left join
 take_rac <- take_rac[!is.na(site_id)]
 
@@ -65,7 +65,7 @@ setkey(take_rac, variable_id)
 setkey(psm_variables, variable_id)
 dd <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 
-# calculate media per variable and take only first 15 hits
+# calculate max per variable and take only first 15 hits
 take_var <- dd[ , list(media = max(rq)) , by = variable_id][order(media, decreasing = TRUE)][1:15, variable_id]
 take_dd <- dd[variable_id %in% take_var]
 
@@ -80,22 +80,77 @@ rak_samples <- length(unique(dd$sample_id))
 # percentake of samples with RQ > 1In 
[…]

### 2016-09-14 article.tex
More discussion & results...

@@
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
-Although, it is known that pesticides pollution and its effect increase with agricultural land-use \todo{citation needed}, studies investigating the shape of this relationships are missing.
+Although, it is known that pesticides pollution and its effect increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of this relationships are missing.
 
@@
 
-Small waters (SW) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
-However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004}.
-It has been shown that SW are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Despite their relevance only a small fraction of studies were conducted on pesticide pollution of SW, with only few large scale studies \citep{lorenz_specifics_2016}.
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+It has been shown that SWB are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Especially, compounds from agricultural use show seasonal rain-event-driven peak concentrations in SWB \citep{wittmer_significance_2010}.
+Despite their biological relevance and potential pesticide exposure only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
+Moreover, it is currently unknown how precipitation might influence measured pesticide concentrations in national monitoring programs.
 
-In this study we try to fill this gap and analyze large scale nation-wide chemical monitoring data from Germany.
+In this study we try to fill these gaps and analyze large scale nation-wide chemical monitoring data from Germany.
 First, we revise the available monitoring data if it is suitable for a large scale description of pesticide pollution.
 Then we analyse the relationship between agricultural land use and catchment size with pesticide pollution and investigate potential thresholds.
-We also investigate whether samples taken during or shortly after heavy rainfall events or different seasons are more likely to indicate chemical pollution.
-Finally, we give an overview on the pollution of SWB in Germany.
+We also investigate whether samples taken during or shortly after heavy rainfall events or different seasons are more likely to indicate pesticide pollution.
+Finally, we give an overview on pesticide pollution of SWB in Germany.
 
@@
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
-Precipitation at the day before ($precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all modeled compounds. 
+Precipitation at the day before ($precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all analysed compounds. 
 The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling ($precip_{0}$)(Figure \ref{fig:fig5}, top).
@@
 Quarters 2-4 showed generally higher $\mu$ than in Q1. 
-Exceptions were Diflufenican, that showed lower RQ in Q2-Q4 compared to Q1.
+Exceptions were Diflufenican, that showed significantly lower RQ in Q2-Q4 compared to Q1.
 Smaller negative effects were found for Clothianidin (Q2) and Isoproturon (Q3) (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
@@
 \subsection{Pesticides in small water bodies}
-The dataset comprised 12,887 samples from 1313 small water bodies.
+The dataset comprised 12,710 samples from 1295 small water bodies.
 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher then 1 were observed.
@@
 
+% current problems in monitoring and possible solutions
+Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
+There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
+For Thiacloprid, Imidaclorpid and Chlorpyrifos LOQ were above RAC.
+For these compounds a lowering of LOQ is essential for reliable assessment.
+Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds.
+
+\todo{unfertig}
+
+\subsection{Thresholds for agricultural land use and catchment size}
+% agriculture
+We found a strong influence of agriculture on the pollution of streams.
+If there is more the 25\% agriculture within a catchment pesticides it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
+To our knowledge this is the first study investigating such thresholds of pesticide exposure.
+Thresholds for agricultural land use have only been investigated in the past for biological communities.
+\citet{feld_response_2013} found change points of biological community metric for agricultural land use at 40\% in low lands. 
+Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
+Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological effects.
+
+% size
+We did not find a relationship between pesticide pollution and catchment size.
+However, previous studies have shown that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
+For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
+Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
+
+\subsection{Effect of precipitation on pesticide exposure}
+Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
+However, we found higher pesticide exposure if samples where taken at the after heavy rainfall events. 
+Interestingly, samples take at they day of a heavy-rainfall-event did not show such a strong relationship.
+This could be explained by a sampling just before the actual rainfall event.
+Pesticide concentration in agricultural SWB show short term peak concentrations \citep{wittmer_significance_2010}.
+Overall, our results indicate that current chemical monitoring strongly underestimates pesticide exposure.
+
+\todo{add alternatives - passive sampler, automatic samplers...}
+
+We found pesticide specific patterns in seasonality of the studied compounds.
+Most pesticides showed an increase from April - September, which coincides with their main application season.
+Hence, we also found the hightest RQ for the herbicide Diflufenican during the winter/spring season, which is application period for which it is registered \todo{cite BVL}.
+
+\subsection{Pesticides in small water bodies}
+Our results suggest that SWB are frequently exposed to biologically relevant pesticide concentrations.
+\citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
+By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1290 and 2113 samples in SWB, Figure \ref{fig:fig6}).
+Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
+Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
+This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
+This might reflect differences in pesticide use between countries and defined RACs.
+
+From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
+Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at RQ of 1.12.
+We found RQ values greater than 1.12 at 25\% of all SWB sites. 
+Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and services they provide. 
+Additional treat arises when taking into account that most pesticides do not occur individually but in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
+
+% Approval / Risk Assessment
+Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
+The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % abstract    :
-% 3 big(600)  : 1800
+% 4 big(600)  : 2400
 % 2 small(300): 600

### 2016-09-14 src/do_pollution.R

@@ -89,6 +89,10 @@ nrow(dd[rq >1]) / nrow(dd) * 100
 nrow(dd[rq >1]) / nrow(dd[rq>0]) * 100
 
 
+length(unique(dd[rq > 1.12, site_id])) / length(unique(dd[, site_id]))* 100
+
+
+
 # calculate percentage of non-detects
 loqd <- take_dd[ , list(tot = length(rq),
                 n_n0 = sum(rq > 0),
@@ -126,6 +130,10 @@ psm_variables[name %like% c('Nicosu') | name %like% c('Diflufe') | name %like% c
 nrow(take_dd[variable_id %in% c(272) & rq > 1]) / nrow( take_dd[variable_id %in% c(272) & rq > 0]) * 100
 nrow(take_dd[variable_id %in% c(282) & rq > 1]) / nrow( take_dd[variable_id %in% c(282) & rq > 0]) * 100
 nrow(take_dd[variable_id %in% c(457) & rq > 1]) / nrow( take_dd[variable_id %in% c(457) & rq > 0]) * 100
+
+# number of samples for Thiaclorpid
+take_samples[variable_id == 588]
+
 
 # detects
 dt <- take_samples[ , list(tot = length(value_fin),

### 2016-09-15 supplement/Supplemental_Materials.tex

@@
 
+%% -------------------------------------------------------------------------
 \title{Supplemental Materials for the paper: \\Pesticides pollution of small streams in
@@
 
+%% -------------------------------------------------------------------------
 \begin{document}
@@
 
+%% -------------------------------------------------------------------------
 \chapter{Data Cleaning}
@@
 
+%% -------------------------------------------------------------------------
 \chapter{Catchment size - stream width relationships}
@@
 
+%% -------------------------------------------------------------------------
 \chapter{Overview on compiled data}
@@
 
+%% -------------------------------------------------------------------------
+\chapter{Thresholds for agricultural land use and catchment size}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
+	\caption{Raw data used for the model in equation 2 and Figure 3 of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
+	\label{fig:ezgagrirac}
+\end{figure}
+
+%% -------------------------------------------------------------------------
 \chapter{Effect of precipitation and season on RQ}
@@
 	\centering
-	\includegraphics[width = 0.6\textwidth]{precip}
+	\includegraphics[width = 0.8\textwidth]{precip}
 	\caption{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
@@
 
-\chapter{Pollution of small streams}
+%% -------------------------------------------------------------------------
+\chapter{Pesticides in small water bodies}
+
 \begin{figure}[h]
 	\centering
-	\includegraphics[width = 0.7\textwidth]{ezgagrirac}
-	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
-	\label{fig:ezgagrirac}
+	\includegraphics[width = 0.8\textwidth]{pmix}
+	\caption{Distribution the number of quantified compounds in the samples from small water bodies.}
+	\label{fig:pmix}
 \end{figure}
-\todo{update refs}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{pdetects}
+	\caption{Proportion of samples with detects in SWB. Only Compounds with more than 100 samples and 15\% of detects are show.}
+	\label{fig:pdetects}
+\end{figure}
 

### 2016-09-15 article.tex
second draft

@@
 
-\usepackage{todonotes}
+% \usepackage{todonotes}
 
@@
 Small streams are important refugia for biodiversity.
-In agricultural areas, they may be at high risk of pesticide pollution.
+In agricultural areas, they may be at high risk from pesticides that enter streams during rainfall. 
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
 In Germany, pesticide monitoring is performed by the federal states as part of water quality surveillance.
+
+We compiled monitoring data focusing on small streams, resulting in a data set of 2,918,604 measurements from 42,236 samples in 3,049 sampling sites covering the years 2005-2014.
+A total of 484 different compounds that can be classified as pesticides were measured.
+We studied the relationships between pesticide exposure and agricultural land use, catchment size, as well as precipitation and seasonal dynamics.
+
+We found a threshold of 25\% agricultural land use, above pesticide exposure becomes apparent. 
+Precipitation increased pesticide concentrations and these showed pesticide specific temporal patterns.
+Especially neonicotinoid insecticides and Chlorpyrifos showed exceedances of regulatory acceptable concentrations (RACs) in small water bodies.
+
+We conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and services they provide.
+
+% Total words: 
 \end{abstract}
@@
 \section{Introduction}
-\todo{Check all references to supplement}.
 More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
@@
 \citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that most European water bodies are at risk from pesticides.
-\citet{stehle_pesticide_2015} compiled 1566 measured concentrations of 23 insecticides in the EU from scientific publications. 
+\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
 They found that many of these measurements exceed regulatory acceptable concentrations (RAC).
@@
 It has been shown that SWB are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Especially, compounds from agricultural use show seasonal rain-event-driven peak concentrations in SWB \citep{wittmer_significance_2010}.
+Especially, compounds from agricultural use show seasonal rain-event-driven short term peak concentrations in SWB \citep{wittmer_significance_2010}.
 Despite their biological relevance and potential pesticide exposure only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
@@
 A clear definition of SWB in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
-The WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further consideration of streams \textless 10km\textsuperscript{2}. 
+The WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2}. 
 \citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
 Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure \ref{fig:fig3}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
-This catchment size corresponds to a stream width of approximately 2~meters (see Supplement Figure S2).
+This catchment size corresponds to a stream width of approximately 2~meters (see Supplement, Figure S2).
 
@@
 We characterized pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
-RACs are derived during pesticide authorization and no unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
+RACs are derived during pesticide authorization as part of the ecological risk assessment.
+No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
 The German Federal Environmental Agency provided RACs for the 105 compounds with highest detection rates (Supplement, Table S2). 
@@
 $\pi_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, where $\mu$ is the mean of the gamma function.
-We used the precipitation at sampling date ($prec_0$) and the day before ($prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\pi$.
-To account for temporal autocorrelation and differences between federal states we used the site nested within state as random intercepts:
+We used the log(x+0.05) transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\pi$.
+To account for temporal autocorrelation and differences between federal $state$ we used the $site$ nested within $state$ as random intercepts.
 
@@
 \begin{split}
-\log(\mu_i) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
-logit(\pi_{ijk}) = prec_{0 ijk} + prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+\log(\mu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+logit(\pi_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
 state_{j} \sim N(0, \sigma^2_{state}) \\
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-Most samples were take in the months April till October, with less samples in the winter (see Supplemental Figure S2).
+Most samples were take in the months April till October, with less samples in the winter (see Supplement, Figure S3).
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
 We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
-Hierarchical clustering revealed three groups (see also Supplement Figure S2):
+Hierarchical clustering revealed three groups (see also Supplemen,t Figure S4):
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to three groups determined by hierarchical clustering (see Supplement Figure S2).}
+  Colors according to three groups determined by hierarchical clustering (see Supplement, Figure S4).}
   \label{fig:fig2}
@@
 \subsection{Effect of precipitation on pesticide exposure}
-The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S3).
-Precipitation at the day before ($precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all analysed compounds. 
-The effects were less pronounced ($\pi$) or not directed ($\mu$) for precipitation at the day of sampling ($precip_{0}$)(Figure \ref{fig:fig5}, top).
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S6).
+Precipitation at the day before ($log~precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all analysed compounds. 
+The effects were less pronounced ($\pi$) or not clearly directed ($\mu$) for precipitation at the day of sampling ($log~precip_{0}$)(Figure \ref{fig:fig5}, top).
 
-Quarters 2-4 showed generally higher $\mu$ than in Q1. 
-Exceptions were Diflufenican, that showed significantly lower RQ in Q2-Q4 compared to Q1.
-Smaller negative effects were found for Clothianidin (Q2) and Isoproturon (Q3) (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
-We found similar effects on $\pi$.
+Q2, Q3 and Q4 showed generally higher $\mu$ than in Q1, with smallest differences for Q4.
+An exception was the herbicide Diflufenican, that showed througout all quarters lower RQ in Q2-Q4 compared to Q1 (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
+We found similar effects of season on $\pi$, but with more compounds showing a reduction in $\pi$.
 
@@
 \subsection{Pesticides in small water bodies}
-The dataset comprised 12,710 samples from 1295 small water bodies.
-1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher then 1 were observed.
+The dataset comprised 12,710 samples from 1,295 small water bodies.
+In 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher then 1 were observed.
 Neonicotinoid Insecticides and Chlorpyrifos showed highest risk quotients (Figure \ref{fig:fig6}).
-For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ \textgreater 1. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin showed also high exceedances of RQ (29.5/14.2/21.4 \% of all samples with detects).
+For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
 Highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Metabolites were most commonly detected in samples (Metazachlor sulfonix acid was in 82\% of all samples were it was analysed) (Supplement Figure S8)
+Metabolites were most commonly detected in samples (Metazachlor sulfonic acid was in 82\% of all samples were it was analysed) (Supplement Figure S8).
 Glyphosate was the compound with highest detection rates, followed by Boscalid and Isoproturon. 
 However, only the latter showed RQ exceedances (Figure \ref{fig:fig6}).
-In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one samples (supplemental figure S7).
+In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one samples (Supplement, Figure S7).
 
@@
 
-\todo{unfertig}
+Given their occurrence in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure \ref{fig:fig3}). 
+Especially, streams below 10~km\textsuperscript{2} are missing, which could be attributed to the missing categorisation in the WFD.
 
@@
 Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
-Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological effects.
+Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.
 
@@
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-However, we found higher pesticide exposure if samples where taken at the after heavy rainfall events. 
-Interestingly, samples take at they day of a heavy-rainfall-event did not show such a strong relationship.
+Nevertheless, we found higher pesticide exposure if samples where taken at the day after heavy rainfall events. 
+Interestingly, samples take at they day of the rainfall-event did not show such a strong relationship.
 This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentration in agricultural SWB show short term peak concentrations \citep{wittmer_significance_2010}.
+Pesticide concentration in agricultural SWB show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after the rain-fall-event might miss peak concentrations.
 Overall, our results indicate that current chemical monitoring strongly underestimates pesticide exposure. 
+Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
 
-\todo{add alternatives - passive sampler, automatic samplers...}
-
-We found pesticide specific patterns in seasonality of the studied compounds.
+We found pesticide specific patterns in the seasonality of exposure.
 Most pesticides showed an increase from April - September, which coincides with their main application season.
-Hence, we also found the hightest RQ for the herbicide Diflufenican during the winter/spring season, which is application period for which it is registered \todo{cite BVL}.
+Hence, we also found the hightest RQ for the herbicide Diflufenican during the winter/spring season, which is application period for which it is registered in Germany \citep{bvl_online_2016}. 
+Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
+Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment.
 
@@
 By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1290 and 2113 samples in SWB, Figure \ref{fig:fig6}).
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure \ref{fig:fig6}).
 Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
@@
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at RQ of 1.12.
+Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at a RQ of 1.12.
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 
@@
 % Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study environmental occurrences of pesticides after approval.
+Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrences of pesticides.
+Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
 The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.

### 2016-09-15 supplement/Supplemental_Materials.tex
second draft

@@
 %% -------------------------------------------------------------------------
-\title{Supplemental Materials for the paper: \\Pesticides pollution of small streams in
-Germany}
+\title{Supplemental Materials for the paper: \\ Pesticide exposure in small streams in Germany}
 \author{Eduard Szöcs, Marvin Brinke, Bilgin Karaoglan, Ralf B. Schäfer}
@@
 	\item Structure: Structure has been adjusted to the database structure.
-	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467) and duplicates merged.
+	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
 	\item Chemicals: Chemical names and identifiers have been unified using the webchem package \citep{szocs_webchem:_2016}.
@@
 	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
-	\caption{Raw data used for the model in equation 2 and Figure 3 of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
+	\caption[Raw data used for the model in equation 2 and Figure 3 of the main article.]{Raw data used for the model in equation 2 and Figure 3 of the main article.
+Color codes the number of RAC exceedances (on a log-scale).
+Grey points denote sites without any exceedance.}
 	\label{fig:ezgagrirac}
@@
 	\includegraphics[width = 0.8\textwidth]{precip}
-	\caption{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
+	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
 	\label{fig:precip}
@@
 	\includegraphics[width = 0.8\textwidth]{pdetects}
-	\caption{Proportion of samples with detects in SWB. Only Compounds with more than 100 samples and 15\% of detects are show.}
+	\caption[Proportion of samples with detects in SWB.]{Proportion of samples with detects in SWB. Only Compounds with more than 100 samples and 15\% of detects are show.}
 	\label{fig:pdetects}

### 2016-09-27 article.tex
title
meta analysis of single compound models.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Pesticide exposure in small streams in Germany}
+\title[Pesticides small streams]{Pesticides in small streams in Germany}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 \begin{align}
-RQ_i \sim ZAGA(\mu_i, \pi_i) = 
+RQ_i \sim ZAGA(\mu_i, \sigma, \nu_i) = 
   \begin{cases}
-    (1 - \pi_i)   & \quad  \text{if } y < LOQ \\
-    \pi_i \times f_{Gamma} (\mu_i) & \quad \text{if } y \ge LOQ \\
+    (1 - \nu_i)   & \quad  \text{if } y < LOQ \\
+    \nu_i \times f_{Gamma} (\mu_i, \sigma) & \quad \text{if } y \ge LOQ \\
   \end{cases}
@@
 
-$\pi_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, where $\mu$ is the mean of the gamma function.
-We used the log(x+0.05) transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\pi$.
-To account for temporal autocorrelation and differences between federal $state$ we used the $site$ nested within $state$ as random intercepts.
+$\nu_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
+We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\nu$.
+$\sigma$ was assumed to be constant. 
+To account for temporal autocorrelation and differences between federal states we used $site$ nested within $state$ as random intercepts:
 
@@
 \log(\mu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
-logit(\pi_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+logit(\nu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
 state_{j} \sim N(0, \sigma^2_{state}) \\
@@
 
-Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\pi$ can be interpreted as changes in the probability that there is any risk.
-We fitted this model separately to each compound with a RAC, with at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
+Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ.
+We fitted this model separately to each compound with a RAC, at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
 We implemented the model using the gamlss package \cite{stasinopoulos_generalized_2007}.
@@
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S6).
-Precipitation at the day before ($log~precip_{-1}$) sampling increased $\mu$ and also $\pi$ for nearly all analysed compounds. 
-The effects were less pronounced ($\pi$) or not clearly directed ($\mu$) for precipitation at the day of sampling ($log~precip_{0}$)(Figure \ref{fig:fig5}, top).
+Precipitation at the day before ($log~precip_{-1}$) sampling increased $\mu$ and also $\nu$ for nearly all analysed compounds. 
+The effects were less pronounced ($\nu$) or not clearly directed ($\mu$) for precipitation at the day of sampling ($log~precip_{0}$)(Figure \ref{fig:fig5}, top).
 
@@
 An exception was the herbicide Diflufenican, that showed througout all quarters lower RQ in Q2-Q4 compared to Q1 (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
-We found similar effects of season on $\pi$, but with more compounds showing a reduction in $\pi$.
+We found similar effects of season on $\nu$, but with more compounds showing a reduction in $\nu$.
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=0.95\textwidth]{figure5.pdf}
   \caption{Estimated coefficients and their 95\% CI for the model described in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Tables S3 and S4 for a list of compounds from top to bottom and their respective coefficients).
-Left column: Effect on mean RQ ($\mu$); Right column: effect on the probability that RQ \textgreater 0 ($\pi$).
+Left column: Effect on mean RQ ($\mu$); Right column: effect on the probability that RQ \textgreater 0 ($\nu$).
   Coefficients where the CI encompasses zero are shown in gray colour.
@@
 \end{figure}
+
+\todo{rewrite + new figures to supplement}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=0.6\textwidth]{figure6.pdf}

### 2016-09-27 supplement/Supplemental_Materials.tex
meta analysis of single compound models.

@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{coefs}
+	\caption[Graphical representation of coefficients from table \ref{tab:var_model_coef}.]{Graphical representation of coefficients from table \ref{tab:var_model_coef}.
+Top row: Effect of precipitation at day before sampling and at day of sampling.
+Bottom row: estimates for the four Quarters.
+Each dot represent one compound.
+Coefficients where the CI encompasses zero are shown in gray colour.
+Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$).}
+	\label{fig:coefs}
+\end{figure}
+
 %% -------------------------------------------------------------------------

### 2016-09-27 src/do_precip.R

@@ -116,24 +116,24 @@ keep_tab_x <- xtable(keep_tab,
                     Order is the same as in Figure 5 of the articles. See Table \\ref{tab:var_model_coef} for model coefficients.',
                     align = 'lp{2.5cm}rlp{1.5cm}p{2cm}p{2cm}')
 
-print(keep_tab_x, 
-      file = file.path(prj, 'supplement/keeptab.tex'),
-      tabular.environment="longtable",
-      floating = FALSE,
-      caption.placement = 'top',
-      comment = FALSE,
-      booktabs = TRUE,
-      hline.after = c(-1, 0, nrow(keep_tab)),
-      sanitize.text.function = identity,
-      size="\\fontsize{8pt}{10pt}\\selectfont"
-)
+# print(keep_tab_x, 
+#       file = file.path(prj, 'supplement/keeptab.tex'),
+#       tabular.environment="longtable",
+#       floating = FALSE,
+#       caption.placement = 'top',
+#       comment = FALSE,
+#       booktabs = TRUE,
+#       hline.after = c(-1, 0, nrow(keep_tab)),
+#       sanitize.text.function = identity,
+#       size="\\fontsize{8pt}{10pt}\\selectfont"
+# )
 
 
 
 take <- take[variable_id %in% keep$variable_id]
 psm_variables[variable_id %in% keep$variable_id, list(variable_id, name, psm_type)]
 
-
+rm(samples_rac, psm_sites, psm_sites_info, take_si, rac, keep_tab, keep_tab_x)
 
 
 # Model -------------------------------------------------------------------
@@ -144,7 +144,7 @@ take_c <- take[variable_id == 457]
 mod_l_m <- gamlss(rq ~ log_precip_1 + log_precip0 + season +
                     re(random=~1|state_fac/s_id_fac),
                   # model also pi with same predictors 
-                  nu.formula =~precip_1_g + precip0_g + season + 
+                  nu.formula =~precip_1 + precip0 + season + 
                     re(random=~1|state_fac/s_id_fac),
                   # sigma is constant
                   data = take_c,
@@ -155,14 +155,30 @@ term.plot(mod_l_m)
 res <- residuals(mod_l_m)
 hist(res)
 
+# random slope model, does not converge?
+# gco <- glim.control(glm.trace = TRUE, bf.trace = TRUE)
[…]

### 2016-09-27 article.tex
meta analysis

@@
 
-% \usepackage{todonotes}
+\usepackage{todonotes}
 
@@
 Therefore, we modeled the effects of precipitation in a separate model.
-RQ showed a skewed distribution with an excess of zeros (no pesticides with RAC detected). 
+RQ showed a skewed distribution with an excess of zeros (no pesticides detected). 
 Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
@@
 $\nu_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
-We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($season_{Q1-Q4}$) as linear predictors for $\mu$ and $\nu$.
+We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
 $\sigma$ was assumed to be constant. 
@@
 \begin{split}
-\log(\mu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
-logit(\nu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + season_{ijk} + state_j + site_{jk}\\
+\log(\mu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + Q1-4_{ijk} + state_j + site_{jk}\\
+logit(\nu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + Q1-4_{ijk} + state_j + site_{jk}\\
 state_{j} \sim N(0, \sigma^2_{state}) \\
@@
 
-Changes in $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ.
-We fitted this model separately to each compound with a RAC, at least 1000 samples and more then 5\% of values above quantification limit (see Supplement, Table S3 for a list of compounds). 
-We implemented the model using the gamlss package \cite{stasinopoulos_generalized_2007}.
+Changes on $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk. 
+We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
+
+We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more then 5\% of values above LOQ (n = 24 compounds, see Supplement, Table S3 for a list of compounds). 
+To summarise the coefficients across the modeled compounds we used a random effect meta-analysis
 

### 2016-09-27 src/load.R

@@ -71,6 +71,7 @@ library(gamlss.dist)
 library(gamlss.add)
 library(lme4)
 library(nlme)
+library(metafor)
 
 ### ----------------------------------------------------------------------------
 ### Source defined functions

### 2016-09-28 article.tex
fixes to abstract and precipitation

@@
 % \abbreviations{mo, neon, ra, tu, fw}
-\keywords{Monitoring, Neonicotinoid, Risk Assessment Toxic Units, Freshwater}
-% RAC,
+\keywords{Monitoring, Neonicotinoid, Risk Assessment, Exposure, Freshwater}
 
@@
 Small streams are important refugia for biodiversity.
-In agricultural areas, they may be at high risk from pesticides that enter streams during rainfall. 
+In agricultural areas they may be at high risk from pesticides that enter streams during rainfall. 
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
-In Germany, pesticide monitoring is performed by the federal states as part of water quality surveillance.
+Data from governmental water quality monitoring provides an opportunity to study occurrences of pesticides on a national scale.
 
-We compiled monitoring data focusing on small streams, resulting in a data set of 2,918,604 measurements from 42,236 samples in 3,049 sampling sites covering the years 2005-2014.
-A total of 484 different compounds that can be classified as pesticides were measured.
-We studied the relationships between pesticide exposure and agricultural land use, catchment size, as well as precipitation and seasonal dynamics.
+We compiled monitoring data focusing on small streams, resulting in a data set of 2,918,604 measurements from 42,236 samples in 3,049 sampling sites and 484 different pesticides. 
+We investigated the relationships between pesticide exposure and agricultural land use, catchment size, as well as precipitation and seasonal dynamics using new modeling techniques.
 
-We found a threshold of 25\% agricultural land use, above pesticide exposure becomes apparent. 
-Precipitation increased pesticide concentrations and these showed pesticide specific temporal patterns.
-Especially neonicotinoid insecticides and Chlorpyrifos showed exceedances of regulatory acceptable concentrations (RACs) in small water bodies.
+We found that the influence of agriculture is detectable even if there is only a small proportion of agriculture with a catchment.
+Precipitation during and before sampling increased measured pesticide concentrations.
+Highest concentrations were found during the summer season, but with differences between compounds.
+Small waters are currently underrepresented in German pesticide monitoring.
+Nevertheless, we show that neonicotinoid and other insecticides frequently exceed regulatory acceptable concentrations.
 
@@
 We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more then 5\% of values above LOQ (n = 24 compounds, see Supplement, Table S3 for a list of compounds). 
-To summarise the coefficients across the modeled compounds we used a random effect meta-analysis
+To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
 
@@
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S6).
-Precipitation at the day before ($log~precip_{-1}$) sampling increased $\mu$ and also $\nu$ for nearly all analysed compounds. 
-The effects were less pronounced ($\nu$) or not clearly directed ($\mu$) for precipitation at the day of sampling ($log~precip_{0}$)(Figure \ref{fig:fig5}, top).
+Precipitation at the day of sampling and the day before increased the probability of exceeding LOQ and value of RQ.
+Precipitation at day before sampling had a stronger effect, than at day of sampling. 
+This difference was less pronounced for the value RQ (Figure \ref{fig:fig5}, top).
 
-Q2, Q3 and Q4 showed generally higher $\mu$ than in Q1, with smallest differences for Q4.
-An exception was the herbicide Diflufenican, that showed througout all quarters lower RQ in Q2-Q4 compared to Q1 (Figure \ref{fig:fig5}, bottom and Supplement Table S4).
-We found similar effects of season on $\nu$, but with more compounds showing a reduction in $\nu$.
+The first quarter showed lowest RQ and probability of exceeding LOQ.
+Both increased during summer months and decreased towards the end of the year.
+The differences were more less pronounced for the value RQ (Figure \ref{fig:fig5}, bottom), with wider CI.
 
 \begin{figure}[ht]
-  \includegraphics[width=0.95\textwidth]{figure5.pdf}
-  \caption{Estimated coefficients and their 95\% CI for the model described in equations \ref{eqn:eqn3} and \ref{eqn:eqn4} for each of the 24 analysed compounds (see Supplement Tables S3 and S4 for a list of compounds from top to bottom and their respective coefficients).
-Left column: Effect on mean RQ ($\mu$); Right column: effect on the probability that RQ \textgreater 0 ($\nu$).
-  Coefficients where the CI encompasses zero are shown in gray colour.
+  \includegraphics[width=0.8\textwidth]{figure5.pdf}
+  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-anaylsis of the 24 modeled compounds.
+Left column: coefficients for $\mu$, right column: coefficients for $\nu$.
+  Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$, see eqn. \ref{eqn:eqn4}).
+  Single compound coefficients are provided in the supplemental table S4 and figure S7).
   }
@@
 \end{figure}
-
-\todo{rewrite + new figures to supplement}
 
@@
 However, only the latter showed RQ exceedances (Figure \ref{fig:fig6}).
-In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one samples (Supplement, Figure S7).
+In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one sample (Supplement, Figure S7).
 
@@
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found higher pesticide exposure if samples where taken at the day after heavy rainfall events. 
-Interestingly, samples take at they day of the rainfall-event did not show such a strong relationship.
+Nevertheless, we found higher RQ if samples where taken after rainfall events. 
+Samples taken at they day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
 This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentration in agricultural SWB show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after the rain-fall-event might miss peak concentrations.
-Overall, our results indicate that current chemical monitoring strongly underestimates pesticide exposure. 
+Pesticide concentration in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+The effects were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
+This could be attributed to a higher variability of absolute concentrations.
+Overall, our results indicate that current pesticide monitoring strongly underestimates pesticide exposure.
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
 
-We found pesticide specific patterns in the seasonality of exposure.
-Most pesticides showed an increase from April - September, which coincides with their main application season.
-Hence, we also found the hightest RQ for the herbicide Diflufenican during the winter/spring season, which is application period for which it is registered in Germany \citep{bvl_online_2016}. 
+We found highest the probability of exceeding LOQ during the summer season (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure \ref{fig:fig5}, bottom right).
+This yearly pattern coincides with their main application season for pesticides.
+Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure \ref{fig:fig5}, bottom left).
+For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplement, Table S4) , which is application period for which it is registered in Germany \citep{bvl_online_2016}.
 Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.

### 2016-09-28 article.tex
typos

@@
 
-We found that the influence of agriculture is detectable even if there is only a small proportion of agriculture with a catchment.
+Our results show that the influence of agriculture is detectable even if there is only a small proportion of agriculture with a catchment.
 Precipitation during and before sampling increased measured pesticide concentrations.
-Highest concentrations were found during the summer season, but with differences between compounds.
+The highest concentrations were found during summer, but with differences between compounds.
 Small waters are currently underrepresented in German pesticide monitoring.
@@
 
-% Total words: 
+% Total words: 199
 \end{abstract}
@@
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
-Although, it is known that pesticides pollution and its effect increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of this relationships are missing.
+Although, it is known that pesticide pollution and its effects increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of these relationships are currently missing.
 
@@
 For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
-For 10\% of the sites we were not able to compile catchment size data.
+We were not able to compile catchment size data for 10\% of the sites.
 For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
@@
 No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
-The German Federal Environmental Agency provided RACs for the 105 compounds with highest detection rates (Supplement, Table S2). 
+The German Federal Environmental Agency provided RACs for 105 compounds with the highest detection rates (Supplement, Table S2). 
 We expressed RACs as Risk Quotient (RQ):
@@
 
-Where $C_i$ is the concentration of a compound $i$ in a sample.
+where $C_i$ is the concentration of a compound $i$ in a sample.
 
@@
   % E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{\kappa} \\
-  log(\mu_i)= \beta_0 + f_1(Agri_i) + f_2(Size_i) + log(n_i) \\
+  log(\mu_i)= \beta_0 + f_1(agri_i) + f_2(size_i) + log(n_i) \\
 \end{split}
@@
 We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
-The proportion of agriculture within the catchment ($Agri_i$) and the catchment size of the site ($Size_i$) were used as predictors. 
+The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors. 
 $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
@@
 
-$\nu_i$ denotes the probability of a observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
+$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
 $\sigma$ was assumed to be constant. 
-To account for temporal autocorrelation and differences between federal states we used $site$ nested within $state$ as random intercepts:
+To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts:
 
@@
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more then 5\% of values above LOQ (n = 24 compounds, see Supplement, Table S3 for a list of compounds). 
+We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more than 5\% of values above LOQ (n = 24 compounds, see Supplement, Table S3 for a list of compounds). 
 To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-Most samples were take in the months April till October, with less samples in the winter (see Supplement, Figure S3).
+Most samples were taken in the months April till October, with fewer samples in the winter (see Supplement, Figure S3).
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
 We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
-Hierarchical clustering revealed three groups (see also Supplemen,t Figure S4):
+Hierarchical clustering revealed three groups (see also Supplement Figure S4):
 
 \begin{enumerate}[i)]
-	\item with less then 100 compounds (SL, ST and TH)
+	\item with less than 100 compounds (SL, ST and TH)
 	\item with medium sized spectra
@@
 Red line marks statistically significant changes.
-Dashed lines denote 95\% pointwise Confidence Intervals.
+Dashed lines denote 95\% point-wise Confidence Intervals.
   }
@@
 
-The first quarter showed lowest RQ and probability of exceeding LOQ.
+The first quarter showed the lowest RQ and probability of exceeding LOQ.
 Both increased during summer months and decreased towards the end of the year.
-The differences were more less pronounced for the value RQ (Figure \ref{fig:fig5}, bottom), with wider CI.
+The differences were less pronounced for the value RQ (Figure \ref{fig:fig5}, bottom), with wider CI.
 
@@
 The dataset comprised 12,710 samples from 1,295 small water bodies.
-In 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher then 1 were observed.
-Neonicotinoid Insecticides and Chlorpyrifos showed highest risk quotients (Figure \ref{fig:fig6}).
+In 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
+Neonicotinoid Insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure \ref{fig:fig6}).
 For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
 The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
-Highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Metabolites were most commonly detected in samples (Metazachlor sulfonic acid was in 82\% of all samples were it was analysed) (Supplement Figure S8).
-Glyphosate was the compound with highest detection rates, followed by Boscalid and Isoproturon. 
+The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
+Metabolites were most commonly detected in samples (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed) (Supplement Figure S8).
+Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
 However, only the latter showed RQ exceedances (Figure \ref{fig:fig6}).
-In 44.8\% of samples more then one compound was quantified, with a maximum of 54 different compounds in one sample (Supplement, Figure S7).
+In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplement, Figure S7).
 
@@
   \includegraphics[width=0.6\textwidth]{figure6.pdf}
-  \caption{15 compounds with highest risk quotients in SWB.
+  \caption{15 compounds with the highest risk quotients in SWB.
 Non-detects are not shown due to the logarithmic axis.
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
 To our knowledge this is the first study investigating such thresholds of pesticide exposure.
@@
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the the unequal distribution of catchment sizes, with less sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
+Another reason might be the unequal distribution of catchment sizes, with fewer sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
 
@@
 Nevertheless, we found higher RQ if samples where taken after rainfall events. 
-Samples taken at they day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
+Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
 This could be explained by a sampling just before the actual rainfall event.
@@
 
-We found highest the probability of exceeding LOQ during the summer season (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure \ref{fig:fig5}, bottom right).
+We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure \ref{fig:fig5}, bottom right).
 This yearly pattern coincides with their main application season for pesticides.
 Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure \ref{fig:fig5}, bottom left).
-For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplement, Table S4) , which is application period for which it is registered in Germany \citep{bvl_online_2016}.
+For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplement, Table S4), which is application period in the German registration \citep{bvl_online_2016}.
 Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
@@
 %%% Word count
-% abstract    :
+% abstract    :  200
 % 4 big(600)  : 2400
 % 2 small(300):  600
-% text body   : 3000 
+% text body   : 3200
 % ackno       :   24
 % ======================================
-% total       :
+% total       : 6424
 

### 2016-09-29 article.tex
minor fixes

@@
 \section{Introduction}
-More than 50\% of the total land area in Germany are used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
+More than 50\% of the total land area in Germany is used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
 In the year 2014 more than 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
@@
 These programs may provide the most comprehensive data that is currently available for pesticide occurrences in German rivers.
-However, currently there is no curated national-wide compilation of this monitoring data available.
+However, currently there is no curated nation-wide compilation of this monitoring data available.
 
@@
 Especially, compounds from agricultural use show seasonal rain-event-driven short term peak concentrations in SWB \citep{wittmer_significance_2010}.
-Despite their biological relevance and potential pesticide exposure only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
+Despite their biological relevance and potential pesticide exposure, only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
 Moreover, it is currently unknown how precipitation might influence measured pesticide concentrations in national monitoring programs.
@@
 First, we revise the available monitoring data if it is suitable for a large scale description of pesticide pollution.
-Then we analyse the relationship between agricultural land use and catchment size with pesticide pollution and investigate potential thresholds.
-We also investigate whether samples taken during or shortly after heavy rainfall events or different seasons are more likely to indicate pesticide pollution.
-Finally, we give an overview on pesticide pollution of SWB in Germany.
-
-% Answering these questions might give important insights for future monitoring planning, ecological risk assessment and water resource management.
+Then we analyse the form of relationship between pesticide exposure, agricultural land use and catchment size.
+Further, we investigate the influence of precipitation and seasonal dynamics of measure pesticide concentrations. 
+Finally, we give an overview on current pesticide pollution of SWB in Germany.
 
@@
 \subsection{Data compilation}
-We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see supplemental table S1 for the abbreviations of federal state names).
-We homogenized and unified all data from the states into a common database.
-We implemented a robust data cleaning workflow (see supplemental figure S1 for details on data processing) \citep{poisot_best_2015}.
+We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+We homogenized and unified all data provided by the federal states into a common database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details on data processing) \citep{poisot_best_2015}.
 Nevertheless, parts of the dataset are proprietary and cannot be shared here.
 
-Given the relevance of precipitation in causing runoff events, we identified chemical samples taken during heavy rainfall events.
-We performed a spatio-temporal intersection of sampling events with gridded daily precipitation data available the German Weather service.
+We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
@@
 We were not able to compile catchment size data for 10\% of the sites.
+
 For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
-We additionally used agricultural cover data provided by authorities (18\% of sites), which resulted to 21\% of sites with missing agricultural cover data. 
+Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
 For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
@@
 \citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
-Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure \ref{fig:fig3}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
-This catchment size corresponds to a stream width of approximately 2~meters (see Supplement, Figure S2).
+Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure~\ref{fig:fig3}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
+This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
 
@@
 No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
-The German Federal Environmental Agency provided RACs for 105 compounds with the highest detection rates (Supplement, Table S2). 
+The German Federal Environmental Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
 We expressed RACs as Risk Quotient (RQ):
@@
 
-While agricultural land use and catchment size vary only between sites, this is not the case for precipitation that changes also with time.
+While agricultural land use and catchment size vary only between sites, this is not the case for precipitation which changes also with time.
 Therefore, we modeled the effects of precipitation in a separate model.
-RQ showed a skewed distribution with an excess of zeros (no pesticides detected). 
+RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
 Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
@@
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
-$\sigma$ was assumed to be constant. 
-To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts:
+We used appropriate link functions for $\mu$ and $\nu$ and $\sigma$ was assumed to be constant. 
+Equation~\ref{eqn:eqn4} summarises the deterministic part of the model:
 
@@
 \begin{split}
-\log(\mu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + Q1-4_{ijk} + state_j + site_{jk}\\
-logit(\nu_{ijk}) = log~prec_{0 ijk} + log~prec_{-1 ijk} + Q1-4_{ijk} + state_j + site_{jk}\\
-state_{j} \sim N(0, \sigma^2_{state}) \\
-site_{jk} \sim N(0, \sigma^2_{site})
+\log(\mu_{i}) = log~prec_{0 i} + log~prec_{-1 i} + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}\\
+logit(\nu_{i}) = log~prec_{0 i} + log~prec_{-1 i} + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}
 \end{split}
@@
 
-Changes on $\mu$ can be interpreted as changes in the absolute value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk. 
+To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
+Changes on $\mu$ can be interpreted as changes in the mean value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk. 
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more than 5\% of values above LOQ (n = 24 compounds, see Supplement, Table S3 for a list of compounds). 
+We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
@@
 The analyzed dataset comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
-We found large differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table S1).
+We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1).
 
@@
   \includegraphics[width=0.6\textwidth]{figure1.pdf}
-  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states, see supplemental table S1 for abbreviations.}
+  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
   \label{fig:fig1}
@@
 
-In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplement, Table S2). 
+In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
-Most samples were taken in the months April till October, with fewer samples in the winter (see Supplement, Figure S3).
+Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-We found substantial differences in the spectra of analyzed compounds between federal states (Figure \ref{fig:fig2}).
-Hierarchical clustering revealed three groups (see also Supplement Figure S4):
+We found substantial differences in the spectra of analyzed compounds between federal states (Figure~\ref{fig:fig2}).
+Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to three groups determined by hierarchical clustering (see Supplement, Figure S4).}
+  Colors according to three groups determined by hierarchical clustering (see Supplemental Figure~S4).}
   \label{fig:fig2}
@@
 
-The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:fig3}).
+The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
 The proportion of agriculture in the catchments decreased with increasing catchment size.
@@
 
-\subsection{Thresholds for agricultural land use and catchment size}
+\subsection{Influence of agricultural land use and catchment size}
 Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
-Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:fig4}, left).
-We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:fig4}, right).
+Above this threshold the exceedances level off followed by a increase above 75\% (Figure~\ref{fig:fig4}, left).
+We could no detect any effect of catchment size on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
 We also found no statistically significant interaction between catchment size and agriculture.
@@
 \subsection{Effect of precipitation on pesticide exposure}
-The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplement, Figure S6).
-Precipitation at the day of sampling and the day before increased the probability of exceeding LOQ and value of RQ.
-Precipitation at day before sampling had a stronger effect, than at day of sampling. 
-This difference was less pronounced for the value RQ (Figure \ref{fig:fig5}, top).
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
+$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
+$prec_{-1}$ had a stronger effect than $prec_{0}$. 
+This difference was less pronounced for the mean value RQ (Figure~\ref{fig:fig5}, top).
 
@@
 Both increased during summer months and decreased towards the end of the year.
-The differences were less pronounced for the value RQ (Figure \ref{fig:fig5}, bottom), with wider CI.
+The differences were less pronounced for the value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
 
@@
   \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-anaylsis of the 24 modeled compounds.
-Left column: coefficients for $\mu$, right column: coefficients for $\nu$.
-  Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$, see eqn. \ref{eqn:eqn4}).
-  Single compound coefficients are provided in the supplemental table S4 and figure S7).
+Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probabilty to exceeed LOQ ($\nu$).
+  Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
+  Single compound coefficients are provided in the Supplemental Table~S4 and Figure~S7).
   }
@@
 In 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
-Neonicotinoid Insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure \ref{fig:fig6}).
-For Thiacloprid, Imidacloprid and Chlorpyrifos RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
+Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
+For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
 The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
 The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Metabolites were most commonly detected in samples (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed) (Supplement Figure S8).
+Metabolites were most commonly detected if analysed (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
 Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
-However, only the latter showed RQ exceedances (Figure \ref{fig:fig6}).
-In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplement, Figure S7).
+However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
+In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).
 
@@
 Non-detects are not shown due to the logarithmic axis.
-Numbers on the right give the percentage of RQ \textgreater 0 and total number of samples.
+Numbers on the right give the percentage of values \textgreater LOQ and the total number of samples were the compound was analysed.
   }
@@
 Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
-There are not only big differences in the spatial distribution and quantity of sampling sites (Figure \ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure \ref{fig:fig2}) and differences in the quality of chemical analyses.
+There are not only big differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and differences in the quality of chemical analyses.
 For Thiacloprid, Imidaclorpid and Chlorpyrifos LOQ were above RAC.
 For these compounds a lowering of LOQ is essential for reliable assessment.
-Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds.
+Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
 
-Given their occurrence in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure \ref{fig:fig3}). 
+Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
 Especially, streams below 10~km\textsuperscript{2} are missing, which could be attributed to the missing categorisation in the WFD.
 
-\subsection{Thresholds for agricultural land use and catchment size}
+\subsection{Influence of agricultural land use and catchment size}
 % agriculture
@@
 However, previous studies have shown that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
-This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure \ref{fig:fig3}, top).
+This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the unequal distribution of catchment sizes, with fewer sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure \ref{fig:fig3}, top).
+Another reason might be the unequal distribution of catchment sizes, with fewer sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure~\ref{fig:fig3}, top).
 
@@
 This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentration in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
-The effects were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
+Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
 This could be attributed to a higher variability of absolute concentrations.
@@
 
-We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure \ref{fig:fig5}, bottom right).
+We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
 This yearly pattern coincides with their main application season for pesticides.
-Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure \ref{fig:fig5}, bottom left).
-For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplement, Table S4), which is application period in the German registration \citep{bvl_online_2016}.
+Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is application period in the German registration \citep{bvl_online_2016}.
 Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
@@
 By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure \ref{fig:fig6}).
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}).
 Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.

### 2016-09-29 src/do_overview.R

@@ -61,6 +61,11 @@ nrow(psm_sites) - sum(table(psm_sites_info$ezg_fin_source))
 
 table(psm_sites_info$agri_fin_source, useNA = 'always')
 # 573 (=18\%) from auhtorities
+# 528 from drain
+#! used for size preferably data from authorities
+#! for agriculture we use preferably derived catchments, 
+#! because of big differences how states calculate/define agriculture
+
 nrow(psm_sites) - sum(table(psm_sites_info$agri_fin_source))
 # 650 (=(without information)
 

### 2016-09-29 article.tex
last fixes

@@
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
-Data from governmental water quality monitoring provides an opportunity to study occurrences of pesticides on a national scale.
+Data from governmental water quality monitoring provides an opportunity to study occurrences of pesticides on a large scale.
 
@@
 
-We conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and services they provide.
+We conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide.
 
@@
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
-Although, it is known that pesticide pollution and its effects increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of these relationships are currently missing.
+Although, it is known that pesticide pollution and its biological effects increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of these relationships are currently missing.
 
@@
 We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-We homogenized and unified all data provided by the federal states into a common database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details on data processing) \citep{poisot_best_2015}.
+We homogenized and unified all data provided by the federal states into a common database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
 Nevertheless, parts of the dataset are proprietary and cannot be shared here.
@@
 \subsection{Characterization of pesticide pollution}
-We characterized pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
+We characterised pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
 RACs are derived during pesticide authorization as part of the ecological risk assessment.
@@
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
-We used appropriate link functions for $\mu$ and $\nu$ and $\sigma$ was assumed to be constant. 
-Equation~\ref{eqn:eqn4} summarises the deterministic part of the model:
+We used appropriate link functions for $\mu$ and $\nu$ and assumed $\sigma$ to be constant. 
+Equation~\ref{eqn:eqn4} summarises the deterministic part of the model.
 
@@
 The dataset comprised 12,710 samples from 1,295 small water bodies.
-In 1173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
+In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
@@
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the unequal distribution of catchment sizes, with fewer sites \textless $10~km^2$ and \textgreater $100~km^2$ (Figure~\ref{fig:fig3}, top).
+Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
 
@@
 Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
-For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is application period in the German registration \citep{bvl_online_2016}.
+For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
 Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
@@
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 
-Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and services they provide. 
+Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide. 
 Additional treat arises when taking into account that most pesticides do not occur individually but in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.

### 2016-09-30 article.tex
Update on Overleaf.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Pesticides in small streams in Germany}
+\title[Pesticides small streams]{Large scale exposure and risk from pesticides in small streams}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 Small streams are important refugia for biodiversity.
-In agricultural areas they may be at high risk from pesticides that enter streams during rainfall. 
+In agricultural areas they may be at high risk from pesticide pollution. 
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
-Data from governmental water quality monitoring provides an opportunity to study occurrences of pesticides on a large scale.
-
-We compiled monitoring data focusing on small streams, resulting in a data set of 2,918,604 measurements from 42,236 samples in 3,049 sampling sites and 484 different pesticides. 
-We investigated the relationships between pesticide exposure and agricultural land use, catchment size, as well as precipitation and seasonal dynamics using new modeling techniques.
-
-Our results show that the influence of agriculture is detectable even if there is only a small proportion of agriculture with a catchment.
-Precipitation during and before sampling increased measured pesticide concentrations.
-The highest concentrations were found during summer, but with differences between compounds.
-Small waters are currently underrepresented in German pesticide monitoring.
-Nevertheless, we show that neonicotinoid and other insecticides frequently exceed regulatory acceptable concentrations.
-
-We conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide.
+We used data from german governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale.
+The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides.
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly  consider the limit of quantification.
+\newline
+\begin{itshape} die naechsten 2 Saetze sind noch schwach - hier muessen auch Zahlen hin - ausserdem sollten die Risiken klarer dargestellt werden inklusive der verantwortlichen Substanzen
+\end{itshape}
+\newline
+Agriculture land use lead to an xxx-fold increase in xxx once the proportion of agriculture in a catchment exceeded yyy.
+Precipitation during and before sampling increased measured pesticide concentrations xxx-fold.
+The highest concentrations were found during summer, though this varied with compound.
+Risk thresholds were exceeded in yyy percent of streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance.
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
+\newline
+\begin{itshape} oder irgendsowas - auf jeden Fall irgendwas spektakulaerers als zu sagen dass pestizide ein risiko darstellen, dass ist nicht sonderlich neu
+\end{itshape}
 
@@
 The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
-Especially run-off after heavy precipitation events has been shown to be one of the major input routes for pesticides \citep{schulz_field_2004}.
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
-Although, it is known that pesticide pollution and its biological effects increase with agricultural land-use \citep{schulz_field_2004}, studies investigating the shape of these relationships are currently missing.
+Although it is known that pesticide pollution and its ecological effects increase with agricultural land-use \citep{schulz_field_2004}, the shape of the relationship is unknown and studies on potential thresholds are lacking.
 
-\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that most European water bodies are at risk from pesticides.
+Two recent studies indicate that pesticides might threaten freshwater biodiversity in the European union. 
+\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
 \citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
 They found that many of these measurements exceed regulatory acceptable concentrations (RAC). 
-Both studies indicate that pesticides might be a threat to biodiversity in the European union. 
-However, theses studies reflect only a small amount of data and it is unclear how representative they are:
-For Germany the study of \citet{malaj_organic_2014} lists only 175 sites and \citet{stehle_pesticide_2015} only 138 measurements.
-%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
-National monitoring programs are setup for the surveillance of water quality.
-In Germany these are setup independently by the federal states in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
-These programs may provide the most comprehensive data that is currently available for pesticide occurrences in German rivers.
-However, currently there is no curated nation-wide compilation of this monitoring data available.
+However, theses studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
+%173 see Schäfer et al.
+2016, Stehle: Table 2.
+Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
+Despite these data providing the opportunity to study research questions that require large scale data with a high spatial density, to date these data have not been compiled and related analyses are lacking.
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}
+%needs a definition here!
+, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
-It has been shown that SWB are more polluted than bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
 Especially, compounds from agricultural use show seasonal rain-event-driven short term peak concentrations in SWB \citep{wittmer_significance_2010}.

### 2016-09-30 article.tex
second draft (with already minor comments from ralf)

@@
 Meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Despite their biological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in SWB, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
+COnse
+Moreover, it is currently unknown how precipitation might influence measured pesticide concentrations in national monitoring programs.
 Especially, compounds from agricultural use show seasonal rain-event-driven short term peak concentrations in SWB \citep{wittmer_significance_2010}.
-Despite their biological relevance and potential pesticide exposure, only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
-Moreover, it is currently unknown how precipitation might influence measured pesticide concentrations in national monitoring programs.
 

### 2016-09-30 article.tex
revise

--- 2016-09-30 article.tex
@@
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
-We used data from german governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale.
+We used data from German governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale. 
 The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. 
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly consider the limit of quantification.
-\newline
-\begin{itshape} die naechsten 2 Saetze sind noch schwach - hier muessen auch Zahlen hin - ausserdem sollten die Risiken klarer dargestellt werden inklusive der verantwortlichen Substanzen
-\end{itshape}
-\newline
-Agriculture land use lead to an xxx-fold increase in xxx once the proportion of agriculture in a catchment exceeded yyy.
-Precipitation during and before sampling increased measured pesticide concentrations xxx-fold.
-The highest concentrations were found during summer, though this varied with compound.
-Risk thresholds were exceeded in yyy percent of streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance.
+Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
+Precipitation increased measured pesticide concentrations xxx-fold. 
+The highest concentrations were found during summer.
+Risk thresholds were exceeded in 27 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
-\newline
-\begin{itshape} oder irgendsowas - auf jeden Fall irgendwas spektakulaerers als zu sagen dass pestizide ein risiko darstellen, dass ist nicht sonderlich neu
-\end{itshape}
 
@@
 However, theses studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
-%173 see Schäfer et al.
-2016, Stehle: Table 2.
-Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
+%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
+Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
+which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
 Despite these data providing the opportunity to study research questions that require large scale data with a high spatial density, to date these data have not been compiled and related analyses are lacking.
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}
-%needs a definition here!
-, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
@@
 The dataset comprised 12,710 samples from 1,295 small water bodies.
-In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
+In 346 (26.7\%) of these streams RQ were exceeded.
+In total 1,173 samples (0.3\% of all and 5.6\% of samples with detects) showed RQ higher than 1 were observed.
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).

### 2016-09-30 cover_letter/cover.tex
revise

@@
 
-\emph{Pesticides in small agricultural streams in Germany}\\
+\emph{Pesticides in small streams in Germany}\\
 
-Pesticides enter freshwater ecosystems from agricultural areas and may have effects on biota.
-However, relatively little is know about their large-scale distribution.
+Pesticides enter freshwater ecosystems from agricultural areas and may have adverse effects on biota.
+However, relatively little is know about their large-scale distribution and dynamics.
 Especially, small streams may be at high risk because of adjacency to fields and small water columns.
@@
 We compiled and homogenized monitoring data from all German Federal States with a focus on small to mid-size streams. 
-Our results show that there is lack of knowledge for small streams. 
-However, the data revealed the pesticides are commonly occurring.
+Our results show that there is currently a lack of knowledge for small streams. 
+Using new and appropriate modeling techniques we show that in the influence of agricultural land use is non-linear.
+Measured pesticide concentrations are influenced by precipitation before sampling and show a temporal pattern with a peak during summer.
 Especially, neonicotinoid pesticides may pose a risk to small agricultural freshwaters.
 
-Our study should be of general interest and add valuable information on chemical monitoring, pesticide occurrences and ecological risk arising from them.
+Our study should be of general interest and provides valuable information on chemical monitoring, pesticide occurrences and ecological risk arising from them.
 

### 2016-09-30 src/do_sizeagri.R

@@ -134,14 +134,7 @@ plot(mod_nb_ti)
 vis.gam(mod_nb_ti, view = c('agri_fin', 'ezg_fin'))
 anova(mod_nb, mod_nb_ti, test = 'Chisq') 
 # smoothing interaction not of interest
-
-
-plot(mod_nb, pages = 1)
-plot(mod_nb, pages = 1, residuals = TRUE) 
-gam.check(mod_nb)
-# overdispersion
-r <- resid(mod_nb, type = "pearson")
-sum(r^2) / (mod_nb$df.res)
+
 # OK
 
 
@@ -237,4 +230,7 @@ ggsave(file.path(prj, "figure4.pdf"),
        p, width = 7, height = 7/1.6,
        units = 'in', dpi = 300, scale = 1)
 
+
+pdat[c(1, 26), ]
+0.436 / 0.126 # = ration risk at zero / risk at no significant
 

### 2016-09-30 article.tex
Update on Overleaf.

--- 2016-09-30 article.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Large scale exposure and risk from pesticides in small streams}
+\title[Pesticides small streams]{Large scale risks from pesticides in small streams}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
-Although it is known that pesticide pollution and its ecological effects increase with agricultural land-use \citep{schulz_field_2004}, the shape of the relationship is unknown and studies on potential thresholds are lacking.
+Although it is known that pesticide pollution and its ecological effects increase with the fraction of agricultural land use in the catchment \citep{schulz_field_2004}, the shape of the relationship is unknown and studies on potential thresholds are lacking.
 
@@
 They found that many of these measurements exceed regulatory acceptable concentrations (RAC). 
-However, theses studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
+However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
 %173 see Schäfer et al.
@@
 Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
-Despite these data providing the opportunity to study research questions that require large scale data with a high spatial density, to date these data have not been compiled and related analyses are lacking.
+Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
 
@@
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
-Meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Especially, compounds from agricultural use show seasonal rain-event-driven short term peak concentrations in SWB \citep{wittmer_significance_2010}.
-Despite their biological relevance and potential pesticide exposure, only a small fraction of studies were conducted on pesticide pollution of SWB, with only few large scale studies \citep{lorenz_specifics_2016}.
-Moreover, it is currently unknown how precipitation might influence measured pesticide concentrations in national monitoring programs.
+Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
+Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in SWB, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
+Consequently, knowledge on the pesticide pollution of SWB on larger scales is scant.
 
-In this study we try to fill these gaps and analyze large scale nation-wide chemical monitoring data from Germany.
-First, we revise the available monitoring data if it is suitable for a large scale description of pesticide pollution.
-Then we analyse the form of relationship between pesticide exposure, agricultural land use and catchment size.
-Further, we investigate the influence of precipitation and seasonal dynamics of measure pesticide concentrations. 
-Finally, we give an overview on current pesticide pollution of SWB in Germany.
+In this study we compiled and analyzed large scale chemical monitoring data from SWB in Germany.
+First, we analysed the shape of the relationship between pesticide risk, agricultural land use and catchment size and examined whether related thresholds for pesticide risks can be derived.
+Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
+Finally, we quantified the current risks from pesticides in SWB in Germany.
 
@@
 We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-We homogenized and unified all data provided by the federal states into a common database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
-Nevertheless, parts of the dataset are proprietary and cannot be shared here.
+% 150 km2 needs justification - can perhaps be linked to introduction, see comment above
+We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
 
@@
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
-We performed the intersection for the actual sampling date and the day before.
+We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling.
+%details required - what is a heavy rainfall event, resolution of data. One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
 
@@
 We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
-Catchment delineation was manually checked for accuracy by comparison with a stream network provided by the government.
-The delineation algorithm produced only for 30\% of the sites accurate results.
-For the rest we were able to compile catchment size data from authorities (47\% of sites) or drainage basins per stream segment provided by authorities (13\% of sites).
-We were not able to compile catchment size data for 10\% of the sites.
+Catchment delineation was manually checked for accuracy by comparison with state stream networks.
+Given that the delineated catchments were accurate only for 30\% of the sites,
+%a reviewer might ask how accuracy was defined/cehcked
+we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
+Ten \% of sites lacked catchment size data and were omitted from the operations outlined below.
 
-For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural areas based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+For each catchment we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+%it may be worth mentioning, what agricultural land cover is
 Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
@@
 % see do_overview.R for numbers
+% this section is puzzling - if you have 90% catchments, why is it only 78% in the end? It is also unclear what is the ATKIS coverage and why additional state-wide??? cover data was used...
 
@@
 This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
+% erstens muss diese definitionsfragen weiter nach oben, da schon über die SWB gesprochen wird.
+Zweitens braucht hier eigentlich keine Definition zu erfolgen - man kann eher untersuchen, ob es eine zusammmenhang zwischen belastung und catchment size gibt.
+War das nicht früher mal drinne.
+Dann könnte man auch in der Einleitung diskutieren, dass das ein weiteres Ziel der Arbeit war.
+Drittens wundert man sich, warum wir Gewässer < 150 analysieren, wenn wir als Definition hier 25 anlegen.
 

### 2016-09-30 article.tex
more rev

@@
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
-We used data from german governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale.
+We used data from German governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale. 
 The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. 
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly consider the limit of quantification.
-\newline
-\begin{itshape} die naechsten 2 Saetze sind noch schwach - hier muessen auch Zahlen hin - ausserdem sollten die Risiken klarer dargestellt werden inklusive der verantwortlichen Substanzen
-\end{itshape}
-\newline
-Agriculture land use lead to an xxx-fold increase in xxx once the proportion of agriculture in a catchment exceeded yyy.
-Precipitation during and before sampling increased measured pesticide concentrations xxx-fold.
-The highest concentrations were found during summer, though this varied with compound.
-Risk thresholds were exceeded in yyy percent of streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance.
+Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
+Precipitation increased measured pesticide concentrations xxx-fold. 
+The highest concentrations were found during summer.
+%%%! Add these numbers also to results section!
+Risk thresholds were exceeded in 27 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
-\newline
-\begin{itshape} oder irgendsowas - auf jeden Fall irgendwas spektakulaerers als zu sagen dass pestizide ein risiko darstellen, dass ist nicht sonderlich neu
-\end{itshape}
 
@@
 However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
-%173 see Schäfer et al.
-2016, Stehle: Table 2.
-Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
+%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
+Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
+which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
 Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}
-%needs a definition here!
-, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+% needs definition of SWB
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
@@
 We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-% 150 km2 needs justification - can perhaps be linked to introduction, see comment above
 We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
@@
 We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
+%%% add resolution 
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
 We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling. 
-%details required - what is a heavy rainfall event, resolution of data. One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
+%%%! One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
 
@@
 Catchment delineation was manually checked for accuracy by comparison with state stream networks.
-Given that the delineated catchments were accurate only for 30\% of the sites,
-%a reviewer might ask how accuracy was defined/cehcked
-we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
-Ten \% of sites lacked catchment size data and were omitted from the operations outlined below.
+Given that the delineated catchments were accurate only for 30\% of the sites, we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
+Ten percent of sites lacked catchment size data and were omitted from the operations outlined below.
 
-For each catchment we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
-%it may be worth mentioning, what agricultural land cover is
+For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
 Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
@@
 % see do_overview.R for numbers
-% this section is puzzling - if you have 90% catchments, why is it only 78% in the end? It is also unclear what is the ATKIS coverage and why additional state-wide??? cover data was used...
 
@@
 This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
-% erstens muss diese definitionsfragen weiter nach oben, da schon über die SWB gesprochen wird.
-Zweitens braucht hier eigentlich keine Definition zu erfolgen - man kann eher untersuchen, ob es eine zusammmenhang zwischen belastung und catchment size gibt.
-War das nicht früher mal drinne.
-Dann könnte man auch in der Einleitung diskutieren, dass das ein weiteres Ziel der Arbeit war.
-Drittens wundert man sich, warum wir Gewässer < 150 analysieren, wenn wir als Definition hier 25 anlegen.
 

### 2016-09-30 article.tex
Update on Overleaf.

--- 2016-09-30 article.tex
@@
 No unacceptable ecological effect are expected if the environmental concentration remains below this concentration. 
+%state here sth.
+like: that Stehle and Schulz showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators.
+Otherwise regulatory values may not be "real" trhesholds because politically influenced and this potential criticism should be taken ahead
 The German Federal Environmental Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
@@
 To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
-We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to identify relationships \citep{fewster_analysis_2000}.
+We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
 We modeled the number of RAC exceedances (RQ \textgreater 1) as:
@@
 where $No_i$ is the observed number of exceedances at site $i$. 
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
-The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors. 
+% die normalen Leser werden SChwierigkeit haben, das zu verstehen, z.B. wo in der FOrmel überhaupt RQ > 1 auftaucht
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) with .....
+The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of .... 
 $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of samples per site ($n_i$) was used as an offset to account different sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
+The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (number of samplings and of analysed compounds) at a site and is equivalent to modeling the rate of exceedances.
+%bleibt unklar, wie beides sampling and compound spectrum in n eingeht
+% beta 0 nicht erklärt - warum eigentlich 0 und nihct i, müssen alle den gleichen parameter haben - und macht es überhaupt sinn einen achsen-schnittpunkt ungleich 0 zu haben?
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 Therefore, we modeled the effects of precipitation in a separate model.
+% begründung bleibt unklar - warum separates modell, denn im modell oben taucht weder raum noch hier zeit auf...
+Entscheidend ist doch auch, dass oben die Anzahl der RAC exdeendances und hier RQ Response variables sind
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
+Therefore, we modeled these as two processes 
+% das gleiche hätte man oben auch shcon erklären können?
+(one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
 
@@
 
-$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
+$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
+%weiß nicht, ob der normale leser kapiert, dass das mean und sd von RQ sind...
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.

### 2016-09-30 article.tex
Update on Overleaf.

@@
 \subsection{Data compilation}
-We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+We compiled pesticide monitoring data from sampling sites in streams with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
 % 150 km2 needs justification - can perhaps be linked to introduction, see comment above
@@
 We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling.
-%details required - what is a heavy rainfall event, resolution of data. One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
+%details required - what is a heavy rainfall event, resolution of data.
+One question with such data is from which precipitation value increases in concentrations are detectable.
+Also interesting and may be added as question in the introduction
 
@@
 Changes on $\mu$ can be interpreted as changes in the mean value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk. 
+%siehe oben - information könnte früher kommen
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds). 
 To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
+%müsste man vielleihct noch 2-3 sätze anfügen, was da genau gemacht wurde
 
@@
 
-The compiled dataset comprised only few standing waters (58 sites) and the majority of samples (91\%) where taken via grab sampling.  % see clean.R for numbers
-9\% of samples from 33 sites were taken as composite samples of different durations.
-Therefore, we restricted the analyses to grab samples from streams. 
-The analyzed dataset comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+These samples were taken via grab sampling.  % see clean.R for numbers
+Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1).
+% ein kritischer reviewer wird hier nachfragen: Wie repräsentativ für deutschland ist das dann.
+Ferner sollte man Erklärungen liefern warum MV und NI, Saarbrücken kaum Probestellen haben und weiß nicht, ob man die überhaupt reinnehmen sollte z.B.
+in Figure 2.
+Lässt sich doch schwer behaupten, dass Niedersachsen ein anderes Spektrum hat, wenn man 2 Stellen mit 500 eines anderen Bundeslandes vergleicht.
+INsgesamt wird auch nicht erwähnt, dass wir nur Stellen kleiner 100 abgefragt haben?!
+Das müsste vielleicht noch irgendwo erwähnt werden
 
@@
 \end{enumerate}
-
+% bei i ist eine zahl gegeben, kann man das für ii un iii auch machen? Was medium und big ist, ist ja recht subjektiv
 \begin{figure}[ht]
@@
 
-The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
-The proportion of agriculture in the catchments decreased with increasing catchment size.
+The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites in catchments below $10~km^2$, with most sampling sites in catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
+% vielleicht müsste man irgendwo nochmal schreiben, dass die zahl der kleinen gewässer zunimmt und man eben diese abnahme sieht in der verteilung
+The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
+% put a statement here like: For example, only x% of sites had more than 75% agriculture in catchments larger 100 km2, whereas in catchments below 100 km2, yyy% of sites had more than 75% agriculture in their catchment.
 
@@
 \subsection{Influence of agricultural land use and catchment size}
-Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
-Above this threshold the exceedances level off followed by a increase above 75\% (Figure~\ref{fig:fig4}, left).
-We could no detect any effect of catchment size on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
+The number of RAC exceedances increased strongly and statistically significant up to 25\% agriculture within the catchment.
+Above this threshold the exceedances levelled followed by an increase above 75\% (Figure~\ref{fig:fig4}, left).
+Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
 We also found no statistically significant interaction between catchment size and agriculture.
@@
   \includegraphics[width=0.95\textwidth]{figure4.pdf}
-  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+  \caption{Effect of \% agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
 Red line marks statistically significant changes.
@@
 The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
+%warum hier die 10 mm, oben scheint es so, als wenn einfach die menge modelliert wurde und dann bekommt man ja auch eher raus, wie der einfluss der precipitation ist...
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-$prec_{-1}$ had a stronger effect than $prec_{0}$. 
-This difference was less pronounced for the mean value RQ (Figure~\ref{fig:fig5}, top).
-
+Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
+This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).
+% wäre vielleicht auch nicht schlecht einen satz zur interpretation der parameter zu geben: z.B. each increase of 10 mm in precipitation corresponded to an increase in ... of ...
 The first quarter showed the lowest RQ and probability of exceeding LOQ.
 Both increased during summer months and decreased towards the end of the year.
-The differences were less pronounced for the value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
+The differences were less pronounced for the mean value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
 
@@
 \subsection{Pesticides in small water bodies}
+% man kann sich natürlich fragen, warum hier der datensatz reduziert wird, wenn catchment size eh kaum einfluss hat.
+dafür braucht es oben eine gute begründung.
+auch warum alle analysen mit dem vollen datensatz gemacht werden, außer dieser
 The dataset comprised 12,710 samples from 1,295 small water bodies.
@@
 The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Metabolites were most commonly detected if analysed (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
+Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
 Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
@@
 In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8). 
+%exceedances per site sollten hier erwähnt werden, nicht erst in Diskussion
 

### 2016-09-30 article.tex
Update on Overleaf.

@@
 %exceedances per site sollten hier erwähnt werden, nicht erst in Diskussion
+% und man könnte sich fragen, ob es da unterschiede zwischen den bundesländern gab
 
@@
 \subsection{Overview on the compiled dataset}
-The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
+The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
+% wäre es nicht besser Stone 2014 ES&T für die USA zu nehmen? Der Link verweist ja gerade auf Daten die nicht aufbereitet sind usw.
 The data compiled and analysed here for Germany is of similar quantity and quality.
+%zumindest mit bezug auf die US Daten stimmt diese aussage nicht
 
@@
 Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
-There are not only big differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and differences in the quality of chemical analyses.
-For Thiacloprid, Imidaclorpid and Chlorpyrifos LOQ were above RAC.
+Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
+% was wiederum die oben aufgeworfene frage verstärkt, ob es bei den ergebnissen unterschiede zwischen den staaten gab...
+For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
 For these compounds a lowering of LOQ is essential for reliable assessment.
-Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
+Moreover, a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
+% ok, aber dann musst du auch irgendwo positiver formulieren, worauf sich die ergebnisse von uns jetzt beziehen - schon mal nicht auf Deutschland...
 
 Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
-Especially, streams below 10~km\textsuperscript{2} are missing, which could be attributed to the missing categorisation in the WFD.
+Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD.
 
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
-To our knowledge this is the first study investigating such thresholds of pesticide exposure.
-Thresholds for agricultural land use have only been investigated in the past for biological communities.
-\citet{feld_response_2013} found change points of biological community metric for agricultural land use at 40\% in low lands. 
-Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
+% likely can be defined (0.5 it seems?) 
+To our knowledge this is the first study investigating such thresholds of pesticide risk.
+Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams. 
+Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use for ....
+%kurze beschreibung welche Gewässer und Region
 Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.
@@
 We did not find a relationship between pesticide pollution and catchment size.
-However, previous studies have shown that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
-This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
+However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments below \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
 Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
+%Argument nicht ganz klar - sample size beeinflusst pollution state? Ansonsten ist das doch das gleiche argument, dass gradient insgesamt nur 10-100 war...
 
 \subsection{Effect of precipitation on pesticide exposure}
-Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found higher RQ if samples where taken after rainfall events. 
-Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
-This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+Our results revealed that most samples for pesticide monitoring in Germany are taken in the absence of prior precipitation events.
+Given that we found higher risk from pesticides (RQ) for samples taken after rainfall events.
+%would be good to be more quantitative here - was meint higher
+Interestingly, samples taken at the day of a rainfall-event showed a slightly lower risk, than samples taken one day after a rainfall-event.
+This can may be explained on the one hand by the sampling preceding the actual rainfall event on the sampling day, which could not be accounted for in analysis due to a lack of data on the hour of sampling and rain event.
+On the other hand, the delay between the start of the precipitation and the peak concentrations in streams varies with context variables such as geology, riparian land cover and slope.
+Thus, even a sampling after the initiation of a precipitation event may not capture the short term peak concentrations \citep{wittmer_significance_2010}.
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
 This could be attributed to a higher variability of absolute concentrations.
-Overall, our results indicate that current pesticide monitoring strongly underestimates pesticide exposure.
+Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
+% ist auch die frage, wie häufig du wirklich SWB, was auch stehende gewässer inkludiert, meinst - unsere ergebnisse sind ja nur für streams - macht eigentlich mehr sinn dann auch small streams zu schreiben oder sowas, da gerade für standgewässer die sache wesentlich komplizierter ist
 
 We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
-This yearly pattern coincides with their main application season for pesticides.
-Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+This yearly pattern coincides with their main application season for pesticides in Central Europe.
+Nevertheless, the annual pattern differed between compounds, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+%etwas seltsam in den sätzen davor bezug auf LOQ und erklärung für RQ
 For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
-Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
-Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment.
+Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
+Currently, little is known on these, but would provide useful information for pesticide monitoring and risk assessment, given that the sensitivity of organisms is often life stage dependent
+%ref needed
+and temporal dynamics would inform on concurrent exposure to multiple pesticides as well as to parameterise toxikokinetik and toxikodynamic models.
+% (1) Ashauer, R.; Albert, C.; Augustine, S.; Cedergreen, N.; Charles, S.; Ducrot, V.; Focks, A.; Gabsi, F.; Gergs, A.; Goussen, B.; et al.
+Modelling survival: exposure pattern, species sensitivity and uncertainty.
+Scientific Reports 2016, 6, 29178.
 
 \subsection{Pesticides in small water bodies}
-Our results suggest that SWB are frequently exposed to biologically relevant pesticide concentrations.
+Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
+%siehe bemerkung oben bzgl SWB
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
-By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
+By contrast, our results that neonicotinoid insecticides showed highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
 This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}).
-Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
-Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
-This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
-Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in pesticide use between countries and defined RACs.
+%unklar warum between - und SWB
+Overall, our results suggest that neonicotonoids may currently pose a high risk to freshwater ecosystems.
+% hier könnte man zumindest auch kurz andere Studien erwähnen, gab ja sehr viel zu neoniks in den letzten Jahren.
+Z.b.
+OUr results add to the growing literature on risks from neonicotinoids in ecosystems such as ...
+Compared to \citet{stehle_pesticide_2015} we found much lower rates of RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
+This may be attributed to two differences: 1) Different objectives of governmental monitoring and scientific studies.
+Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+2) DIfferent sampling
+% ich meine dass bei Stehle und Schulz das meiste event driven war - hier diesen zweiten punkt nennen
+
+By contrast, \citet{knauer_pesticides_2016} analysed Suisse pesticide monitoring data and found exceedances mainly for herbicides and fungicides and for only one insecticide, Chlorpyrifos.
+% hier auch unterschiede in % der samples?
+This might reflect differences in pesticide use between countries and, albeit to a smaller extent, in the RAC values.
+%dieses argument torpediert die generelle nutzung von RACs als indikator für Risiko! Wenn er variiert würde es bedeutetn, dass es keinen festen Bezug zum ökoloigischen Risiko gibt...
 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at a RQ of 1.12.
+% wie gesagt widerspricht der aussage davor - vielleicht das als grund für unterschied einfach weglassen?
+Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
+%give number - how much reduced
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 
-Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide. 
-Additional treat arises when taking into account that most pesticides do not occur individually but in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
+%kannst du nicht einfach oben dann RQ 1 in den resultaten angeben?
+Zudem sie auch schon abnahme bei 1/10 RAC fanden!
+das könnte man hier besser darstellen und vielleicht hier zusätzlich geben, a RAC exceedance of 1/10 was observed at xx5 of sites.
+Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
+This threat may exacerbate because esticides often occur in mixtures \cite{schreiner_pesticide_2016} and co-occur with other stressors \citep{schafer_contribution_2016}.
 
 % Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrences of pesticides.
-Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
-The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
+Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrence patterns of pesticides.
+Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
+The high exceedances of RAC indicate that the authorisation process for pesticides may require further development and refinement.
 

### 2016-10-02 article.tex
more revisions

--- 2016-09-30 article.tex
@@
 No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
-%state here sth.
-like: that Stehle and Schulz showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators.
-Otherwise regulatory values may not be "real" trhesholds because politically influenced and this potential criticism should be taken ahead
 The German Federal Environmental Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
@@
 
-where $No_i$ is the observed number of exceedances at site $i$. 
-% die normalen Leser werden SChwierigkeit haben, das zu verstehen, z.B. wo in der FOrmel überhaupt RQ > 1 auftaucht
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) with .....
-The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of .... 
+where $No_i$ is the observed number of RAC exceedances at site $i$. 
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
+The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors. 
 $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (number of samplings and of analysed compounds) at a site and is equivalent to modeling the rate of exceedances.
-%bleibt unklar, wie beides sampling and compound spectrum in n eingeht
-% beta 0 nicht erklärt - warum eigentlich 0 und nihct i, müssen alle den gleichen parameter haben - und macht es überhaupt sinn einen achsen-schnittpunkt ungleich 0 zu haben?
+The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 Therefore, we modeled the effects of precipitation in a separate model.
-% begründung bleibt unklar - warum separates modell, denn im modell oben taucht weder raum noch hier zeit auf...
-Entscheidend ist doch auch, dass oben die Anzahl der RAC exdeendances und hier RQ Response variables sind
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes 
-% das gleiche hätte man oben auch shcon erklären können?
-(one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
+Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
 
@@
 
-$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
-%weiß nicht, ob der normale leser kapiert, dass das mean und sd von RQ sind...
+$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
@@
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
@@
 
-The compiled dataset comprised only few standing waters (58 sites) and the majority of samples (91\%) where taken via grab sampling.  % see clean.R for numbers
-9\% of samples from 33 sites were taken as composite samples of different durations.
-Therefore, we restricted the analyses to grab samples from streams. 
-The analyzed dataset comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
+These samples were taken via grab sampling.  
+Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1).
@@
 
-The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
-The proportion of agriculture in the catchments decreased with increasing catchment size.
+The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites in catchments below $10~km^2$, which deviated from the expected distribution \citep{nadeau_hydrological_2007}.
+%%%! check expected distribution (exponential distribution?)
+Most sampling sites had catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}). 
+The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
 
@@
   \includegraphics[width=0.95\textwidth]{figure4.pdf}
-  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
 Red line marks statistically significant changes.

### 2016-10-02 article.tex
more revisions

@@
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-$prec_{-1}$ had a stronger effect than $prec_{0}$. 
-This difference was less pronounced for the mean value RQ (Figure~\ref{fig:fig5}, top).
+Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
+This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).
 
@@
 Both increased during summer months and decreased towards the end of the year.
-The differences were less pronounced for the value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
+The differences were less pronounced for the value of RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
 
@@
 \subsection{Overview on the compiled dataset}
-The compiled dataset of governmental monitoring data represents currently the most comprehensive one available for Germany.
+The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
@@
 Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
-There are not only big differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), but also the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and differences in the quality of chemical analyses.
-For Thiacloprid, Imidaclorpid and Chlorpyrifos LOQ were above RAC.
+Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
+For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
 For these compounds a lowering of LOQ is essential for reliable assessment.
-Moreover, would a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
+Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.
 
 Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
-Especially, streams below 10~km\textsuperscript{2} are missing, which could be attributed to the missing categorisation in the WFD.
+Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD.
 
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that RAC will be exceeded with an increase in fully agricultural catchments (above 75 \% agriculture).
-To our knowledge this is the first study investigating such thresholds of pesticide exposure.
-Thresholds for agricultural land use have only been investigated in the past for biological communities.
-\citet{feld_response_2013} found change points of biological community metric for agricultural land use at 40\% in low lands. 
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
+To our knowledge this is the first study investigating such thresholds of pesticide risk.
+Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams.
 Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
@@
 We did not find a relationship between pesticide pollution and catchment size.
-However, previous studies have shown that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
-This could be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams being \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
+However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
-Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
 
@@
 This could be attributed to a higher variability of absolute concentrations.
-Overall, our results indicate that current pesticide monitoring strongly underestimates pesticide exposure.
+Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
@@
 We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
-This yearly pattern coincides with their main application season for pesticides.
+This yearly pattern coincides with their main application season for pesticides in Central Europe.
 Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
 For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
-Our study suggest that pesticide exposure shows compound specific spatio-temporal dynamics.
+Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
 Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment.
@@
 \subsection{Pesticides in small water bodies}
-Our results suggest that SWB are frequently exposed to biologically relevant pesticide concentrations.
+Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
-By contrast, our results revealed that neonicotinoid insecticides show high exceedances, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}).
-Our results shows that this particular class of insecticides may currently pose a high risk to freshwater ecosystems.
+By contrast, our results that neonicotinoid insecticides showed highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 2,113 samples of Thiacloprid, Figure~\ref{fig:fig6}).
+Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
 Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
 This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
-Contrary, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
+By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
 This might reflect differences in pesticide use between countries and defined RACs.
@@
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-Accordingly, \citet{stehle_agricultural_2015} found that biological diversity is significantly reduced at a RQ of 1.12.
+Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 
-Consequently, we conclude that pesticides from agricultural land use are a major treat to small water bodies, the biodiversity they host and the services they provide. 
-Additional treat arises when taking into account that most pesticides do not occur individually but in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
+Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
+This threat may exacerbate because esticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
 
 % Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrences of pesticides.
-Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
-The high exceedances of RAC indicate that the approval process for pesticides must be checked and refined.
+Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrence patterns of pesticides.
+Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
+The high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
 

### 2016-10-10 article.tex
more revisions

@@
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
-We used data from German governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale. 
+We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
 The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. 
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly consider the limit of quantification.
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
 Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
@@
 \subsection{Data compilation}
-We compiled pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+We requested pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+Additionally, we used data available from previous projects with bigger catchment sizes. 
 We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
@@
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances (RQ \textgreater 1) as:
+We modeled the number of RAC exceedances as:
 
@@
 \begin{split}
-  No_i \sim NB(\mu_i, \kappa) \\
-  % E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{\kappa} \\
+  No(RQ > 1)_i \sim NB(\mu_i, \kappa) \\
   log(\mu_i)= \beta_0 + f_1(agri_i) + f_2(size_i) + log(n_i) \\
@@
 
-where $No_i$ is the observed number of RAC exceedances at site $i$. 
-We modeled $No_i$ as resulting from a negative binomial distribution ($NB$).
-The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors. 
-$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
+where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
+We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
+The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
+$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006} and $\beta_0$ is the intercept.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
+The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 
-While agricultural land use and catchment size vary only between sites, this is not the case for precipitation which changes also with time.
-Therefore, we modeled the effects of precipitation in a separate model.
+To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as response variable.
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
+Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution (Equation~\ref{eqn:eqn3}) \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}.
+These two processes can be interpreted as changes in the mean value of RQ (change in $\mu$) and changes in the probability of exceeding LOQ and showing any risk (change in $\nu$).
 
@@
 To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
-Changes on $\mu$ can be interpreted as changes in the mean value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk. 
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
@@
 We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
+To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}. 
+The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
+The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.
 
@@
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
+The number of sampling sites per state ranged from 3 (Lower Saxonia, NI) to 1320 (North Rhine-Westphalia, NW). % Lower Saxony (NI) provided
 
@@
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-We found substantial differences in the spectra of analyzed compounds between federal states (Figure~\ref{fig:fig2}).
+We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
+The number of different pesticides per state ranged from 57 (Saarland, SL) to 278 (Rhineland-Palatinate, RP). 
 Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):
@@
 \begin{enumerate}[i)]
-	\item with less than 100 compounds (SL, ST and TH)
-	\item with medium sized spectra
-	\item with a big and distinct spectrum (RP and NI)
+	\item SL, ST and TH with a small (less than 100 compounds) compound spectra.
+	\item RP and NI with a big (278 and 226 compounds) and distinct compound spectra
+  \item the other states showed a similar spectra of medium to big size (between 127 and 227 compounds)
 \end{enumerate}
@@
 
-The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites in catchments below $10~km^2$, which deviated from the expected distribution \citep{nadeau_hydrological_2007}.
-%%%! check expected distribution (exponential distribution?)
-Most sampling sites had catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}). 
+The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites of catchments below $10~km^2$, with
+most sampling sites having catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}). 
 The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
+For example, only 3 sites had more than 75\% agriculture in catchments larger 100 $km^2$, whereas in catchments below 100 km2, 245 sites had more than 75\% agriculture in their catchment.
 
@@
 
-\subsection{Effect of precipitation on pesticide exposure}
-The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
+\subsection{Effect of precipitation on pesticide risk}
+The spatio-temporal intersection revealed that most samples were taken during periods with low precipitation.
+For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
+
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
+An increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 6\% higher mean RQ in $Q2$.
+The probability to exceed LOQ increases in $Q2$ from  12\% to 16\%.
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
@@
 Both increased during summer months and decreased towards the end of the year.
+There was a probability of 4\% to exceed LOQ in $Q1$ and 10\% in $Q2$.
 The differences were less pronounced for the value of RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
@@
 To our knowledge this is the first study investigating such thresholds of pesticide risk.
-Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
 \citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams.
@@
 
-\subsection{Effect of precipitation on pesticide exposure}
+\subsection{Effect of precipitation on pesticide risk}
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
@@
 Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
-This threat may exacerbate because esticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
+This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
 

### 2016-10-10 src/do_overview.R

@@ -18,6 +18,8 @@ psm_variables <- fread(file.path(cachedir, 'psm_variables.csv'))
 psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
 var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
+
+
 # some numbers
 # no. observations
 (tot <- nrow(psm_samples))
@@ -29,6 +31,8 @@ length(unique(psm_samples$variable_id))
 length(psm_samples[value_fin > 0, value_fin])
 length(psm_samples[value_fin > 0, value_fin]) / tot * 100 
 
+nrow(psm_sites_info[agri_fin > 0.75 & ezg_fin > 100]) 
+nrow(psm_sites_info[agri_fin > 0.75 & ezg_fin < 100]) 
 
 
 # Temporal overview of samples --------------------------------------------

### 2016-10-11 article.tex
more revisions
restricted analyses to <100km2, revised ms accordingly

@@
 We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
-The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides. 
+The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  %see do_overview.R
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
-Precipitation increased measured pesticide concentrations xxx-fold. 
+Agriculture land use lead to an 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
+Precipitation increased measured pesticide risk xxx-fold. 
 The highest concentrations were found during summer.
@@
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
+
+%%% Check all numbers
 
@@
 Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
-% needs definition of SWB
+Nevertheless, a clear definition of SWB in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
+For example, the WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
+
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
@@
 \subsection{Data compilation}
-We requested pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-Additionally, we used data available from previous projects with bigger catchment sizes. 
+We requested pesticide monitoring data from sampling sites that can be classified as SWB (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+The request encompassed only the years 2005 to 2015.
 We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
 
-We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
-%%% add resolution 
+We identified precipitation at sampling sites by spatio-temporal intersection of sampling events with gridded daily precipitation data (60$\times$30 arcsec resolution) available from the German Weather service (DWD).
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
-We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling. 
-%%%! One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
+We performed the intersection for the actual sampling date and the day before, an extracted precipitation during and up to 48 hours before sampling.
 
 \subsection{Characterization of catchments}
-We compiled a total of 3,049 sampling sites with pesticide measurements.
-We delineated catchments upstream for each of the sampling sites using a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012}.
-Catchment delineation was manually checked for accuracy by comparison with state stream networks.
-Given that the delineated catchments were accurate only for 30\% of the sites, we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
-Ten percent of sites lacked catchment size data and were omitted from the operations outlined below.
+We compiled a total of 2,369 sampling sites in SWB with pesticide measurements. %see do_overview.R
+Alongside, we also requested catchment sizes (available for 59\% of sites) and percentage agricultural land use within the catchment (available for 24\% of sites) for the sampling sites from the federal states.
+%see do_overview.R
+Additionally, we delineated upstream catchments for each of the sampling sites using (i) a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012} and (ii) from drainage basins provided by the Federal Institute of Hydrology (BfG).
+Delineated catchments were visually checked for accuracy by comparison with state stream networks and combined (with priorisation: authorities~\textgreater~DEM~\textgreater drainage basins).
+Thus, catchment size information was available for 99\% of all sites (59\% from authorities, 24\% from DEM and 16\% from drainage basins).
 
-For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
-Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
-For 78\% of the sites both, the proportion of agricultural land use and catchment size were available.
-% see do_overview.R for numbers
-
-A clear definition of SWB in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
-The WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2}. 
-\citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
-Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure~\ref{fig:fig3}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
-This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
+For each derived catchment (either from DEM or drainage basins) we calculated the percentage agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
+In total, for 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
+Sites lacking these two informations were om 
+%see do_overview.R
 
@@
 No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
+\citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
 The German Federal Environmental Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
@@
 To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+We determined the optimum number of clusters using the average silhouette width \citep{rousseeuw1987silhouettes}.
+
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
@@
 
-The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
-These samples were taken via grab sampling.  
-Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
+The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples in 2,301 SWB sampling sites.  %see do_overview.R for numbers.
+These samples were all taken via grab sampling.  
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
-The number of sampling sites per state ranged from 3 (Lower Saxonia, NI) to 1320 (North Rhine-Westphalia, NW). % Lower Saxony (NI) provided
+The number of sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
 
@@
   \includegraphics[width=0.6\textwidth]{figure1.pdf}
-  \caption{Spatial distribution of the 3109 sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
+  \caption{Spatial distribution of the 2,301 SWB sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
   \label{fig:fig1}
@@
 
-In total 484 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
-Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
+In total 478 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
+Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
 Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
-Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
+Only 4\% (71,113) of all measurements were detects above the limit of quantification (LOQ).
 We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
-The number of different pesticides per state ranged from 57 (Saarland, SL) to 278 (Rhineland-Palatinate, RP). 
-Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):
-
-\begin{enumerate}[i)]
-	\item SL, ST and TH with a small (less than 100 compounds) compound spectra.
-	\item RP and NI with a big (278 and 226 compounds) and distinct compound spectra
-  \item the other states showed a similar spectra of medium to big size (between 127 and 227 compounds)
-\end{enumerate}
+The number of different pesticides per state ranged from 57 (Saarland, SL) to 236 (Rhineland-Palatinate, RP) (Supplemental Table~S1).
+Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
+Although, there was high variation within the remaining states these could not be separated (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
+%%% unschöner satz: bearbeiten!
 
@@
 
+%%%! Use a black only plot!?
+
 The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites of catchments below $10~km^2$, with
 most sampling sites having catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
-The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
-For example, only 3 sites had more than 75\% agriculture in catchments larger 100 $km^2$, whereas in catchments below 100 km2, 245 sites had more than 75\% agriculture in their catchment.
 
@@
   \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
-  Only sampling sites with catchment area < 150 km\textsuperscript{2} are displayed. 
   Colour codes the 2-dimensional density of points.}
@@
 The number of RAC exceedances increased strongly and statistically significant up to 25\% agriculture within the catchment.
-Above this threshold the exceedances levelled followed by an increase above 75\% (Figure~\ref{fig:fig4}, left).
+Above this threshold the exceedances levelled followed by an increase above 70\% (Figure~\ref{fig:fig4}, left).
 Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
@@
 \subsection{Pesticides in small water bodies}
-The dataset comprised 12,710 samples from 1,295 small water bodies.
+Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure~\ref{fig:fig3}) and the low influence of catchment size on RAC exceedances (Figure~\ref{fig:fig4}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
+This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
+
+The dataset comprised 12,710 samples from 1,295 SWB.
 In 346 (26.7\%) of these streams RQ were exceeded.

### 2016-10-11 src/do_overview.R

@@ -19,25 +19,80 @@ psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
 var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 
+# restrict to sites < 100km² (or unknown) and > 2005
+range(psm_samples$date)
+take_sites <- psm_sites_info[ezg_fin <= 100 | is.na(ezg_fin), site_id]
+psm_sites_info <- psm_sites_info[site_id %in% take_sites]
+psm_sites <- psm_sites[site_id %in% take_sites]
+psm_samples<-psm_samples[site_id %in% take_sites]
 
-# some numbers
+
+
+# overview on Catchment derivation ----------------------------------------
+table(psm_sites_info$ezg_fin_source, useNA = 'always')
+# NA = no information available
+# auth = from authorities direct
+# drain = from river segments
+# TRUE = from DEM
+table(psm_sites_info$ezg_fin_source, useNA = 'always') / nrow(psm_sites_info) * 100
+# 1401 (=59%) with data from authorities
+# 372 (=16%) with data from stream segments
+# 568 (=24%) with data from dem
+# 28 (=1%) without any information
+
+# sites with data from auth
+nrow(psm_sites_info[!is.na(ezg_auth)]) / nrow(psm_sites_info)
+# sites with good delienation
+nrow(psm_sites_info[!is.na(ezg_gis) & use == TRUE]) / nrow(psm_sites_info)
+# use == TRUE: catchment from DEM good
+# use == drain: catchment from DEM bad, but river segment good
+# use == NA: no good shapes available
+nrow(psm_sites_info[!is.na(ezg_gis) & (use == 'drain')]) / nrow(psm_sites_info) * 100
+
+
+table(psm_sites_info$agri_fin_source, useNA = 'always')
+table(psm_sites_info$agri_fin_source, useNA = 'always') / nrow(psm_sites_info) * 100
+# 571 (=24%) from authorities
+# 522 (=22%) from drain
+# 1231 (=52%) from DEM
+# 45 (=2%) missing
+
+
+nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))])
+nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))]) / nrow(psm_sites_info)
+# 2301 (=97%) with both informations
+
+
[…]

### 2016-10-11 src/do_precip.R

@@ -402,8 +402,24 @@ exp(0.05582678) - 1
 # (ten-one) / one
 
 # nu 
+int <- -2.21435302
+a <- exp(int + 0.27717549)
+a / (1+a) # prob at 1mm)
+(one <- plogis(int + 0.27717549))
+(ten <- plogis(int + 0.27717549*2))
+(hun <- plogis(int + 0.27717549*3))
+(ten-one) 
+(hun-ten)
+(ten-one) / one 
+(hun-ten) / ten
 
-
+# Q1 mu
+exp(-3.80443384)
+# Q3 mu
+exp(-3.40903988)
+# Q1 nu
+plogis(-3.15180443)
+plogis(-2.21435302)
 
 # coefplot
 resmd$type <- gsub('^(.*)\\.(.*)$', '\\1', resmd$term)

### 2016-10-11 src/load.R

@@ -64,6 +64,7 @@ library(xtable)
 
 # modelling
 library(vegan)
+library(cluster)
 library(mgcv)
 library(gamlss)
 library(gamlss.cens)
@@ -96,7 +97,7 @@ ld <- TRUE
 
 # keep these objects
 load_obj <- ls()
-load_obj <- c(load_obj, 'load_obj')
+load_obj <- c(load_obj, 'load_obj', 'take_site_id')
 
 # use utf8 encoding
 options(encoding = 'utf8')

### 2016-10-12 article.tex
reanalysed with only data for < 100km2

@@
 %%%! Add these numbers also to results section!
-Risk thresholds were exceeded in 27 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
+Risk thresholds were exceeded in 26 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
+%%%! Add these numbers also to results section!
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
@@
 \end{abstract}
+
+%%%! Check all numbers
+%%%! Check all references to supplement
 
@@
 We requested pesticide monitoring data from sampling sites that can be classified as SWB (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+%%%! Hier vielleicht noch auf ganz kleine Gewässer eingehen?
 The request encompassed only the years 2005 to 2015.
@@
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}. 
@@
 Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
-Although, there was high variation within the remaining states these could not be separated (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
-%%% unschöner satz: bearbeiten!
+Although, there was high variation within the remaining cluster this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to three groups determined by hierarchical clustering (see Supplemental Figure~S4).}
+  Colors according to two clusters determined by hierarchical clustering (see Supplemental Figure~S4).}
   \label{fig:fig2}
 \end{figure}
-
-%%%! Use a black only plot!?
 
@@
 
-\subsection{Pesticides in small water bodies}
-Because of data scarcity of streams \textless 10~km\textsuperscript{2} (Figure~\ref{fig:fig3}) and the low influence of catchment size on RAC exceedances (Figure~\ref{fig:fig4}) we define in this study all streams below 25~km\textsuperscript{2} as SWB.
-This catchment size corresponds to a stream width of approximately 2~meters (see Supplemental Figure~S2).
-
-The dataset comprised 12,710 samples from 1,295 SWB.
-In 346 (26.7\%) of these streams RQ were exceeded.
-In total 1,173 samples (0.3\% of all and 5.6\% of samples with detects) showed RQ higher than 1 were observed.
-Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
-For Thiacloprid, Imidacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (29.5, 14.2 and 21.4 \% of all samples with detects).
-The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 82\% of all samples were it was analysed, see also Supplemental Figure~S9).
-Glyphosate was the compound with the highest detection rates, followed by Boscalid and Isoproturon. 
+\subsection{Pesticide risk in small water bodies}
+We found RAC exceedances in 26\% of sampling sites. 
+Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}).
+For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of all samples~\textgreater~LOQ).
+The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
+Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples were it was analysed (n = 3038, see also Supplemental Figure~S9).
+Glyphosate was the compound with the highest detection rates (41\%, n = 3557 samples), followed by Boscalid (23\%, n = 9886) and Isoproturon (22\%, n = 19112). 
 However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
-In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).
+In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).
 

### 2016-10-12 supplement/Supplemental_Materials.tex
fixes to abstract and precipitation
restricted analyses to <100km2, revised ms accordingly
reanalysed with only data for < 100km2

@@
 %% -------------------------------------------------------------------------
-\chapter{Catchment size - stream width relationships}
-We studied the relationship between catchment size based on three datasets containing this informations:
-Data delivered by the federal state Thuringia, \citet{vos_organic_2015} and \citet{fernandez_effects_2015} (both from Rhineland-Palatinate).
-We fitted to each dataset separately and to the combined dataset a power-function.
-The resulting models are shown in Figure \ref{fig:size_width}.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width = 0.8\textwidth]{width_size}
-	\caption[Relationship between catchment size and stream width.]{Relationship between catchment size and stream width.
-A power function has been fitted to each dataset separately and the combined dataset (black line and equation).
-The gray rectangle marks the estimated with for a catchment size of 25km\textsuperscript{2}.}
-	\label{fig:size_width}
-\end{figure}
-
-%% -------------------------------------------------------------------------
 \chapter{Overview on compiled data}
@@
 Abbreviations of state names according to ISO 3166-2:DE.}
+	\label{fig:varclus}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width = 0.6\textwidth]{silhouette}
+	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes. Two clusters showed maximum silhouette width.}
 	\label{fig:varclus}
@@
 Bottom row: estimates for the four Quarters.
-Each dot represent one compound.
+Each dot represent one compound (in the order described in table \ref{tab:var_model}).
 Coefficients where the CI encompasses zero are shown in gray colour.

### 2016-10-12 src/do_overview.R

@@ -97,12 +97,12 @@ monthly <- psm_samples[ , length(unique(sample_id)), by = month(date)]
 ggplot(monthly, aes(x = factor(month), y = V1)) + 
   geom_bar(stat = 'identity')
 yearlymonthly <- psm_samples[ , length(unique(sample_id)), by = list(year(date), month(date))]
-p <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) + 
+p_temp <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) + 
   geom_bar(stat = 'identity') +
   facet_grid(~year) +
   labs(x = 'Month', y = 'No. samples')
 # p
-ggsave(file.path(prj, 'supplement/temporal.pdf'), p, width = 20, height = 7)
+ggsave(file.path(prj, 'supplement/temporal.pdf'), p_temp, width = 20, height = 7)
 
 
 
@@ -326,7 +326,7 @@ bl_groups <- cutree(hc, k = 2)
 ## nicer plot
 # colors (red, blue or green)
 # colo <- c("#E41A1C", "#4DAF4A", "#377EB8")
-colo <- c("#E41A1C", "#4DAF4A")
+colo <- c("#E41A1C", "#377EB8")
 hcd <- as.dendrogram(hc)
 
 # function to get color labels

### 2016-10-12 src/do_pollution.R

@@ -18,6 +18,13 @@ psm_maxtu <- fread(file.path(cachedir, 'psm_maxtu.csv'))
 var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 
+# restrict to sites < 100km² (or unknown)  and both data available (see do_overview.R)
+take_site_id <- readRDS(file = file.path(cachedir, 'take_site_id.rds'))
+psm_sites_info <- psm_sites_info[site_id %in% take_site_id]
+psm_sites <- psm_sites[site_id %in% take_site_id]
+psm_samples <- psm_samples[site_id %in% take_site_id]
+
+
 ## n neonics
 psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin')]
 neo_id <- psm_variables[name %chin% c('Thiacloprid', 'Imidacloprid', 'Clothianidin'), variable_id]
@@ -29,22 +36,24 @@ psm_samples[variable_id %in% psm_variables[!is.na(rak_uba), variable_id]]
 
 # Classify stream types ---------------------------------------------------
 
-# Analyze only small streams (catchment < 25km2)
-take_sites <- psm_sites_info[ezg_fin < 25, site_id]  
-take_samples <- psm_samples[site_id %chin% take_sites]
+# # Analyze only small streams (catchment < 25km2)
+# take_sites <- psm_sites_info[ezg_fin < 25, site_id]  
+# take_samples <- psm_samples[site_id %chin% take_sites]
 
-# no. observations
-(tot <- nrow(take_samples))
-# no. sites
-length(unique(take_samples$site_id))
-# no. samples
-length(unique(take_samples$sample_id))
-# no. compounds
-length(unique(take_samples$variable_id))
-# values > loq
-length(take_samples[value_fin > 0, value_fin])
+take_samples <- psm_samples
 
-length(take_samples[variable_id == 457, value_fin])
+# # no. observations
+# (tot <- nrow(take_samples))
+# # no. sites
+# length(unique(take_samples$site_id))
+# # no. samples
+# length(unique(take_samples$sample_id))
+# # no. compounds
+# length(unique(take_samples$variable_id))
+# # values > loq
+# length(take_samples[value_fin > 0, value_fin])
+# 
+# length(take_samples[variable_id == 457, value_fin])
 
 
[…]

### 2016-10-12 src/do_precip.R

@@ -19,11 +19,19 @@ precip_dates <- readRDS(file.path(cachedir, 'precip_dates.rds'))
 precip_dates1 <- readRDS(file.path(cachedir, 'precip_dates1.rds'))
 
 
+# restrict to sites < 100km² (or unknown)  and both data available (see do_overview.R)
+take_site_id <- readRDS(file = file.path(cachedir, 'take_site_id.rds'))
+psm_sites_info <- psm_sites_info[site_id %in% take_site_id]
+psm_sites <- psm_sites[site_id %in% take_site_id]
+psm_samples <- psm_samples[site_id %in% take_site_id]
+
+
+
 
 # Risk Quotients ----------------------------------------------------------
 # restrict samples to variables with rac
 rac <- psm_variables[!is.na(rak_uba), list(variable_id, name, cas, pgroup, rak_uba)]
-# join eqs
+# join with racs
 setkey(psm_samples, variable_id)
 setkey(rac, variable_id)
 samples_rac <- psm_samples[rac]
@@ -74,10 +82,6 @@ with(take, cor(log_precip_1, log_precip0))
 #! use them separetely in modelling
 
 
-# split data into two groups (below and above 10mm)
-take$precip_1_g <- factor(ifelse(take$precip_1 <= 15, 'low', 'high'), levels = c('low', 'high'))
-take$precip0_g <- factor(ifelse(take$precip0 <= 15, 'low', 'high'), levels = c('low', 'high'))
-
 # set season / quarters
 take[ , season := cut(month(date), breaks = c(0.5, 3.5, 6.5, 9.5, 12.5), 
                       labels = c('Q1', 'Q2', 'Q3', 'Q4'))]
@@ -97,7 +101,7 @@ hist(props[prop < 0.2 , prop])
 # and with at least 500 observations
 hist(props[tot < 1000, tot])
 
-keep <- props[prop > 0.05 & tot > 1000]
+(keep <- props[prop > 0.05 & tot > 1000])
 # 25 compounds left
 
 
@@ -120,17 +124,17 @@ keep_tab_x <- xtable(keep_tab,
                     Order is the same as in Figure 5 of the articles. See Table \\ref{tab:var_model_coef} for model coefficients.',
                     align = 'lp{2.5cm}rlp{1.5cm}p{2cm}p{2cm}')
 
-# print(keep_tab_x, 
-#       file = file.path(prj, 'supplement/keeptab.tex'),
-#       tabular.environment="longtable",
[…]

### 2016-10-12 src/load.R

@@ -22,7 +22,7 @@ cachedir <- file.path(prj, "cache")	# intermediate data
 #! use developmental version of data.table
 #! CRAN version should also work...
 # install.packages('devtools')
-library(devtools)
+# library(devtools)
 # install_github("Rdatatable/data.table")
 library(data.table)
 library(xlsx)    # to read .xls files

### 2016-10-17 article.tex
Update on Overleaf.

--- 2016-10-10 article.tex
@@
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales.
-We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
+We used data from german governmental water quality monitoring to quantify the drivers of pesticide exposure and to assess pesticide risk in small streams on a large scale.
 The data set comprised of 2,918,604 measurements related to 42,236 samples in 3,049 sampling sites and of 484 pesticides.
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agriculture land use lead to an 3.5-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
-Precipitation increased measured pesticide concentrations xxx-fold. 
-The highest concentrations were found during summer.
-%%%! Add these numbers also to results section!
-Risk thresholds were exceeded in 27 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide exposure using new statistical modeling techniques that explicitly  consider the limit of quantification.
+\newline
+\begin{itshape} die naechsten 2 Saetze sind noch schwach - hier muessen auch Zahlen hin - ausserdem sollten die Risiken klarer dargestellt werden inklusive der verantwortlichen Substanzen
+\end{itshape}
+\newline
+Agriculture land use lead to an xxx-fold increase in xxx once the proportion of agriculture in a catchment exceeded yyy.
+Precipitation during and before sampling increased measured pesticide concentrations xxx-fold.
+The highest concentrations were found during summer, though this varied with compound.
+Risk thresholds were exceeded in yyy percent of streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance.
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
+\newline
+\begin{itshape} oder irgendsowas - auf jeden Fall irgendwas spektakulaerers als zu sagen dass pestizide ein risiko darstellen, dass ist nicht sonderlich neu
+\end{itshape}
 
@@
 However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
-%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
-Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
-which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
+%173 see Schäfer et al.
+2016, Stehle: Table 2.
+Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality, which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs.
 Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
-% needs definition of SWB
+Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}
+, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
@@
 \subsection{Data compilation}
-We requested pesticide monitoring data from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-Additionally, we used data available from previous projects with bigger catchment sizes. 
+We compiled pesticide monitoring data from sampling sites in streams with catchment sizes $\mathrm{< 150km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
+
 We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
@@
 We identified chemical samples taken during heavy rainfall events by spatio-temporal intersection of sampling events with gridded daily precipitation data available from the German Weather service (DWD).
-%%% add resolution 
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
 We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling.
-%%%! One question with such data is from which precipitation value increases in concentrations are detectable. Also interesting
+%details required - what is a heavy rainfall event, resolution of data.
+One question with such data is from which precipitation value increases in concentrations are detectable.
+Also interesting and may be added as question in the introduction
 
@@
 Catchment delineation was manually checked for accuracy by comparison with state stream networks.
-Given that the delineated catchments were accurate only for 30\% of the sites, we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
-Ten percent of sites lacked catchment size data and were omitted from the operations outlined below.
+Given that the delineated catchments were accurate only for 30\% of the sites,
+%a reviewer might ask how accuracy was defined/cehcked
+we compiled complementary catchment size data from authorities that provided catchments for an additional (60\% of sites).
+Ten \% of sites lacked catchment size data and were omitted from the operations outlined below.
 
-For each derived catchment (either from DEM or drainage basins) we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+For each catchment we calculated the relative cover (in \%) with agricultural area based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+%it may be worth mentioning, what agricultural land cover is
 Additionally, we used agricultural cover data provided by authorities (19\% of sites). 
@@
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances as:
+We modeled the number of RAC exceedances (RQ \textgreater 1) as:
 
@@
 \begin{split}
-  No(RQ > 1)_i \sim NB(\mu_i, \kappa) \\
+  No_i \sim NB(\mu_i, \kappa) \\
+  % E(No_i) = \mu_i~and~Var(No_i) = \mu_i + \frac{\mu_i^2}{\kappa} \\
   log(\mu_i)= \beta_0 + f_1(agri_i) + f_2(size_i) + log(n_i) \\
@@
 
-where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
-We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
-The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
-$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006} and $\beta_0$ is the intercept.
+where $No_i$ is the observed number of exceedances at site $i$.
+
+We modeled $No_i$ as resulting from a negative binomial distribution ($NB$) with .....
+The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of .... 
+$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}.
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
+The number of samples per site ($n_i$) was used as an offset to account for differences in sampling efforts (number of samplings and of analysed compounds) at a site and is equivalent to modeling the rate of exceedances.
+
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
@@
 
-To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as response variable.
+While agricultural land use and catchment size vary only between sites, this is not the case for precipitation which changes also with time.
+Therefore, we modeled the effects of precipitation in a separate model.
+
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution (Equation~\ref{eqn:eqn3}) \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}.
-These two processes can be interpreted as changes in the mean value of RQ (change in $\mu$) and changes in the probability of exceeding LOQ and showing any risk (change in $\nu$).
+Therefore, we modeled these as two processes
+
+(one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}:
 
@@
 
-$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
+$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation.
+%weiß nicht, ob der normale leser kapiert, dass das mean und sd von RQ sind...
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
@@
 To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
+Changes on $\mu$ can be interpreted as changes in the mean value of RQ, whereas changes in $\nu$ can be interpreted as changes in the probability of exceeding LOQ and showing any risk.
+
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}. 
-The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
-The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.
+We fitted this model to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 24 compounds, see Supplemental Table~S3 for a list of compounds). 
+To summarise the coefficients across the modeled compounds we used a random effect meta-analysis \citep{harrison_getting_2011}.
+%müsste man vielleihct noch 2-3 sätze anfügen, was da genau gemacht wurde
 
@@
 The compiled dataset used in analysis comprised 2,918,604 pesticide measurements of 42,236 samples in 3,049 sampling sites.  %see do_overview.R for numbers.
-These samples were taken via grab sampling.  
+These samples were taken via grab sampling.  % see clean.R for numbers
 Composite samples of different durations that were taken in 33 sites were not considered in analysis, given a potential methodical bias and the low sample size.
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1).
-The number of sampling sites per state ranged from 3 (Lower Saxonia, NI) to 1320 (North Rhine-Westphalia, NW). % Lower Saxony (NI) provided
 
@@
 Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
-We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
-The number of different pesticides per state ranged from 57 (Saarland, SL) to 278 (Rhineland-Palatinate, RP). 
+We found substantial differences in the spectra of analyzed compounds between federal states (Figure~\ref{fig:fig2}).
 Hierarchical clustering revealed three groups (see also Supplemental Figure~ S4):
@@
 \begin{enumerate}[i)]
-	\item SL, ST and TH with a small (less than 100 compounds) compound spectra.
-	\item RP and NI with a big (278 and 226 compounds) and distinct compound spectra
-  \item the other states showed a similar spectra of medium to big size (between 127 and 227 compounds)
+	\item with less than 100 compounds (SL, ST and TH)
+	\item with medium sized spectra
+	\item with a big and distinct spectrum (RP and NI)
 \end{enumerate}
@@
 
-The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites of catchments below $10~km^2$, with
-most sampling sites having catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}). 
+The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites in catchments below $10~km^2$, with most sampling sites in catchments between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
+% vielleicht müsste man irgendwo nochmal schreiben, dass die zahl der kleinen gewässer zunimmt und man eben diese abnahme sieht in der verteilung
 The proportion of agriculture in the catchments tended to decrease with increasing catchment size.
-For example, only 3 sites had more than 75\% agriculture in catchments larger 100 $km^2$, whereas in catchments below 100 km2, 245 sites had more than 75\% agriculture in their catchment.
+% put a statement here like: For example, only x% of sites had more than 75% agriculture in catchments larger 100 km2, whereas in catchments below 100 km2, yyy% of sites had more than 75% agriculture in their catchment.
 
@@
   \includegraphics[width=0.95\textwidth]{figure4.pdf}
-  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+  \caption{Effect of \% agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
 Red line marks statistically significant changes.
@@
 
-\subsection{Effect of precipitation on pesticide risk}
-The spatio-temporal intersection revealed that most samples were taken during periods with low precipitation.
-For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
-
+\subsection{Effect of precipitation on pesticide exposure}
+The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-An increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 6\% higher mean RQ in $Q2$.
-The probability to exceed LOQ increases in $Q2$ from  12\% to 16\%.
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
@@
 Both increased during summer months and decreased towards the end of the year.
-There was a probability of 4\% to exceed LOQ in $Q1$ and 10\% in $Q2$.
-The differences were less pronounced for the value of RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
+The differences were less pronounced for the mean value RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
 
@@
 \subsection{Pesticides in small water bodies}
+
 The dataset comprised 12,710 samples from 1,295 small water bodies.
-In 346 (26.7\%) of these streams RQ were exceeded.
-In total 1,173 samples (0.3\% of all and 5.6\% of samples with detects) showed RQ higher than 1 were observed.
+In 1,173 samples (0.3\% of all and 5.6\% of samples with detects) RQ higher than 1 were observed.
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ exceedances (Figure~\ref{fig:fig6}).
@@
 In 44.8\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8). 
+% und man könnte sich fragen, ob es da unterschiede zwischen den bundesländern gab
 
@@
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
+
 The data compiled and analysed here for Germany is of similar quantity and quality.
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
+% was wiederum die oben aufgeworfene frage verstärkt, ob es bei den ergebnissen unterschiede zwischen den staaten gab...
 For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
 For these compounds a lowering of LOQ is essential for reliable assessment.
-Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.
+Moreover, a nation-wide assessment benefit from a harmonized spectrum of analysed compounds between federal states.
+% ok, aber dann musst du auch irgendwo positiver formulieren, worauf sich die ergebnisse von uns jetzt beziehen - schon mal nicht auf Deutschland...
 
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
+
 To our knowledge this is the first study investigating such thresholds of pesticide risk.
-Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+Previous studies examined thresholds for \%agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
 \citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams. 
-Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
+Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use for ....
 Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.
@@
 However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
-This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
+This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments below \textless $100~km^2$ (Figure~\ref{fig:fig3}, top).
 For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
+Another reason might be the unequal distribution of catchment sizes, with fewer sites below $10~km^2$ and above $100~km^2$ catchment size (Figure~\ref{fig:fig3}, top).
 
-\subsection{Effect of precipitation on pesticide risk}
-Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found higher RQ if samples where taken after rainfall events. 
-Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
-This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+\subsection{Effect of precipitation on pesticide exposure}
+Our results revealed that most samples for pesticide monitoring in Germany are taken in the absence of prior precipitation events.
+Given that we found higher risk from pesticides (RQ) for samples taken after rainfall events.
+%would be good to be more quantitative here - was meint higher
+Interestingly, samples taken at the day of a rainfall-event showed a slightly lower risk, than samples taken one day after a rainfall-event.
+This can may be explained on the one hand by the sampling preceding the actual rainfall event on the sampling day, which could not be accounted for in analysis due to a lack of data on the hour of sampling and rain event.
+On the other hand, the delay between the start of the precipitation and the peak concentrations in streams varies with context variables such as geology, riparian land cover and slope.
+Thus, even a sampling after the initiation of a precipitation event may not capture the short term peak concentrations \citep{wittmer_significance_2010}.
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
@@
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
+% ist auch die frage, wie häufig du wirklich SWB, was auch stehende gewässer inkludiert, meinst - unsere ergebnisse sind ja nur für streams - macht eigentlich mehr sinn dann auch small streams zu schreiben oder sowas, da gerade für standgewässer die sache wesentlich komplizierter ist
 
@@
 This yearly pattern coincides with their main application season for pesticides in Central Europe.
-Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+Nevertheless, the annual pattern differed between compounds, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+%etwas seltsam in den sätzen davor bezug auf LOQ und erklärung für RQ
 For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
-Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment.
+Currently, little is known on these, but would provide useful information for pesticide monitoring and risk assessment, given that the sensitivity of organisms is often life stage dependent
+%ref needed
+and temporal dynamics would inform on concurrent exposure to multiple pesticides as well as to parameterise toxikokinetik and toxikodynamic models.
+% (1) Ashauer, R.; Albert, C.; Augustine, S.; Cedergreen, N.; Charles, S.; Ducrot, V.; Focks, A.; Gabsi, F.; Gergs, A.; Goussen, B.; et al.
+Modelling survival: exposure pattern, species sensitivity and uncertainty.
+Scientific Reports 2016, 6, 29178.
 
@@
 Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
+%siehe bemerkung oben bzgl SWB
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
 By contrast, our results that neonicotinoid insecticides showed highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 2,113 samples of Thiacloprid, Figure~\ref{fig:fig6}).
-Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
-Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
-This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
-By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in pesticide use between countries and defined RACs.
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}).
+%unklar warum between - und SWB
+Overall, our results suggest that neonicotonoids may currently pose a high risk to freshwater ecosystems.
+% hier könnte man zumindest auch kurz andere Studien erwähnen, gab ja sehr viel zu neoniks in den letzten Jahren.
+Z.b.
+OUr results add to the growing literature on risks from neonicotinoids in ecosystems such as ...
+Compared to \citet{stehle_pesticide_2015} we found much lower rates of RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
+This may be attributed to two differences: 1) Different objectives of governmental monitoring and scientific studies.
+Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+2) DIfferent sampling
+% ich meine dass bei Stehle und Schulz das meiste event driven war - hier diesen zweiten punkt nennen
+
+By contrast, \citet{knauer_pesticides_2016} analysed Suisse pesticide monitoring data and found exceedances mainly for herbicides and fungicides and for only one insecticide, Chlorpyrifos.
+% hier auch unterschiede in % der samples?
+This might reflect differences in pesticide use between countries and, albeit to a smaller extent, in the RAC values.
+%dieses argument torpediert die generelle nutzung von RACs als indikator für Risiko! Wenn er variiert würde es bedeutetn, dass es keinen festen Bezug zum ökoloigischen Risiko gibt...
 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
+% wie gesagt widerspricht der aussage davor - vielleicht das als grund für unterschied einfach weglassen?
 Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
+%give number - how much reduced
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 
+%kannst du nicht einfach oben dann RQ 1 in den resultaten angeben?
+Zudem sie auch schon abnahme bei 1/10 RAC fanden!
+das könnte man hier besser darstellen und vielleicht hier zusätzlich geben, a RAC exceedance of 1/10 was observed at xx5 of sites.
 Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
-This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
+This threat may exacerbate because esticides often occur in mixtures \cite{schreiner_pesticide_2016} and co-occur with other stressors \citep{schafer_contribution_2016}.
 
@@
 Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
-The high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
+The high exceedances of RAC indicate that the authorisation process for pesticides may require further development and refinement.
 

### 2016-10-18 article.tex
Update on Overleaf.

@@
 We performed the intersection for the actual sampling date and the day before, to identify rain events during and up to 48 hours before sampling.
-%details required - what is a heavy rainfall event, resolution of data.
-One question with such data is from which precipitation value increases in concentrations are detectable.
-Also interesting and may be added as question in the introduction
 
@@
 Given that we found higher risk from pesticides (RQ) for samples taken after rainfall events.
-%would be good to be more quantitative here - was meint higher
 Interestingly, samples taken at the day of a rainfall-event showed a slightly lower risk, than samples taken one day after a rainfall-event.
@@
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
-% ist auch die frage, wie häufig du wirklich SWB, was auch stehende gewässer inkludiert, meinst - unsere ergebnisse sind ja nur für streams - macht eigentlich mehr sinn dann auch small streams zu schreiben oder sowas, da gerade für standgewässer die sache wesentlich komplizierter ist
 
@@
 Nevertheless, the annual pattern differed between compounds, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
-%etwas seltsam in den sätzen davor bezug auf LOQ und erklärung für RQ
 For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
@@
 Currently, little is known on these, but would provide useful information for pesticide monitoring and risk assessment, given that the sensitivity of organisms is often life stage dependent
-%ref needed
 and temporal dynamics would inform on concurrent exposure to multiple pesticides as well as to parameterise toxikokinetik and toxikodynamic models.
-% (1) Ashauer, R.; Albert, C.; Augustine, S.; Cedergreen, N.; Charles, S.; Ducrot, V.; Focks, A.; Gabsi, F.; Gergs, A.; Goussen, B.; et al.
-Modelling survival: exposure pattern, species sensitivity and uncertainty.
-Scientific Reports 2016, 6, 29178.
 
@@
 Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
-%siehe bemerkung oben bzgl SWB
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
@@
 This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (between 1,290 and 2,113 samples in SWB, Figure~\ref{fig:fig6}).
-%unklar warum between - und SWB
 Overall, our results suggest that neonicotonoids may currently pose a high risk to freshwater ecosystems.
-% hier könnte man zumindest auch kurz andere Studien erwähnen, gab ja sehr viel zu neoniks in den letzten Jahren.
-Z.b.
-OUr results add to the growing literature on risks from neonicotinoids in ecosystems such as ...
 Compared to \citet{stehle_pesticide_2015} we found much lower rates of RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
@@
 2) DIfferent sampling
-% ich meine dass bei Stehle und Schulz das meiste event driven war - hier diesen zweiten punkt nennen
 
@@
 This might reflect differences in pesticide use between countries and, albeit to a smaller extent, in the RAC values.
-%dieses argument torpediert die generelle nutzung von RACs als indikator für Risiko! Wenn er variiert würde es bedeutetn, dass es keinen festen Bezug zum ökoloigischen Risiko gibt...
 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-% wie gesagt widerspricht der aussage davor - vielleicht das als grund für unterschied einfach weglassen?
 Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
-%give number - how much reduced
 We found RQ values greater than 1.12 at 25\% of all SWB sites. 

### 2016-10-18 article.tex
more revisions

--- 2016-10-12 article.tex
@@
 We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
-The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  %see do_overview.R
+The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agriculture land use lead to an 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent.
-Precipitation increased measured pesticide risk xxx-fold. 
-The highest concentrations were found during summer.
-%%%! Add these numbers also to results section!
-Risk thresholds were exceeded in 26 percent of small streams, with neonicotinoid and other insecticides exhibiting highest frequencies of exceedance. 
-%%%! Add these numbers also to results section!
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events in summer.
+Agricultural land use lead to an 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
+Precipitation increased measured pesticide risk by 5\% and was highest risk  during summer months.
+Risk thresholds were exceeded in 26 percent of small streams, with highest risk emerging from neonicotinoid insecticides. 
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events.
 
-%%% Check all numbers
-
-% Total words: 199
+% Total words: 198
 \end{abstract}
@@
 However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
-%175 estimated from digitized figure C.3 in Malaj 2014, Stehle: Table 2.
+% 173 estimated from digitized figure C.3 in Malaj 2014; Stehle: Table 2.
 Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
@@
 
-Small water bodies (SWB) comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
-Nevertheless, a clear definition of SWB in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
-For example, the WFD defines SWB with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines SWB with catchment size \textless 10km\textsuperscript{2} as SWB.
+Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Nevertheless, a clear definition of SS in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
+For example, the WFD defines SS with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines SS with catchment size \textless 10km\textsuperscript{2}.
 
-However, SWB might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+However, small streams might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in SWB, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
-Consequently, knowledge on the pesticide pollution of SWB on larger scales is scant.
+Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
+Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant.
 
-In this study we compiled and analyzed large scale chemical monitoring data from SWB in Germany. 
+In this study we compiled and analyzed large scale chemical monitoring data from small streams in Germany. 
 First, we analysed the shape of the relationship between pesticide risk, agricultural land use and catchment size and examined whether related thresholds for pesticide risks can be derived. 
 Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
-Finally, we quantified the current risks from pesticides in SWB in Germany.
+Finally, we quantified the current risks from pesticides in small streams in Germany.
 
@@
 \subsection{Data compilation}
-We requested pesticide monitoring data from sampling sites that can be classified as SWB (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-%%%! Hier vielleicht noch auf ganz kleine Gewässer eingehen?
+We requested pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
 The request encompassed only the years 2005 to 2015.
@@
 \subsection{Characterization of catchments}
-We compiled a total of 2,369 sampling sites in SWB with pesticide measurements. %see do_overview.R
+We compiled a total of 2,369 sampling sites in small streams with pesticide measurements. %see do_overview.R
 Alongside, we also requested catchment sizes (available for 59\% of sites) and percentage agricultural land use within the catchment (available for 24\% of sites) for the sampling sites from the federal states.
@@
 Additionally, we delineated upstream catchments for each of the sampling sites using (i) a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012} and (ii) from drainage basins provided by the Federal Institute of Hydrology (BfG).
-Delineated catchments were visually checked for accuracy by comparison with state stream networks and combined (with priorisation: authorities~\textgreater~DEM~\textgreater drainage basins).
+Delineated catchments were visually checked for accuracy by comparison with state stream networks and derived information amalgamated with existing data.
 Thus, catchment size information was available for 99\% of all sites (59\% from authorities, 24\% from DEM and 16\% from drainage basins).
@@
 Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
-In total, for 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
-Sites lacking these two informations were om 
+For 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
+68 sites (3\%) lacking these two informations were omitted from the operations outlined below.
 %see do_overview.R
@@
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances as:
+We modeled the number of RAC exceedances at a site as:
 
@@
 
-$\nu_i$ denotes the probability of an observation i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
+$\nu_i$ denotes the probability of a measurement i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
 We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
@@
 We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}. 
+To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
 The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
@@
 
-The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples in 2,301 SWB sampling sites.  %see do_overview.R for numbers.
+The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples at 2,301 sampling sites in small streams.  %see do_overview.R for numbers.
 These samples were all taken via grab sampling.  
 We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
-The number of sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
+The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW) and no data from Brandenburg.
 
@@
   \includegraphics[width=0.6\textwidth]{figure1.pdf}
-  \caption{Spatial distribution of the 2,301 SWB sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
+  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
   \label{fig:fig1}
@@
 Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
-Only 4\% (71,113) of all measurements were detects above the limit of quantification (LOQ).
+4\% (71,113) of all measurements were detects above LOQ.
 We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
-The number of different pesticides per state ranged from 57 (Saarland, SL) to 236 (Rhineland-Palatinate, RP) (Supplemental Table~S1).
+The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1).
 Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
@@
 The number of RAC exceedances increased strongly and statistically significant up to 25\% agriculture within the catchment.
-Above this threshold the exceedances levelled followed by an increase above 70\% (Figure~\ref{fig:fig4}, left).
+The mean number of RAC exceedances per site increased strongly and statistically significant from 0.13 (no agriculture) to 0.44 (25\% agriculture within the catchment). 
+Above this threshold the exceedances levelled.
+Above 75\% agriculture within the catchment the number of exceedances further increased, but the increase was not statistically significant (Figure~\ref{fig:fig4}, left). 
 Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
-We also found no statistically significant interaction between catchment size and agriculture.
+We also could not detect a statistically significant interaction between catchment size and agriculture.
 
@@
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-An increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 6\% higher mean RQ in $Q2$.
-The probability to exceed LOQ increases in $Q2$ from  12\% to 16\%.
+An increase of precipitation in $Q2$ before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% on average (Figure~\ref{fig:fig5}, top).
+% do_precip.R
+Effect on individual compounds are provided in the Supplemental Table~S4.
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
@@
 There was a probability of 4\% to exceed LOQ in $Q1$ and 10\% in $Q2$.
-The differences were less pronounced for the value of RQ and with higher variation (Figure~\ref{fig:fig5}, bottom).
+The differences were less pronounced for the mean value of RQ and with less precision (Figure~\ref{fig:fig5}, bottom). 
+Individual compounds showed different temporal patterns (see Supplemental Table~S4).
 
@@
 
-\subsection{Pesticide risk in small water bodies}
-We found RAC exceedances in 26\% of sampling sites. 
-Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}).
+\subsection{Pesticide risk in small streams}
+We found RQ > 1 in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
+In 23\% of sites chemicals with RAC where never detected (see also Supplemental Figure~S8).
+Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
 For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of all samples~\textgreater~LOQ).
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ).
+RAC exceedances were found in 15\% of samples with detects (and 7.7\% of all samples with RAC).
+
 The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
@@
   \includegraphics[width=0.6\textwidth]{figure6.pdf}
-  \caption{15 compounds with the highest risk quotients in SWB.
+  \caption{15 compounds with the highest risk quotients in small streams.
 Non-detects are not shown due to the logarithmic axis.
@@
 The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
-Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States (Water Quality Portal (WQP) \url{www.waterqualitydata.us}).
-The data compiled and analysed here for Germany is of similar quantity and quality.
+Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
+While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here for Germany, the compilation used in \citet{stone2014pesticides} is much smaller.
+Nevertheless, there might more data for the United States available (e.g. from Water Quality Portal (WQP), \url{www.waterqualitydata.us}).
 
 % current problems in monitoring and possible solutions
-Nevertheless, a nationwide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
+A nation-wide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-For Thiacloprid, Imidaclorpid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
+Although, the outlined differences between states, all ecoregions occuring Germany \citep{illies1978limnofauna,abell2008freshwater} are covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
+For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
 For these compounds a lowering of LOQ is essential for reliable assessment.
@@
 
-Given their high abundance in the landscape \citep{nadeau_hydrological_2007} SWB are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
+Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
 Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD. 
+Clearly, there is currently a lack of knowledge for these small streams.
+We analysed only data from small streams, however, for lentic small water bodies this lack might be even greater \citep{lorenz_specifics_2016}.
 
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase the likelihood of RAC exceedance in fully agricultural catchments (above 75 \% agriculture).
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
 To our knowledge this is the first study investigating such thresholds of pesticide risk.
 Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
-\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use  in low land streams.
-Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\%.
+\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in low land streams in Europe.
+Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use in wadeable streams in the United States.
 Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.
@@
 We did not find a relationship between pesticide pollution and catchment size.
-However, previous studies showed that SWB are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+However, previous studies showed that small streams are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
 This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
@@
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found higher RQ if samples where taken after rainfall events. 
+Nevertheless, we found a 5\% higher RQ if samples where taken after rainfall events. 
 Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
 This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentrations in agricultural SWB generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+Pesticide concentrations in agricultural small streams generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
 This could be attributed to a higher variability of absolute concentrations.
-Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
-Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for SWB \citep{lorenz_specifics_2016}.
+Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, considerably underestimates pesticide risks.
+Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for small water bodies \citep{lorenz_specifics_2016}.
 
-We found highest the probability of exceeding LOQ during summer (9.9\% for Q2) and lowest in the first Quarter (2.8\%, Figure~\ref{fig:fig5}, bottom right).
+We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first Quarter (4\%, Figure~\ref{fig:fig5}, bottom right).
 This yearly pattern coincides with their main application season for pesticides in Central Europe.
-Nevertheless, there are compound specific differences in the yearly pattern, which explain the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
-For example, the herbicide Diflufenican showed highest RQ during the winter quarters (Supplemental Table~S4), which is the application period it is registered for in Germany \citep{bvl_online_2016}.
+Nevertheless, there are compound specific differences in the yearly pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+For example, the herbicide Diflufenican showed highest RQ and probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
 Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment. 
+For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.
 
-\subsection{Pesticides in small water bodies}
-Our results suggest that SWB are frequently exposed to ecologically relevant pesticide concentrations.
+\subsection{Pesticides in small streams}
+Our results suggest that small streams are frequently exposed to ecologically relevant pesticide concentrations.
+In one quarter of small streams RAC were exceeded at least once.
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
-By contrast, our results that neonicotinoid insecticides showed highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
-This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 2,113 samples of Thiacloprid, Figure~\ref{fig:fig6}).
+By contrast, we found that that neonicotinoid insecticides have highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
+This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 3,540 samples of Thiacloprid, Figure~\ref{fig:fig6}).
 Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
-Compared to \citet{stehle_pesticide_2015} we found much lower rates RAC exceedances (0.3\% of 372,304 samples with RAC vs 44\% of 1,566 samples). 
-This can be attributed to different aims of the data sources: scientific research aims at finding pollutants, whereas monitoring aims mainly at surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial ecosystems \citep{pisa2015effects}.
+
+Compared to \citet{stehle_pesticide_2015} we found lower rates RAC exceedances (15\% (n = 12,615) vs 44\% (n = 1,566) of samples with measurements above LOQ). 
+This may be attributed to two differences: i) Different objectives of governmental monitoring and scientific studies.
+Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+ii) Different sampling strategies.
+The dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013}.
+
 By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-This might reflect differences in pesticide use between countries and defined RACs.
-
+However, RAC exceedances were generally lower and less abundant compared to our results from Germany. 
+This might reflect differences in pesticide use between countries and ecoregions. 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced at a RQ of 1.12.
-We found RQ values greater than 1.12 at 25\% of all SWB sites. 
-Consequently, we conclude that agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
+Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
+We found RQ values greater than 1.12 in one quarter of small streams and RQ at 1/10 of RAC in more the half of streams. 
+Consequently, we conclude that on a large scale agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
 This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
@@
 Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
-The high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
+Our results suggest that exceedances of RACs are landscape dependent % = Zusammenhang RAC<->Agriculture
+and therefore, landscape features should be taken into account in the registration process of pesticides.
+Moreover, the high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
 
@@
 % 2 small(300):  600
+% acknowledg  :   27
+%              =====
+% subtotal      3227
 % text body   : 3200
-% ackno       :   24
 % ======================================
-% total       : 6424
+% total       : 6427
 

### 2016-10-18 supplement/Supplemental_Materials.tex
more revisions

@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes. Two clusters showed maximum silhouette width.}
-	\label{fig:varclus}
+	\label{fig:silhouette}
 \end{figure}
@@
 \chapter{Thresholds for agricultural land use and catchment size}
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 \chapter{Effect of precipitation and season on RQ}
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 %% -------------------------------------------------------------------------
-\chapter{Pesticides in small water bodies}
+\chapter{Pesticides in small streams}
 
-\begin{figure}[h]
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{prac_ex}
+	\caption[Cumulative distribution of the number sites exceeding RAC.]{Cumulative distribution of sites exceeding RAC.
+Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1.
+23\% of sites showed no detection of compounds with RAC and are not shown due to logarithmic x-axis.}
+	\label{fig:pdetects}
+\end{figure}
+
+\begin{figure}[ht]
 	\centering
 	\includegraphics[width = 0.8\textwidth]{pmix}
-	\caption{Distribution the number of quantified compounds in the samples from small water bodies.}
+	\caption{Distribution the number of quantified compounds in the samples from small streams.}
 	\label{fig:pmix}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
 	\includegraphics[width = 0.8\textwidth]{pdetects}
-	\caption[Proportion of samples with detects in SWB.]{Proportion of samples with detects in SWB. Only Compounds with more than 100 samples and 15\% of detects are show.}
+	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are show.}
 	\label{fig:pdetects}

### 2016-10-18 src/do_overview.R

@@ -29,6 +29,9 @@ psm_samples<-psm_samples[site_id %in% take_sites]
 
 
 # overview on Catchment derivation ----------------------------------------
+# total number of sites
+nrow(psm_sites_info)
+
 table(psm_sites_info$ezg_fin_source, useNA = 'always')
 # NA = no information available
 # auth = from authorities direct
@@ -50,6 +53,7 @@ nrow(psm_sites_info[!is.na(ezg_gis) & use == TRUE]) / nrow(psm_sites_info)
 nrow(psm_sites_info[!is.na(ezg_gis) & (use == 'drain')]) / nrow(psm_sites_info) * 100
 
 
+
 table(psm_sites_info$agri_fin_source, useNA = 'always')
 table(psm_sites_info$agri_fin_source, useNA = 'always') / nrow(psm_sites_info) * 100
 # 571 (=24%) from authorities
@@ -58,7 +62,9 @@ table(psm_sites_info$agri_fin_source, useNA = 'always') / nrow(psm_sites_info) *
 # 45 (=2%) missing
 
 
+
 nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))])
+
 nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))]) / nrow(psm_sites_info)
 # 2301 (=97%) with both informations
 
@@ -66,6 +72,7 @@ nrow(psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin))]) / nrow(psm_sites_info)
 # restrict to sites < 100km² (or unknown)  and both data available
 take_site_id <- psm_sites_info[!(is.na(ezg_fin) | is.na(agri_fin)), site_id]
 saveRDS(take_site_id, file = file.path(cachedir, 'take_site_id.rds'))
+
 # removed sites
 psm_sites_info[!site_id %in% take_site_id]
 psm_sites_info <- psm_sites_info[site_id %in% take_site_id]
@@ -247,7 +254,15 @@ psm_variables[!is.na(rak_uba)]
 ## total number of substances
 length(unique(psm_samples[ , variable_id]))
 # no. of compounds per group
-table(psm_variables$pgroup, useNA = 'always')
+sort(table(psm_variables$pgroup, useNA = 'always'), decreasing = TRUE)
+# 179 herbicides
+# 117 insecticides
+# 109 fungicides
+
+# values > loq
[…]

### 2016-10-18 src/do_pollution.R

@@ -70,57 +70,111 @@ take_rac[ , rq := value_fin / rak_uba]
 # join variable name and type
 setkey(take_rac, variable_id)
 setkey(psm_variables, variable_id)
-dd <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
+meas <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 
 # remove variable with less then 1000 samples
-dd <- dd[variable_id %in% dd[ , list(n = length(value)), by = variable_id][n > 1000, variable_id]]
+meas <- meas[variable_id %in% meas[ , list(n = length(value)), 
+                                  by = variable_id][n > 1000, variable_id]]
+
+
 
 # calculate max per variable and take only first 15 hits
-take_var <- dd[ , list(media = max(rq)) , by = variable_id][order(media, decreasing = TRUE)][1:15, variable_id]
-take_dd <- dd[variable_id %in% take_var]
+take_var <- meas[ , list(media = max(rq)) , 
+                  by = variable_id][order(media, 
+                                          decreasing = TRUE)][1:15, variable_id]
+take_meas <- meas[variable_id %in% take_var]
 
-# highest RQs
-take_dd[ , list(media = max(rq)) , by = name][order(media, decreasing = TRUE)][1:15]
+# highest RQs (15 subs, all samples)
+take_meas[ , list(media = max(rq)) , by = name][order(media, decreasing = TRUE)]
 
-# percentage of RQ > 1
-nrow(dd[rq > 1]) / nrow(dd) * 100
 
-# take max per sample
-rak_samples <- length(unique(dd$sample_id))
-# percentake of samples with RQ > 1In 
-nrow(dd[ ,list(rq_max = max(rq)) , by = sample_id][rq_max > 1]) / rak_samples * 100
 
-# number of SWB sites
-length(unique(dd$site_id))
+## % of samples
+# for each sample, take maximum RQ (from different compounds) and check exceedance
+nrow(meas[ ,list(rq_max = max(rq)) , 
+           by = sample_id][rq_max > 1]) /  length(unique(meas$sample_id)) * 100
+# in 7.7\% of samples taken there was was RQ > 1 found
+nrow(meas[ ,list(rq_max = max(rq)) , 
[…]

### 2016-10-18 src/do_precip.R

@@ -60,7 +60,8 @@ rm(precip_dates1, precip_dates, psm_samples)
 
 
 # join catchment size
-take_si <- psm_sites_info[!is.na(ezg_fin) | is.na(agri_fin) , list(site_id, agri_fin, ezg_fin)]
+take_si <- psm_sites_info[!is.na(ezg_fin) | is.na(agri_fin) , 
+                          list(site_id, agri_fin, ezg_fin)]
 setkey(take_si, 'site_id')
 setkey(samples_rac, 'site_id')
 take <- take_si[samples_rac][!is.na(agri_fin)]
@@ -95,7 +96,8 @@ take <- psm_variables[ , list(variable_id, psm_type)][take]
 # remove compounds with no risk
 props <- take[  , list(prop = sum(rq>0) / length(rq),
               abs = sum(rq>0),
-              tot = length(rq)), by = variable_id][order(prop, decreasing = TRUE)]
+              tot = length(rq)), by = variable_id][order(prop, 
+                                                         decreasing = TRUE)]
 hist(props[prop < 0.2 , prop])
 # keep compounds that were found in at least 5% of samples
 # and with at least 500 observations
@@ -114,7 +116,8 @@ setkey(psm_variables, variable_id)
 setkey(keep, variable_id)
 keep_tab <- psm_variables[ , list(variable_id, name, cas, psm_type)][keep]
 keep_tab[ , prop := round(prop*100, 2)]
-setnames(keep_tab, c('id', 'Compound', 'CAS', 'Group', '\\%>LOQ', 'no. > LOQ', 'total no.'))
+setnames(keep_tab, c('id', 'Compound', 'CAS', 'Group', '\\%>LOQ', 
+                     'no. > LOQ', 'total no.'))
 keep_tab$id <- NULL
 keep_tab
 
@@ -139,14 +142,15 @@ print(keep_tab_x,
 
 
 take <- take[variable_id %in% keep$variable_id]
-psm_variables[variable_id %in% keep$variable_id, list(variable_id, name, psm_type)]
+psm_variables[variable_id %in% keep$variable_id, list(variable_id, name, 
+                                                      psm_type)]
 
 rm(samples_rac, psm_sites, psm_sites_info, take_si, rac, keep_tab, keep_tab_x)
 
 
 # Model -------------------------------------------------------------------
 # one single compouns
[…]

### 2016-10-18 src/functions.R

@@ -461,3 +461,10 @@ read_regnie <- function(file, crs = NULL){
 # fix for ggsave [error: " plot should be a ggplot2 plot"]
 ggsave_fix <- ggplot2::ggsave; body(ggsave_fix) <- body(ggplot2::ggsave)[-2]
 
+#' log spaced sequence
+#' @source http://grokbase.com/t/r/r-help/099nykfnnf/r-logarithmic-seq
+#' 
+logspace <- function( d1, d2, length.out) {
+  exp(log(10)*seq(d1, d2, length.out=length.out))
+}
+

### 2016-10-18 article.tex
Update on Overleaf.
typos+ grammar

@@
 The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agricultural land use lead to an 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
-Precipitation increased measured pesticide risk by 5\% and was highest risk  during summer months.
-Risk thresholds were exceeded in 26 percent of small streams, with highest risk emerging from neonicotinoid insecticides. 
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
+Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
+Precipitation increased measured pesticide risk by 5\% and was thw highest during summer months.
+Risk thresholds were exceeded in 26 percent of small streams, with the highest risk emerging from neonicotinoid insecticides. 
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events.
 
-% Total words: 198
+% Total words: 197
 \end{abstract}
@@
 Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
-which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state specific needs. 
+which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state-specific needs. 
 Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
@@
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}.
-Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies were conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
+Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies was conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
 Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant.
 
-In this study we compiled and analyzed large scale chemical monitoring data from small streams in Germany. 
+In this study we compiled and analyzed large-scale chemical monitoring data from small streams in Germany. 
 First, we analysed the shape of the relationship between pesticide risk, agricultural land use and catchment size and examined whether related thresholds for pesticide risks can be derived. 
-Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
+Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
 Finally, we quantified the current risks from pesticides in small streams in Germany.
@@
 
-We identified precipitation at sampling sites by spatio-temporal intersection of sampling events with gridded daily precipitation data (60$\times$30 arcsec resolution) available from the German Weather service (DWD).
+We identified precipitation at sampling sites by a spatio-temporal intersection of sampling events with gridded daily precipitation data (60$\times$30 arcsec resolution) available from the German Weather service (DWD).
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
-We performed the intersection for the actual sampling date and the day before, an extracted precipitation during and up to 48 hours before sampling.
+We performed the intersection for the actual sampling date and the day before and extracted precipitation during and up to 48 hours before sampling.
 
@@
 We compiled a total of 2,369 sampling sites in small streams with pesticide measurements. %see do_overview.R
-Alongside, we also requested catchment sizes (available for 59\% of sites) and percentage agricultural land use within the catchment (available for 24\% of sites) for the sampling sites from the federal states.
-%see do_overview.R
+Alongside, we also requested catchment sizes and percentage agricultural land use within the catchment for the sampling sites from the federal states. %see do_overview.R
+Catchment size was available for 59\% of sites from authorities. 
 Additionally, we delineated upstream catchments for each of the sampling sites using (i) a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012} and (ii) from drainage basins provided by the Federal Institute of Hydrology (BfG).
@@
 
-To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as response variable.
+To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as the response variable.
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
@@
 We used appropriate link functions for $\mu$ and $\nu$ and assumed $\sigma$ to be constant. 
-Equation~\ref{eqn:eqn4} summarises the deterministic part of the model.
+Equation~\ref{eqn:eqn4} summarises the deterministic part of the model for a measurement $i$.
 
@@
 
-We fitted this model separately to each compound with a RAC, measured in least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
@@
 Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
-Although, there was high variation within the remaining cluster this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
+Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
 
@@
 The mean number of RAC exceedances per site increased strongly and statistically significant from 0.13 (no agriculture) to 0.44 (25\% agriculture within the catchment). 
-Above this threshold the exceedances levelled.
+Above this threshold the exceedances leveled.
 Above 75\% agriculture within the catchment the number of exceedances further increased, but the increase was not statistically significant (Figure~\ref{fig:fig4}, left). 
@@
 \subsection{Effect of precipitation on pesticide risk}
-The spatio-temporal intersection revealed that most samples were taken during periods with low precipitation.
+The spatio-temporal intersection revealed that most samples were taken during periods of low precipitation.
 For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
@@
 % do_precip.R
-Effect on individual compounds are provided in the Supplemental Table~S4.
+Effects on individual compounds are provided in the Supplemental Table~S4.
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
@@
 We found RQ > 1 in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
-In 23\% of sites chemicals with RAC where never detected (see also Supplemental Figure~S8).
+In 23\% of sites chemicals with RAC were never detected (see also Supplemental Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
-For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
+For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have an RQ~\textgreater~1. 
 The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ).
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Although, the outlined differences between states, all ecoregions occuring Germany \citep{illies1978limnofauna,abell2008freshwater} are covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
+Although, the outlined differences between states, all ecoregions occurring Germany \citep{illies1978limnofauna,abell2008freshwater} are covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
 For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
@@
 If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
-To our knowledge this is the first study investigating such thresholds of pesticide risk.
+To our knowledge, this is the first study investigating such thresholds of pesticide risk.
 Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
-\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in low land streams in Europe.
+\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
 Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use in wadeable streams in the United States.
@@
 This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
-For example the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
+For example, the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude.
 
@@
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found a 5\% higher RQ if samples where taken after rainfall events. 
-Samples taken at the day of a rainfall-event showed less risk, than samples taken one day after a rainfall-event.
+Nevertheless, we found a 5\% higher RQ if samples were taken after rainfall events. 
+Samples taken at the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
 This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentrations in agricultural small streams generally show short term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+Pesticide concentrations in agricultural small streams generally show short-term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ.
@@
 
-We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first Quarter (4\%, Figure~\ref{fig:fig5}, bottom right).
+We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter (4\%, Figure~\ref{fig:fig5}, bottom right).
 This yearly pattern coincides with their main application season for pesticides in Central Europe.
@@
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
-Currently, little is known on these and further research on those might provide useful information for future ecological risk assessment. 
+Currently, little is known about these and further research on those might provide useful information for future ecological risk assessment. 
 For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.
@@
 Our results suggest that small streams are frequently exposed to ecologically relevant pesticide concentrations.
-In one quarter of small streams RAC were exceeded at least once.
+In one quarter of small streams RACs were exceeded at least once.
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
@@
 This may be attributed to two differences: i) Different objectives of governmental monitoring and scientific studies.
-Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods with lower pesticide usage and at natural sites.
+Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods of lower pesticide usage and at natural sites.
 ii) Different sampling strategies.
@@
 We found RQ values greater than 1.12 in one quarter of small streams and RQ at 1/10 of RAC in more the half of streams. 
-Consequently, we conclude that on a large scale agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
+Consequently, we conclude that on a large-scale agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
 This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
@@
 % Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study large scale environmental occurrence patterns of pesticides.
+Monitoring data, despite the outlined limitations, provides an opportunity to study large-scale environmental occurrence patterns of pesticides.
 Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
@@
 % 2 small(300):  600
-% acknowledg  :   27
+% acknowledg  :   30
 %              =====
 % subtotal      3227
-% text body   : 3200
+% text body   : 3700
 % ======================================
-% total       : 6427
+% total       : 6930
 

### 2016-10-18 cover_letter/cover.tex
typos+ grammar

@@
     % Include the logos
-    \includegraphics[width=3in]{Uni-Logo-2c.jpg}\hspace{\fill}\includegraphics[width=2in]{logo_institut.jpg}
+    \includegraphics[width=3in]{Uni-Logo-2c.jpg}\hspace{\fill}\includegraphics[width=1.5in]{logo_institut.jpg}
     ~\\[-0.11in] % Reduce the whitespace above the horizontal rule
@@
 {\raggedright \toname \\ \toaddress \par} % Print the to name and address
-\vspace{2em}
+\vspace{3em}
 % Subject line
@@
 \long\def\closing#1{
-    \vspace{0.1in} % Some whitespace after the letter content and before the signature
+    \vspace{0.4in} % Some whitespace after the letter content and before the signature
     \noindent % Stop paragraph indentation
@@
     #1 % Print the signature text
-    \vskip 0.65in % Whitespace between the signature text and your name
+    \vskip 0.1in % Whitespace between the signature text and your name
     \fromsig}
@@
 \opening{Dear Editor,}
-\vspace{-1em}\\
-enclosed is a manuscript to be considered for publication in Environmental Science \& Technology titled \\
+\vspace{-2em}\\
+enclosed is a manuscript to be considered for publication in Environmental Science \& Technology titled \\[-1em]
 
-\emph{Pesticides in small streams in Germany}\\
+\emph{Large scale risks from pesticides in small streams}\\[-1em]
 
@@
 
-We compiled and homogenized monitoring data from all German Federal States with a focus on small to mid-size streams. 
-Our results show that there is currently a lack of knowledge for small streams. 
-Using new and appropriate modeling techniques we show that in the influence of agricultural land use is non-linear.
-Measured pesticide concentrations are influenced by precipitation before sampling and show a temporal pattern with a peak during summer.
-Especially, neonicotinoid pesticides may pose a risk to small agricultural freshwaters.
+We compiled and homogenized monitoring data from all German Federal States with a focus on small streams. 
+Our results show that:
+\begin{enumerate} 
+\item there is currently a lack of knowledge for small streams
+\item the influence of agricultural land use is non-linear with a threshold at 25\%
+\item measured pesticide concentrations increase by precipitation before sampling and therefore current monitoring underestimates risks
+\item show compound specific yearly dynamics that corresponds to applications
+\item especially, neonicotinoids pose a risk to small agricultural freshwaters.
+\end{enumerate}
 
-Our study should be of general interest and provides valuable information on chemical monitoring, pesticide occurrences and ecological risk arising from them.
+Our study should be of general interest and adds valuable information on chemical monitoring, pesticide dynamics and ecological risk arising from them.
 

### 2016-10-18 src/do_overview.R

@@ -20,11 +20,11 @@ var_props <- fread(file.path(cachedir, 'var_props.csv'))
 
 
 # restrict to sites < 100km² (or unknown) and > 2005
-range(psm_samples$date)
 take_sites <- psm_sites_info[ezg_fin <= 100 | is.na(ezg_fin), site_id]
+
 psm_sites_info <- psm_sites_info[site_id %in% take_sites]
 psm_sites <- psm_sites[site_id %in% take_sites]
-psm_samples<-psm_samples[site_id %in% take_sites]
+psm_samples <- psm_samples[site_id %in% take_sites]
 
 
 

### 2016-10-19 article.tex
more fixes, typos etc

@@
 \usepackage{enumerate}
-
-\usepackage{todonotes}
 
@@
 We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
-The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites and of 478 pesticides.  
+The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites of 478 pesticides.  
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
 Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
-Precipitation increased measured pesticide risk by 5\% and was thw highest during summer months.
+Precipitation increased measured pesticide risk by 5\% and was the highest during summer months.
 Risk thresholds were exceeded in 26 percent of small streams, with the highest risk emerging from neonicotinoid insecticides. 
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity, and that governmental pesticide sampling should be adapted to precipitation events.
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that governmental pesticide sampling should be adapted to precipitation events.
 
@@
 Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
-Nevertheless, a clear definition of SS in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
-For example, the WFD defines SS with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines SS with catchment size \textless 10km\textsuperscript{2}.
+Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
+For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}.
 
@@
 In this study we compiled and analyzed large-scale chemical monitoring data from small streams in Germany. 
-First, we analysed the shape of the relationship between pesticide risk, agricultural land use and catchment size and examined whether related thresholds for pesticide risks can be derived. 
+First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
 Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
@@
 The request encompassed only the years 2005 to 2015.
-We homogenized and unified all data provided by the federal states into a database and implemented a robust data cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
+We homogenized and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
 
@@
 For 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
-68 sites (3\%) lacking these two informations were omitted from the operations outlined below.
+68 sites (3\%) lacking these two pieces of information were omitted from the operations outlined below.
 %see do_overview.R
@@
 We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances at a site as:
+We modeled the number of RAC exceedances (RQ \textgreater 1) at a site as:
 
@@
 The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
-$f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006} and $\beta_0$ is the intercept.
+$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
 The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
@@
 
-To account for temporal auto-correlation and differences between federal states we used $site$ nested within $state$ as random intercepts.
+To account for temporal autocorrelation and differences between federal states we used $site$ nested within $state$ as random intercepts.
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
 
-We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with an RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
@@
 These samples were all taken via grab sampling.  
-We found large differences in the number of sampling sites between federal states and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
-The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW) and no data from Brandenburg.
+We found large differences between federal states in the number of sampling sites and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
+The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
+No data was available from Brandenburg.
 
@@
   \includegraphics[width=0.6\textwidth]{figure1.pdf}
-  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states, see Supplemental Table~S1 for abbreviations.}
+  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states (see Supplemental Table~S1 for abbreviations).}
   \label{fig:fig1}
@@
 Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
-4\% (71,113) of all measurements were detects above LOQ.
+4\% (=71,113) of all measurements were concentrations above LOQ.
 We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
@@
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-An increase of precipitation in $Q2$ before sampling ($prec_{-1}$) from 1mm to 10mm precipitation lead to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% on average (Figure~\ref{fig:fig5}, top).
+In $Q2$ an increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm lead on average to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% (Figure~\ref{fig:fig5}, top).
 % do_precip.R
-Effects on individual compounds are provided in the Supplemental Table~S4.
-Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$). 
+Effects differed between individual compounds and are provided in the Supplemental Table~S4.
+Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
 This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).
@@
   Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
-  Single compound coefficients are provided in the Supplemental Table~S4 and Figure~S7).
+  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
   }
@@
 \subsection{Pesticide risk in small streams}
-We found RQ > 1 in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
-In 23\% of sites chemicals with RAC were never detected (see also Supplemental Figure~S8).
+We found RAC exceedances in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
+In 23\% of sites no chemicals with RAC were detected (see also Supplemental Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
 For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have an RQ~\textgreater~1. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ).
-RAC exceedances were found in 15\% of samples with detects (and 7.7\% of all samples with RAC).
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
+RAC exceedances were found in 15\% of samples with concentrations \textgreater LOQ (and 7.7\% of all samples).
 
 The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples were it was analysed (n = 3038, see also Supplemental Figure~S9).
+Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supplemental Figure~S9).
 Glyphosate was the compound with the highest detection rates (41\%, n = 3557 samples), followed by Boscalid (23\%, n = 9886) and Isoproturon (22\%, n = 19112). 
-However, only the latter showed RQ exceedances (Figure~\ref{fig:fig6}).
+However, only the latter showed RAC exceedances (Figure~\ref{fig:fig6}).
 In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).
@@
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
-While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here for Germany, the compilation used in \citet{stone2014pesticides} is much smaller.
+While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here, the compilation used in \citet{stone2014pesticides} is much smaller.
 Nevertheless, there might more data for the United States available (e.g. from Water Quality Portal (WQP), \url{www.waterqualitydata.us}).
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Although, the outlined differences between states, all ecoregions occurring Germany \citep{illies1978limnofauna,abell2008freshwater} are covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
+Although the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
 For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
@@
 
-Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring (Figure~\ref{fig:fig3}). 
-Especially, sampling sites in streams below 10~km\textsuperscript{2} are disproportionally low, which may be attributed to the missing categorisation in the WFD. 
+Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring. 
+Given their abundance, small streams below 10~km\textsuperscript{2} are disproportionally less sampled (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD. 
 Clearly, there is currently a lack of knowledge for these small streams.
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
+If there is more the 25\% agriculture within a catchment pesticides, it is likely that an RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
 To our knowledge, this is the first study investigating such thresholds of pesticide risk.
@@
 
-We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter (4\%, Figure~\ref{fig:fig5}, bottom right).
+We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
 This yearly pattern coincides with their main application season for pesticides in Central Europe.
 Nevertheless, there are compound specific differences in the yearly pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
-For example, the herbicide Diflufenican showed highest RQ and probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
+For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics.
@@
 Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
-Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial ecosystems \citep{pisa2015effects}.
+Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.
 
@@
 By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-However, RAC exceedances were generally lower and less abundant compared to our results from Germany. 
+Moreover, RAC exceedances in Switzerland were generally lower and less abundant compared to our results from Germany. 
 This might reflect differences in pesticide use between countries and ecoregions. 
@@
 Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
-We found RQ values greater than 1.12 in one quarter of small streams and RQ at 1/10 of RAC in more the half of streams. 
-Consequently, we conclude that on a large-scale agricultural pesticides are a major threat to small streams, the biodiversity they host and the services they provide. 
+We found RQ values greater than 1.12 in one-quarter of small streams and RQ at 1/10 of RAC in more then half of small streams. 
+Consequently, we conclude that agricultural pesticides are on a large scale a major threat to small streams, the biodiversity they host and the services they provide. 
 This threat may exacerbate because pesticides often occur in mixtures \cite{schreiner_pesticide_2016} and may co-occur with other stressors \citep{schafer_contribution_2016}.
@@
 The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
+We thank Wolfram König and Alexandra Müller (German Federal Environmental Protection Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
 \end{acknowledgement}
@@
 % 2 small(300):  600
-% acknowledg  :   30
+% acknowledg  :   70
 %              =====
-% subtotal      3227
+% subtotal      3270
 % text body   : 3700
 % ======================================
-% total       : 6930
+% total       : 6970
 

### 2016-10-19 supplement/Supplemental_Materials.tex
more fixes, typos etc

@@
 \usepackage{float} % for [H]
+\usepackage{array, ragged2e}
+\newcolumntype{R}[1]{>{\RaggedLeft\arraybackslash}p{#1}} % right aling column with fixed width
 
@@
 	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
-	\item Chemicals: Chemical names and identifiers have been unified using the webchem package \citep{szocs_webchem:_2016}.
+	\item Chemicals: Chemical names and identifiers have been unified using the webchem package (\url{https://github.com/ropensci/webchem}).
 	\item  Identifiers: Unique identifiers have been assigned.
@@
 
+\input{racdat}
+
 \begin{figure}[ht]
@@
 23\% of sites showed no detection of compounds with RAC and are not shown due to logarithmic x-axis.}
-	\label{fig:pdetects}
+	\label{fig:prac_ex}
 \end{figure}
@@
 
-\bibliographystyle{apalike}      % basic style, author-year citations
-\bibliography{references}
+% \bibliographystyle{apalike}      % basic style, author-year citations
+% \bibliography{references}
 

### 2016-10-19 src/do_overview.R

@@ -167,9 +167,10 @@ names(psm_samples_tab)[6] <- 'no.compounds \\textsuperscript{b}'
 
 xtab <- xtable(psm_samples_tab, 
        label = 'tab:phch_overview',
-      caption = 'Overview on chemical samples. Only data from running waters and grab
+      caption = c('Overview on chemical samples. Only data from running waters and grab
 sampling is shown. \\textsuperscript{a}: Abbreviations according to ISO 3166-2:DE. 
       \\textsuperscript{b}: Including metabolites',
+                  'Overview on chemical samples.'),
       align = 'llllrrr'
 )
 print(xtab,
@@ -224,10 +225,11 @@ var_tab[Name == 'Benzoesäure', CAS := '65-85-0']
 
 var_tab_x <- xtable(var_tab, 
                     label = 'tab:phch_var',
-                    caption = 'Overview on pesticides in the database. \\
+                    caption = c('Overview on pesticides in the database. \\
                     \\textsuperscript{a} Authorized in Germany (Source: BVL, 2015). 
                     \\textsuperscript{b} Authorized in the EU (Source: EU).
                     \\textsuperscript{c} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).',
+                    'Overview on pesticides in the database.'),
                     align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1cm}')
                     
 print(var_tab_x,

### 2016-10-19 src/do_pollution.R

@@ -153,10 +153,41 @@ length(unique(meas$site_id))   # 2270 (from 2301 in total)
 length(unique(meas$sample_id)) # 24344 samples with rac (from 24743 in total)
 
 
+### table for comppunds with more the 1000 measurements
+rac_dat <- psm_variables[ , list(variable_id, psm_type, name)][  # join with variable_names
+  meas[ ,list(n_meas = .N,      # number of measurements
+            n_detects = sum(value_fin > 0),    # number of detects
+            p_detects = round(sum(value_fin > 0) / .N * 100, 1), # proportion of detects
+            n_racex = sum(value_fin > rak_uba),  # number of exceedances
+            p_racex = round(sum(value_fin > rak_uba) / .N * 100, 1), # prop of exceedances
+            p_racex_d = round(sum(value_fin > rak_uba) / sum(value_fin > 0) * 100, 1) # prop of exceedances of detects
+            ) , by = variable_id]
+  ]
+rac_dat[ , variable_id := NULL]
+rac_dat[ , psm_type := NULL]
+rac_dat <- rac_dat[order(Name)]
+setnames(rac_dat, c('Name', 'No. ', 'No. \\textgreater LOQ', 
+                    '\\% \\textgreater LOQ', 'No. RQ \\textgreater 1', 
+                    '\\% RQ \\textgreater 1', '\\% RQ \\textgreater 1 | \\textgreater LOQ'))
 
-
-
-
+rac_dat_x <- xtable(rac_dat, 
+               label = 'tab:rac_dat',
+               caption = c('Overview on RAC exceedances of the 76 compounds with more the 1000 measurements. No. = number of measurements;  \\% RQ \\textgreater 1 = RAC exceedances; \\% RQ \\textgreater 1 | \\textgreater LOQ= RAC exceedances as fraction of detects.',
+                           'Overview on RAC exceedances of the 76 compounds with more the 1000 measurements.'),
+               align = 'lp{3cm}rR{1.4cm}R{1.4cm}R{1.4cm}R{1.4cm}R{1.4cm}',
+               digits = 1
+)
+print(rac_dat_x,
+      file = file.path(prj, 'supplement/racdat.tex'),
+      tabular.environment="longtable",
+      floating = FALSE,
+      caption.placement = 'top',
+      include.rownames = FALSE,
+      comment = FALSE,
+      booktabs = TRUE,
+      hline.after = c(-1, 0, nrow(rac_dat)),
+      sanitize.text.function = identity
+)
 
 
 

### 2016-10-19 src/do_precip.R

@@ -116,15 +116,16 @@ setkey(psm_variables, variable_id)
 setkey(keep, variable_id)
 keep_tab <- psm_variables[ , list(variable_id, name, cas, psm_type)][keep]
 keep_tab[ , prop := round(prop*100, 2)]
-setnames(keep_tab, c('id', 'Compound', 'CAS', 'Group', '\\%>LOQ', 
+setnames(keep_tab, c('id', 'Name', 'CAS', 'Group', '\\%>LOQ', 
                      'no. > LOQ', 'total no.'))
 keep_tab$id <- NULL
 keep_tab
 
 keep_tab_x <- xtable(keep_tab, 
                     label = 'tab:var_model',
-                    caption = '24 pesticides for which we modelled the relationship with precipitation and seasonality.
+                    caption = c('24 pesticides for which we modelled the relationship with precipitation and seasonality.
                     Order is the same as in Figure 5 of the articles. See Table \\ref{tab:var_model_coef} for model coefficients.',
+                                '24 pesticides for which we modelled the relationship with precipitation and seasonality.'),
                     align = 'lp{2.5cm}rlp{1.5cm}p{2cm}p{2cm}')
 
 print(keep_tab_x,
@@ -313,9 +314,10 @@ rownames(keep_tab2_w)<-NULL
 
 keep_tab2_x <- xtable(keep_tab2_w, 
                      label = 'tab:var_model_coef',
-                     caption = 'Coefficients and CI from per compound models. 
+                     caption = c('Coefficients and CI from per compound models. 
                      Bold values denote coefficients where the CI for precipitation encompasses zero.
                      Coefficients are on the link scale (log for $\\mu$ and logit for $\\nu$).',
+                     'Coefficients and CI from per compound models.'),
                      align = 'lp{2cm}p{0.6cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}p{1.8cm}')
 
 print(keep_tab2_x, 
@@ -461,6 +463,7 @@ resmd <- rbindlist(resm)
 # log_preciü,p_1: nu 
 # intercept
 int <- resmd$est[resmd$term ==  'mu.seasonQ2']
+# int<-0
 # by hand:
 # p <- 1
[…]

### 2016-10-19 article.tex
up8 word count
update wc
fix supplement and refs to supplement

@@
 We fitted this model separately to each compound with an RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the 24 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 24 compounds.
+To summarise the coefficients across the 23 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 23 compounds.
 The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
-Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S3).
+Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S2).
 4\% (=71,113) of all measurements were concentrations above LOQ.
@@
 Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
-Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S4 and S5).
+Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
 
@@
   \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to two clusters determined by hierarchical clustering (see Supplemental Figure~S4).}
+  Colors according to two clusters determined by hierarchical clustering (see Supplemental Figure~S3 and S4).}
   \label{fig:fig2}
@@
 However, only the latter showed RAC exceedances (Figure~\ref{fig:fig6}).
-In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S8).
+In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S10).
 
@@
 %%% Word count
-% abstract    :  200
+% abstract    :  196
 % 4 big(600)  : 2400
 % 2 small(300):  600
-% acknowledg  :   70
+% acknowledg  :   69
 %              =====
 % subtotal      3270
-% text body   : 3700
+% text body   : 3734
 % ======================================
-% total       : 6970
+% total       : 6999
 

### 2016-10-19 src/do_overview.R

@@ -161,8 +161,15 @@ tot <- c('Total',
          length(unique(psm_samples$variable_id)))
 psm_samples_tab <- rbind(data.frame(psm_samples_tab), tot)
 psm_samples_tab <- psm_samples_tab[order(psm_samples_tab$state), ]
-names(psm_samples_tab)[1] <- 'state \\textsuperscript{a}'
-names(psm_samples_tab)[6] <- 'no.compounds \\textsuperscript{b}'
+psm_samples_tab$name <- c('Baden-Württemberg', 'Bavaria', 'Hesse', 
+                          'Mecklenburg Western Pomerania', 'Lower Saxony',
+                          'Northrhine-Westphalia', 'Rhineland Palatinate',
+                          'Schleswig Holstein', 'Saarland', 'Saxony', 
+                          'Saxony-Anhalt', 'Thuringia', ''
+                          )
+setcolorder(psm_samples_tab, c(7, 1:6))
+names(psm_samples_tab)[2] <- 'state \\textsuperscript{a}'
+names(psm_samples_tab)[7] <- 'no.compounds \\textsuperscript{b}'
 
 
 xtab <- xtable(psm_samples_tab, 
@@ -171,7 +178,7 @@ xtab <- xtable(psm_samples_tab,
 sampling is shown. \\textsuperscript{a}: Abbreviations according to ISO 3166-2:DE. 
       \\textsuperscript{b}: Including metabolites',
                   'Overview on chemical samples.'),
-      align = 'llllrrr'
+      align = 'lp{3cm}lllrrr'
 )
 print(xtab,
       file = file.path(prj, 'supplement/phchoverview.tex'),

### 2016-10-20 cover_letter/cover.tex
revised supplement + cover

@@
 \begin{enumerate} 
-\item there is currently a lack of knowledge for small streams
-\item the influence of agricultural land use is non-linear with a threshold at 25\%
-\item measured pesticide concentrations increase by precipitation before sampling and therefore current monitoring underestimates risks
-\item show compound specific yearly dynamics that corresponds to applications
+\item regulatory threshold exceedances depend non-linearly on agricultural land use within the catchment
+\item measured pesticide concentrations depend on precipitation events before sampling and therefore current monitoring underestimates risks
+\item the annual dynamic of concentrations is compound-specific and corresponds to applications
 \item especially, neonicotinoids pose a risk to small agricultural freshwaters.

### 2016-10-20 supplement/Supplemental_Materials.tex
fix supplement and refs to supplement
revised supplement + cover

@@
 \author{Eduard Szöcs, Marvin Brinke, Bilgin Karaoglan, Ralf B. Schäfer}
+\date{}
 
@@
 	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
-	\item Chemicals: Chemical names and identifiers have been unified using the webchem package (\url{https://github.com/ropensci/webchem}).
+	\item Chemicals: Chemical names and identifiers have been unified using the webchem package (https://github.com/ropensci/webchem).
 	\item  Identifiers: Unique identifiers have been assigned.
-	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification have be set to zero.
+	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification were set to zero (and can be used to identity non-detects).
 	\item Other meta-data: meta-data has been standardised.
@@
 	\centering
-	\includegraphics[width = 0.8\textwidth]{pmix}
-	\caption{Distribution the number of quantified compounds in the samples from small streams.}
-	\label{fig:pmix}
+	\includegraphics[width = 0.8\textwidth]{pdetects}
+	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are show.}
+	\label{fig:pdetects}
 \end{figure}
@@
 	\centering
-	\includegraphics[width = 0.8\textwidth]{pdetects}
-	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are show.}
-	\label{fig:pdetects}
+	\includegraphics[width = 0.8\textwidth]{pmix}
+	\caption{Distribution of the number of quantified compounds in the samples.}
+	\label{fig:pmix}
 \end{figure}

### 2016-10-20 article.tex
abstract

@@
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds once the proportion of agriculture in a catchment exceeded 25 percent. 
+Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 25 percent. 
 Precipitation increased measured pesticide risk by 5\% and was the highest during summer months.
-Risk thresholds were exceeded in 26 percent of small streams, with the highest risk emerging from neonicotinoid insecticides. 
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that governmental pesticide sampling should be adapted to precipitation events.
+Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven precipitation events.
 

### 2016-10-20 article.tex
main text

@@
 The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites of 478 pesticides. 
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics of pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
 Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 25 percent. 
@@
 Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven precipitation events.
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven by q1Qprecipitation events.
 
@@
 
-However, small streams might be also at high risk of pesticide contamination from adjacent agricultural areas and lower dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+However, small streams might also be at high risk of pesticide contamination in case of adjacent agricultural areas given their dilution potential \citep{schulz_field_2004,liess_determination_1999}.
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
@@
 The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
-We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to check if there are regions of statistically significant changes.
+We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
 GAMs were fitted using the mgcv package \citep{wood_fast_2011}.

### 2016-10-20 article.tex
methods text

@@
 Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven by q1Qprecipitation events.
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven by precipitation events.
 
-% Total words: 197
+% Total words: 195
 \end{abstract}
@@
 
-However, small streams might also be at high risk of pesticide contamination in case of adjacent agricultural areas given their dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+However, small streams might also be at high risk of pesticide contamination in case of adjacent agricultural areas given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
@@
 \subsection{Data compilation}
-We requested pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names).
-The request encompassed only the years 2005 to 2015.
+We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names) for 2005 to 2015.
 We homogenized and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
@@
 We compiled a total of 2,369 sampling sites in small streams with pesticide measurements. %see do_overview.R
-Alongside, we also requested catchment sizes and percentage agricultural land use within the catchment for the sampling sites from the federal states. %see do_overview.R
-Catchment size was available for 59\% of sites from authorities. 
+Alongside, we also queried catchment sizes and agricultural land use within the catchment for the sampling sites from the federal states. %see do_overview.R
+Catchment size was provided for 59\% of sites. 
 Additionally, we delineated upstream catchments for each of the sampling sites using (i) a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012} and (ii) from drainage basins provided by the Federal Institute of Hydrology (BfG).
@@
 
-For each derived catchment (either from DEM or drainage basins) we calculated the percentage agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+For each derived catchment (either from DEM or drainage basins) we calculated the \% agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
 Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
-For 97\% of the sites both, the proportion of agricultural land use and catchment size were available. 
-68 sites (3\%) lacking these two pieces of information were omitted from the operations outlined below.
+68 sites (3\%) that lacked catchment size or land use information were omitted from the analysis.
 %see do_overview.R

### 2016-10-20 article.tex
results text

@@
 
-The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites of catchments below $10~km^2$, with
-most sampling sites having catchment sizes between 10 and 25 $km^2$ (Figure~\ref{fig:fig3}).
+The distribution of sampling sites across catchment sizes indicated a disproportionally low number of sites with catchments below $10~km^2$, with
+most sampling sites having catchment sizes between 10 and 25~$km^2$ (Figure~\ref{fig:fig3}).
 

### 2016-10-21 article.tex
fix supp + minor fixes

@@
 Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 25 percent. 
-Precipitation increased measured pesticide risk by 5\% and was the highest during summer months. 
+Precipitation increased measured pesticide risk by 5\% and risk was the highest during summer months. 
 Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
@@
 First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
-Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010}\citep{schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
-Finally, we quantified the current risks from pesticides in small streams in Germany.
+Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010,schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
+Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.
 
@@
 
-We fitted this model separately to each compound with an RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
 To summarise the coefficients across the 23 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 23 compounds.
@@
 Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S2).
-4\% (=71,113) of all measurements were concentrations above LOQ.
 We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
@@
 Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
+4\% (=71,113) of all measurements were concentrations above LOQ.
 
@@
 The number of RAC exceedances increased strongly and statistically significant up to 25\% agriculture within the catchment.
-The mean number of RAC exceedances per site increased strongly and statistically significant from 0.13 (no agriculture) to 0.44 (25\% agriculture within the catchment). 
+The mean number of RAC exceedances per site increased 3.4-fold from 0.13 (no agriculture) to 0.44 (25\% agriculture within the catchment). 
 Above this threshold the exceedances leveled.
@@
 The spatio-temporal intersection revealed that most samples were taken during periods of low precipitation.
-For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day (Supplemental Figure~S6).
+For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day that may lead to run-off (Supplemental Figure~S6).
 
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-In $Q2$ an increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm lead on average to a 5\% higher mean RQ and the probability to exceed LOQ increases from 4\% to 6\% (Figure~\ref{fig:fig5}, top).
+In $Q2$ an increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm lead on average to a 5\% higher mean RQ and the probability to exceed LOQ increases 1.5-fold from 4\% to 6\% (Figure~\ref{fig:fig5}, top).
 % do_precip.R
@@
 Both increased during summer months and decreased towards the end of the year.
-There was a probability of 4\% to exceed LOQ in $Q1$ and 10\% in $Q2$.
+There was a 2.5-folder higher probability of exceeding LOQ in $Q2$ (10\%) than in $Q1$ (4\%).
 The differences were less pronounced for the mean value of RQ and with less precision (Figure~\ref{fig:fig5}, bottom). 
@@
 We found RAC exceedances in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
-In 23\% of sites no chemicals with RAC were detected (see also Supplemental Figure~S8).
+In 23\% of sites none of the chemicals, for which RAC were available, were detected (see also Supplemental Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
-For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have an RQ~\textgreater~1. 
+For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
 The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
@@
 \subsection{Overview on the compiled dataset}
-The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive one available for Germany.
+The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive available for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
-While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here, the compilation used in \citet{stone2014pesticides} is much smaller.
-Nevertheless, there might more data for the United States available (e.g. from Water Quality Portal (WQP), \url{www.waterqualitydata.us}).
+While the compilations from Europe are of similar quantity and quality to the  data compiled and analysed here, the compilation used in \citet{stone2014pesticides} is much smaller, though these data may be complemented by more data in future analyses.
 
 % current problems in monitoring and possible solutions
-A nation-wide assessment of pesticide pollution is hampered by the inhomogeneity of monitoring data between federal states:
+A nation-wide assessment of pesticide pollution is hampered by inhomogeneous data across federal states:
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Although the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
-For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that some exceedances were likely not detected.
-For these compounds a lowering of LOQ is essential for reliable assessment.
+Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
+For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that exceedances are likely underestimated.
+For these compounds a lowering of LOQ through an improvement of chemical analysis is essential for reliable assessment.
 Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.
 
-Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams are underrepresented in the current monitoring. 
-Given their abundance, small streams below 10~km\textsuperscript{2} are disproportionally less sampled (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD. 
-Clearly, there is currently a lack of knowledge for these small streams.
+Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams below 10~km\textsuperscript{2} are disproportionally less sampled in current monitoring (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD.
+Clearly, there is currently a lack of knowledge on stressor effects in small streams.
 We analysed only data from small streams, however, for lentic small water bodies this lack might be even greater \citep{lorenz_specifics_2016}.
@@
 We found a strong influence of agriculture on the pollution of streams.
-If there is more the 25\% agriculture within a catchment pesticides, it is likely that an RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
+Above 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
 To our knowledge, this is the first study investigating such thresholds of pesticide risk.
@@
 We did not find a relationship between pesticide pollution and catchment size.
-However, previous studies showed that small streams are more polluted that bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
+However, previous studies showed that small streams are more polluted than bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
 This can be explained by the relatively short gradient of catchment sizes in our dataset, with most of the streams with catchments above $10~km^2$ and below $100~km^2$ (Figure~\ref{fig:fig3}, top).
@@
 Samples taken at the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
-This could be explained by a sampling just before the actual rainfall event.
-Pesticide concentrations in agricultural small streams generally show short-term peak concentrations \citep{wittmer_significance_2010}, therefore, a sampling at the day after a rainfall-event might miss peak concentrations.
+This could be explained by the sampling preceding the rainfall event and the delay between the start of a rain event and the peak in discharge or runoff. 
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ. 
 This could be attributed to a higher variability of absolute concentrations.
-Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, considerably underestimates pesticide risks.
+Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
 Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for small water bodies \citep{lorenz_specifics_2016}.
 
-We found highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
-This yearly pattern coincides with their main application season for pesticides in Central Europe.
-Nevertheless, there are compound specific differences in the yearly pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+We found the highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
+This annual pattern coincides with the main application season for pesticides in Central Europe.
+Nevertheless, there are compound specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
 For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
@@
 
-Compared to \citet{stehle_pesticide_2015} we found lower rates RAC exceedances (15\% (n = 12,615) vs 44\% (n = 1,566) of samples with measurements above LOQ). 
+Compared to \citet{stehle_pesticide_2015} we found lower rates of RAC exceedances (15\% (n = 12,615) vs 44\% (n = 1,566) of samples with measurements above LOQ). 
 This may be attributed to two differences: i) Different objectives of governmental monitoring and scientific studies.
@@
 
-By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos.
-Moreover, RAC exceedances in Switzerland were generally lower and less abundant compared to our results from Germany. 
-This might reflect differences in pesticide use between countries and ecoregions. 
+By contrast, \citet{knauer_pesticides_2016} found exceedances from monitoring data mainly for herbicides and fungicides and only one insecticide Chlorpyrifos-methyl.
+Moreover, RAC exceedances in Switzerland were generally lower and less abundant (for example 6 exceedances (=0.2\%) for Isoproturon with a maximum RQ of 2) compared to our results for Germany. 
+This might reflect differences in pesticivar_tde use between countries, ecoregions and RACs used. 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
@@
 Our results suggest that exceedances of RACs are landscape dependent % = Zusammenhang RAC<->Agriculture
-and therefore, landscape features should be taken into account in the registration process of pesticides.
+and therefore, pesticide regulation should account for landscape features. 
 Moreover, the high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
@@
 \begin{acknowledgement}
-The authors thank the federal state authorities for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
-We thank Wolfram König and Alexandra Müller (German Federal Environmental Protection Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
+The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
+We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Federal Environmental Protection Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
 \end{acknowledgement}

### 2016-10-21 src/do_overview.R

@@ -238,7 +238,8 @@ var_tab_x <- xtable(var_tab,
                     \\textsuperscript{b} Authorized in the EU (Source: EU).
                     \\textsuperscript{c} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).',
                     'Overview on pesticides in the database.'),
-                    align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1cm}')
+                    align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1.5cm}',
+                    digits = 4)
                     
 print(var_tab_x,
       file = file.path(prj, 'supplement/phchvar.tex'),

### 2016-11-07 article.tex
revised according to marvins comments

@@
 \author{Bilgin Karaoglan}
-\affiliation[German Federal Environmental Agency]{Federal Environmental Agency (UBA), Dessau-Roßlau, Germany}
+\affiliation[German Federal Environment Agency]{German Federal Environment Agency (UBA), Dessau-Roßlau, Germany}
 
@@
 We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
-The data set comprised of 1,766,104 measurements related to 24,743 samples in 2,301 sampling sites of 478 pesticides. 
+The data set comprised of 1,766,104 measurements of 478 pesticides (including metabolites) related to 24,743 samples from 2,301 sampling sites. 
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
@@
 
-% Total words: 195
+% Total words: 198
 \end{abstract}
@@
 More than 50\% of the total land area in Germany is used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more than 45,000 tonnes of 766 authorized pesticides were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+In the year 2014 more than 45,000 tonnes of 766 authorized plant protection products were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
 The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
@@
 which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state-specific needs. 
-Despite these data providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
+Despite that these data are providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
 
@@
 For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}.
-
-However, small streams might also be at high risk of pesticide contamination in case of adjacent agricultural areas given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+Moreover, small streams might particularly be at high risk of pesticide contamination in case of adjacent agricultural areas given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
@@
 Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant. 
+In European law, the Directive 2009/128/EC \citep{European_Union_2009} places an obligation on the EU Member States to adopt National Action Plans (NAP) for the Sustainable Use of Plant Protection Products and the German NAP also addresses the knowledge gap concerning pesticide impact on small steams, specifically including those with catchment size \textless 10km\textsuperscript{2}.
 
@@
 RACs are derived during pesticide authorization as part of the ecological risk assessment.
-No unacceptable ecological effect are expected if the environmental concentration remains below this concentration.
+No unacceptable ecological effects are expected if the environmental concentration remains below this concentration.
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
-The German Federal Environmental Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
-We expressed RACs as Risk Quotient (RQ):
+The German Federal Environment Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
+Based on theses RACs, we calculated Risk Quotients (RQ):
 
@@
 
-where $C_i$ is the concentration of a compound $i$ in a sample.
+where $C_i$ is the concentration of a compound $i$ in a sample and $RAC_i$ the respective RAC.
 
@@
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution (Equation~\ref{eqn:eqn3}) \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016}.
+Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016} (Equation~\ref{eqn:eqn3}).
 These two processes can be interpreted as changes in the mean value of RQ (change in $\mu$) and changes in the probability of exceeding LOQ and showing any risk (change in $\nu$).
@@
 
-The compiled dataset used in analysis comprised 1,766,104 pesticide measurements of 24,743 samples at 2,301 sampling sites in small streams.  %see do_overview.R for numbers.
+The compiled dataset used for analysis comprised 1,766,104 pesticide measurements in 24,743 samples from 2,301 sampling sites in small streams.  %see do_overview.R for numbers.
 These samples were all taken via grab sampling.  
@@
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
-Most samples were taken in the months April till October, with fewer samples during winter (see Supplemental Figure~S2).
+Most samples were taken in the months April till October, while fewer samples were taken during winter (see Supplemental Figure~S2).
 We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
 The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1). 
-Hierarchical clustering revealed that RP and NI analysed a distinct compound spectra compared to the cluster of other states.
+Hierarchical clustering revealed that RP and NI analysed distinct compound spectra compared to the cluster of other states. 
+However, it has to be noted that both states surveyed these distinct spectra in special monitoring programs from only a few sites. 
 Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and therefore, may represent a sample of all small streams in Central Europe.
+Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and thus it might nonetheless represent a sample covering all types of small streams in Germany.
 For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that exceedances are likely underestimated.
@@
 We found a strong influence of agriculture on the pollution of streams.
-Above 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in fully agricultural catchments (above 75 \% agriculture).
+Above 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
 To our knowledge, this is the first study investigating such thresholds of pesticide risk.
-Previous studies examined thresholds for percent agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+Previous studies examined thresholds for percentage of agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
 \citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
@@
 This could be explained by the sampling preceding the rainfall event and the delay between the start of a rain event and the peak in discharge or runoff. 
-The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for RQ. 
+The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for the absolute value of RQ. 
 This could be attributed to a higher variability of absolute concentrations.
@@
 This annual pattern coincides with the main application season for pesticides in Central Europe.
-Nevertheless, there are compound specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left).
+Nevertheless, there are compound-specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left). 
 For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
@@
 Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
-Our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.
+Moreover, our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.
 
@@
 Moreover, RAC exceedances in Switzerland were generally lower and less abundant (for example 6 exceedances (=0.2\%) for Isoproturon with a maximum RQ of 2) compared to our results for Germany. 
-This might reflect differences in pesticivar_tde use between countries, ecoregions and RACs used. 
+This might reflect differences in pesticide use between countries, ecoregions and RACs used. 
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
@@
 Monitoring data, despite the outlined limitations, provides an opportunity to study large-scale environmental occurrence patterns of pesticides.
-Nevertheless, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}, though the sampling design need to account for precipitation events to provide robust data.
+Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
+However, the sampling design needs to account for precipitation events to provide robust data. 
 Our results suggest that exceedances of RACs are landscape dependent % = Zusammenhang RAC<->Agriculture
@@
 \begin{acknowledgement}
-The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Federal Environmental Protection Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
-We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Federal Environmental Protection Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
+The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Federal Environment Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
+We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Federal Environment Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
 \end{acknowledgement}
@@
 %%% Word count
-% abstract    :  196
+% abstract    :  198
 % 4 big(600)  : 2400
 % 2 small(300):  600
-% acknowledg  :   69
+% acknowledg  :   82
 %              =====
-% subtotal      3270
-% text body   : 3734
+% subtotal      3280
+% text body   : 3800
 % ======================================
-% total       : 6999
+% subtotal    : 7080
+% - captions  :  155
+% =========================================
+% total       : 6925
 

### 2016-11-07 supplement/Supplemental_Materials.tex
revised according to marvins comments

@@
 \chapter{Data Cleaning}
-More then 30 datasets have been cleaned and homogenized separately, before combing into a common database.
-Cleaning steps comprised (Figure~\ref{fig:data_cleaning} gives a graphical overview).
+Before combining into a common database, more then 30 datasets have been cleaned and homogenized separately.
+Cleaning steps comprised the following steps (Figure~\ref{fig:data_cleaning} gives a graphical overview):
 
 \begin{enumerate}
-	\item Structure: Structure has been adjusted to the database structure.
+	\item Structure: Datasets have been adjusted to the database structure.
 	\item Coordinates: Coordinates have been transformed to a common Coordinate Reference System (DHDN / 3-Grad Gauss-Krüger Zone 3 (EPSG:31467)) and duplicates merged.
@@
 	\item Other meta-data: meta-data has been standardised.
-	\item Temporal resolution: The temporal resolution of the database is 1 day. Date below this resolution has been aggregated by maximum.
-	\item Validity Checks: Simple rules for validity checks have been implemented (e.g. no negative concentrations).
+	\item Temporal resolution: The temporal resolution of the database is 1 day. Samplings below this resolution have been aggregated by the maximum daily value.
+	\item Validity Checks: Simple rules for validity checks have been implemented.
 \end{enumerate}
@@
 	\includegraphics[width = 0.8\textwidth]{data_cleaning}
-	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning data has been stored in a relational spatial PostgreSQL database.}
+	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning, data has been stored in a relational spatial PostgreSQL database.}
 	\label{fig:data_cleaning}
@@
 	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.
-Abbreviations of state names according to ISO 3166-2:DE.}
+Abbreviations of state names according to ISO 3166-2:DE (see also Table~\ref{tab:phch_overview}).}
 	\label{fig:varclus}
@@
 	\includegraphics[width = 0.6\textwidth]{silhouette}
-	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes. Two clusters showed maximum silhouette width.}
+	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes of complete linkage clustering of jaccard similarity of analysed compound spectra between federal states.
+Two clusters showed maximum silhouette width.}
 	\label{fig:silhouette}
@@
 % table generated in do_overview.R
+\newpage
 \input{phchvar}
@@
 
+%%% from do_precip.R
 \input{keeptab}
@@
 Bottom row: estimates for the four Quarters.
-Each dot represent one compound (in the order described in table \ref{tab:var_model}).
+Each dot represents one compound (in the order described in table \ref{tab:var_model}).
 Coefficients where the CI encompasses zero are shown in gray colour.
@@
 \chapter{Pesticides in small streams}
-
+%% from do_pollution.R
 \input{racdat}
@@
 Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1.
-23\% of sites showed no detection of compounds with RAC and are not shown due to logarithmic x-axis.}
+23\% of the sites showed no detection of compounds with available RAC values and are not shown due to logarithmic x-axis.}
 	\label{fig:prac_ex}
@@
 	\includegraphics[width = 0.8\textwidth]{pdetects}
-	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are show.}
+	\caption[Proportion of samples with detects in small streams.]{Proportion of samples with detects in small streams. Only Compounds with more than 100 samples and 15\% of detects are shown.}
 	\label{fig:pdetects}

### 2016-11-07 src/do_overview.R

@@ -162,9 +162,9 @@ tot <- c('Total',
 psm_samples_tab <- rbind(data.frame(psm_samples_tab), tot)
 psm_samples_tab <- psm_samples_tab[order(psm_samples_tab$state), ]
 psm_samples_tab$name <- c('Baden-Württemberg', 'Bavaria', 'Hesse', 
-                          'Mecklenburg Western Pomerania', 'Lower Saxony',
-                          'Northrhine-Westphalia', 'Rhineland Palatinate',
-                          'Schleswig Holstein', 'Saarland', 'Saxony', 
+                          'Mecklenburg-Western Pomerania', 'Lower Saxony',
+                          'North Rhine-Westphalia', 'Rhineland-Palatinate',
+                          'Schleswig-Holstein', 'Saarland', 'Saxony', 
                           'Saxony-Anhalt', 'Thuringia', ''
                           )
 setcolorder(psm_samples_tab, c(7, 1:6))
@@ -228,22 +228,25 @@ var_tab$Group <- gsub('organics, psm, ', '', var_tab$Group)
 # fix bug with encoding (extra space or so....)
 var_tab[Name == 'Benzoesäure', CAS := '65-85-0']
 
+# order by name
+var_tab <- var_tab[order(Name)]
+
 
 
 
 var_tab_x <- xtable(var_tab, 
                     label = 'tab:phch_var',
-                    caption = c('Overview on pesticides in the database. \\
-                    \\textsuperscript{a} Authorized in Germany (Source: BVL, 2015). 
-                    \\textsuperscript{b} Authorized in the EU (Source: EU).
-                    \\textsuperscript{c} Regulatory Acceptable Concentration [ug/L] (Source: German EPA).',
+                    caption = c('Overview on pesticides (and metabolites) in the database. \\
+                    \\textsuperscript{a} Authorized in Germany (Source: German Federal Office of Consumer Protection and Food Safety (BVL) as at March 2015). 
+                    \\textsuperscript{b} Authorized in the European union (Source: EU Pesticides database as at March 2015).
+                    \\textsuperscript{c} Regulatory Acceptable Concentration [ug/L] (Source: German Federal Environment Agency (UBA) as at November 2015).',
                     'Overview on pesticides in the database.'),
-                    align = 'lp{3cm}rlp{0.5cm}p{0.5cm}p{1.5cm}',
+                    align = 'lp{4cm}rlp{1.3cm}p{1.3cm}p{1.5cm}',
                     digits = 4)
[…]

### 2016-11-12 article.tex
reanalysed everthing with new/fixed racs and comments from BK

@@
 \author{Bilgin Karaoglan}
-\affiliation[German Federal Environment Agency]{German Federal Environment Agency (UBA), Dessau-Roßlau, Germany}
+\affiliation[German Environment Agency]{German Environment Agency (UBA), Dessau-Roßlau, Germany}
 
@@
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
-Agricultural land use lead to a 3.4-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 25 percent. 
+Agricultural land use lead to a 3.7-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 28 percent. 
 Precipitation increased measured pesticide risk by 5\% and risk was the highest during summer months. 
@@
 More than 50\% of the total land area in Germany is used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more than 45,000 tonnes of 766 authorized plant protection products were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
-The applied pesticides may enter surface waters via spray-drift, edge-off-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
+In the year 2014 more than 45,000 tonnes of 776 authorized plant protection products were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+The applied pesticides may enter surface waters via spray-drift, edge-of-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
 Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
@@
 
-In this study we compiled and analyzed large-scale chemical monitoring data from small streams in Germany. 
+In this study, we compiled and analysed large-scale chemical monitoring data from small streams in Germany. 
 First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
@@
 
-We identified precipitation at sampling sites by a spatio-temporal intersection of sampling events with gridded daily precipitation data (60$\times$30 arcsec resolution) available from the German Weather service (DWD).
+We identified precipitation at sampling sites by a spatio-temporal intersection of sampling events with gridded daily precipitation data (60$\times$30 arcsec resolution) available from the German Meteorological Service (DWD).
 This data spatially interpolates daily precipitation values from local weather stations \citep{rauthe_central_2013}. 
@@
 
-For each derived catchment (either from DEM or drainage basins) we calculated the \% agricultural land-use within the catchment based on Official Topographical Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
+For each derived catchment (either from DEM or drainage basins) we calculated the \% agricultural land-use within the catchment based on the Authoritative Topographic-Cartographic Information System (ATKIS) of the land survey authorities \citep{adv_atkis_2016}.
 Thus, agricultural land use information was available for 98\% of all sites (24\% from authorities, 52\% from DEM and 22\% from drainage basins). 
-68 sites (3\%) that lacked catchment size or land use information were omitted from the analysis. 
+68 sites (3\%) that lacked catchment size or land use information were omitted from the analysis, resulting in 2301 sites used in the analyses outlined below.
 %see do_overview.R
@@
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
-The German Federal Environment Agency (UBA) provided RACs for the 105 compounds with highest detection rates (Supplemental Table~S2). 
+The German Environment Agency (UBA) provided RACs for 107 compounds including those with highest detection rates (Supplemental Table~S2). 
 Based on theses RACs, we calculated Risk Quotients (RQ):
@@
 
-We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 23 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the 23 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 23 compounds.
+We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 22 compounds, see Supplemental Table~S3 for a list of compounds).
+To summarise the coefficients across the 22 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 22 compounds.
 The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
@@
 \subsection{Influence of agricultural land use and catchment size}
-The number of RAC exceedances increased strongly and statistically significant up to 25\% agriculture within the catchment.
-The mean number of RAC exceedances per site increased 3.4-fold from 0.13 (no agriculture) to 0.44 (25\% agriculture within the catchment). 
+The number of RAC exceedances increased strongly and statistically significant up to 28\% agriculture within the catchment.
+The mean number of RAC exceedances per site increased 3.7-fold from 0.10 (no agriculture) to 0.39 (28\% agriculture within the catchment). 
 Above this threshold the exceedances leveled.
@@
 \subsection{Pesticide risk in small streams}
-We found RAC exceedances in 26\% of sampling sites and RQ > 0.1 in 54\% of sites. 
+We found RAC exceedances in 25.5\% of sampling sites and RQ > 0.1 in 54\% of sites. 
 In 23\% of sites none of the chemicals, for which RAC were available, were detected (see also Supplemental Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
-For Thiacloprid and Chlorpyrifos the RAC was less than LOQ, therefore, all detections have a RQ~\textgreater~1. 
+For Thiacloprid and Chlorpyrifos the RAC was equal or less than LOQ, therefore, all detections have a $RQ \ge 1$. 
 The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
-RAC exceedances were found in 15\% of samples with concentrations \textgreater LOQ (and 7.7\% of all samples).
+RAC exceedances were found in 14\% of samples with concentrations \textgreater LOQ (and 7.3\% of all samples).
 
-The highest RQ were observed for Chlorpyrifos (max(RQ) = 244), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
+The highest RQ were observed for Chlorpyrifos (max(RQ) = 220), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
 Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supplemental Figure~S9).
@@
 % size
-We did not find a relationship between pesticide pollution and catchment size.
+We did not find a statistically significant relationship between pesticide pollution and catchment size.
 However, previous studies showed that small streams are more polluted than bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
@@
 Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
-Automatic event-drive samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for small water bodies \citep{lorenz_specifics_2016}.
+Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for small water bodies \citep{lorenz_specifics_2016}.
 
@@
 
-Compared to \citet{stehle_pesticide_2015} we found lower rates of RAC exceedances (15\% (n = 12,615) vs 44\% (n = 1,566) of samples with measurements above LOQ). 
-This may be attributed to two differences: i) Different objectives of governmental monitoring and scientific studies.
+%%%! den absatz hier nochmal überarbeiten
+Compared to \citet{stehle_pesticide_2015} we found lower rates of RAC exceedances (5\% (n = 39,609) vs 37.1\% (n = 1,352) of measurements above LOQ). 
+This may be attributed to several differences: i) Different compound spectra.
+\citet{stehle_pesticide_2015} analysed 23 insectices, we 107 pesticides of different groups.
+ii) Different objectives of governmental monitoring and scientific studies.
 Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods of lower pesticide usage and at natural sites.
-ii) Different sampling strategies.
+iii) Different sampling strategies.
 The dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013}.
+%%%! den absatz hier nochmal überarbeiten
 
@@
 Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
-We found RQ values greater than 1.12 in one-quarter of small streams and RQ at 1/10 of RAC in more then half of small streams. 
+We found RQ values greater than 1.12 in 25\% of small streams and RQ at 1/10 of RAC in 54\% of small streams. 
 Consequently, we conclude that agricultural pesticides are on a large scale a major threat to small streams, the biodiversity they host and the services they provide. 
@@
 and therefore, pesticide regulation should account for landscape features. 
-Moreover, the high exceedances of RAC indicate that the authorisation process for pesticides may require further development.
+Moreover, the high exceedances of RACs indicate that greater efforts are needed to describe causal links, which may lead to further developments of the current authorisation procedure.
 
@@
 \begin{acknowledgement}
-The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Federal Environment Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
-We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Federal Environment Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
+The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Environment Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
+We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Environment Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
 \end{acknowledgement}

### 2016-11-12 src/clean.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       e.g. prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))

### 2016-11-12 src/do_overview.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -115,13 +115,14 @@ ggsave(file.path(prj, 'supplement/temporal.pdf'), p_temp, width = 20, height = 7
 
 # Spatial distribution ----------------------------------------------------
 # (=Figure 1)
-coordinates(psm_sites) <- ~easting + northing
-proj4string(psm_sites) <- CRS("+init=epsg:31467")
-psm_sites <- spTransform(psm_sites, CRS('+init=epsg:4326'))
-psm_sites_bb <- psm_sites@bbox
-psm_sites <- cbind(coordinates(psm_sites), psm_sites@data)
-psm_sites$state_ab <- gsub('(.*?)_.*', '\\1', psm_sites$site_id)
-setDT(psm_sites)
+psm_sites_p <- as.data.frame(psm_sites)
+coordinates(psm_sites_p) <- ~easting + northing
+proj4string(psm_sites_p) <- CRS("+init=epsg:31467")
+psm_sites_p <- spTransform(psm_sites_p, CRS('+init=epsg:4326'))
+psm_sites_bb <- psm_sites_p@bbox
+psm_sites_p <- cbind(coordinates(psm_sites_p), psm_sites_p@data)
+psm_sites_p$state_ab <- gsub('(.*?)_.*', '\\1', psm_sites_p$site_id)
+setDT(psm_sites_p)
 
 # get administrative borders
 adm1 <- raster::getData('GADM', country = 'DE', level = 1)
@@ -129,7 +130,7 @@ adm1 <- fortify(adm1)
 p_map <- ggplot() +
   geom_polygon(data = adm1, aes(x = long, y = lat, group = group), fill = "grey90") +
   geom_path(data = adm1, aes(x = long, y = lat, group = group), size = .3) +
-  geom_point(data = psm_sites, aes(x = easting, y = northing, col = state_ab), 
+  geom_point(data = psm_sites_p, aes(x = easting, y = northing, col = state_ab), 
              size = 1) +
[…]

### 2016-11-12 src/do_pollution.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -72,10 +72,12 @@ setkey(take_rac, variable_id)
 setkey(psm_variables, variable_id)
 meas <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 
-# remove variable with less then 1000 samples
-meas <- meas[variable_id %in% meas[ , list(n = length(value)), 
-                                  by = variable_id][n > 1000, variable_id]]
 
+
+# # remove variables with less then 1000 samples
+# meas <- meas[variable_id %in% meas[ , list(n = length(value)), 
+#                                   by = variable_id][n > 1000, variable_id]]
+# 
 
 
 # calculate max per variable and take only first 15 hits
@@ -93,34 +95,40 @@ take_meas[ , list(media = max(rq)) , by = name][order(media, decreasing = TRUE)]
 # for each sample, take maximum RQ (from different compounds) and check exceedance
 nrow(meas[ ,list(rq_max = max(rq)) , 
            by = sample_id][rq_max > 1]) /  length(unique(meas$sample_id)) * 100
-# in 7.7\% of samples taken there was was RQ > 1 found
+# in 7.3\% of samples taken there was was RQ > 1 found
 nrow(meas[ ,list(rq_max = max(rq)) , 
            by = sample_id][rq_max > 1]) /  length(unique(meas$sample_id[meas$rq > 0])) * 100
-# in 14.6\% of samples with detecs was was RQ > 1 found
+# in 14\% of samples with detecs was was RQ > 1 found
 
 
 ## % of measurements
-mean(meas[ , rq >1])*100
-mean(meas[rq > 0 , rq >1])* 100
+mean(meas[ , rq > 1])*100 # 0.2 % of all measurements
+nrow(meas[rq > 0])
+mean(meas[rq > 0 , rq > 1]) * 100 # 5% of all detects
+
+# number of detects
+
[…]

### 2016-11-12 src/do_precip.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -94,8 +94,8 @@ take <- psm_variables[ , list(variable_id, psm_type)][take]
 
 
 # remove compounds with no risk
-props <- take[  , list(prop = sum(rq>0) / length(rq),
-              abs = sum(rq>0),
+props <- take[  , list(prop = sum(rq > 0) / length(rq),
+              abs = sum(rq > 0),
               tot = length(rq)), by = variable_id][order(prop, 
                                                          decreasing = TRUE)]
 hist(props[prop < 0.2 , prop])
@@ -104,7 +104,7 @@ hist(props[prop < 0.2 , prop])
 hist(props[tot < 1000, tot])
 
 (keep <- props[prop > 0.05 & tot > 1000])
-# 25 compounds left
+# 23 compounds left
 
 
 
@@ -151,7 +151,8 @@ rm(samples_rac, psm_sites, psm_sites_info, take_si, rac, keep_tab, keep_tab_x)
 
 # Model -------------------------------------------------------------------
 # one single compouns
-take_c <- take[variable_id == 191]
+# take_c <- take[variable_id == 191]
+take_c <- take[variable_id == 378]
 
 
 gco <- glim.control(glm.trace = TRUE, bf.trace = FALSE)
@@ -165,9 +166,9 @@ glco <- gamlss.control(c.crit = 0.01, mu.step = 0.5)
 # gamma hurdle model with precipitation as groups, season and psm_type as preds
 # state as random effect
 mod_l_m <- gamlss(rq ~ 0  + log_precip_1 + log_precip0 + season +
-                    re(random=~1|state_fac/s_id_fac),
-                  nu.formula =~ 0 + precip_1 + precip0 + season + 
-                    re(random=~1|state_fac/s_id_fac),
[…]

### 2016-11-12 src/do_sizeagri.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Paper/ms_est' or
+       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -65,7 +65,7 @@ rak_exceed
 rm(samples_raks, samples_exceed)
 # exclude NA (=sites wihtout ezg & agri info)
 rak_exceed <- rak_exceed[!(is.na(agri_fin) | is.na(ezg_fin)), ]
-# 2343 sites with ezg and agri data
+# 2270 sites with ezg and agri data
 
 
 # logn used as offset
@@ -205,8 +205,8 @@ pdat <- transform(pdat,
                   sig_value = ifelse(sig, fit, NA))
 
 mylabeller <- as_labeller(c(
-  'agri_fin'="Agriculture [%]",
-  'ezg_fin'="Catchment Size [km2]"
+  'agri_fin' = "Agriculture [%]",
+  'ezg_fin' = "Catchment Size [km2]"
 ))
 
 
@@ -214,8 +214,8 @@ mylabeller <- as_labeller(c(
 odat <- melt(rak_exceed[ , list(ezg_fin, agri_fin)])
 p <- ggplot() +
   geom_line(data = pdat, aes(x = value, y = fit, group = variable)) +
-  geom_line(data = pdat, aes(x = value, y = up, group = variable), lty ='dashed') +
-  geom_line(data = pdat, aes(x = value, y = low, group = variable), lty ='dashed') +
+  geom_line(data = pdat, aes(x = value, y = up, group = variable), lty = 'dashed') +
+  geom_line(data = pdat, aes(x = value, y = low, group = variable), lty = 'dashed') +
   geom_line(data = pdat, aes(x = value, y = sig_value, group = variable), 
             colour = 'red', lwd = 1.5) +
   facet_wrap(~variable, scales = 'free_x', labeller = mylabeller) +
@@ -231,5 +231,5 @@ ggsave(file.path(prj, "figure4.pdf"),
 
 pdat
 pdat[c(1, 29), ]
[…]

### 2016-11-12 src/load.R

@@ -4,6 +4,7 @@ rm(list = ls()[!ls() %in% 'prj'])
 ## Project Path
 # prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project' # EDiLD
 # prj <- 'C:\\Users\\Edi\\Documents\\BFG\\monitoring_ger' # EDiLD2
+# prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est'
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
        e.g. prj <- '/home/edisz/Documents/Uni/Projects/PHD/4BFG/Project'!")

### 2016-11-13 article.tex
revised comparison with stehle & Schulz

@@
 
-%%%! den absatz hier nochmal überarbeiten
-Compared to \citet{stehle_pesticide_2015} we found lower rates of RAC exceedances (5\% (n = 39,609) vs 37.1\% (n = 1,352) of measurements above LOQ). 
-This may be attributed to several differences: i) Different compound spectra.
-\citet{stehle_pesticide_2015} analysed 23 insectices, we 107 pesticides of different groups.
-ii) Different objectives of governmental monitoring and scientific studies.
-Scientific studies may be slightly biased towards streams with high pollution to detect effects, whereas monitoring aims mainly at spatially representative surveillance of water quality, also during periods of lower pesticide usage and at natural sites.
-iii) Different sampling strategies.
-The dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013}.
-%%%! den absatz hier nochmal überarbeiten
+Compared to \citet{stehle_pesticide_2015} we found higher rates of RAC exceedances for insecticides.
+They found exceedances in 37.1\% of measurements \textgreater LOQ (n = 1352, 23 insecticides), whereas, we found exceedances in 67\% of measurements \textgreater LOQ (n = 1855, 22 insecticides).
+This could be attributed to different insecticides considered and different underlying RACs.
+Our study has only 7 insecticides with RACs in common with the insecticides investigated by \citet{stehle_pesticide_2015}.
+Moreover, all RACs were lower in our study (mean = -0.71 $\mu g/L$, range = [-2.7570; -0.0045]). 
+Nevertheless, it must be noted that the dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013, Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
 

### 2016-11-13 src/do_pollution.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
+       e.g. prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -78,6 +78,36 @@ meas <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 # meas <- meas[variable_id %in% meas[ , list(n = length(value)), 
 #                                   by = variable_id][n > 1000, variable_id]]
 # 
+
+
+# restrict to insecticides compunds as Stehle & Schulz
+rsvar <- c(290, # only alpha endosulfan, rs did no differenciate
+           146, 213, 257, 397, 487, 488, 192, 1216, 167, 650,
+           # beta cyfluthrin no in our datababse
+           237, 872, 648, 645,
+           # fenvalerate not in db
+           905, 1195, 728, 365, 588, 589
+           )
+# compounds in common
+com <- psm_variables[variable_id %in% rsvar & !is.na(rak_uba), list(variable_id, name, rak_uba)]
+# add rac_sw from stehle
+com$rs <- c(0.005, 0.1, 0.025, 0.3, 1.57, 2.8, 0.5)
+diff <- com$rak_uba - com$rs
+mean(diff)
+range(diff)
+
+# restrict to only insecticides
+insvar <- psm_variables[pgroup == 'organics, psm, insecticide', variable_id]
+psm_variables[variable_id %in% insvar]
+# meas_rvar <- meas[variable_id %in% rsvar]
+meas_ins <- meas[variable_id %in% insvar]
+
+# % of exceedances as proportion > loq
+# nrow(meas_rvar[rq > 1]) / nrow(meas_rvar[rq > 0])
+nrow(meas_ins[rq > 0])
+length(unique(meas_ins$variable_id))
+nrow(meas_ins[rq > 1]) / nrow(meas_ins[rq > 0]) * 100
+
 
 
 # calculate max per variable and take only first 15 hits

### 2016-11-13 article.tex
revised ms after comments of bk

@@
 The data set comprised of 1,766,104 measurements of 478 pesticides (including metabolites) related to 24,743 samples from 2,301 sampling sites. 
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modeling techniques that explicitly consider the limit of quantification.
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modelling techniques that explicitly consider the limit of quantification.
 Agricultural land use lead to a 3.7-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 28 percent. 
-Precipitation increased measured pesticide risk by 5\% and risk was the highest during summer months. 
+Precipitation increased pesticide risk by 36\% and risk was the highest during summer months. 
 Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
@@
 
-% Total words: 198
+% Total words: 197
 \end{abstract}
@@
 More than 50\% of the total land area in Germany is used by agriculture \citep{statistisches_bundesamt_bodenflache_2014}.
-In the year 2014 more than 45,000 tonnes of 776 authorized plant protection products were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
+In the year 2014 more than 45,000 tonnes of 776 authorised plant protection products were sold for application on this area \citep{bundesamt_fur_verbraucherschutz_und_lebensmittelsicherheit_absatz_2015}.
 The applied pesticides may enter surface waters via spray-drift, edge-of-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
@@
 Two recent studies indicate that pesticides might threaten freshwater biodiversity in the European union.
-\citet{malaj_organic_2014} analyzed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
+\citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
 \citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
@@
 
-Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
@@
 Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant. 
-In European law, the Directive 2009/128/EC \citep{European_Union_2009} places an obligation on the EU Member States to adopt National Action Plans (NAP) for the Sustainable Use of Plant Protection Products and the German NAP also addresses the knowledge gap concerning pesticide impact on small steams, specifically including those with catchment size \textless 10km\textsuperscript{2}.
+In European law, the Directive 2009/128/EC \citep{European_Union_2009} places an obligation on the EU Member States to adopt National Action Plans (NAP) for the Sustainable Use of Plant Protection Products and the German NAP also addresses the knowledge gap concerning pesticide impact on small streams, specifically including those with catchment size \textless 10km\textsuperscript{2}.
 
@@
 We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names) for 2005 to 2015.
-We homogenized and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
+We homogenised and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
 
@@
 We characterised pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
-RACs are derived during pesticide authorization as part of the ecological risk assessment.
+RACs are derived during pesticide authorisation as part of the ecological risk assessment.
 No unacceptable ecological effects are expected if the environmental concentration remains below this concentration.
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
-The German Environment Agency (UBA) provided RACs for 107 compounds including those with highest detection rates (Supplemental Table~S2). 
+The German Environment Agency (UBA) provided RACs for 107 compounds, including those with the highest detection rates (Supplemental Table~S2). 
 Based on theses RACs, we calculated Risk Quotients (RQ):
@@
 All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
-To display differences in the spectra of analyzed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
+To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
 We determined the optimum number of clusters using the average silhouette width \citep{rousseeuw1987silhouettes}.
 
-We expected non-linear responses to agriculture and catchment size and therefore, used generalized additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
-We modeled the number of RAC exceedances (RQ \textgreater 1) at a site as:
+We expected non-linear responses to agriculture and catchment size and therefore, used generalised additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
+We modelled the number of RAC exceedances (RQ \textgreater 1) at a site as:
 
@@
 where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
-We modeled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
+We modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
 The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
 $\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
-The degree of smoothness was estimated using restricted maximum likelihood (REML) during model fitting process \citep{wood_fast_2011}.
-The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modeling the rate of exceedances.
+The degree of smoothness was estimated using restricted maximum likelihood (REML) during the model fitting process \citep{wood_fast_2011}.
+The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modelling the rate of exceedances.
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
@@
 
-To assess the influence of precipitation and seasonality we modeled the RQ of individual compounds as the response variable.
+To assess the influence of precipitation and seasonality, we modelled the RQ of individual compounds as the response variable.
 RQ and concentrations show a skewed distribution with an excess of zeros (no pesticides detected and quantified). 
-Therefore, we modeled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016} (Equation~\ref{eqn:eqn3}).
+Therefore, we modelled these as two processes (one generating values below the limit of quantification (LOQ) and one generating values above LOQ) using a Zero-Adjusted Gamma (ZAGA) distribution \cite{rigby_generalized_2005,stasinopoulos_gamlss.dist:_2016} (Equation~\ref{eqn:eqn3}).
 These two processes can be interpreted as changes in the mean value of RQ (change in $\mu$) and changes in the probability of exceeding LOQ and showing any risk (change in $\nu$).
@@
 We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 22 compounds, see Supplemental Table~S3 for a list of compounds).
-To summarise the coefficients across the 22 modeled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 22 compounds.
+To summarise the coefficients across the 22 modelled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 22 compounds.
 The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
@@
 The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
-No data was available from Brandenburg.
+No data were available from Brandenburg.
 
@@
 Most samples were taken in the months April till October, while fewer samples were taken during winter (see Supplemental Figure~S2).
-We found substantial differences in the spectra of analyzed pesticides between federal states (Figure~\ref{fig:fig2}).
+We found substantial differences in the spectra of analysed pesticides between federal states (Figure~\ref{fig:fig2}).
 The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1). 
@@
 However, it has to be noted that both states surveyed these distinct spectra in special monitoring programs from only a few sites. 
-Although there was high variation within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
+Although there was high variability within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
 4\% (=71,113) of all measurements were concentrations above LOQ.
@@
 The mean number of RAC exceedances per site increased 3.7-fold from 0.10 (no agriculture) to 0.39 (28\% agriculture within the catchment). 
-Above this threshold the exceedances leveled.
+Above this threshold the exceedances levelled.
 Above 75\% agriculture within the catchment the number of exceedances further increased, but the increase was not statistically significant (Figure~\ref{fig:fig4}, left). 
@@
 
-$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and the mean value of RQ.
-In $Q2$ an increase of precipitation before sampling ($prec_{-1}$) from 1mm to 10mm lead on average to a 5\% higher mean RQ and the probability to exceed LOQ increases 1.5-fold from 4\% to 6\% (Figure~\ref{fig:fig5}, top).
-% do_precip.R
+$prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and RQ.
+In $Q2$ an increase from 0.1~mm to 10~mm of precipitation before sampling ($prec_{-1}$) lead on average to a 36\% higher mean RQ of 0.05.
+The probability to exceed LOQ increases 1.6-fold from 8.7\% to 13.5\% (Figure~\ref{fig:fig5}, top). % do_precip.R
 Effects differed between individual compounds and are provided in the Supplemental Table~S4. 
@@
 Both increased during summer months and decreased towards the end of the year.
-There was a 2.5-folder higher probability of exceeding LOQ in $Q2$ (10\%) than in $Q1$ (4\%).
+There was a 2.5-folder higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%).
 The differences were less pronounced for the mean value of RQ and with less precision (Figure~\ref{fig:fig5}, bottom). 
@@
   \includegraphics[width=0.8\textwidth]{figure5.pdf}
-  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-anaylsis of the 24 modeled compounds.
-Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probabilty to exceeed LOQ ($\nu$).
+  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 24 modelled compounds.
+Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probability to exceed LOQ ($\nu$).
   Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
@@
 We found RAC exceedances in 25.5\% of sampling sites and RQ > 0.1 in 54\% of sites. 
-In 23\% of sites none of the chemicals, for which RAC were available, were detected (see also Supplemental Figure~S8).
+In 23\% of sites none of the chemicals, for which RACs were available, were detected (see also Supplemental Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
@@
 
-The highest RQ were observed for Chlorpyrifos (max(RQ) = 220), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
+The highest RQs were observed for Chlorpyrifos (max(RQ) = 220), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
 Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supplemental Figure~S9).
@@
 % current problems in monitoring and possible solutions
-A nation-wide assessment of pesticide pollution is hampered by inhomogeneous data across federal states:
-Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analyzed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
+A nationwide assessment of pesticide pollution is hampered by inhomogeneous data across federal states:
+Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analysed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
 Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and thus it might nonetheless represent a sample covering all types of small streams in Germany.
@@
 For these compounds a lowering of LOQ through an improvement of chemical analysis is essential for reliable assessment.
-Moreover, a nation-wide assessment would benefit from a harmonized spectrum of analysed compounds between federal states.
+Moreover, a nationwide assessment would benefit from a harmonised spectrum of analysed compounds between federal states.
 
 Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams below 10~km\textsuperscript{2} are disproportionally less sampled in current monitoring (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD.
-Clearly, there is currently a lack of knowledge on stressor effects in small streams.
+Clearly, there is currently a lack of knowledge on stressor effects on small streams.
 We analysed only data from small streams, however, for lentic small water bodies this lack might be even greater \citep{lorenz_specifics_2016}.
@@
 We found a strong influence of agriculture on the pollution of streams.
-Above 25\% agriculture within a catchment pesticides, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
+Above 25\% agriculture within a catchment, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
 To our knowledge, this is the first study investigating such thresholds of pesticide risk.
-Previous studies examined thresholds for percentage of agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
+Previous studies examined thresholds for the percentage of agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
 \citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
@@
 Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found a 5\% higher RQ if samples were taken after rainfall events. 
-Samples taken at the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
+Nevertheless, we found a 36\% higher RQ if samples were taken after rainfall events. 
+Samples taken on the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
 This could be explained by the sampling preceding the rainfall event and the delay between the start of a rain event and the peak in discharge or runoff. 
@@
 Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
-Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation, especially for small water bodies \citep{lorenz_specifics_2016}.
+Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation of risks, especially for small water bodies \citep{lorenz_specifics_2016}.
 
@@
 For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.
+Moreover, our results show that analysing absolute concentrations and probabilities of LOQ together might deliver valuable insights into risk dynamics.
 
@@
 Our results suggest that small streams are frequently exposed to ecologically relevant pesticide concentrations.
-In one quarter of small streams RACs were exceeded at least once.
+In one-quarter of small streams RACs were exceeded at least once.
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
@@
 Overall, our results suggest that neonicotinoids may currently pose a high risk to freshwater ecosystems. 
-Moreover, our results add further evidence to the growing literature on risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.
+Moreover, our results add further evidence to the growing literature on the risks arising from neonicotinoids for aquatic \citep{morrissey2015neonicotinoid} and terrestrial \citep{pisa2015effects} ecosystems.
 
 Compared to \citet{stehle_pesticide_2015} we found higher rates of RAC exceedances for insecticides.
-They found exceedances in 37.1\% of measurements \textgreater LOQ (n = 1352, 23 insecticides), whereas, we found exceedances in 67\% of measurements \textgreater LOQ (n = 1855, 22 insecticides).
+They found exceedances in 37.1\% of insecticide measurements \textgreater LOQ (n = 1352, 23 insecticides), whereas, we found exceedances in 67\% of insecticide measurements with RACs \textgreater LOQ (n = 1855, 22 insecticides).
 This could be attributed to different insecticides considered and different underlying RACs.
 Our study has only 7 insecticides with RACs in common with the insecticides investigated by \citet{stehle_pesticide_2015}.
-Moreover, all RACs were lower in our study (mean = -0.71 $\mu g/L$, range = [-2.7570; -0.0045]). 
+Moreover, all RACs were lower in our study (average = -0.71 $\mu g/L$, range = [-2.757; -0.005]). 
 Nevertheless, it must be noted that the dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013, Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
@@
 From the definition of RAC it follows that if the concentration of a compound exceeds its RAC ecological effects are expected.
-Indeed, \citet{stehle_agricultural_2015} found that biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
+Indeed, \citet{stehle_agricultural_2015} found that the biological diversity of stream invertebrates was significantly reduced by 30\% at RQ = 1.12 and by 10\% at 1/10 of RAC.
 We found RQ values greater than 1.12 in 25\% of small streams and RQ at 1/10 of RAC in 54\% of small streams. 
@@
 Monitoring data, despite the outlined limitations, provides an opportunity to study large-scale environmental occurrence patterns of pesticides.
-Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modeling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
+Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modelling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
 However, the sampling design needs to account for precipitation events to provide robust data. 
@@
 The authors thank the federal state authorities and the German Working Group on water issues of the Federal States (LAWA) for providing chemical monitoring data and the German Environment Agency (UBA) for funding a related project (FKZ 3714 67 4040 / 1).
-We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Environment Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their valuable contributions to this project.
+We thank Alexandra Müller, Wolfram König and Volker Mohaupt (German Environment Agency (UBA)), Martin Keller and Beate Bänsch-Baltruschat (German Federal Institute of Hydrology (BfG)), Matthias Liess and Kaarina Foit (Center for Environmental Research (UFZ)) for their contributions to this project.
 \end{acknowledgement}
@@
 %%% Word count
-% abstract    :  198
+% abstract    :  197
 % 4 big(600)  : 2400
 % 2 small(300):  600
-% acknowledg  :   82
+% acknowledg  :   80
 %              =====
-% subtotal      3280
-% text body   : 3800
+% subtotal      3277
+% text body   : 3927
 % ======================================
-% subtotal    : 7080
-% - captions  :  155
+% subtotal    : 7205
+% - captions  :  191
 % =========================================
-% total       : 6925
+% total       : 7013
 

### 2016-11-13 supplement/Supplemental_Materials.tex
reanalysed everthing with new/fixed racs and comments from BK
revised ms after comments of bk

@@
 \chapter{Data Cleaning}
-Before combining into a common database, more then 30 datasets have been cleaned and homogenized separately.
+Before combining into a common database, more than 30 datasets have been cleaned and homogenised separately.
 Cleaning steps comprised the following steps (Figure~\ref{fig:data_cleaning} gives a graphical overview):
@@
 	\item  Identifiers: Unique identifiers have been assigned.
-	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification were set to zero (and can be used to identity non-detects).
+	\item Units: All concentrations have been converted to $\mu g/L$. Values below limit of quantification were set to zero (and can be used to identify non-detects).
 	\item Other meta-data: meta-data has been standardised.
@@
 	\includegraphics[width = 0.8\textwidth]{data_cleaning}
-	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning, data has been stored in a relational spatial PostgreSQL database.}
+	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning, data have been stored in a relational spatial PostgreSQL database.}
 	\label{fig:data_cleaning}
@@
 	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes of complete linkage clustering of jaccard similarity of analysed compound spectra between federal states.
-Two clusters showed maximum silhouette width.}
+Two clusters showed the maximum silhouette width.}
 	\label{fig:silhouette}
@@
 	\includegraphics[width = 0.8\textwidth]{precip}
-	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at day before sampling.}
+	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at the day before sampling.}
 	\label{fig:precip}
@@
 %%% from do_precip.R
+\newpage
 \input{keeptab}
+\newpage
 \input{keeptab2}
@@
 	\caption[Graphical representation of coefficients from table \ref{tab:var_model_coef}.]{Graphical representation of coefficients from table \ref{tab:var_model_coef}.
-Top row: Effect of precipitation at day before sampling and at day of sampling.
+Top row: Effect of precipitation at the day before sampling and at day of sampling.
 Bottom row: estimates for the four Quarters.

### 2016-11-13 src/do_pollution.R

@@ -80,6 +80,8 @@ meas <- psm_variables[ , list(variable_id, name, psm_type)][take_rac]
 # 
 
 
+
+# Compare with Stehle & Schulz ---------------------------------------
 # restrict to insecticides compunds as Stehle & Schulz
 rsvar <- c(290, # only alpha endosulfan, rs did no differenciate
            146, 213, 257, 397, 487, 488, 192, 1216, 167, 650,
@@ -92,6 +94,7 @@ rsvar <- c(290, # only alpha endosulfan, rs did no differenciate
 com <- psm_variables[variable_id %in% rsvar & !is.na(rak_uba), list(variable_id, name, rak_uba)]
 # add rac_sw from stehle
 com$rs <- c(0.005, 0.1, 0.025, 0.3, 1.57, 2.8, 0.5)
+# differences in RAC between studies
 diff <- com$rak_uba - com$rs
 mean(diff)
 range(diff)
@@ -110,7 +113,9 @@ nrow(meas_ins[rq > 1]) / nrow(meas_ins[rq > 0]) * 100
 
 
 
-# calculate max per variable and take only first 15 hits
+
+# Maximum RQ --------------------------------------------------------------
+# calculate maxRQ per variable and take only first 15 hits
 take_var <- meas[ , list(media = max(rq)) , 
                   by = variable_id][order(media, 
                                           decreasing = TRUE)][1:15, variable_id]
@@ -121,7 +126,7 @@ take_meas[ , list(media = max(rq)) , by = name][order(media, decreasing = TRUE)]
 
 
 
-## % of samples
+## Exceedance as % of samples
 # for each sample, take maximum RQ (from different compounds) and check exceedance
 nrow(meas[ ,list(rq_max = max(rq)) , 
            by = sample_id][rq_max > 1]) /  length(unique(meas$sample_id)) * 100
@@ -131,13 +136,12 @@ nrow(meas[ ,list(rq_max = max(rq)) ,
 # in 14\% of samples with detecs was was RQ > 1 found
 
 
-## % of measurements
+## Exceedances as % of measurements
 mean(meas[ , rq > 1])*100 # 0.2 % of all measurements
 nrow(meas[rq > 0])
 mean(meas[rq > 0 , rq > 1]) * 100 # 5% of all detects
 
-# number of detects
-
+# % of detects
 nrow(meas[rq > 1]) / nrow(meas[rq > 0]) * 100
 
 
[…]

### 2016-11-13 src/do_precip.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
+       e.g. prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -110,6 +110,7 @@ hist(props[tot < 1000, tot])
 
 
 
+#  ------------------------------------------------------------------------
 # export overview for supplement
 keep
 setkey(psm_variables, variable_id)
@@ -150,7 +151,7 @@ rm(samples_rac, psm_sites, psm_sites_info, take_si, rac, keep_tab, keep_tab_x)
 
 
 # Model -------------------------------------------------------------------
-# one single compouns
+# single compouds (for testing)
 # take_c <- take[variable_id == 191]
 take_c <- take[variable_id == 378]
 
@@ -179,6 +180,14 @@ term.plot(mod_l_m)
 term.plot(mod_l_m, what = 'nu')
 res <- residuals(mod_l_m)
 hist(res)
+
+foo <- function(x) log10(x + 0.05) # transformation
+newdata = data.frame(log_precip0 = foo(0.1),
+                     log_precip_1 = foo(0.1),
+                     season = 'Q2')
+predict(mod_l_m,  what = 'nu',  type = 'response')
+
+
 
 # plot marginal data
 plot(log(rq) ~ log_precip_1 , data = take_c[take_c$rq > 0, ], pch = 16, col = rgb(0, 0, 0, 0.2))
@@ -225,6 +234,8 @@ if (run_model) {
   file.remove(file.path(cachedir, 'lmodels', paste0('mod_', keep$variable_id, '.rds')))
   lapply(keep$variable_id, model_foo)
 }
+
+
 
 
 # function to extract the needed model components,
@@ -279,6 +290,8 @@ resdf$cisig <- ifelse(sign(resdf$upci)  == sign(resdf$lowci),
 resdf$termind <- substr(resdf$term, 1, 2)
 
 
+
[…]

### 2016-11-14 article.tex
final fixes before submission

@@
 Both increased during summer months and decreased towards the end of the year.
-There was a 2.5-folder higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%).
+There was a 2.5-fold higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%).
 The differences were less pronounced for the mean value of RQ and with less precision (Figure~\ref{fig:fig5}, bottom). 
@@
   \includegraphics[width=0.8\textwidth]{figure5.pdf}
-  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 24 modelled compounds.
-Left column: coefficients for mean RQ ($\mu$), right column: coefficients for probability to exceed LOQ ($\nu$).
+  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 22 modelled compounds.
+Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$).
   Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
@@
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for the absolute value of RQ. 
-This could be attributed to a higher variability of absolute concentrations.
+This may be explained by a higher variability of absolute concentrations.
 Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
@@
 \citet{stehle_pesticide_2015} found the highest percentage of RAC exceedances for organophosphate insecticides. 
-By contrast, we found that that neonicotinoid insecticides have highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
+By contrast, we found that neonicotinoid insecticides have highest exceedances of RACs, followed by the organophosphate chlorpyrifos. 
 This difference can be attributed to the low sample size for neonicotinoid insecticides in their study (n = 33) compared to the dataset presented here (for example 3,540 samples of Thiacloprid, Figure~\ref{fig:fig6}).
@@
 Our study has only 7 insecticides with RACs in common with the insecticides investigated by \citet{stehle_pesticide_2015}.
-Moreover, all RACs were lower in our study (average = -0.71 $\mu g/L$, range = [-2.757; -0.005]). 
+Moreover, all RACs were lower in our study (average difference = -0.71 $\mu g/L$, range = [-2.757; -0.005]). 
 Nevertheless, it must be noted that the dataset compiled here comprised only samples from grab sampling, which may considerably underestimate pesticide exposure \citep{stehle_probabilistic_2013, Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.

### 2016-11-15 src/do_overview.R
many changes for phd thesis

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
+       e.g. prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -107,9 +107,26 @@ yearlymonthly <- psm_samples[ , length(unique(sample_id)), by = list(year(date),
 p_temp <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) + 
   geom_bar(stat = 'identity') +
   facet_grid(~year) +
-  labs(x = 'Month', y = 'No. samples')
+  labs(x = 'Month', y = 'No. samples') 
 # p
 ggsave(file.path(prj, 'supplement/temporal.pdf'), p_temp, width = 20, height = 7)
+
+# # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/appendix/smallstreams/one/'
+# p_temp <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) + 
+#   geom_bar(stat = 'identity') +
+#   facet_grid(~year) +
+#   labs(x = 'Month', y = 'No. samples') +
+#   phdtheme +
+#   theme(panel.grid.major = element_blank()) +
+#   scale_x_discrete(labels = c('2' = '',
+#                               '3' = '',
+#                               '5' = '',
+#                               '6' = '',
+#                               '8' = '',
+#                               '9' = '',
+#                               '11' = '',
+#                               '12' = ''))
+# ggsave(file.path(phd, 'temporal.pdf'), p_temp, width = 250, units = 'mm')
 
 
 
@@ -142,7 +159,9 @@ p_map <- ggplot() +
 ggsave(file.path(prj, "figure1.pdf"),
        p_map, width = 3.5, height = 3, units = 'in', dpi = 300, scale = 2)
 
-
[…]

### 2016-11-15 src/do_pollution.R

@@ -193,6 +193,27 @@ pdf(file.path(prj, "supplement/prac_ex.pdf"))
   abline(h = prec[2], lty = 'dotted')
 dev.off()
 
+# phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/appendix/smallstreams/one/'
+# pdf(file.path(phd, "prac_ex.pdf"))
+#   par(mar = c(5,5,4,2))
+#   plot(rqs[3:length(rqs)], prec[3:length(prec)],
+#        log = 'x',
+#        pch = 16,
+#        ylim = c(0, 100),
+#        cex = 1.2,
+#        ylab = 'Fraction of sites',
+#        xlab = 'max(RQ)',
+#        xaxt = 'n', cex.lab = 1.5, cex.axis = 1.5)
+#   axis(side = 1, at = c(0.001, 0.01, 0.1, 1,  10, 100),
+#        labels = c("0.001", "0.01", "0.1", "1", "10", "100"),
+#          cex.axis = 1.5)
+#   abline(v = 1, lty = 'dotted')
+#   abline(v = 0.1, lty = 'dotted')
+#   abline(h = prec[1], lty = 'dotted')
+#   abline(h = prec[2], lty = 'dotted')
+# dev.off()
+
+
 
 # number of SWB sites with RAC 
 length(unique(meas$site_id))   # 2270 (from 2301 in total)
@@ -254,6 +275,9 @@ levs <- levels(reorder(take_meas[rq > 0, name], take_meas[rq > 0, rq], median))
 loqd[ , name := factor(name, levels = levs[levs %in% loqd$name])]
 loqd
 
+
+# Violin Plot of RQ -------------------------------------------------------
+
 prac <- ggplot() +
   geom_violin(data = take_meas[rq > 0 & name %in% loqd$name],
               aes(x = reorder(name, rq, FUN = median), y = rq, fill = psm_type)) +
@@ -274,6 +298,8 @@ prac <- ggplot() +
 # prac
 ggsave(file.path(prj, "figure6.pdf"), prac, width = 7, height = 6.5)
 
+# phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/chapters/smallstreams/'
+# ggsave(file.path(phd, "figure6.pdf"), prac, width = 7, height = 6.5)
 
 # RQ exceedances for other compounds
[…]

### 2016-11-15 src/functions.R

@@ -365,6 +365,18 @@ mytheme <- theme_bw(base_size = 12, base_family = "Helvetica") +
         strip.background = element_blank(),
         strip.text = element_text(size = 14, face = 'bold'))
 
+phdtheme <- theme_bw(base_size = 12, base_family = "Helvetica") + 
+  theme(
+    # panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(), 
+    text = element_text(size = 14),
+    axis.text = element_text(size = 12),
+    axis.title.x = element_text(size = 14, face = "bold", vjust = 0),
+    axis.title.y = element_text(size = 14, face = "bold", vjust = 1),
+    legend.key = element_blank(),
+    strip.background = element_blank(),
+    strip.text = element_text(size = 14, face = 'bold'))
+
 theme_end <- theme_bw(base_size = 12, base_family = "Helvetica") + 
   theme(panel.grid.minor = element_blank(), 
         axis.title.x = element_text(face = "bold", vjust = 0),

### 2016-11-15 src/load.R

@@ -52,6 +52,7 @@ library(ggmap)
 
 # plotting
 library(ggplot2)
+library(scales)
 library(gridExtra)
 library(ggExtra)
 library(cowplot)

### 2016-12-28 supplement/Supplemental_Materials.tex
final fixes before submission
revised technical notes

@@
-\documentclass[pdftex,a4paper]{scrreprt}
+\documentclass[pdftex,
+	a4paper,
+	titlepage=false]{scrreprt}
 \usepackage[utf8]{inputenc}       
@@
 \renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thepage}{S\arabic{page}}
 
@@
 
+% affiliations
+\usepackage[affil-it]{authblk}
+
+% counts
+% http://ctan.mackichan.com/macros/latex/contrib/caption/totalcount.pdf
+\usepackage[figure,table, page]{totalcount}
+
 %% -------------------------------------------------------------------------
-\title{Supplemental Materials for the paper: \\ Pesticide exposure in small streams in Germany}
-\author{Eduard Szöcs, Marvin Brinke, Bilgin Karaoglan, Ralf B. Schäfer}
-\date{}
+\title{Supplemental Materials for: \\ Large scale risks from pesticides in small streams}
+\author{Eduard Szöcs%
+  \thanks{Corresponding author, Email: \texttt{szoecs@uni-landau.de}}}
+\affil[1]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}
+
+\author{Marvin Brinke}
+\affil{German Federal Institute of Hydrology (BfG), Koblenz, Germany}
+
+\author{Bilgin Karaoglan}
+\affil{German Environment Agency (UBA), Dessau-Roßlau, Germany}
+
+\author[1]{Ralf B. Schäfer}
 
@@
 \maketitle
+
+\vfill
+\begin{description}
+	\item[Number of pages: ] \totalpages
+	\item[Number of figures: ] \totalfigures
+	\item[Number of tables: ] \totaltables
+\end{description}
+% increase page counter by 1
+\setcounter{page}{2}
 

### 2017-01-17 src/do_overview.R
fix 2d sitribution + map crs; new precipitation plot;

@@ -152,16 +152,24 @@ p_map <- ggplot() +
   theme(legend.key = element_rect(fill = 'white')) +
   # guides(colour = FALSE) +
   labs(x = 'Lon.', y = 'Lat.') +
-  scale_color_hue(name = 'state', l = 50) +
+  scale_color_hue(name = 'State', l = 50) +
   mytheme +
-  coord_equal()
-# p_map
+  coord_map()
+p_map
 ggsave(file.path(prj, "figure1.pdf"),
        p_map, width = 3.5, height = 3, units = 'in', dpi = 300, scale = 2)
 
+# for thesis
 # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/chapters/smallstreams/'
 # ggsave(file.path(phd, "figure1.pdf"),
 #        p_map, width = 176, height = 150, units = 'mm')
+#        
+# for defense
+# def <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/'
+# p <- p_map +
+#   theme(legend.position = 'none')
+# ggsave(file.path(def, "map_phch.pdf"),
+#        p, width = 176, height = 150, units = 'mm')
 
 
 # Tabular overview --------------------------------------------------------
@@ -467,16 +475,19 @@ options(stringsAsFactors = TRUE) # to fix bug in stat_density2d with polygons
 ezg_lu <- ggplot(psm_sites_info[ezg_fin < 150 & !is.na(agri_fin) & !is.na(ezg_fin)], 
                  aes(x = ezg_fin, y = agri_fin * 100)) +
   stat_density2d(aes(alpha = ..level.., fill = ..level..), geom = "polygon") +
+  scale_x_continuous(limits = c(-0, 100), # extend to plot full polygons
+                     breaks = c(0, 10, 25, 50, 100)) + 
+  scale_y_continuous(limits = c(-10, 100)) +
+  scale_fill_viridis() +
   geom_point(size = 0.5) +
   guides(alpha = FALSE, fill = FALSE) +
   mytheme +
   # phdtheme + 
-  labs(x = expression('Catchment area ['~km^2~']'), y = expression('Agriculture [%]')) +
-  # scale_fill_gradient(low = "yellow", high = "red") +
-  scale_fill_viridis() +
-  scale_x_continuous(breaks = c(0, 10, 25, 50, 100))
[…]

### 2017-01-17 src/do_precip.R

@@ -482,10 +482,14 @@ foo <- function(x) log10(x + 0.05) # transformation
 # log_precip_1: nu 
 # intercept
 int <- resmd$est[resmd$term ==  'nu.seasonQ2']
+plogis(int) *100
+
 # prop of exceeding loq in q2 at 0.1 mm
 (zerop1 <- plogis(int + foo(0.1)*resmd$est[resmd$term ==  'nu.log_precip_1'])) * 100
+(one <- plogis(int + foo(1)*resmd$est[resmd$term ==  'nu.log_precip_1'])) * 100
 # prop of exceeding loq in q2 at 10 mm
-(ten <- plogis(int + foo(10)*resmd$est[resmd$term ==  'nu.log_precip_1'])) * 100
+(ten <- plogis(int + foo(15)*resmd$est[resmd$term ==  'nu.log_precip_1'])) * 100
+
 (ten - zerop1) 
 (ten - zerop1) / zerop1 * 100
 
@@ -501,6 +505,7 @@ range(tena)
 plogis(resmd$est[resmd$term ==  'nu.seasonQ1'])
 # Q2 nu
 plogis(resmd$est[resmd$term ==  'nu.seasonQ2'])
+
 
 
 
@@ -570,8 +575,47 @@ p_precip <- ggplot(resmd[resmd$coeftype == 'precip', ]) +
   theme(strip.text.x = element_text(size = 22))
 
 p <- arrangeGrob(p_precip, p_season, ncol = 1, heights = c(1, 1.25))
+plot(p)
 # plot(p)
 ggsave(file.path(prj, "figure5.pdf"), p, width = 7, height = 5.5)
 
 # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/chapters/smallstreams/'
 # ggsave(file.path(phd, "figure5.pdf"), p, width = 7, height = 5.5)
+
+
+
+# plot for defense
+# plot data
+pdata <- resmd[resmd$coeftype == 'season' & resmd$type == 'nu~(LOQ)', ]
+# calculate for high precipitation
+int <- pdata$est
+pdata$high <- plogis(int + foo(15) * resmd[resmd$term == 'nu.log_precip_1' & resmd$type == 'nu~(LOQ)', ]$est) * 100
+
+
+
+library(tikzDevice)
+p <- ggplot() +
+  geom_pointrange(data = pdata,
+                  aes(x = substr(term, nchar(term) - 1, nchar(term)),
[…]

### 2017-02-01 src/do_overview.R
additions for defense

@@ -475,22 +475,24 @@ options(stringsAsFactors = TRUE) # to fix bug in stat_density2d with polygons
 ezg_lu <- ggplot(psm_sites_info[ezg_fin < 150 & !is.na(agri_fin) & !is.na(ezg_fin)], 
                  aes(x = ezg_fin, y = agri_fin * 100)) +
   stat_density2d(aes(alpha = ..level.., fill = ..level..), geom = "polygon") +
-  scale_x_continuous(limits = c(-0, 100), # extend to plot full polygons
-                     breaks = c(0, 10, 25, 50, 100)) + 
-  scale_y_continuous(limits = c(-10, 100)) +
-  scale_fill_viridis() +
   geom_point(size = 0.5) +
+  scale_x_continuous(limits = c(-0, 100), # extend to plot full polygons
+                     breaks = c(0, 10, 25, 50, 100)) +
+  scale_fill_viridis() +
+  scale_alpha_continuous(limits = c(0.00015, 0.0004), na.value = 0, range = c(0.3, 1)) +
   guides(alpha = FALSE, fill = FALSE) +
   mytheme +
-  # phdtheme + 
-  labs(x = expression('Catchment area ['~km^2~']'), y = expression('Agriculture [%]')) 
-  # scale_fill_gradient(low = "yellow", high = "red") 
-# ezg_lu
-ezg_lu <- ggMarginal(ezg_lu, type = 'histogram', binwidth = 5)
+  labs(x = expression('Catchment area ['~km^2~']'), y = expression('Agriculture [%]')) +
+  coord_cartesian(ylim = c(0, 100), xlim = c(0, 100))
 ezg_lu
+
+# add marginal distr
+ezg_p <- ggMarginal(ezg_lu, type = 'histogram', binwidth = 5)
+ezg_p
+
 # ezg_lu
 ggsave(file.path(prj, 'figure3.pdf'), 
-       ezg_lu,
+       ezg_p,
        width = 3.3, height = 3.3/1.2,
        units = 'in', dpi = 300, scale = 2.5)
 
@@ -502,6 +504,15 @@ ggsave(file.path(prj, 'figure3.pdf'),
 options(stringsAsFactors = FALSE)
 
 
+
+# # for defense
+# def <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/'
+# p <- ezg_lu + 
+#   theme_edi(base_size = 16)
+# p <- ggMarginal(p, type = 'histogram', binwidth = 5)
+# p
[…]

### 2017-02-01 src/do_precip.R

@@ -143,6 +143,10 @@ print(keep_tab_x,
 
 
 
+
+
+
+# subset
 take <- take[variable_id %in% keep$variable_id]
 psm_variables[variable_id %in% keep$variable_id, list(variable_id, name, 
                                                       psm_type)]
@@ -164,36 +168,36 @@ glco <- gamlss.control(c.crit = 0.01, mu.step = 0.5)
 
 
 
-# gamma hurdle model with precipitation as groups, season and psm_type as preds
-# state as random effect
-mod_l_m <- gamlss(rq ~ 0  + log_precip_1 + log_precip0 + season +
-                    re(random = ~1|state_fac/s_id_fac),
-                  nu.formula = ~ 0 + precip_1 + precip0 + season + 
-                    re(random = ~1|state_fac/s_id_fac),
-                  data = take_c,
-                  family = ZAGA,
-                  control = glco,
-                  i.control = gco)
-plot(mod_l_m)
-summary(mod_l_m)
-term.plot(mod_l_m)
-term.plot(mod_l_m, what = 'nu')
-res <- residuals(mod_l_m)
-hist(res)
-
-foo <- function(x) log10(x + 0.05) # transformation
-newdata = data.frame(log_precip0 = foo(0.1),
-                     log_precip_1 = foo(0.1),
-                     season = 'Q2')
-predict(mod_l_m,  what = 'nu',  type = 'response')
-
-
-
-# plot marginal data
-plot(log(rq) ~ log_precip_1 , data = take_c[take_c$rq > 0, ], pch = 16, col = rgb(0, 0, 0, 0.2))
-plot(I(ifelse(rq > 0, 1, 0)) ~ log_precip_1 , data = take_c, 
-     col = rgb(0, 0, 0, 0.1), pch = 16)
-plot(log(rq) ~ season , data = take_c[take_c$rq > 0, ])
+# # gamma hurdle model with precipitation as groups, season and psm_type as preds
+# # state as random effect
+# mod_l_m <- gamlss(rq ~ 0  + log_precip_1 + log_precip0 + season +
+#                     re(random = ~1|state_fac/s_id_fac),
+#                   nu.formula = ~ 0 + precip_1 + precip0 + season + 
+#                     re(random = ~1|state_fac/s_id_fac),
+#                   data = take_c,
[…]

### 2017-02-01 article.tex
new count
many changes for phd thesis
revised technical notes
fix 2d sitribution + map crs; new precipitation plot;
started with revisions

@@
 \usepackage{enumerate}
+\pdfminorversion=4
+\usepackage{lineno}
+\usepackage{todonotes}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\listfiles
+\listfiles
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{tocentry}
-
-\includegraphics[width=0.7\textwidth]{abstract.pdf}
-
-\end{tocentry}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\linenumbers
+
 \begin{abstract}
@@
 Small streams are important refugia for biodiversity.
-In agricultural areas they may be at high risk from pesticide pollution. 
+In agricultural areas they may be at risk from pesticide pollution. 
 However, most related studies have been limited to a few streams on the regional level, hampering extrapolation to larger scales. 
-We used data from German governmental water quality monitoring to quantify the drivers of pesticide risk and to assess pesticide risk in small streams on a large scale. 
+We quantified risks as exceedances of regulatory acceptable concentrations (RACs) and used German monitoring data to quantify the drivers thereof and to assess current risks in small streams on a large scale.
 The data set comprised of 1,766,104 measurements of 478 pesticides (including metabolites) related to 24,743 samples from 2,301 sampling sites. 
-We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk using new statistical modelling techniques that explicitly consider the limit of quantification.
-Agricultural land use lead to a 3.7-fold increase in exceedance of risk thresholds when the proportion of agriculture in a catchment exceeded 28 percent. 
+We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk taking also concentrations below the limit of quantification into account.
+Agricultural land use lead to a 3.7-fold increase in exceedance of risk thresholds when the proportion of agricultural land use in a catchment exceeded 28 percent. 
 Precipitation increased pesticide risk by 36\% and risk was the highest during summer months. 
-Risk thresholds were exceeded in 26 \% of streams, with the highest risk related to neonicotinoid insecticides. 
-We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity and that a realistic pesticide sampling would be driven by precipitation events.
+RACs were exceeded in 26\% of streams.
+We found the highest exceedances for neonicotinoid insecticides. 
+We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity. 
+To reflect peak concentrations, current pesticide monitoring needs to be refined.
 
@@
 \end{abstract}
+
+% \begin{tocentry}
+
+\section{TOC Art}
+
+\includegraphics[width=3.33in, height=1.875in]{abstract.pdf}
+
+% \end{tocentry}
 
@@
 
-Two recent studies indicate that pesticides might threaten freshwater biodiversity in the European union.
+Two recent studies indicate that pesticides concentrations in streams might threaten freshwater biodiversity in the European union.
 \citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
-\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications. 
-They found that many of these measurements exceed regulatory acceptable concentrations (RAC).
+\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications and considerable exceedances of regulatory acceptable concentrations (RAC).
 However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
@@
 Much more comprehensive data on thousands of sites are available from national monitoring programs that are setup for the surveillance of water quality,
-which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008} and additional state-specific needs. 
-Despite that these data are providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled and related analyses are lacking.
+which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008}. 
+Despite that these data are providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled.
 
-Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger freshwater systems \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to bigger streams \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
 For example, the WFD defines small streams with a catchment size between 10 and 100 km\textsuperscript{2}, without further categorisation of streams \textless 10km\textsuperscript{2} and \citet{lorenz_specifics_2016} defines small streams with catchment size \textless 10km\textsuperscript{2}.
-Moreover, small streams might particularly be at high risk of pesticide contamination in case of adjacent agricultural areas given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
+Moreover, small streams might particularly be at high risk of pesticide contamination in case of adjacent agricultural areas and given their low dilution potential \citep{schulz_field_2004,liess_determination_1999}.
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
@@
 Additionally, we delineated upstream catchments for each of the sampling sites using (i) a digital elevation model (DEM) \citep{eea_digital_2013} and the multiple flow direction algorithm \citep{holmgren_multiple_1994} as implemented in GRASS GIS 7 \citep{neteler_grass_2012} and (ii) from drainage basins provided by the Federal Institute of Hydrology (BfG).
-Delineated catchments were visually checked for accuracy by comparison with state stream networks and derived information amalgamated with existing data.
+Delineated catchments were visually checked for accuracy by comparison of coverage with stream networks provided by the federal states.
 Thus, catchment size information was available for 99\% of all sites (59\% from authorities, 24\% from DEM and 16\% from drainage basins).
@@
 \subsection{Statistical analyses}
-All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
-To display differences in the spectra of analysed compounds between federal states we used Multidimensional Scaling (MDS) based on Jaccard dissimilarity in conjunction with complete linkage hierarchical clustering using the vegan package \citep{oksanen_vegan:_2016}.
-We determined the optimum number of clusters using the average silhouette width \citep{rousseeuw1987silhouettes}.
-
 We expected non-linear responses to agriculture and catchment size and therefore, used generalised additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
@@
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
+All data-processing and analyses were performed using R \citep{r_core_team_r:_2016}.
 GAMs were fitted using the mgcv package \citep{wood_fast_2011}.
@@
 We found substantial differences in the spectra of analysed pesticides between federal states (Figure~\ref{fig:fig2}).
-The number of different pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1). 
-Hierarchical clustering revealed that RP and NI analysed distinct compound spectra compared to the cluster of other states. 
-However, it has to be noted that both states surveyed these distinct spectra in special monitoring programs from only a few sites. 
-Although there was high variability within the remaining cluster, this could not be further split (Figure~\ref{fig:fig2}, also Supplemental Figures~S3 and S4).
+The number of analysed pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1). 
 4\% (=71,113) of all measurements were concentrations above LOQ.
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure2.pdf}
-  \caption{Compound spectra of the different federal states. Left: Barcode plot - each vertical line is an analysed compound. Right: MDS ordination. 
-  Colors according to two clusters determined by hierarchical clustering (see Supplemental Figure~S3 and S4).}
+  \includegraphics[width=0.6\textwidth]{figure2.pdf}
+  \caption{Barcode plot of compound spectra of the federal states. Each vertical line is an analysed compound.}
   \label{fig:fig2}
@@
   \includegraphics[width=.8\textwidth]{figure3.pdf}
-  \caption{Distribution of catchment area and agriculture within the catchment area across the sampling sites.
-  Colour codes the 2-dimensional density of points.}
+  \caption{Distribution of catchment area across the sampling sites.}
   \label{fig:fig3}
@@
   \includegraphics[width=0.95\textwidth]{figure4.pdf}
-  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances.
+  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the mean number of RAC exceedances per site.
 Red line marks statistically significant changes.
@@
 We found a strong influence of agriculture on the pollution of streams.
-Above 25\% agriculture within a catchment, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
+Above 28\% agriculture within a catchment, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
 To our knowledge, this is the first study investigating such thresholds of pesticide risk.
-Previous studies examined thresholds for the percentage of agricultural land use with respect to the response of biological communities, integrating different agricultural stressors.
-\citet{feld_response_2013} found change points of biological community metrics at 40\% agricultural land use in lowland streams in Europe.
-Similarly, \citet{waite_agricultural_2014} found a threshold for aquatic diatoms at 40\% agricultural land use in wadeable streams in the United States.
-Our results coincide with these thresholds and suggest that pesticides might contribute to the observed biological changes.
+\todo{more dicussion here?;drop shows the agri is important; compare to linear fit}
 
@@
 % 2 small(300):  600
-% acknowledg  :   80
+% acknowledg  :   79
 %              =====
-% subtotal      3277
-% text body   : 3927
+% subtotal      3276
+% text body   : 3846
 % ======================================
-% subtotal    : 7205
-% - captions  :  191
+% subtotal    : 7122
+% - captions  :  189
 % =========================================
-% total       : 7013
+% total       : 6933
 

### 2017-02-01 supplement/Supplemental_Materials.tex
started with revisions

@@
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width = 0.8\textwidth]{varclus}
-	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.
+% \begin{figure}[ht]
+% 	\centering
+% 	\includegraphics[width = 0.8\textwidth]{varclus}
+% 	\caption[Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.]{Complete Linkage Cluster Dendrogram of Jaccard Similarity of analysed compound spectra between federal states.
 Abbreviations of state names according to ISO 3166-2:DE (see also Table~\ref{tab:phch_overview}).}
-	\label{fig:varclus}
-\end{figure}
+% 	\label{fig:varclus}
+% \end{figure}
 
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width = 0.6\textwidth]{silhouette}
-	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes of complete linkage clustering of jaccard similarity of analysed compound spectra between federal states.
+% \begin{figure}[ht]
+% 	\centering
+% 	\includegraphics[width = 0.6\textwidth]{silhouette}
+% 	\caption[Average silhouette width for different cluster sizes.]{Average silhouette width for different cluster sizes of complete linkage clustering of jaccard similarity of analysed compound spectra between federal states.
 Two clusters showed the maximum silhouette width.}
-	\label{fig:silhouette}
-\end{figure}
+% 	\label{fig:silhouette}
+% \end{figure}
 

### 2017-02-01 src/do_overview.R

@@ -108,7 +108,7 @@ p_temp <-  ggplot(yearlymonthly, aes(x = factor(month), y = V1)) +
   geom_bar(stat = 'identity') +
   facet_grid(~year) +
   labs(x = 'Month', y = 'No. samples') 
-# p
+p_temp
 ggsave(file.path(prj, 'supplement/temporal.pdf'), p_temp, width = 20, height = 7)
 
 # # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/appendix/smallstreams/one/'
@@ -157,7 +157,7 @@ p_map <- ggplot() +
   coord_map()
 p_map
 ggsave(file.path(prj, "figure1.pdf"),
-       p_map, width = 3.5, height = 3, units = 'in', dpi = 300, scale = 2)
+       p_map, width = 4, height = 5, units = 'in', dpi = 300, scale = 1.5)
 
 # for thesis
 # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/chapters/smallstreams/'
@@ -323,37 +323,40 @@ makeone <-  function(DT) {
 }
 makeone(vw)
 
+
 # calculte jaccard distance between BL
 rownames(vw) <- vw[ ,site_id]
-dp <- vegdist(vw[ , -1, with = FALSE], method = 'jaccard')
+# dp <- vegdist(vw[ , -1, with = FALSE], method = 'jaccard')
 
-# hierarchical clustering
-hc <- hclust(dp, method = 'complete')
-plot(hc, labels = vw$site_id)
 
-# # fusion level plot
-plot(hc$height, nrow(vw[ , -1, with = FALSE]):2, type = 'S')
-text(hc$height, nrow(vw[ , -1, with = FALSE]):2, nrow(vw[ , -1, with = FALSE]):2, col = 'red')
-# not easy to spot
 
-#  #average silhoutte with to choose be number of clusters
-nr <- nrow(vw[ , -1, with = FALSE])
-
-# number of clusters to check
-ks <- 2:(nr - 1)
-asw <- numeric(length(ks))
-for (i in seq_along(ks)) {
-  sil <- silhouette(cutree(hc, k = ks[i]), dp)
-  asw[i] <- summary(sil)$avg.width
-}
-ks[which.max(asw)]
-
-pdf(file = file.path(prj, 'supplement/silhouette.pdf'))
[…]

### 2017-02-01 src/do_sizeagri.R

@@ -1,6 +1,6 @@
 if (!exists("prj")) {
   stop("You need to create a object 'prj' that points to the top folder, 
-       e.g. prj <- 'prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
+       e.g. prj <- '/home/edisz/Documents/work/research/projects/2016/4BFG/Paper/ms_est' or
        prj <- '/home/user/Documents/projects_git/ms_est'!")
 } else {
   source(file.path(prj, "src", "load.R"))
@@ -96,7 +96,7 @@ p_raw <- ggplot(rak_exceed, aes(y = agri_fin * 100, x = ezg_fin, col = log(n_exc
   scale_color_gradient(name = 'log n(RQ>1)', low = 'blue', high = 'red') +
   ylab("Agriculture [%]") +
   xlab("Catchment Size [km2]") 
-# p_raw
+p_raw
 ggsave(file.path(prj, "supplement/ezgagrirac.pdf"),
        p_raw)
 
@@ -225,9 +225,9 @@ p <- ggplot() +
   geom_rug(data = odat, aes(x = value), alpha = 0.05) +
   mytheme +
   xlab('') +
-  ylab('No. RAC exceedances') +
+  ylab('Mean no. RAC exceedances') +
   ylim(c(0, 1.3))
-# p
+p
 ggsave(file.path(prj, "figure4.pdf"),
        p, width = 7, height = 7/1.6,
        units = 'in', dpi = 300, scale = 1)

### 2017-02-02 article.tex
More revisions

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[Pesticides small streams]{Large scale risks from pesticides in small streams}
+\title[Pesticides small streams]{Large scale risks from agricultural pesticides in small streams}
 % \abbreviations{mo, neon, ra, tu, fw}
@@
 Indeed, meta-analyses using data from studies with a few sites reported higher pesticide pollution in smaller streams compared to bigger streams \citep{stehle_pesticide_2015,schulz_field_2004}. 
-Despite their ecological relevance and potentially higher pesticide exposure, a recent analysis of pesticide studies showed that a disproportionally small fraction of studies was conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
+Despite their ecological relevance and potentially higher pesticide exposure, a recent review of pesticide studies showed that a disproportionally small fraction of studies was conducted in small water bodies, and these were largely limited to a few sites \citep{lorenz_specifics_2016}.
 Consequently, knowledge on the pesticide pollution of small streams on larger scales is scant. 
@@
 
-In this study, we compiled and analysed large-scale chemical monitoring data from small streams in Germany. 
-First, we analysed the shape of the relationship between pesticide risk, agricultural land use, and catchment size and examined whether related thresholds for pesticide risks can be derived. 
-Second, we investigated the influence of precipitation and seasonal dynamics on pesticide detections, given that precipitation proved an important driver of pesticide exposure in several small-scale studies \citep{wittmer_significance_2010,schulz_field_2004}, but it is unknown whether a precipitation signal prevails on large scales.
+In this study, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany in order to identify drivers and dynamics of pesticide concentrations.
+We expect that the landscape is a determinant of measured pesticide concentrations.
+Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with high proportion of agriculture. 
+Moreover, if agriculture is a main source for pesticides in streams, we expect that concentrations drop to zero if there is not agriculture in the catchment. 
+Moreover, these relationships may show thresholds that could be to define reference streams without pollution.
+Given their low dilution potential we expected that small streams show highest concentrations. 
+However, also the timing of sampling may influence measured concentrations:
+A sampling directly after a precipitation might show higher concentrations because of run-off.
+Furthermore, pesticides are not applied throughout the whole year and we expected highest concentrations during the main growing season.
 Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.
@@
 where $No(RQ > 1)_i$ is the observed number of RAC exceedances at site $i$. 
-We modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
+Because of overdispersion, we modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
 The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 

### 2017-02-02 src/do_sizeagri.R

@@ -113,6 +113,10 @@ plot(mod_p, pages = 1)
 plot(mod_p, pages = 1, residuals = TRUE) 
 gam.check(mod_p)
 
+# ggplot(rak_exceed, aes(x = agri_fin, y = n_exceed+1)) +
+#   geom_point(alpha = 0.1) +
+#   scale_y_log10()
+
 # overdispersion
 r <- resid(mod_p, type = "pearson")
 sum(r^2) / (mod_p$df.res)
@@ -120,11 +124,49 @@ sum(r^2) / (mod_p$df.res)
 
 # try negative binomial model
 # with offset, automatic theta search and REML
-mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'cr') + s(ezg_fin, bs = 'cr') + offset(logn), 
+# increase number of knots from 9/18 to 27/54 (make no differnce)
+mod_nb <- gam(n_exceed ~ s(agri_fin, bs = 'cr', k = 18) + s(ezg_fin, bs = 'cr', k = 18) + offset(logn), 
               data = rak_exceed,
               family = nb(),
               method = 'REML')
+summary(mod_nb)
 plot(mod_nb, pages = 1)
+gam.check(mod_nb) #OK
+
+
+# linear fit for comparison
+mod_nb_lin <- gam(n_exceed ~ agri_fin + ezg_fin + offset(logn), 
+              data = rak_exceed,
+              family = nb(),
+              method = 'REML')
+mod_nb_lin
+plot(mod_nb_lin, pages = 1)
+gam.check(mod_nb_lin) 
+
+
+# check / show mean variance relationship
+cp <- rak_exceed
+# bin predictors
+cp$bins <- with(rak_exceed, 
+     paste('a_',
+     cut(agri_fin, breaks = seq(min(agri_fin) - 0.1, max(agri_fin) + 0.1, length.out = 12)),
+     ' - e_',
+     cut(ezg_fin, breaks = seq(min(ezg_fin) - 0.1, max(ezg_fin) + 0.1, length.out = 12))
+     ))
+setDT(cp)
+mv <- cp[  ,list(m = mean(n_exceed),
+           v = var(n_exceed)) , by = bins]
+length(unique(cp$bins))
+pdf(file.path(prj, "revision1/response/figs/mv.pdf"))
+  plot(v ~ m, data = mv, xlab = 'mean', ylab = 'variance'
+       )
+  # add estimate quadrativ relationship
[…]

### 2017-02-07 article.tex
more revisions

@@
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk taking also concentrations below the limit of quantification into account.
-Agricultural land use lead to a 3.7-fold increase in exceedance of risk thresholds when the proportion of agricultural land use in a catchment exceeded 28 percent. 
-Precipitation increased pesticide risk by 36\% and risk was the highest during summer months. 
+The exceedances of risk thresholds dropped 3.7-fold at sites with no agriculture, indicating that agricultural land use is a major contributor of pesticides in streams.
+Precipitation increased detection probability by 43\% and concentrations were the highest during summer months. 
 RACs were exceeded in 26\% of streams.
@@
 
-% Total words: 197
+\todo{überarbeiten}
+% Total words: 199
 \end{abstract}
@@
 
-In this study, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany in order to identify drivers and dynamics of pesticide concentrations.
+The aim of this study is to identify drivers and dynamics of pesticide concentrations in streams.
+To achieve this, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany in order to identify drivers and dynamics of pesticide concentrations. 
 We expect that the landscape is a determinant of measured pesticide concentrations.
 Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with high proportion of agriculture. 
-Moreover, if agriculture is a main source for pesticides in streams, we expect that concentrations drop to zero if there is not agriculture in the catchment. 
-Moreover, these relationships may show thresholds that could be to define reference streams without pollution.
-Given their low dilution potential we expected that small streams show highest concentrations. 
+Moreover, if agricultural land use is a main source for pesticides in streams, we expect that exceedances drop to zero if there is not agricultural land use in the catchment. 
+Moreover, these relationships may show thresholds that could be to define reference streams without pollution for a future monitoring. 
+Given their low dilution potential and direct adjacency to agricultural fields we expected that small streams show highest pesticide concentrations. 
 However, also the timing of sampling may influence measured concentrations:
@@
 \subsection{Statistical analyses}
-We expected non-linear responses to agriculture and catchment size and therefore, used generalised additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
+As outlined in the introduction, we expected non-linear responses to agricultural land use and catchment size and search for potential thresholds.
+Therefore, we used generalised additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
 We modelled the number of RAC exceedances (RQ \textgreater 1) at a site as:
@@
 Because of overdispersion, we modelled $No(RQ > 1)_i$ as resulting from a negative binomial distribution ($NB$) with mean $\mu_i$ and a quadratic mean-variance-relationship ($Var(No(RQ > 1)_i) = \mu_i + \frac{\mu_i^2}{\kappa}$).
-The proportion of agriculture within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
-$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006}. 
-The degree of smoothness was estimated using restricted maximum likelihood (REML) during the model fitting process \citep{wood_fast_2011}.
+The proportion of agricultural land use within the catchment ($agri_i$) and the catchment size of the site ($size_i$) were used as predictors of the number of RAC exceedances. 
+$\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006, wood_fast_2011}. 
 The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modelling the rate of exceedances.
@@
 \begin{figure}[ht]
-  \includegraphics[width=0.6\textwidth]{figure1.pdf}
+  \includegraphics[width=3.33in]{figure1.pdf}
   \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states (see Supplemental Table~S1 for abbreviations).}
@@
 \begin{figure}[ht]
-  \includegraphics[width=0.6\textwidth]{figure2.pdf}
+  \includegraphics[width=3.33in]{figure2.pdf}
   \caption{Barcode plot of compound spectra of the federal states. Each vertical line is an analysed compound.}
@@
 \begin{figure}[ht]
-  \includegraphics[width=.8\textwidth]{figure3.pdf}
+  \includegraphics[width=3.33in]{figure3.pdf}
   \caption{Distribution of catchment area across the sampling sites.}
@@
 \subsection{Influence of agricultural land use and catchment size}
-The number of RAC exceedances increased strongly and statistically significant up to 28\% agriculture within the catchment.
-The mean number of RAC exceedances per site increased 3.7-fold from 0.10 (no agriculture) to 0.39 (28\% agriculture within the catchment). 
-Above this threshold the exceedances levelled.
-Above 75\% agriculture within the catchment the number of exceedances further increased, but the increase was not statistically significant (Figure~\ref{fig:fig4}, left). 
+We found a positive relationship between agricultural land use and the number of RAC exceedances. 
+The non-linear model showed, that below 28\% agriculture the mean number of RAC exceedances dropped statistically significant 3.7-fold from 0.39 (28\% agriculture within the catchment) to 0.10 (no agriculture) (Figure~\ref{fig:fig4}, left).
+\todo{discussion anpassen}.
 Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
@@
 \begin{figure}[ht]
-  \includegraphics[width=0.95\textwidth]{figure4.pdf}
+  \includegraphics[width=6.5in]{figure4.pdf}
   \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the mean number of RAC exceedances per site.
@@
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and RQ.
-In $Q2$ an increase from 0.1~mm to 10~mm of precipitation before sampling ($prec_{-1}$) lead on average to a 36\% higher mean RQ of 0.05.
-The probability to exceed LOQ increases 1.6-fold from 8.7\% to 13.5\% (Figure~\ref{fig:fig5}, top). % do_precip.R
-Effects differed between individual compounds and are provided in the Supplemental Table~S4. 
+In $Q2$ an increase from 0.1~mm to 15~mm of precipitation before sampling ($prec_{-1}$) lead on average to a 43\% higher mean RQ of 0.05 (Supplemental Figure~S7).
+The probability to exceed LOQ increases in $Q2$ 1.6-fold from 8.7\% to 13.5\% (Figure~\ref{fig:fig5}). % do_precip.R
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
-This difference was less pronounced for the mean value of RQ (Figure~\ref{fig:fig5}, top).
+This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7). 
+Moreover, effects differed between individual compounds and are provided in the Supplemental Table~S4.
 
@@
 Both increased during summer months and decreased towards the end of the year.
-There was a 2.5-fold higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%).
-The differences were less pronounced for the mean value of RQ and with less precision (Figure~\ref{fig:fig5}, bottom). 
+There was a 2.5-fold higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%) (Figure~\ref{fig:fig5}).
+The differences were less pronounced for the mean value of RQ and with less precision  (see Supplemental Figure~S7, left).
 Individual compounds showed different temporal patterns (see Supplemental Table~S4).
@@
 \begin{figure}[ht]
-  \includegraphics[width=0.8\textwidth]{figure5.pdf}
-  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 22 modelled compounds.
-Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$).
-  Coefficients are shown on the link scale (see Eq.~\ref{eqn:eqn4}).
-  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
+  \includegraphics[width=3.33in]{figure5.pdf}
+  \caption{Summarised model predictions for the probability to exceed LOQ throughout the year. Black points indicate the probabilities at 0.1 mm precipitation (and their 95\% CI). 
+  Orange points indicated the probabilities at 15 mm precipitation.
+  Probabilities have been summaries from a meta-analysis of the 22 modelled compounds.
+  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7.
   }
@@
 \begin{figure}[ht]
-  \includegraphics[width=0.6\textwidth]{figure6.pdf}
-  \caption{15 compounds with the highest risk quotients in small streams.
+  \includegraphics[width=3.33in]{figure6.pdf}
+  \caption{15 compounds with the highest observed risk quotients in small streams.
 Non-detects are not shown due to the logarithmic axis.
@@
 For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that exceedances are likely underestimated.
-For these compounds a lowering of LOQ through an improvement of chemical analysis is essential for reliable assessment.
+For compounds with low RACs a lowering of LOQ through an improvement of chemical analysis is essential for reliable assessment.
 Moreover, a nationwide assessment would benefit from a harmonised spectrum of analysed compounds between federal states.
@@
 %%% Word count
-% abstract    :  197
-% 4 big(600)  : 2400
-% 2 small(300):  600
-% acknowledg  :   79
+% abstract    :  200
+% 1 big(600)  :  600
+% 5 small(300): 1500
+% acknowledg  :   80
 %              =====
-% subtotal      3276
+% subtotal      2380
 % text body   : 3846

### 2017-02-07 supplement/Supplemental_Materials.tex
More revisions
more revisions

@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width = \textwidth]{smooth_lin.pdf}
+	\caption[Smooth and linear fit]{Smooth and linear fits for the data presented in figure 3 of the main article.}
+	\label{fig:smooth_lin}
+\end{figure}
+
 %% -------------------------------------------------------------------------
@@
 	\label{fig:coefs}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+  \includegraphics[width=0.8\textwidth]{mean_coef.pdf}
+  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6).
+Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$).
+  Coefficients are shown on the link scale.
+  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
+  }
+  \label{fig:mean_coefs}
 \end{figure}

### 2017-02-07 src/do_precip.R

@@ -357,9 +357,9 @@ p_loglik <- ggplot(lliks, aes(y = value - 0.001, x = name, col = model, group = 
   # scale_y_continuous(breaks = NULL) +
   labs(y = 'Likelihood') 
 p_loglik
-ggsave('/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/logliks.pdf',
-       p_loglik,
-       width = 7, height = 8)
+# ggsave('/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/logliks.pdf',
+#        p_loglik,
+#        width = 7, height = 8)
 
 # example model verification:
 # for glyphosate
@@ -382,10 +382,11 @@ p_ftd_vs_obs <- ggplot(pdat, aes(x = rq, y = ftd)) +
   scale_fill_viridis('n') +
   labs(x = 'Observed RQ', y = 'Estimated RQ') +
   ggtitle(label = 'Observed vs. Fitted', subtitle = 'Glyphosate, only RQ > 0 shown, n = 1389')
+p_ftd_vs_obs
 # 
-ggsave('/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/pftdvsobs.pdf',
-       p_ftd_vs_obs,
-       width = 8, height = 7)
+# ggsave('/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/pftdvsobs.pdf',
+#        p_ftd_vs_obs,
+#        width = 8, height = 7)
 
 
 #  ------------------------------------------------------------------------
@@ -615,10 +616,10 @@ exp(resmd$est[resmd$term ==  'mu.seasonQ1'])
 exp(resmd$est[resmd$term ==  'mu.seasonQ2'])
 
 int <- resmd$est[resmd$term ==  'mu.seasonQ2']
-# prop of exceeding loq in q2 at 0.1 mm
+# predictions in q2 at 0.1 mm
 (zerop1 <- exp(int + foo(0.1)*resmd$est[resmd$term ==  'mu.log_precip_1']))
-# prop of exceeding loq in q2 at 10 mm
-(ten <- exp(int + foo(10)*resmd$est[resmd$term ==  'nu.log_precip_1']))
+# predictions in q2 at 15 mm
+(ten <- exp(int + foo(15)*resmd$est[resmd$term ==  'nu.log_precip_1']))
[…]

### 2017-02-07 src/do_sizeagri.R

@@ -355,5 +355,8 @@ p <- ggplot() +
   mytheme +
   xlab('') +
   ylab('Mean no. RAC exceedances') +
-  ylim(c(0, 1.3))
+  coord_cartesian(ylim = c(0, 1.3))
 p
+ggsave(file.path(prj, "supplement/smooth_lin.pdf"),
+       p, width = 7, height = 7/1.6,
+       units = 'in', dpi = 300, scale = 1)

### 2017-02-08 article.tex
revised up to discussion

@@
 \subsection{Statistical analyses}
-As outlined in the introduction, we expected non-linear responses to agricultural land use and catchment size and search for potential thresholds.
+As outlined in the introduction, we expected non-linear responses to agricultural land use and catchment size and searched for potential thresholds (defined as abrupt changes).
 Therefore, we used generalised additive models (GAM) to establish relationships \citep{fewster_analysis_2000}.
@@
 $\beta_0$ is the intercept and $f_1$ and $f_2$ are smoothing functions using penalized cubic regression splines \citep{wood_generalized_2006, wood_fast_2011}. 
-The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts (sampling interval and analysed compound spectrum) at a site and is equivalent to modelling the rate of exceedances.
+The number of measurements per site ($n_i$) was used as an offset to account for differences in sampling efforts  at a site (in terms of number of samples and analysed compounds) and is equivalent to modelling the rate of exceedances.
 We used point-wise 95\% Confidence Intervals (CI) of the first derivative of the fitted smooth to identify regions of statistically significant changes.
@@
 
-To account for temporal autocorrelation and differences between federal states we used $site$ nested within $state$ as random intercepts.
+To account for differences between federal states we used $site$ nested within $state$ as random intercepts.
 We implemented this model using the gamlss package \cite{stasinopoulos_generalized_2007}.
@@
 \subsection{Effect of precipitation on pesticide risk}
-The spatio-temporal intersection revealed that most samples were taken during periods of low precipitation.
-For example, only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day that may lead to run-off (Supplemental Figure~S6).
-
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and RQ.
@@
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
-This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7). 
-Moreover, effects differed between individual compounds and are provided in the Supplemental Table~S4.
+This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7, topleft). 
+Moreover, effects differed between individual compounds (see Supplemental Table~S4).
 
@@
 
-\subsection{Pesticide risk in small streams}
+\subsection{Current pesticide risks in small streams}
 We found RAC exceedances in 25.5\% of sampling sites and RQ > 0.1 in 54\% of sites. 

### 2017-02-10 article.tex
first draft revisions

@@
 
-\todo{überarbeiten}
 % Total words: 199
@@
 
-Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to bigger streams \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
+Small streams comprise a major fraction of streams \citep{nadeau_hydrological_2007}, accommodate a higher proportion of biodiversity compared to larger streams \citep{davies_comparison_2008, biggs_report_2014} and play an important role in the recolonization of disturbed downstream reaches \citep{liess_analyzing_2005, orlinskiy_forested_2015}.
 Nevertheless, a clear definition of small streams in terms of catchment or stream size is currently lacking \citep{lorenz_specifics_2016}. 
@@
 We expect that the landscape is a determinant of measured pesticide concentrations.
-Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with high proportion of agriculture. 
-Moreover, if agricultural land use is a main source for pesticides in streams, we expect that exceedances drop to zero if there is not agricultural land use in the catchment. 
+Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with a high proportion of agriculture. 
+Moreover, if agricultural land use is the main source for pesticides in streams, we expect that exceedances drop to zero if there is not agricultural land use in the catchment. 
 Moreover, these relationships may show thresholds that could be to define reference streams without pollution for a future monitoring. 
@@
 We characterised pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
-RACs are derived during pesticide authorisation as part of the ecological risk assessment.
-No unacceptable ecological effects are expected if the environmental concentration remains below this concentration.
+RACs are derived during pesticide authorisation as part of the ecological risk assessment (ERA).
+According to the goals of ERA, exceedances of RACs should not occur after pesticide authorisation. 
+No unacceptable ecological effects are expected if the environmental concentration remains below the RAC. 
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
@@
 \begin{split}
-\log(\mu_{i}) = log~prec_{0 i} + log~prec_{-1 i} + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}\\
-logit(\nu_{i}) = log~prec_{0 i} + log~prec_{-1 i} + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}
+\log(\mu_{i}) = log(prec_{0 i}) + log(prec_{-1 i}) + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}\\
+logit(\nu_{i}) = log(prec_{0 i}) + log(prec_{-1 i}) + Q1_{i} + Q2_{i}+Q3_{i}+Q4_{i}
 \end{split}
@@
 To summarise the coefficients across the 22 modelled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 22 compounds.
-The results of individual compounds are provided in the Supplemental Table~S4 and Figure~S7.
+The results of individual compounds are provided in the Supplemental Table~S4 and Figures~S6 and S7.
 The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.
@@
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
-This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7, topleft). 
+This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7, top left). 
 Moreover, effects differed between individual compounds (see Supplemental Table~S4).
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analysed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset and thus it might nonetheless represent a sample covering all types of small streams in Germany.
+Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset (see Supplemental Figure~S3) and thus it might nonetheless represent a major fraction of small stream types in Germany.
+
+Nevertheless, the differences in monitoring efforts among states hamper a spatial interpolation. 
+We accounted for differences in sampling efforts per site, by taking the total number of measurement into account. 
+However, we acknowledge that this takes only partly differences between sampling intervals into account which might influence our results \citep{stehle_probabilistic_2013,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
+Moreover, it is known that differences in analytical quality might influence estimated effects \citep{Martin_Eberle_Nakagaki}.
+However, the used model (Equation~\ref{eqn:eqn3}) takes explicitly LOQs and differences therein into account.
+
 For Thiacloprid and Chlorpyrifos the LOQs were above the RAC, which means that exceedances are likely underestimated.
@@
 % agriculture
-We found a strong influence of agriculture on the pollution of streams.
-Above 28\% agriculture within a catchment, it is likely that a RAC will be exceeded, with a further increase in entirely agricultural catchments (above 75 \% agriculture).
-To our knowledge, this is the first study investigating such thresholds of pesticide risk.
-\todo{more dicussion here?;drop shows the agri is important; compare to linear fit}
+As hypothesised, we found a positive relationship between agricultural land use and the number of RAC exceedances.
+Especially, we found a statistically significant drop below 28\% of agricultural land use (Figure~\ref{fig:fig4}, left).
+This drop indicates that agricultural land use might be a major contributor to the observed exceedances.
+The absence of such a drop would have indicated that there are exceedances that cannot be explained by agricultural land use, like inputs from urban land uses.
 
@@
 \subsection{Effect of precipitation on pesticide risk}
-Our results revealed that pesticide sampling for chemical monitoring in Germany is mainly performed when no precipitation occurs. 
-Nevertheless, we found a 36\% higher RQ if samples were taken after rainfall events. 
-Samples taken on the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
-This could be explained by the sampling preceding the rainfall event and the delay between the start of a rain event and the peak in discharge or runoff. 
+We found a 36\% higher RQ if samples were taken after rainfall events, that correspond to the hypothesis that pesticides enter streams via run-off.
+However, samples taken on the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
+This discrepancy could be explained by a sampling preceding the rainfall event because the temporal resolution of our dataset was 1 day.
+Additionally, this might be also explained by a delay between the start of a rain event and the peak in discharge or runoff.
+
 The effects of precipitation were more pronounced for the probability to exceed LOQ, with smaller effect sizes for the absolute value of RQ. 
@@
 We found the highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
-This annual pattern coincides with the main application season for pesticides in Central Europe.
+This annual pattern coincides, as expected, with the main application season for pesticides in Central Europe.
 Nevertheless, there are compound-specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left). 
 For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
+Moreover, compound properties, like half-life or water solubility, might influence compound dynamics. 
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics. 
@@
 % subtotal      2380
-% text body   : 3846
+% text body   : 3912
 % ======================================
-% subtotal    : 7122
-% - captions  :  189
+% subtotal    : 6290
+% - captions  :  200
 % =========================================
-% total       : 6933
+% total       : 6090
 

### 2017-02-10 supplement/Supplemental_Materials.tex
revised up to discussion
first draft revisions

@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{fig_map}
+	\caption{Spatial distribution of the 2,301 small stream sampling sites.
+Colour codes different federal states (see Supplemental Table S1 for abbreviations).
+Overlayed are ecoregions.
+ER4: Alps (no label), ER8: Western highlands,
+	ER9: Central highlands, ER 13: Western plains , ER14: Central plains. }
+	\label{fig:ecoreg}
+\end{figure}
+
 % \begin{figure}[ht]
@@
 %% -------------------------------------------------------------------------
-\chapter{Thresholds for agricultural land use and catchment size}
+\chapter{Influence of agricultural land use and catchment size}
 \begin{figure}[ht]
@@
 \chapter{Effect of precipitation and season on RQ}
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width = 0.8\textwidth]{precip}
-	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at the day before sampling.}
-	\label{fig:precip}
-\end{figure}
+% \begin{figure}[ht]
+% 	\centering
+% 	\includegraphics[width = 0.8\textwidth]{precip}
+% 	\caption[Distribution of precipitation at sampling occasions.]{Distribution of precipitation at sampling occasions. top: at sampling date. bottom: at the day before sampling.}
+% 	\label{fig:precip}
+% \end{figure}
 

### 2017-02-10 src/do_overview.R

@@ -135,6 +135,7 @@ ggsave(file.path(prj, 'supplement/temporal.pdf'), p_temp, width = 20, height = 7
 psm_sites_p <- as.data.frame(psm_sites)
 coordinates(psm_sites_p) <- ~easting + northing
 proj4string(psm_sites_p) <- CRS("+init=epsg:31467")
+bb <- psm_sites_p@bbox
 psm_sites_p <- spTransform(psm_sites_p, CRS('+init=epsg:4326'))
 psm_sites_bb <- psm_sites_p@bbox
 psm_sites_p <- cbind(coordinates(psm_sites_p), psm_sites_p@data)
@@ -142,7 +143,7 @@ psm_sites_p$state_ab <- gsub('(.*?)_.*', '\\1', psm_sites_p$site_id)
 setDT(psm_sites_p)
 
 # get administrative borders
-adm1 <- raster::getData('GADM', country = 'DE', level = 1)
+adm1_r <- adm1 <- raster::getData('GADM', country = 'DE', level = 1)
 adm1 <- fortify(adm1)
 p_map <- ggplot() +
   geom_polygon(data = adm1, aes(x = long, y = lat, group = group), fill = "grey90") +
@@ -163,13 +164,79 @@ ggsave(file.path(prj, "figure1.pdf"),
 # phd <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_thesis/chapters/smallstreams/'
 # ggsave(file.path(phd, "figure1.pdf"),
 #        p_map, width = 176, height = 150, units = 'mm')
-#        
+#
 # for defense
 # def <- '/home/edisz/Documents/work/research/projects/2016/1PHD/phd_defense/figs/'
 # p <- p_map +
 #   theme(legend.position = 'none')
 # ggsave(file.path(def, "map_phch.pdf"),
 #        p, width = 176, height = 150, units = 'mm')
+#
+
+
+# with ecoregions
+eco_reg <- readShapePoly('/home/edisz/Documents/work/research/projects/2016/4BFG/Project/data/spatial/wfd_ecoregions/Ecoregions_31467.shp')
+proj4string(eco_reg) <- CRS("+init=epsg:31467")
+
+psm_sites_b <- as.data.frame(psm_sites)
+coordinates(psm_sites_b) <- ~easting + northing
+proj4string(psm_sites_b) <- CRS("+init=epsg:31467")
+
+
[…]

### 2017-02-13 article.tex
coauthor comments

@@
 \usepackage{lineno}
-\usepackage{todonotes}
+% \usepackage{todonotes}
 
@@
 The exceedances of risk thresholds dropped 3.7-fold at sites with no agriculture, indicating that agricultural land use is a major contributor of pesticides in streams.
-Precipitation increased detection probability by 43\% and concentrations were the highest during summer months. 
+Precipitation increased detection probability by 43\% and concentrations were the highest from April to June. 
 RACs were exceeded in 26\% of streams.
@@
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity. 
-To reflect peak concentrations, current pesticide monitoring needs to be refined.
+To reflect peak concentrations, current pesticide monitoring needs refinement.
 
-% Total words: 199
 \end{abstract}
@@
 
-Two recent studies indicate that pesticides concentrations in streams might threaten freshwater biodiversity in the European union.
+Two recent studies indicate that pesticide concentrations in streams might threaten freshwater biodiversity in the European union.
 \citet{malaj_organic_2014} analysed data supplied to the European Union (EU) in the context of the Water Framework Directive (WFD) and showed that almost half of European water bodies are at risk from pesticides.
-\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications and considerable exceedances of regulatory acceptable concentrations (RAC).
+\citet{stehle_pesticide_2015} compiled 1,566 measured concentrations of 23 insecticides in the EU from scientific publications and found considerable exceedances of regulatory acceptable concentrations (RAC).
 However, these studies reflect only a small amount of potentially available data (173 sites in predominantly mid-sized and large rivers in \citet{malaj_organic_2014} and 138 measurements in \citet{stehle_pesticide_2015}), and it is unclear how representative they are for Germany.
@@
 The aim of this study is to identify drivers and dynamics of pesticide concentrations in streams.
-To achieve this, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany in order to identify drivers and dynamics of pesticide concentrations. 
-We expect that the landscape is a determinant of measured pesticide concentrations.
-Because a major fraction of pesticides is applied to agricultural fields, we hypothesised highest concentrations and possible exceedances of RACs in streams with a high proportion of agriculture. 
-Moreover, if agricultural land use is the main source for pesticides in streams, we expect that exceedances drop to zero if there is not agricultural land use in the catchment. 
-Moreover, these relationships may show thresholds that could be to define reference streams without pollution for a future monitoring. 
-Given their low dilution potential and direct adjacency to agricultural fields we expected that small streams show highest pesticide concentrations. 
-However, also the timing of sampling may influence measured concentrations:
-A sampling directly after a precipitation might show higher concentrations because of run-off.
-Furthermore, pesticides are not applied throughout the whole year and we expected highest concentrations during the main growing season.
+To achieve this, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany and examined four hypotheses:
+1) A major fraction of pesticides is applied to agricultural fields.
+Therefore, the highest concentrations and possible exceedances of RACs should be found in streams with a high proportion of agricultural land use in a catchment. 
+Moreover, if agricultural land use was indeed the main source for pesticides in streams, we expect that RAC exceedances drop to negligible levels if there is no agricultural land use in the catchment.
+Additionally, these relationships may unravel thresholds because of non-linearity which may guide the definition of reference streams without pesticide pollution in future monitoring. 
+2) Given a higher dilution potential, larger streams should show lower concentrations.
+3) However, also the timing of sampling may influence measured concentrations,
+as local and regional studies reported higher pesticide concentrations after precipitation events \citep{liess_determination_1999,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}. 
+Therefore, highest concentrations should be found after precipitation. 
+4) Furthermore, pesticides are not applied throughout the whole year and highest concentrations should be found during the main growing season.
 Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.
@@
 RACs are derived during pesticide authorisation as part of the ecological risk assessment (ERA).
-According to the goals of ERA, exceedances of RACs should not occur after pesticide authorisation. 
+According to the goals of ERA, exceedances of RACs should not occur after pesticide authorisation \citep{stehle_pesticide_2015}. 
 No unacceptable ecological effects are expected if the environmental concentration remains below the RAC. 
@@
 $\nu_i$ denotes the probability of a measurement i being above LOQ and $f_{Gamma}$ denotes the gamma function and is used for values equal to or greater LOQ, with $\mu$ being the mean and $\sigma$ the standard deviation of RQ.
-We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1-Q4$) as linear predictors for $\mu$ and $\nu$.
+We used the $log(x+0.05)$ transformed precipitation at sampling date ($log~prec_0$) and the day before ($log~prec_{-1}$), as well as quarters of the year ($Q1$: Jan-Mar, $Q2$: Apr-Jun, $Q3$: Jul-Sep, $Q4$: Oct-Dec) as linear predictors for $\mu$ and $\nu$.
 We used appropriate link functions for $\mu$ and $\nu$ and assumed $\sigma$ to be constant. 
@@
 The non-linear model showed, that below 28\% agriculture the mean number of RAC exceedances dropped statistically significant 3.7-fold from 0.39 (28\% agriculture within the catchment) to 0.10 (no agriculture) (Figure~\ref{fig:fig4}, left).
-\todo{discussion anpassen}.
 Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
@@
 The first quarter showed the lowest RQ and probability of exceeding LOQ.
-Both increased during summer months and decreased towards the end of the year.
+Both increased in $Q2$ and decreased towards the end of the year.
 There was a 2.5-fold higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%) (Figure~\ref{fig:fig5}).
@@
   \caption{Summarised model predictions for the probability to exceed LOQ throughout the year. Black points indicate the probabilities at 0.1 mm precipitation (and their 95\% CI). 
-  Orange points indicated the probabilities at 15 mm precipitation.
-  Probabilities have been summaries from a meta-analysis of the 22 modelled compounds.
+  Orange points indicate the probabilities at 15 mm precipitation.
+  Probabilities have been summarised from a meta-analysis of the 22 modelled compounds.
   Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7.
@@
 Beside large differences in the spatial distribution and quantity of sampling sites (Figure~\ref{fig:fig1}), the spectrum of analysed compounds (Figure~\ref{fig:fig2}) and the quality of chemical analyses differed between states.
-Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} were covered by the presented dataset (see Supplemental Figure~S3) and thus it might nonetheless represent a major fraction of small stream types in Germany.
-
-Nevertheless, the differences in monitoring efforts among states hamper a spatial interpolation. 
-We accounted for differences in sampling efforts per site, by taking the total number of measurement into account. 
-However, we acknowledge that this takes only partly differences between sampling intervals into account which might influence our results \citep{stehle_probabilistic_2013,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
-Moreover, it is known that differences in analytical quality might influence estimated effects \citep{Martin_Eberle_Nakagaki}.
-However, the used model (Equation~\ref{eqn:eqn3}) takes explicitly LOQs and differences therein into account.
+Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} and all major stream types were covered by the data set.
+The unequal distribution of sampling sites and in sampling strategies hamper inference on the total population of small streams in Germany.
+We accounted for differences in sampling efforts per site by including the total number of measurement into the statistical models. 
+However, we acknowledge that additional differences such as sampling frequency and temporal distribution of the sampling might incur bias between states  \citep{stehle_probabilistic_2013,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
+Consequently, we did not compare the results between states.
+Moreover, it is known that differences in analytical quality can influence estimated effects \citep{Martin_Eberle_Nakagaki, Schreiner_Szocs_Bhowmik_Vijver_Schafer_2016}.
+However, the used model (Equation~\ref{eqn:eqn3}) explicitly accounts for LOQs and differences therein.
 
@@
 Especially, we found a statistically significant drop below 28\% of agricultural land use (Figure~\ref{fig:fig4}, left).
-This drop indicates that agricultural land use might be a major contributor to the observed exceedances.
-The absence of such a drop would have indicated that there are exceedances that cannot be explained by agricultural land use, like inputs from urban land uses.
+This drop indicates that agricultural land use might be a major contributor to the observed RAC exceedances.
+The absence of such a drop would have indicated that other inputs such as from urban gardening are relevant contributors to RAC exceedances.  
+This drop would have been missed by linear modeling (Supplemental Figure~S5).
 
@@
 \subsection{Effect of precipitation on pesticide risk}
-We found a 36\% higher RQ if samples were taken after rainfall events, that correspond to the hypothesis that pesticides enter streams via run-off.
+We found a 36\% higher RQ if samples were taken after rainfall events, which conforms to the hypothesis that run-off is a major input path on the large scale.
 However, samples taken on the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
 This discrepancy could be explained by a sampling preceding the rainfall event because the temporal resolution of our dataset was 1 day.
-Additionally, this might be also explained by a delay between the start of a rain event and the peak in discharge or runoff.
+Additionally, this might be explained by a delay between the start of a rain event and the peak in discharge or runoff.
 
@@
 Overall, our results indicate that current pesticide monitoring relying on grab sampling, largely disconnected from precipitation events, underestimates pesticide risks.
-Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation of risks, especially for small water bodies \citep{lorenz_specifics_2016}.
+Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation of risks.
+Our results demonstrate that future monitoring of small water bodies should also capture precipitation events, which is in agreement with other studies \citet{lorenz_specifics_2016}.
 
-We found the highest the probability of exceeding LOQ during summer (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
+We found the highest the probability of exceeding LOQ from April to June (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
 This annual pattern coincides, as expected, with the main application season for pesticides in Central Europe.
@@
 % subtotal      2380
-% text body   : 3912
+% text body   : 3962
 % ======================================
-% subtotal    : 6290
+% subtotal    : 6342
 % - captions  :  200
 % =========================================
-% total       : 6090
+% total       : 6142
 

### 2017-02-13 src/do_overview.R

@@ -196,9 +196,7 @@ int_df$area_reg <- gArea(int, byid = TRUE)
 setDT(int_df)
 
 # areal proportion of ecoregions in D and in samples
-props_eco #<-
-  int_df[ , sum(area_reg), by = AREA_ID]
-  setDT(aggregate(area_reg ~ AREA_ID, data = int_df, function(y) sum(y) / sum(int_df$area_reg)))
+props_eco <- setDT(aggregate(area_reg ~ AREA_ID, data = int_df, function(y) sum(y) / sum(int_df$area_reg)))
 
 props_samples <- setDT(data.frame(psm_sites_b))
 props_samples <- props_samples[ , list(pp = length(site_id) / nrow(props_samples)),

### 2017-02-14 article.tex
few more changes from rs

@@
 
-The aim of this study is to identify drivers and dynamics of pesticide concentrations in streams.
+The aim of this study is to identify drivers and dynamics of pesticide concentrations in streams on large spatial scales.
 To achieve this, we compiled and analysed large-scale pesticide monitoring data from small streams in Germany and examined four hypotheses:
 1) A major fraction of pesticides is applied to agricultural fields.
-Therefore, the highest concentrations and possible exceedances of RACs should be found in streams with a high proportion of agricultural land use in a catchment. 
-Moreover, if agricultural land use was indeed the main source for pesticides in streams, we expect that RAC exceedances drop to negligible levels if there is no agricultural land use in the catchment.
-Additionally, these relationships may unravel thresholds because of non-linearity which may guide the definition of reference streams without pesticide pollution in future monitoring. 
-2) Given a higher dilution potential, larger streams should show lower concentrations.
+Therefore, we hypothesised that the most frequent exceedances of RACs occur in streams with a high proportion of agricultural land use in the catchment. 
+If agricultural land use was indeed the main source for pesticides in streams, the RAC exceedances should drop to negligible levels in the absence of agricultural land use in the catchment. 
+Given this drop we expected a non-linear relationship between exceedances and agricultural land use.
+In case of a non-linear relationship, our analyses might guide the definition of reference streams without pesticide pollution in future monitoring.
+
+2) We hypothesised that an increase in stream size is associated with an decrease in RAC exceedances, given a higher dilution potential. 
 3) However, also the timing of sampling may influence measured concentrations,
 as local and regional studies reported higher pesticide concentrations after precipitation events \citep{liess_determination_1999,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}. 
-Therefore, highest concentrations should be found after precipitation. 
-4) Furthermore, pesticides are not applied throughout the whole year and highest concentrations should be found during the main growing season.
+Therefore, we hypothesised the highest RAC exceedances to be found after precipitation events. 
+4) Pesticides are not applied throughout the whole year and highest RAC exceedances should be found during the main growing season.
 Finally, we quantified the current risks from pesticides in small streams in Germany and the compounds accountable for the risk.
@@
 \subsection{Overview on the compiled dataset}
-The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive available for Germany.
+The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive for Germany.
 Similar nationwide datasets have been compiled for the Netherlands \citep{vijver_spatial_2008}, Switzerland \citep{munz_pestizidmessungen_2011} and the United States \citep{stone2014pesticides}.
@@
 Despite the outlined differences between states, all ecoregions occurring in Germany \citep{illies1978limnofauna,abell2008freshwater} and all major stream types were covered by the data set.
-The unequal distribution of sampling sites and in sampling strategies hamper inference on the total population of small streams in Germany.
+The unequal distribution of sampling sites and the different sampling strategies hamper inference on the total population of small streams in Germany.
 We accounted for differences in sampling efforts per site by including the total number of measurement into the statistical models. 
-However, we acknowledge that additional differences such as sampling frequency and temporal distribution of the sampling might incur bias between states  \citep{stehle_probabilistic_2013,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
+However, we acknowledge that additional differences such as sampling frequency and temporal distribution of the samplings might incur bias between states  \citep{stehle_probabilistic_2013,Xing_Chow_Rees_Meng_Li_Ernst_Benoy_Zha_Hewitt_2013}.
 Consequently, we did not compare the results between states.
 Moreover, it is known that differences in analytical quality can influence estimated effects \citep{Martin_Eberle_Nakagaki, Schreiner_Szocs_Bhowmik_Vijver_Schafer_2016}.
-However, the used model (Equation~\ref{eqn:eqn3}) explicitly accounts for LOQs and differences therein.
+However, the used model (Equation~\ref{eqn:eqn3}) explicitly accounted for LOQs and differences therein.
 
@@
 Especially, we found a statistically significant drop below 28\% of agricultural land use (Figure~\ref{fig:fig4}, left).
-This drop indicates that agricultural land use might be a major contributor to the observed RAC exceedances.
-The absence of such a drop would have indicated that other inputs such as from urban gardening are relevant contributors to RAC exceedances.  
-This drop would have been missed by linear modeling (Supplemental Figure~S5).
+This drop indicates that agricultural land use is a major contributor to the observed RAC exceedances.
+We note that this drop would have been missed by linear modeling (Supplemental Figure~S5). 
+The absence of such a drop would have suggested that other inputs such as from urban gardening are relevant contributors to RAC exceedances.
 
 % size
-We did not find a statistically significant relationship between pesticide pollution and catchment size. 
+We did not find a statistically significant relationship between pesticide pollution and catchment size (Hypothesis 2). 
 However, previous studies showed that small streams are more polluted than bigger streams \citep{schulz_field_2004,stehle_pesticide_2015,knauer_pesticides_2016}.
@@
 We found the highest the probability of exceeding LOQ from April to June (10\% for Q2) and lowest in the first quarter of the year (4\%, Figure~\ref{fig:fig5}, bottom right).
-This annual pattern coincides, as expected, with the main application season for pesticides in Central Europe.
+This annual pattern coincides, as hypothesised, with the main application season for pesticides in Central Europe.
 Nevertheless, there are compound-specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left). 

### 2017-02-20 article.tex
finalized revision

@@
 Precipitation increased detection probability by 43\% and concentrations were the highest from April to June. 
-RACs were exceeded in 26\% of streams.
-We found the highest exceedances for neonicotinoid insecticides. 
+RACs were exceeded in 26\% of streams, with the highest exceedances found for neonicotinoid insecticides. 
 We conclude that pesticides from agricultural land use are a major threat to small streams and their biodiversity. 
@@
 Therefore, we hypothesised that the most frequent exceedances of RACs occur in streams with a high proportion of agricultural land use in the catchment. 
-If agricultural land use was indeed the main source for pesticides in streams, the RAC exceedances should drop to negligible levels in the absence of agricultural land use in the catchment. 
+If agricultural land use was indeed the main source for pesticides in streams, the RAC exceedances should drop to negligible levels in the absence of agricultural land use in their catchments. 
 Given this drop we expected a non-linear relationship between exceedances and agricultural land use.
 In case of a non-linear relationship, our analyses might guide the definition of reference streams without pesticide pollution in future monitoring. 
-
 2) We hypothesised that an increase in stream size is associated with an decrease in RAC exceedances, given a higher dilution potential. 
@@
 \subsection{Data compilation}
-We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names) for 2005 to 2015.
+We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names) for the years 2005 to 2015.
 We homogenised and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
@@
 We characterised pesticide pollution using regulatory acceptable concentrations (RAC) \citep{brock_linking_2010}.
-RACs are derived during pesticide authorisation as part of the ecological risk assessment (ERA).
-According to the goals of ERA, exceedances of RACs should not occur after pesticide authorisation \citep{stehle_pesticide_2015}. 
-No unacceptable ecological effects are expected if the environmental concentration remains below the RAC. 
+RACs are derived during pesticide authorisation as part of the environmental risk assessment (ERA).
+According to the goals of ERA, exceedances of RACs should not occur after pesticide authorisation \citep{stehle_pesticide_2015} and thus no unacceptable ecological effects are expected if the environmental concentration remains below the RAC.
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
@@
   \includegraphics[width=3.33in]{figure3.pdf}
-  \caption{Distribution of catchment area across the sampling sites.}
+  \caption{Distribution of catchment sizes across the sampling sites.}
   \label{fig:fig3}
@@
 For Thiacloprid and Chlorpyrifos the RAC was equal or less than LOQ, therefore, all detections have a $RQ \ge 1$. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RQ (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RACs (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
 RAC exceedances were found in 14\% of samples with concentrations \textgreater LOQ (and 7.3\% of all samples).
@@
 \subsection{Effect of precipitation on pesticide risk}
-We found a 36\% higher RQ if samples were taken after rainfall events, which conforms to the hypothesis that run-off is a major input path on the large scale.
+We found a 43\% higher mean RQ if samples were taken after rainfall events, which conforms to the hypothesis that run-off is a major entry path way for pesticides into streams on the large scale.
 However, samples taken on the day of a rainfall event showed less risk than samples taken one day after a rainfall event.
@@
 Automatic event-driven samplers \citep{stehle_probabilistic_2013} and passive samplers \citep{fernandez_calibration_2014,moschet_evaluation_2015} may help overcome these shortcomings and provide a better representation of risks.
-Our results demonstrate that future monitoring of small water bodies should also capture precipitation events, which is in agreement with other studies \citet{lorenz_specifics_2016}.
+Our results demonstrate that future monitoring of small water bodies should also capture precipitation events, which is in agreement with other studies, such as \citet{lorenz_specifics_2016}.
 
@@
 Our study suggests that pesticide risks display compound specific spatio-temporal dynamics. 
-Currently, little is known about these and further research on those might provide useful information for future ecological risk assessment. 
+Currently, little is known about these and further research on those might provide useful information for future environmental risk assessment. 
 For example, the sensitivity of organisms is often life stage dependent \citep{hutchinson1998analysis} and knowledge on temporal dynamics could inform on concurrent exposure to multiple pesticides, as well as assist to parameterise toxicokinetic and toxicodynamic models \citep{ashauer2016modelling}.

### 2017-02-20 supplement/Supplemental_Materials.tex
coauthor comments
finalized revision

@@
 %%% from do_precip.R
-\newpage
 \input{keeptab}
@@
 Top row: Effect of precipitation at the day before sampling and at day of sampling.
-Bottom row: estimates for the four Quarters.
+Bottom row: estimates for the four quarters of the year.
 Each dot represents one compound (in the order described in table \ref{tab:var_model}).
@@
   \includegraphics[width=0.8\textwidth]{mean_coef.pdf}
-  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6).
+  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter of the year (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6).
 Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$).
   Coefficients are shown on the link scale.
-  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7).
+  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7.
   }
@@
 	\caption[Cumulative distribution of the number sites exceeding RAC.]{Cumulative distribution of sites exceeding RAC.
-Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1.
+Dotted lines indicate fraction of sites exceeding a RQ of 1 and 0.1, respectively.
 23\% of the sites showed no detection of compounds with available RAC values and are not shown due to logarithmic x-axis.}

### 2017-05-01 article.tex
2nd round of revisions

@@
 \author{Eduard Szöcs}
-\affiliation[Institute for Environmental Sciences]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}
-\email{szoecs@uni-landau.de}
-\phone{+49 (0)6341 280 31552}
+\affiliation[University Koblenz-Landau]{Institute for Environmental Sciences, University of Koblenz-Landau, Fortstraße 7, 76829 Landau, Germany}
 
@@
 \author{Ralf B. Schäfer}
-\affiliation[University Koblenz-Landau]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}
+\affiliation[University Koblenz-Landau]{Institute for Environmental Sciences, University of Koblenz-Landau, Fortstraße 7, 76829 Landau, Germany}
+\email{schaefer-ralf@uni-landau.de}
+\phone{+49 (0)6341 280 31536}
+\fax{+49 (0)6341 280 31326}
 
@@
 We investigated the influence of agricultural land use, catchment size, as well as precipitation and seasonal dynamics on pesticide risk taking also concentrations below the limit of quantification into account.
-The exceedances of risk thresholds dropped 3.7-fold at sites with no agriculture, indicating that agricultural land use is a major contributor of pesticides in streams.
+The exceedances of risk thresholds dropped 3.7-fold at sites with no agriculture.
 Precipitation increased detection probability by 43\% and concentrations were the highest from April to June. 
+Overall, this indicates that agricultural land use is a major contributor of pesticides in streams.
 RACs were exceeded in 26\% of streams, with the highest exceedances found for neonicotinoid insecticides. 
@@
 If agricultural land use was indeed the main source for pesticides in streams, the RAC exceedances should drop to negligible levels in the absence of agricultural land use in their catchments. 
-Given this drop we expected a non-linear relationship between exceedances and agricultural land use.
+Given this possible drop we expected a non-linear relationship between exceedances and agricultural land use.
 In case of a non-linear relationship, our analyses might guide the definition of reference streams without pesticide pollution in future monitoring. 
-2) We hypothesised that an increase in stream size is associated with an decrease in RAC exceedances, given a higher dilution potential. 
+2) Based on previouse studies, we hypothesised that an increase in catchment size is associated with an decrease in RAC exceedances \citep{stehle_pesticide_2015,schulz_field_2004}. 
 3) However, also the timing of sampling may influence measured concentrations,
@@
 
-Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams below 10~km\textsuperscript{2} are disproportionally less sampled in current monitoring (Figure~\ref{fig:fig3}), which may be attributed to the missing categorisation in the WFD.
+Given their high abundance in the landscape \citep{nadeau_hydrological_2007} small streams below 10~km\textsuperscript{2} are disproportionally less sampled in current monitoring (Figure~\ref{fig:fig3}).
+This may be attributed to the missing categorisation in the WFD, but also to technical (e.g. sampling at low flow) and natural reasons (e.g. emphemeral streams).
+
 Clearly, there is currently a lack of knowledge on stressor effects on small streams.
@@
 For example, the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude. 
+Although there is a relationship between catchment size and stream size (Supplemental Figure~S11), this might be also influenced by geology and precipitation regime. 
+Therefore, our results cannot be directly translated to stream size.
 
@@
 Moreover, our results show that analysing absolute concentrations and probabilities of LOQ together might deliver valuable insights into risk dynamics.
+The influence of agricultural land use within the catchment area and the coincidence with the growing season indicates that agricultural land use a major contributor of pesticides in streams.
 
@@
 Monitoring data, despite the outlined limitations, provides an opportunity to study large-scale environmental occurrence patterns of pesticides.
-Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modelling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015}or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
+Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modelling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015} or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
 However, the sampling design needs to account for precipitation events to provide robust data. 
-Our results suggest that exceedances of RACs are landscape dependent % = Zusammenhang RAC<->Agriculture
-and therefore, pesticide regulation should account for landscape features. 
+Therefore, non-linear modeling can provide additional insights to risk assessment compared to linear modeling \citep{fewster_analysis_2000}. 
+Our results suggest that exceedances of RACs are landscape dependent and therefore, pesticide regulation should account for landscape features. 
 Moreover, the high exceedances of RACs indicate that greater efforts are needed to describe causal links, which may lead to further developments of the current authorisation procedure.

### 2017-05-01 supplement/Supplemental_Materials.tex
2nd round of revisions

@@
 % increase page counter by 1
-\setcounter{page}{2}
+\setcounter{page}{1}
 
@@
 	\includegraphics[width = 0.8\textwidth]{fig_map}
-	\caption{Spatial distribution of the 2,301 small stream sampling sites.
+	\caption[Spatial distribution of the 2,301 small stream sampling sites.]{Spatial distribution of the 2,301 small stream sampling sites.
 Colour codes different federal states (see Supplemental Table S1 for abbreviations).
@@
 \chapter{Influence of agricultural land use and catchment size}
+
 \begin{figure}[ht]
@@
 	\includegraphics[width = 0.8\textwidth]{ezgagrirac}
-	\caption[Raw data used for the model in equation 2 and Figure 3 of the main article.]{Raw data used for the model in equation 2 and Figure 3 of the main article.
+	\caption[Raw data used for the model in Eqn.
+2 and Fig.
+3 of the main article.]{Raw data used for the model in equation 2 and Figure 3 of the main article.
 Color codes the number of RAC exceedances (on a log-scale).
@@
   \includegraphics[width=0.8\textwidth]{mean_coef.pdf}
-  \caption{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter of the year (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6).
+  \caption[Summarised coefficients (and their 95\% CI) for precipitation and quarter of the year]{Summarised coefficients (and their 95\% CI) for precipitation (top row) and quarter of the year (bottom row) from a meta-analysis of the 22 modelled compounds (see Figure S6).
 Left: coefficients for mean RQ ($\mu$), right: coefficients for probability of exceeding LOQ ($\nu$).
@@
 
-% \bibliographystyle{apalike}      % basic style, author-year citations
-% \bibliography{references}
+%% -------------------------------------------------------------------------
+\chapter{Additional material}
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width = 0.8\textwidth]{width_size}
+	\caption[Relationship between catchment size and stream width.]{Relationship between catchment size and stream width.
+A power function has been fitted to each dataset separately (colored dashed lines) and the combined dataset (black line and equation).
+The gray rectangle marks the estimated catchment size for a width of 1 m (=7 km\textsuperscript{2}).
+Data has been compiled from three different sources (\citet{Fern2015}, \citet{Voss2015} and the federal state of Thuringia)}
+	\label{fig:size_width}
+\end{figure}
+
+\bibliographystyle{apalike}      % basic style, author-year citations
+\bibliography{references}
 

### 2017-05-02 article.tex
comments from rs

@@
 For example, the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude. 
-Although there is a relationship between catchment size and stream size (Supplemental Figure~S11), this might be also influenced by geology and precipitation regime. 
-Therefore, our results cannot be directly translated to stream size.
+Although catchment size and stream size were strongly correlated (Supplemental Figure~S11), other factors such as geology, hydrology and precipitation regime also determine the stream size. 
+Therefore, translating our results to stream size bears uncertainties.
 

### 2017-06-04 article.tex
started revision 3

@@
 The applied pesticides may enter surface waters via spray-drift, edge-of-field run-off or drainage \citep{stehle_probabilistic_2013,schulz_comparison_2001,liess_determination_1999}.
-Once entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
+Once having entered the surface waters they may have adverse effects on biota and ecosystem functioning \citep{schafer_thresholds_2012}. 
 Although it is known that pesticide pollution and its ecological effects increase with the fraction of agricultural land use in the catchment \citep{schulz_field_2004}, the shape of the relationship is unknown and studies on potential thresholds are lacking.
@@
 which is done independently by the federal states in Germany in compliance with the WFD \citep{quevauviller_water_2008}. 
-Despite that these data are providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled.
+Although, these data are providing the opportunity to study pesticide risks and other research questions on a large scale with high spatial density, to date these data have not been compiled.
 
@@
 If agricultural land use was indeed the main source for pesticides in streams, the RAC exceedances should drop to negligible levels in the absence of agricultural land use in their catchments. 
-Given this possible drop we expected a non-linear relationship between exceedances and agricultural land use.
+Given this possible drop we expectedppl a non-linear relationship between exceedances and agricultural land use.
 In case of a non-linear relationship, our analyses might guide the definition of reference streams without pesticide pollution in future monitoring. 
-2) Based on previouse studies, we hypothesised that an increase in catchment size is associated with an decrease in RAC exceedances \citep{stehle_pesticide_2015,schulz_field_2004}. 
+2) Based on previous studies, we hypothesised that an increase in catchment size is associated with an decrease in RAC exceedances \citep{stehle_pesticide_2015,schulz_field_2004}. 
 3) However, also the timing of sampling may influence measured concentrations,
@@
 \subsection{Data compilation}
-We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supplemental Table~S1 for the abbreviations of federal state names) for the years 2005 to 2015.
-We homogenised and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supplemental Figure~S1 for details) \citep{poisot_best_2015}.
+We queried pesticide monitoring data from sampling sites that can be classified as small streams (catchment sizes $\mathrm{< 100~km^2}$ according to the WFD) from all 13 non-city federal states of Germany (see Supporting Information Table~S1 for the abbreviations of federal state names) for the years 2005 to 2015.
+We homogenised and unified all data provided by the federal states into a database and implemented a robust data-cleaning workflow (see Supporting Information Figure~S1 for details) \citep{poisot_best_2015}.
 
@@
 \citet{stehle_pesticide_2015} showed that RAC exceedances reflect a decrease in biodiversity and from this perspective are ecologically relevant indicators. 
-The German Environment Agency (UBA) provided RACs for 107 compounds, including those with the highest detection rates (Supplemental Table~S2). 
+The German Environment Agency (UBA) provided RACs for 107 compounds, including those with the highest detection rates (Supporting Information Table~S2). 
 Based on theses RACs, we calculated Risk Quotients (RQ):
@@
 
-We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 22 compounds, see Supplemental Table~S3 for a list of compounds).
+We fitted this model separately to each compound with a RAC, measured in at least 1000 samples and with more than 5\% of values above LOQ (n = 22 compounds, see Supporting Information Table~S3 for a list of compounds).
 To summarise the coefficients across the 22 modelled compounds we used a random effect meta-analysis for each model coefficient separately \citep{harrison_getting_2011}, resulting in an averaged effect of the 22 compounds.
-The results of individual compounds are provided in the Supplemental Table~S4 and Figures~S6 and S7.
+The results of individual compounds are provided in the Supporting Information Table~S4 and Figures~S6 and S7.
 The meta-analysis was performed using the metafor package \citep{Viechtbauer_2010}.
@@
 These samples were all taken via grab sampling.  
-We found large differences between federal states in the number of sampling sites and their spatial distribution (Figure~\ref{fig:fig1} and Supplemental Table~S1). 
+We found large differences between federal states in the number of sampling sites and their spatial distribution (Figure~\ref{fig:fig1} and Supporting Information Table~S1). 
 The number of small stream sampling sites per state ranged from 1 (Lower Saxonia, NI) to 1139 (North Rhine-Westphalia, NW).
@@
   \includegraphics[width=3.33in]{figure1.pdf}
-  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states (see Supplemental Table~S1 for abbreviations).}
+  \caption{Spatial distribution of the 2,301 small stream sampling sites. Colour codes different federal states (see Supporting Information Table~S1 for abbreviations).}
   \label{fig:fig1}
@@
 
-In total 478 different compounds used as pesticides and their metabolites were measured at least once (Supplemental Table~S2). 
+In total 478 different compounds used as pesticides and their metabolites were measured at least once (Supporting Information Table~S2). 
 Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109). % see overview.R
-Most samples were taken in the months April till October, while fewer samples were taken during winter (see Supplemental Figure~S2).
+Most samples were taken in the months April till October, while fewer samples were taken during winter (see Supporting Information Figure~S2).
 We found substantial differences in the spectra of analysed pesticides between federal states (Figure~\ref{fig:fig2}).
-The number of analysed pesticides per state ranged from 57 (SL) to 236 (RP) (Supplemental Table~S1). 
+The number of analysed pesticides per state ranged from 57 (SL) to 236 (RP) (Supporting Information Table~S1). 
 4\% (=71,113) of all measurements were concentrations above LOQ.
@@
 We found a positive relationship between agricultural land use and the number of RAC exceedances. 
-The non-linear model showed, that below 28\% agriculture the mean number of RAC exceedances dropped statistically significant 3.7-fold from 0.39 (28\% agriculture within the catchment) to 0.10 (no agriculture) (Figure~\ref{fig:fig4}, left).
+The non-linear model revealed, that below 28\% agriculture the mean number of RAC exceedances show a 3.7-fold decrease from 0.39 (28\% agriculture within the catchment) to 0.10 (no agriculture) (Figure~\ref{fig:fig4}, left).
 Catchment size had no statistically significant effect on the number of RAC exceedances (Figure~\ref{fig:fig4}, right).
@@
   \includegraphics[width=6.5in]{figure4.pdf}
-  \caption{Effect of percent agriculture within the catchment (left) and catchment size (right) on the mean number of RAC exceedances per site.
+  \caption{Effect of percent agriculture within the catchment (left) and of catchment size (right) on the mean number of RAC exceedances per site.
 Red line marks statistically significant changes.
@@
 $prec_{0}$ and $prec_{-1}$ increased the probability of exceeding LOQ and RQ.
-In $Q2$ an increase from 0.1~mm to 15~mm of precipitation before sampling ($prec_{-1}$) lead on average to a 43\% higher mean RQ of 0.05 (Supplemental Figure~S7).
+In $Q2$ an increase from 0.1~mm to 15~mm of precipitation before sampling ($prec_{-1}$) lead on average to a 43\% higher mean RQ of 0.05 (Supporting Information Figure~S7).
 The probability to exceed LOQ increases in $Q2$ 1.6-fold from 8.7\% to 13.5\% (Figure~\ref{fig:fig5}). % do_precip.R
 Precipitation before sampling ($prec_{-1}$) had a stronger effect than precipitation during sampling ($prec_{0}$) on the probability of exceeding LOQ. 
-This difference was less pronounced for the mean value of RQ (Supplemental Figure~S7, top left). 
-Moreover, effects differed between individual compounds (see Supplemental Table~S4).
+This difference was less pronounced for the mean value of RQ (Supporting Information Figure~S7, top left). 
+Moreover, effects differed between individual compounds (see Supporting Information Table~S4).
 
@@
 There was a 2.5-fold higher probability of exceeding LOQ in $Q2$ (10.6\%) than in $Q1$ (4.6\%) (Figure~\ref{fig:fig5}).
-The differences were less pronounced for the mean value of RQ and with less precision  (see Supplemental Figure~S7, left).
-Individual compounds showed different temporal patterns (see Supplemental Table~S4).
+The differences were less pronounced for the mean value of RQ and with less precision  (see Supporting Information Figure~S7, left).
+Individual compounds showed different temporal patterns (see Supporting Information Table~S4).
 
@@
   \includegraphics[width=3.33in]{figure5.pdf}
-  \caption{Summarised model predictions for the probability to exceed LOQ throughout the year. Black points indicate the probabilities at 0.1 mm precipitation (and their 95\% CI). 
-  Orange points indicate the probabilities at 15 mm precipitation.
-  Probabilities have been summarised from a meta-analysis of the 22 modelled compounds.
-  Single compound coefficients are provided in Supplemental Table~S4 and Figure~S7.
+  \caption{Summarised model predictions for the probability to exceed LOQ throughout the year. 
+  Black dots indicate the probabilities at 0.1 mm precipitation (and their 95\% CI). 
+  Orange dots indicate the probabilities at 15 mm precipitation.
+  Probabilities were summarised from a meta-analysis of the 22 modelled compounds.
+  Single compound coefficients are provided in Supporting Information Table~S4 and Figure~S7.
   }
@@
 We found RAC exceedances in 25.5\% of sampling sites and RQ > 0.1 in 54\% of sites. 
-In 23\% of sites none of the chemicals, for which RACs were available, were detected (see also Supplemental Figure~S8).
+In 23\% of sites none of the chemicals for which RACs were available were detected (see also Supporting Information Figure~S8).
 Neonicotinoid insecticides and Chlorpyrifos showed the highest RQ (Figure~\ref{fig:fig6}). %do_pollution.R
 For Thiacloprid and Chlorpyrifos the RAC was equal or less than LOQ, therefore, all detections have a $RQ \ge 1$. 
-The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RACs (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supplemental Table~S5).
+The herbicides Nicosulfuron and Diflufenican, as well as the fungicide Dimoxystrobin also showed high exceedances of RACs (26.7, 14.1 and 21.1 \% of measurements~\textgreater~LOQ), see also Supporting Information Table~S5).
 RAC exceedances were found in 14\% of samples with concentrations \textgreater LOQ (and 7.3\% of all samples).
@@
 The highest RQs were observed for Chlorpyrifos (max(RQ) = 220), Clothianidin (max(RQ) = 157), Dimoxystrobin(max(RQ) = 117) and Isoproturon (max(RQ) = 80). 
-Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supplemental Figure~S9).
+Where analysed, metabolites exhibited the highest detection rates (for example, Metazachlor sulfonic acid was detected in 84\% of all samples where it was analysed (n = 3038, see also Supporting Information Figure~S9).
 Glyphosate was the compound with the highest detection rates (41\%, n = 3557 samples), followed by Boscalid (23\%, n = 9886) and Isoproturon (22\%, n = 19112). 
 However, only the latter showed RAC exceedances (Figure~\ref{fig:fig6}).
-In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supplemental Figure~S10).
+In 45.9\% of samples more than one compound was quantified, with a maximum of 54 different compounds in one sample (Supporting Information Figure~S10).
 
@@
 Non-detects are not shown due to the logarithmic axis.
-Numbers on the right give the percentage of values \textgreater LOQ and the total number of samples were the compound was analysed.
+Numbers on the right give the percentage of values \textgreater LOQ and the total number of samples where the compound was analysed.
   }
@@
 \section{Discussion}
-\subsection{Overview on the compiled dataset}
+\subsection{Overview of the compiled dataset}
 The compiled dataset of governmental monitoring data, with a particular focus on small streams, represents currently the most comprehensive for Germany.
@@
 This drop indicates that agricultural land use is a major contributor to the observed RAC exceedances. 
-We note that this drop would have been missed by linear modeling (Supplemental Figure~S5). 
+We note that this drop would have been missed by linear modeling (Supporting Information Figure~S5). 
 The absence of such a drop would have suggested that other inputs such as from urban gardening are relevant contributors to RAC exceedances.
@@
 For example, the gradient of \citet{schulz_field_2004} covered 6 orders of magnitude. 
-Although catchment size and stream size were strongly correlated (Supplemental Figure~S11), other factors such as geology, hydrology and precipitation regime also determine the stream size. 
+Although catchment size and stream size were strongly correlated (Supporting Information Figure~S11), other factors such as geology, hydrology and precipitation regime also determine the stream size. 
 Therefore, translating our results to stream size bears uncertainties.
@@
 Nevertheless, there are compound-specific differences in the annual pattern, which explains the wide CI for the absolute RQ (Figure~\ref{fig:fig5}, bottom left). 
-For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supplemental Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
+For example, the herbicide Diflufenican showed the highest RQ and the highest probability of exceeding LOQ during the winter quarters Q1 and Q4 (Supporting Information Table~S4), which coincides with the application period it is registered for in Germany \citep{bvl_online_2016}.
 Moreover, compound properties, like half-life or water solubility, might influence compound dynamics. 
@@
 % Approval / Risk Assessment
-Monitoring data, despite the outlined limitations, provides an opportunity to study large-scale environmental occurrence patterns of pesticides.
+Monitoring data, despite the outlined limitations, provide an opportunity to study large-scale environmental occurrence patterns of pesticides.
 Furthermore, such nationwide compilations, may not only be used for governmental surveillance, but also to answer other questions, like validation of exposure modelling \cite{knabel_fungicide_2014}, retrospective evaluation of regulatory risk assessment \citep{knauer_pesticides_2016,stehle_pesticide_2015} or occurrences of pesticide mixtures \cite{schreiner_pesticide_2016}.
@@
 \begin{itemize}
-  \item Supplemental\_Materials.pdf : Supplemental Materials (Figures, Tables, Models).
+  \item Supplemental\_Materials.pdf : Supporting Information (Figures, Tables, Models).
 \end{itemize}

### 2017-06-04 supplement/Supplemental_Materials.tex
comments from rs
first draft

@@
 %% -------------------------------------------------------------------------
-\title{Supplemental Materials for: \\ Large scale risks from pesticides in small streams}
-\author{Eduard Szöcs%
-  \thanks{Corresponding author, Email: \texttt{szoecs@uni-landau.de}}}
+\title{Supporting Information for: \\ Large scale risks from agricultural pesticides in small streams}
+\author{Eduard Szöcs}
 \affil[1]{Institute for Environmental Sciences, University of Koblenz-Landau, Germany}
@@
 
-\author[1]{Ralf B. Schäfer}
+\author[1]{Ralf B. Schäfer   
+\thanks{Corresponding author, Email: \texttt{schaefer-ralf@uni-landau.de}}}
+
+\date{}
 
@@
 	\includegraphics[width = 0.8\textwidth]{data_cleaning}
-	\caption[Overview on data cleaning steps.]{Overview on data cleaning steps. After cleaning, data have been stored in a relational spatial PostgreSQL database.}
+	\caption[Overview of data cleaning steps.]{Overview of data cleaning steps. After cleaning, data have been stored in a relational spatial PostgreSQL database.}
 	\label{fig:data_cleaning}
@@
 %% -------------------------------------------------------------------------
-\chapter{Overview on compiled data}
+\chapter{Overview of compiled data}
 % Overview samples
@@
 Color codes the number of RAC exceedances (on a log-scale).
-Grey points denote sites without any exceedance.}
+Grey dots denote sites without any exceedance.}
 	\label{fig:ezgagrirac}
@@
 Each dot represents one compound (in the order described in table \ref{tab:var_model}).
-Coefficients where the CI encompasses zero are shown in gray colour.
+Coefficients where the CI encompasses zero are shown in grey colour.
 Coefficients are shown on the link scale (log for $\mu$ and logit for $\nu$).}
@@
 	\caption[Relationship between catchment size and stream width.]{Relationship between catchment size and stream width.
-A power function has been fitted to each dataset separately (colored dashed lines) and the combined dataset (black line and equation).
-The gray rectangle marks the estimated catchment size for a width of 1 m (=7 km\textsuperscript{2}).
-Data has been compiled from three different sources (\citet{Fern2015}, \citet{Voss2015} and the federal state of Thuringia)}
+A power function was fitted to each dataset separately (colored dashed lines) and the combined dataset (black line and equation).
+The grey rectangle marks the estimated catchment size for a width of 1 m (=7 km\textsuperscript{2}).
+Data was compiled from three different sources: \citet{Fern2015}, \citet{Voss2015} and the federal state of Thuringia.}
 	\label{fig:size_width}
@@
 
-\bibliographystyle{apalike}      % basic style, author-year citations
+\bibliographystyle{spbasic}      % basic style, author-year citations
 \bibliography{references}

### 2017-06-04 supplement/Supplemental_Materials.tex
final draft of revision

@@
 	round,	%(defaultage in the main file and \input ) for round parentheses;
-	authoryear,% (default) for author-year citations;
+	numbers,
 	sort,		% orders multiple citations into the sequence in which they 
 ]{natbib}	
+\renewcommand{\bibname}{References}
 
@@
 
-\bibliographystyle{spbasic}      % basic style, author-year citations
+\bibliographystyle{plainnat}      
 \bibliography{references}
